### 2014-01-19 proposal/proposal.tex
Skeleton for proposal and .gitignore.

\documentclass[10pt]{article}

% =========================================================================
% document style changes
% =========================================================================

\usepackage{amsmath}                    % AMS math packages
\usepackage{amssymb}                    %
\usepackage[]{graphpap}
\usepackage[T1]{fontenc}                % for \mathrm{}
\usepackage{courier}                    % for \texttt{}
\usepackage{bbm}                        % for \mathbbm{1} (indicator function)
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}

%\setlength{\parskip}{\baselineskip}     % skip line following paragraphs
\setlength{\parskip}{0.0in}
\setlength{\parindent}{0.3in}             % Control margins and amount of text
\setlength{\topmargin}{-0.8in}
\setlength{\oddsidemargin}{-.15in}       % changed from {-.15in}
\setlength{\textheight}{9.5in}
\setlength{\textwidth}{6.8in}
\pagestyle{empty}                       % No page numbers

\newcommand{\spc}{\vspace{0.25in}}      % Shortcut commands
\newcommand{\ds}{\displaystyle}         %\newcommand{\ds}[1]{\displaystyle{#1}}
\newcommand{\ra}{\rightarrow}

\begin{document}                        % This is where the document begins

{\LARGE\bf
\begin{tabbing}
\hspace{2.8in} \= \hspace{1.3in} \= \hspace{1.2in} \= \\

%=========    Heading   ==================================================
CS 399 \> PA4 \> Bill Chickering (bchick)\\
\normalsize Jan 18, 2014 \> \> Jamie Irvine (jirvine)
% =========================================================================
\end{tabbing}
}
\vspace{.4in}

The abstract goes here. . . . 
Blah blah blah.  Blah blah blah.  Blah blah blah.  Blah blah blah.  Blah blah
blah.  Blah blah blah.  Blah blah blah.  Blah blah blah.  Blah blah blah.  Blah
blah blah.  Blah blah blah.  Blah blah blah.  Blah blah Blah blah blah.  Blah
blah blah.  Blah blah blah.  Blah blah blah.  Blah blah Blah blah blah.  Blah
blah blah.  Blah blah blah.  Blah blah blah.  Blah blah Blah blah blah.  Blah
blah blah.  Blah blah blah.  Blah blah blah.  Blah blah Blah blah blah.  Blah
blah blah.  Blah blah blah.  Blah blah blah.  Blah blah Blah blah blah.  Blah
blah blah.  Blah blah blah.  Blah blah blah.  Blah blah blah.  Blah blah blah.
Blah blah blah.  Blah blah blah.  blah.  Blah blah blah.  Blah blah blah.  Blah
blah blah.  blah.  Blah blah blah.  Blah blah blah.  Blah blah blah.  blah.
Blah blah blah.  Blah blah blah.  Blah blah blah.  blah.  Blah blah blah.  Blah
blah blah.  Blah blah blah.  blah.  Blah blah blah.  Blah blah blah.  Blah blah
blah.

\section*{Vision}

\section*{Implementation Ideas}

\end{document}

### 2014-01-26 proposal/proposal.tex
Add language from email to Anand.

@@
 
+Shopping for lifestyle items like apparel, shoes, glasses, watches, etc. calls
+for a distinct user experience. The shopper often has a fuzzy idea of what they
+want--Emma wants a black party dress or Paul wants printed T-shirts--so they'd
+like to peruse many similar items before deciding what to purchase. Today, it's
+difficult to achieve this without having 1) a huge product catalog 2) an
+excellent search engine 3) an exceptionally well articulated query. Certainly
+(1) exists. Arguably (2) exists. But (3) is an onerous proposition. Instead,
+one typically wades through a large number of heterogenous items hoping to find
+something they like.
+
+We're debating the details of the user interface, but the shopping app we want
+on our phones would resemble the following. You begin by either entering a
+query or clicking on a highlighted item. You are then shown an item. If you
+like the item and want to see more like it you swipe it to the right. If you
+don't like it, you swipe it to the left. If you really like it, you swipe it
+down into your cart/list. You can swipe through items quickly--less than a
+second per item. As you shop, the system learns in real time about your likes
+and dislikes, showing you items you like with increasing frequency. At any
+time, you can toggle between your search, your cart/list, or even the lists of
+everything you liked or disliked. Click any item anywhere, and you see product
+details along with an option to buy. Lists of items can be posted on Pinterest,
+Tumblr, Facebook, or emailed to yourself or a friend. Or save a list for later
+and continue a shopping session at any time.
+
+We believe a critical ingredient to maximal user-engagement is rapid feedback
+and responsiveness. People need less than a second or two to decide whether
+they're interested in a pair shoes. And every time a user indicates interest or
+disinterest in an item the system learns about the user's immediate desires and
+is therefore better equipped to identify desirable items. Meanwhile, the system
+can leverage user-sessions in aggregate to learn about item popularity and
+item-item associations.  We believe this last idea is key. Learning item-item
+associations from user activity is what our CS 341 work was all about. In that
+project, we combined aggregated user activity along with a particular user's
+click history to improve search results. The app described here would work
+similarly. Importantly, however, this app idea attempts to address a data
+sparsity issue by creating a user interface that by its nature will collect far
+more user activity than a traditional search engine.
+
 \section*{Implementation Ideas}

### 2014-01-26 proposal/proposal.tex
Added Experiment section

@@
 
+\section*{Experiment}
+Unfortunately, we do not have user data on which to test our methods, since the system
+has never been used before. Instead, we will simulate user behavior to get a rough
+estimate of performance. To do this, we make the following assumptions:
+
+\begin{enumerate}
+\item There are a number of universal hidden features that describe each product.
+These features could be something as tangible as color or as intangible as
+`hipness'. Note that like latent topics, these features need not be
+representable by an explicit, expressible concept.
+
+\item Each user has a preference for products based on these hidden features. For
+instance, a user could care a lot that a product is `hip', not care
+at all what color it is, and have a mild preference for slim-fitting products,
+assuming these are all hidden features of products.
+\end{enumerate}
+
+For testing, we simplify the model by assuming that there is only one hidden
+feature. We model this as color. Since the learning algorithm is blind to color
+data, this should be as difficult to determine as a less tangible feature, such
+as style, but color is easier for us to use, since it is metadata that we
+have about each product. Next, we create a number of fake users, defined by their color
+preferences. These preferences are modeled as multinomial distributions. For instance, if
+there are only the three colors red, white, and blue, a user could care deeply
+that their product is red by having a (1, 0, 0) distribution over those colors.
+Or, if the user doesn't care what color their product is, their distribution
+would be (0.33, 0.33, 0.33).
+
+We can then construct a number of users with different distributions, using an
+arbitrary Dirichlet distribution or our creation. This way, we can model the
+phenomenon that preferences are not uniform; many users have the same style
+preferences as each other and some styles are less popular than others.
+
+With this model, we can construct a number of fake users with different
+preference distributions. Based on these distributions, we can
+probabilistically choose which items each user likes and then simulate their
+interaction with the system. Then we can train the system on a number of fake
+users to learn from their behavior. For testing, we then introduce new fake users,
+and measure how well the system can predict the underlying color dtistributions of
+the users. We will aim to increase the accuracy of the predicted distributions and
+reduce the number of shown products needed to learn it.
+
 \end{document}

### 2014-01-27 macys/filterBots.py
Add Macys scraper and proto-LDA stuff.

@@ -1,0 +1,57 @@
+import sqlite3
+
+# params
+botsPerParentCategoryFilter = 1
+botsPerCategoryFilter = 3
+
+# db params
+db_fname = 'macys.db'
+createSessionsTableStmt =\
+    ('CREATE TABLE IF NOT EXISTS Sessions(sessionId INT, productId INT, '
+     'PRIMARY KEY (sessionId, productId))')
+selectCategoriesStmt = 'SELECT distinct Category FROM Categories'
+selectParentCategoriesStmt = 'SELECT distinct ParentCategory FROM Categories'
+insertParentCategoryFilterStmt =\
+    ('INSERT INTO Sessions(sessionId, productId) '
+     'SELECT :sessionId, Id FROM Categories '
+     'WHERE ParentCategory = :ParentCategory GROUP BY Id')
+insertCategoryFilterStmt =\
+    ('INSERT INTO Sessions(sessionId, productId) '
+     'SELECT :sessionId, Id FROM Categories WHERE Category = :Category '
+     'GROUP BY Id')
+
+def main():
+    # connect to db
+    db_conn = sqlite3.connect(db_fname)
+    with db_conn:
+        db_curs = db_conn.cursor()
+
+        # create table if not already exists
+        db_curs.execute(createSessionsTableStmt)
+
+        sessionId = 1
+
+        # do parent category filterbots
+        db_curs.execute(selectParentCategoriesStmt)
+        for row in db_curs.fetchall():
+            parentCategory = row[0]
+            print 'ParentCategory filterBots for: %s' % parentCategory
+            for i in range(botsPerParentCategoryFilter):
+                db_curs.execute(insertParentCategoryFilterStmt,\
+                                (sessionId, parentCategory))
+                print 'sessionId = %d' % sessionId
+                sessionId += 1
+
+        # do category filterbots
+        db_curs.execute(selectCategoriesStmt)
+        for row in db_curs.fetchall():
+            category = row[0]
+            print 'Category filterBots for: %s' % category
+            for i in range(botsPerCategoryFilter):
+                db_curs.execute(insertCategoryFilterStmt,\
+                                (sessionId, category))
+                print 'sessionId = %d' % sessionId
+                sessionId += 1
+
+if __name__ == '__main__':
+    main()

### 2014-01-27 macys/genCatSessions.py

@@ -1,0 +1,87 @@
+#!/usr/local/bin/python
+
+from optparse import OptionParser
+import csv
+import logging
+import random
+
+logging.basicConfig(format='%(asctime)s : %(levelname)s : %(message)s', level=logging.INFO)
+
+def getParser():
+    parser = OptionParser()
+    parser.add_option('-f', '--file', dest='filename', default='categories.csv',
+        help='Input CSV file containing <category, productId> on each line.',
+        metavar='FILE')
+    parser.add_option('-o', '--output', dest='outfilename', default='trainSessions.csv',
+        help='Output CSV file containing <productId, productId, productId, ...> on each line.',
+        metavar='OUT_FILE')
+    parser.add_option('-s', '--session-size', dest='session_size', type="int", default=100,
+        help='Max number of products per session.', metavar='SIZE')
+    parser.add_option('--num-sessions-factor', dest='num_sessions_factor', type="int",
+        default=10, help='NUM_SESSIONS(category) = NUM_PRODUCTS(category)/FACTOR',
+        metavar='FACTOR')
+    return parser
+    
+def randomPop(data):
+	if data != []:
+		pos = random.randrange( len(data) )
+		elem = data[pos]
+		data[pos] = data[-1]
+		del data[-1]
+		return elem
+	else:
+		return None
+
+def generateSessions(category, productIds, session_size,
+                     num_sessions_factor, writer):
+    logging.info('=============================================')
+    logging.info('Generating sessions for %s' % category)
+    num_products = len(productIds)
+    logging.info('len_products = %d' % num_products)
+    num_sessions = 1 + int(num_products/num_sessions_factor)
+    logging.info('num_sessions = %d' % num_sessions)
+    _session_size = int(min(num_products, session_size))
+    logging.info('session_size = %d' % _session_size)
+
+    for i in range(num_sessions):
+        copyIds = list(productIds)
+        sessionIds = []
+        for j in range(_session_size):
+            sessionIds.append(randomPop(copyIds));
+        writer.writerow(sessionIds)
+
+
+def main():
+    parser = getParser()
+    (options, args) = parser.parse_args()
+    random.seed()
+    logging.info('Reading data from %s' % options.filename)
+    with open(options.filename, 'r') as csvIn:
+        reader = csv.reader(csvIn)
+        logging.info('Writing sessions to %s' % options.outfilename)
+        with open(options.outfilename, 'w') as csvOut:
+            writer = csv.writer(csvOut)
+            category = None
+            productIds = None
+            for row in reader:
[…]

### 2014-01-27 macys/lda_test.py

@@ -1,0 +1,94 @@
+#!/usr/local/bin/python
+
+from gensim import corpora, models, similarities
+from gensim.models import ldamodel
+import sqlite3
+import logging
+import pickle
+import csv
+
+# params
+db_fname = 'macys.db'
+inFile = 'trainSessions.csv'
+testFile = 'testSessions.csv'
+corpusFile = 'categorySessions.corp'
+dictFile = 'products.dict'
+ldaFile = 'lda.pickle'
+
+logging.basicConfig(format='%(asctime)s : %(levelname)s : %(message)s', level=logging.INFO)
+
+def getDictionary(filename):
+    try:
+        dictionary = corpora.Dictionary.load_from_text(dictFile)
+        logging.info('Reading dictionary from %s.' % dictFile)
+    except IOError:
+        with open(filename) as csvIn:
+            reader = csv.reader(csvIn)
+            records = [record for record in reader]
+            dictionary = corpora.Dictionary(records)
+            dictionary.save_as_text(dictFile)
+    return dictionary
+            
+def getCorpus(dictionary, filename):
+    logging.info('Reading corpus from %s.' % filename)
+    with open(filename) as csvIn:
+        reader = csv.reader(csvIn)
+        records = [record for record in reader]
+        return [dictionary.doc2bow(record) for record in records]
+
+def getLDA():
+    try:
+        with open(ldaFile, 'r') as f:
+            lda = pickle.load(f)
+            logging.info('Reading LDA model from %s.' % ldaFile)
+    except IOError:
+        dictionary = getDictionary(inFile)
+        corpus = getCorpus(dictionary, inFile)
+        lda = ldamodel.LdaModel(corpus=corpus,
+                                id2word=dictionary,
+                                num_topics=100,
+                                passes=40,
+                                update_every=1,
+                                chunksize=800)
+        pickle.dump(lda, open(ldaFile, 'w'))
+    return lda
+
+def getSessionCats(db_curs, dictionary, session):
+    selectCategoriesStmt = 'SELECT Category FROM Categories WHERE Id = :Id'
+    sessionCats = {}
+    for product in session:
+        productId = dictionary[product[0]]
+        db_curs.execute(selectCategoriesStmt, (productId,))
+        for row in db_curs.fetchall():
+            category = row[0]
+            if category in sessionCats:
+                sessionCats[category] += 1
+            else:
+                sessionCats[category] = 1
+    return sessionCats
[…]

### 2014-01-27 macys/macys.py

@@ -1,0 +1,191 @@
+import urllib
+import urllib2
+from bs4 import BeautifulSoup, NavigableString
+import sqlite3
+import json
+import re
+import unicodedata
+import math
+import sys
+import os
+
+# params
+categories_fname = 'categories.json'
+imageFileTemplate = 'images/%d.jpg'
+metaUrlTemplate = 'http://www1.macys.com/catalog/category/facetedmeta?sortBy=NAME&productsPerPage=%d&pageIndex=%d&parentCategoryId=%d&categoryId=%d'
+productsUrlTemplate = 'http://www1.macys.com/shop/catalog/product/thumbnail/1?edge=hybrid&limit=none&suppressColorSwatches=false&categoryId=%d&ids=%s'
+productsPerPage = 100
+
+# db params
+db_fname = 'macys.db'
+createProductTableStmt =\
+    ('CREATE TABLE IF NOT EXISTS Products(Id INT PRIMARY KEY, Url TEXT, '
+     'ImgFile TEXT, Description TEXT, Prices TEXT, Available INT, '
+     'Time TIMESTAMP)')
+createCategoryTableStmt =\
+    ('CREATE TABLE IF NOT EXISTS Categories(Id INT, ParentCategory TEXT, '
+     'Category TEXT, PRIMARY KEY (Id, ParentCategory, Category), '
+     'FOREIGN KEY(Id) REFERENCES Products(Id))')
+updateAllProductsStmt = 'UPDATE Products SET Available = 0'
+updateProductStmt = 'UPDATE Products SET Available = 1 WHERE Id = ?'
+selectProductStmt = 'SELECT * FROM Products WHERE Id = ?'
+selectCategoryStmt =\
+    ('SELECT * FROM Categories WHERE Id = :Id AND '
+     'ParentCategory = :ParentCategory AND Category = :Category')
+insertProductStmt =\
+    ('INSERT INTO Products(Id, Url, ImgFile, Description, Prices, Available, '
+     'Time) VALUES (:Id, :Url, :ImgFile, :Description, :Prices, 1, '
+     'CURRENT_TIMESTAMP)')
+insertCategoryStmt =\
+    ('INSERT INTO Categories(Id, ParentCategory, Category) '
+     'VALUES (:Id, :ParentCategory, :Category)')
+
+def normalizeStr(s):
+    decodedStr = s.decode("utf8")
+    u = unicode(decodedStr)
+    return unicodedata.normalize('NFKD', u).encode('ascii','ignore')
+
[…]

### 2014-01-27 macys/BayesPredictor.py
Add buildModel.py, LDAPredictor.py, and BayesPredictor.py.

@@ -1,0 +1,111 @@
+#!/usr/local/bin/python
+
+from LDAPredictor import LDAPredictor
+from gensim import corpora, models, similarities
+from gensim.models import ldamodel
+from numpy.random import multinomial
+import logging
+
+class BayesPredictor(LDAPredictor):
+    """An implementation of an LDAPredictor that uses Bayes' rule to generate
+    item predictions.
+    """
+
+    def __init__(self, topn_items=1000, **kwds):
+        """Constructs a new BayesPredictor.
+
+        Note that predictions associated with a topic are limited to the most
+        likely topn_items items.
+        """
+        super(BayesPredictor, self).__init__(**kwds)
+        self.topn_items = topn_items=1000
+
+    def initSession(self):
+        self.session_likes = set()
+        self.session_dislikes = set()
+        self.session_content = set()
+
+    def feedback(self, likes, dislikes):
+        print 'likes =', likes
+        print 'dislikes =', dislikes
+        negative_content = set([self.dictionary.token2id[str(-x)] for x in dislikes])
+        print 'negative_content =', negative_content
+        likes = set([self.dictionary.token2id[str(item)] for item in likes])
+        print 'likes =', likes
+        dislikes = set([self.dictionary.token2id[str(item)] for item in dislikes])
+        print 'dislikes =', dislikes
+        assert likes.isdisjoint(dislikes),\
+            'likes and dislikes are not disjoint'
+        assert self.session_likes.isdisjoint(likes),\
+            'new likes are not disjoint with old likes'
+        assert self.session_dislikes.isdisjoint(dislikes),\
+            'new dislikes are not disjoint with old dislikes'
+        assert self.session_content.isdisjoint(negative_content),\
+            'new negative content is not disjoint with old content'
+        self.session_likes = self.session_likes.union(likes)
+        print 'self.session_likes =', self.session_likes
+        self.session_dislikes = self.session_dislikes.union(dislikes)
+        print 'self.session_dislikes =', self.session_dislikes
+        self.session_content = self.session_content.union(likes)
+        self.session_content = self.session_content.union(negative_content)
+        print 'self.session_content =', self.session_content
+
+    def predict(self, num_items):
+        print 'self.session_content =', self.session_content
+        content_vector = [(item, 1) for item in self.session_content]
+        print 'content_vector =', content_vector
+        topic_dist = self.model[content_vector]
+        print 'topic_dist =', topic_dist
+        topic_probs = [topic[1] for topic in topic_dist]
[…]

### 2014-01-27 macys/LDAPredictor.py

@@ -1,0 +1,55 @@
+#!/usr/local/bin/python
+
+from gensim import corpora, models, similarities
+from gensim.models import ldamodel
+import pickle
+
+class LDAPredictor(object):
+    """An abstract class that loads an LDA model from disk and utilizes it to
+    predict new items for a session given previously liked and disliked items.
+    """
+
+    def __init__(self, dict_fname=None, model_fname=None):
+        """Constructs a new LDAPredictor.
+
+        Arguments are the filenames that store the dictionary and LDA model.
+        """
+        self.dict_fname = dict_fname
+        self.model_fname = model_fname
+        self.initSession()
+
+    def loadModel(self):
+        """Loads a dictionary that maps item tokens to LDA Ids as well as a
+        corresonding LDA model from disk.
+        """
+        self.dictionary = self._loadDictionary(self.dict_fname)
+        self.model = self._loadModel(self.model_fname)
+
+    def initSession(self):
+        """Initializes a new session. This function does not return anything."""
+        raise NotImplementedError
+
+    def feedback(self, likes, dislikes):
+        """Informs the LDAPredictor that items in the `likes' set are liked and
+        the items in the `dislikes' set are disliked in the present session. The
+        provided sets must be disjoint with each other and with all sets previously
+        provided. This function does not return anything.
+        """
+        raise NotImplementedError
+
+    def predict(self, num_items):
+        """Returns `num_items' new items that are disjoint with all `likes' and
+        `dislikes'  sets previous provided to the feedback() function.
+        """
+        raise NotImplementedError
+
+    def _loadDictionary(self, filename):
+        """Loads dictionary from a text file."""
+        dictionary = corpora.Dictionary.load_from_text(filename)
+        return dictionary
+    
+    def _loadModel(self, filename):
+        """Loads a pickled LDA model."""
+        with open(filename, 'r') as f:
+            lda = pickle.load(f)
+        return lda

### 2014-01-27 macys/buildModel.py

@@ -1,0 +1,75 @@
+#!/usr/local/bin/python
+
+"""
+Construct and save LDA model using Radim Rehurek's gensim module.
+"""
+
+from gensim import corpora, models, similarities
+from gensim.models import ldamodel
+from optparse import OptionParser
+import logging
+import pickle
+import csv
+
+def getParser():
+    parser = OptionParser()
+    parser.add_option('-f', '--file', dest='filename', default='trainSessions.csv',
+        help=('CSV file containing training data. Each line consists of tokens '
+              'liked in a single session.'), metavar='FILE')
+    parser.add_option('-d', '--dict-fname-in', dest='dict_fname_in', default='tokens.csv',
+        help='File containing all tokens in the vocabulary.', metavar='FILE')
+    parser.add_option('--dict-fname-out', dest='dict_fname_out', default='tokens.dict',
+        help='Save file for dictionary that maps itemids to tokens.', metavar='FILE')
+    parser.add_option('-l', '--lda-file', dest='lda_filename', default='lda.pickle',
+        help='Save file for LDA model.', metavar='FILE')
+    parser.add_option('-t', '--num-topics', dest='num_topics', type='int', default=100,
+        help='Number of topics in LDA model.', metavar='NUM')
+    return parser
+
+def buildDictionary(fname_in, fname_out):
+    logging.info('Building dictionary from %s.' % fname_in)
+    with open(fname_in) as csvIn:
+        reader = csv.reader(csvIn)
+        records = [record for record in reader]
+        dictionary = corpora.Dictionary(records)
+        logging.info('Saving dictionary to %s.' % fname_out)
+        dictionary.save_as_text(fname_out)
+    return dictionary
+            
+def getCorpus(dictionary, filename):
+    logging.info('Building training corpus from %s.' % filename)
+    with open(filename) as csvIn:
+        reader = csv.reader(csvIn)
+        records = [record for record in reader]
+    return [dictionary.doc2bow(record) for record in records]
+
+def buildLDA(dictionary, corpus, num_topics):
+    lda = ldamodel.LdaModel(corpus=corpus,
+                            id2word=dictionary,
+                            num_topics=num_topics,
+                            passes=40,
+                            update_every=1,
+                            chunksize=800)
+    return lda
+
+def main():
+    # Setup logging
[…]

### 2014-01-27 macys/genCatSessions.py

@@ -1,24 +1,26 @@
 #!/usr/local/bin/python
 
+"""
+Generates session information based on category listings.
+"""
+
 from optparse import OptionParser
 import csv
 import logging
 import random
 
-logging.basicConfig(format='%(asctime)s : %(levelname)s : %(message)s', level=logging.INFO)
-
 def getParser():
     parser = OptionParser()
     parser.add_option('-f', '--file', dest='filename', default='categories.csv',
-        help='Input CSV file containing <category, productId> on each line.',
+        help='Input CSV file containing <category, item> on each line.',
         metavar='FILE')
     parser.add_option('-o', '--output', dest='outfilename', default='trainSessions.csv',
-        help='Output CSV file containing <productId, productId, productId, ...> on each line.',
+        help='Output CSV file containing <item, item, item, ...> on each line.',
         metavar='OUT_FILE')
     parser.add_option('-s', '--session-size', dest='session_size', type="int", default=100,
-        help='Max number of products per session.', metavar='SIZE')
+        help='Max number of items per session.', metavar='SIZE')
     parser.add_option('--num-sessions-factor', dest='num_sessions_factor', type="int",
-        default=10, help='NUM_SESSIONS(category) = NUM_PRODUCTS(category)/FACTOR',
+        default=10, help='NUM_SESSIONS(category) = NUM_ITEMS(category)/FACTOR',
         metavar='FACTOR')
     return parser
     
@@ -32,29 +34,36 @@ def randomPop(data):
 	else:
 		return None
 
-def generateSessions(category, productIds, session_size,
+def generateSessions(category, items, session_size,
                      num_sessions_factor, writer):
     logging.info('=============================================')
     logging.info('Generating sessions for %s' % category)
-    num_products = len(productIds)
-    logging.info('len_products = %d' % num_products)
-    num_sessions = 1 + int(num_products/num_sessions_factor)
+    num_items = len(items)
+    logging.info('num_items = %d' % num_items)
+    num_sessions = 1 + int(num_items/num_sessions_factor)
     logging.info('num_sessions = %d' % num_sessions)
-    _session_size = int(min(num_products, session_size))
+    _session_size = int(min(num_items, session_size))
     logging.info('session_size = %d' % _session_size)
 
     for i in range(num_sessions):
-        copyIds = list(productIds)
+        items_copy = list(items)
         sessionIds = []
         for j in range(_session_size):
[…]

### 2014-01-27 macys/lda_test.py

@@ -1,94 +1,0 @@
-#!/usr/local/bin/python
-
-from gensim import corpora, models, similarities
-from gensim.models import ldamodel
-import sqlite3
-import logging
-import pickle
-import csv
-
-# params
-db_fname = 'macys.db'
-inFile = 'trainSessions.csv'
-testFile = 'testSessions.csv'
-corpusFile = 'categorySessions.corp'
-dictFile = 'products.dict'
-ldaFile = 'lda.pickle'
-
-logging.basicConfig(format='%(asctime)s : %(levelname)s : %(message)s', level=logging.INFO)
-
-def getDictionary(filename):
-    try:
-        dictionary = corpora.Dictionary.load_from_text(dictFile)
-        logging.info('Reading dictionary from %s.' % dictFile)
-    except IOError:
-        with open(filename) as csvIn:
-            reader = csv.reader(csvIn)
-            records = [record for record in reader]
-            dictionary = corpora.Dictionary(records)
-            dictionary.save_as_text(dictFile)
-    return dictionary
-            
-def getCorpus(dictionary, filename):
-    logging.info('Reading corpus from %s.' % filename)
-    with open(filename) as csvIn:
-        reader = csv.reader(csvIn)
-        records = [record for record in reader]
-        return [dictionary.doc2bow(record) for record in records]
-
-def getLDA():
-    try:
-        with open(ldaFile, 'r') as f:
-            lda = pickle.load(f)
-            logging.info('Reading LDA model from %s.' % ldaFile)
-    except IOError:
-        dictionary = getDictionary(inFile)
-        corpus = getCorpus(dictionary, inFile)
-        lda = ldamodel.LdaModel(corpus=corpus,
-                                id2word=dictionary,
-                                num_topics=100,
-                                passes=40,
-                                update_every=1,
-                                chunksize=800)
-        pickle.dump(lda, open(ldaFile, 'w'))
-    return lda
-
-def getSessionCats(db_curs, dictionary, session):
-    selectCategoriesStmt = 'SELECT Category FROM Categories WHERE Id = :Id'
-    sessionCats = {}
-    for product in session:
-        productId = dictionary[product[0]]
-        db_curs.execute(selectCategoriesStmt, (productId,))
-        for row in db_curs.fetchall():
-            category = row[0]
-            if category in sessionCats:
-                sessionCats[category] += 1
-            else:
-                sessionCats[category] = 1
-    return sessionCats
[…]

### 2014-01-27 macys/BayesPredictor.py
Remove debug statements from BayesPredictor.

@@ -4,7 +4,6 @@ from LDAPredictor import LDAPredictor
 from gensim import corpora, models, similarities
 from gensim.models import ldamodel
 from numpy.random import multinomial
-import logging
 
 class BayesPredictor(LDAPredictor):
     """An implementation of an LDAPredictor that uses Bayes' rule to generate
@@ -26,14 +25,13 @@ class BayesPredictor(LDAPredictor):
         self.session_content = set()
 
     def feedback(self, likes, dislikes):
-        print 'likes =', likes
-        print 'dislikes =', dislikes
-        negative_content = set([self.dictionary.token2id[str(-x)] for x in dislikes])
-        print 'negative_content =', negative_content
-        likes = set([self.dictionary.token2id[str(item)] for item in likes])
-        print 'likes =', likes
-        dislikes = set([self.dictionary.token2id[str(item)] for item in dislikes])
-        print 'dislikes =', dislikes
+        negative_content =\
+            set([self.dictionary.token2id[str(-token)] for token in dislikes])
+        likes =\
+            set([self.dictionary.token2id[str(token)] for token in likes])
+        dislikes =\
+            set([self.dictionary.token2id[str(token)] for token in dislikes])
+
         assert likes.isdisjoint(dislikes),\
             'likes and dislikes are not disjoint'
         assert self.session_likes.isdisjoint(likes),\
@@ -42,24 +40,17 @@ class BayesPredictor(LDAPredictor):
             'new dislikes are not disjoint with old dislikes'
         assert self.session_content.isdisjoint(negative_content),\
             'new negative content is not disjoint with old content'
+
         self.session_likes = self.session_likes.union(likes)
-        print 'self.session_likes =', self.session_likes
         self.session_dislikes = self.session_dislikes.union(dislikes)
-        print 'self.session_dislikes =', self.session_dislikes
-        self.session_content = self.session_content.union(likes)
-        self.session_content = self.session_content.union(negative_content)
-        print 'self.session_content =', self.session_content
+        self.session_content =\
+            self.session_content.union(likes).union(negative_content)
 
     def predict(self, num_items):
-        print 'self.session_content =', self.session_content
         content_vector = [(item, 1) for item in self.session_content]
-        print 'content_vector =', content_vector
         topic_dist = self.model[content_vector]
-        print 'topic_dist =', topic_dist
         topic_probs = [topic[1] for topic in topic_dist]
-        print 'topic_probs =', topic_probs
[…]

### 2014-01-27 macys/SessionTranslator.py
Add SessionTranslator.py.

@@ -1,0 +1,54 @@
+#!/usr/local/bin/python
+
+import sqlite3
+
+class SessionTranslator(object):
+   """Translates sets of tokens into either sets of item descriptions or
+   sparse category histograms.
+   """
+
+   selectCatStmt = 'SELECT Category FROM Categories WHERE Id = :Id'
+   selectDescTemplate = 'SELECT Description FROM Products WHERE Id IN (%s)'
+
+   def __init__(self, db_fname):
+       """Constructs SessionTranslator, which connects to sqlite3 db."""
+       db_conn = sqlite3.connect(db_fname)
+       with db_conn:
+           self.cursor = db_conn.cursor()
+
+   def sessionToDesc(self, session):
+       """Takes a list of itemsids and returns a list of item descriptions."""
+       selectDescStmt = self.selectDescTemplate % ','.join('?'*len(session))
+       self.cursor.execute(selectDescStmt, session)
+       return [desc for desc in self.cursor.fetchall()]
+
+   def sessionToCats(self, session):
+       """Takes a list of itemids and returns a dictionary of category
+       populations.
+       """
+       cats = {}
+       for itemid in session:
+           self.cursor.execute(self.selectCatStmt, (itemid,))
+           for row in self.cursor.fetchall():
+               cat = row[0]
+               if cat in cats:
+                   cats[cat] += 1
+               else:
+                   cats[cat] = 1
+       return cats
+
+def main():
+    """A simple, sanity-checking test."""
+    print 'Constructing SessionTranslator (and hence, connecting to db). . .'
+    translator = SessionTranslator('macys.db')
+    session = [960598,666644,632709,551024,932073,960606,824431,914853,1093859,1053814]
+    print 'session =', session
+    print 'Translating to category populations. . .'
+    cats = translator.sessionToCats(session)
+    print cats
+    print 'Translating to item descriptions. . .'
+    descriptions = translator.sessionToDesc(session)
+    print descriptions
+
+if __name__ == '__main__':
+    main()

### 2014-01-27 macys/SessionTranslator.py
Add simpleSim.py.

@@ -13,14 +13,13 @@ class SessionTranslator(object):
    def __init__(self, db_fname):
        """Constructs SessionTranslator, which connects to sqlite3 db."""
        db_conn = sqlite3.connect(db_fname)
-       with db_conn:
-           self.cursor = db_conn.cursor()
+       self.cursor = db_conn.cursor()
 
    def sessionToDesc(self, session):
        """Takes a list of itemsids and returns a list of item descriptions."""
        selectDescStmt = self.selectDescTemplate % ','.join('?'*len(session))
        self.cursor.execute(selectDescStmt, session)
-       return [desc for desc in self.cursor.fetchall()]
+       return [row[0] for row in self.cursor.fetchall()]
 
    def sessionToCats(self, session):
        """Takes a list of itemids and returns a dictionary of category

### 2014-01-27 macys/simpleSim.py

@@ -1,0 +1,82 @@
+#!/usr/local/bin/python
+
+"""
+Conduct a stupidly simple simulation of user-predictor interaction.
+"""
+
+from SessionTranslator import SessionTranslator
+from BayesPredictor import BayesPredictor
+import sqlite3
+
+# params
+db_fname = 'macys.db'
+dict_fname = 'tokens.dict'
+model_fname = 'lda.pickle'
+NUM_INIT_ITEMS = 10
+NUM_ITEMS_PER_ROUND = 10
+#CATEGORY = 'T-Shirts'            # ~0.3 accurancy
+#CATEGORY = 'Sunglasses'          # ~0.9
+#CATEGORY = 'Calvin Klein'        # ~0.3
+#CATEGORY = 'Swimwear'            # ~0.03 huh?
+#CATEGORY = 'Lucky Brand Jeans'   # ~0.2
+CATEGORY = 'Dress Shirts'        # ~0.3
+ROUNDS = 100
+
+selectItemsStmt = 'SELECT Id FROM Categories WHERE Category = :category LIMIT :limit'
+selectCatsStmt = 'SELECT Category FROM Categories WHERE Id = :id'
+
+def getInitItems(cursor, category):
+    cursor.execute(selectItemsStmt, (category, NUM_INIT_ITEMS))
+    return [row[0] for row in cursor.fetchall()]
+
+def judgeItems(cursor, category, items):
+    """Partition items into likes and dislikes using the simple-minded protocol
+    that items in CATEGORY are liked, all other items are disliked.
+    """
+    judgements = {}
+    judgements['likes'] = []
+    judgements['dislikes'] = []
+    for item in items:
+        cursor.execute(selectCatsStmt, (item,))
+        cats = [row[0] for row in cursor.fetchall()]
+        if category in cats:
+            judgements['likes'].append(item)
+        else:
+            judgements['dislikes'].append(item)
+    return judgements
+
+def main():
+    print 'Connecting to db. . .'
+    db_conn = sqlite3.connect(db_fname)
+    with db_conn:
+        cursor = db_conn.cursor()
+        translator = SessionTranslator(db_fname)
+        predictor = BayesPredictor(dict_fname=dict_fname, model_fname=model_fname)
+        print 'Loading LDA Model. . .'
+        predictor.loadModel()
+        print 'Generating initial items (%s). . .' % CATEGORY
+        items = getInitItems(cursor, CATEGORY)
+        print items
[…]

### 2014-01-28 macys/ArgMaxPredictor.py
Add ArgMaxPredictor and fix BayesPredictor and simpleSim.

@@ -4,8 +4,9 @@ from LDAPredictor import LDAPredictor
 from gensim import corpora, models, similarities
 from gensim.models import ldamodel
 from numpy.random import multinomial
+import operator
 
-class BayesPredictor(LDAPredictor):
+class ArgMaxPredictor(LDAPredictor):
     """An implementation of an LDAPredictor that uses Bayes' rule to generate
     item predictions.
     """
@@ -16,7 +17,7 @@ class BayesPredictor(LDAPredictor):
         Note that predictions associated with a topic are limited to the most
         likely topn_items items.
         """
-        super(BayesPredictor, self).__init__(**kwds)
+        super(ArgMaxPredictor, self).__init__(**kwds)
         self.ignore_dislikes = ignore_dislikes
         self.topn_items = topn_items=1000
 
@@ -26,6 +27,8 @@ class BayesPredictor(LDAPredictor):
         self.session_content = set()
 
     def feedback(self, likes, dislikes):
+        print 'likes =', likes
+        print 'dislikes =', dislikes
         if not self.ignore_dislikes:
             negative_content =\
                 set([self.dictionary.token2id[str(-token)] for token in dislikes])
@@ -53,38 +56,41 @@ class BayesPredictor(LDAPredictor):
     def predict(self, num_items):
         content_vector = [(item, 1) for item in self.session_content]
         topic_dist = self.model[content_vector]
-        topic_probs = [topic[1] for topic in topic_dist]
-        topic_sample = multinomial(num_items, topic_probs)
-        items = set()
+        candidates = {}
+        for topic_cpd in topic_dist:
+            topic = topic_cpd[0]
+            topic_prob = topic_cpd[1]
+            item_dist = self.model.show_topic(topic, topn=self.topn_items)
+            for item_cpd in item_dist:
+                item_prob = item_cpd[0]
+                token_str = item_cpd[1]
+                if not self.ignore_dislikes and int(token_str) < 0:
+                    continue
+                if token_str in candidates:
+                    candidates[token_str] += topic_prob*item_prob
+                else:
+                    candidates[token_str] = topic_prob*item_prob
+        sorted_candidates = sorted(candidates.iteritems(),
+                                   key=operator.itemgetter(1),
+                                   reverse=True)
         predictions = []
-        for i in range(len(topic_sample)):
-            draws = topic_sample[i]
-            if draws == 0:
-                continue
-            topicid = topic_dist[i][0]
[…]

### 2014-01-28 macys/simpleSim.py

@@ -6,7 +6,9 @@ Conduct a stupidly simple simulation of user-predictor interaction.
 
 from SessionTranslator import SessionTranslator
 from BayesPredictor import BayesPredictor
+#from ArgMaxPredictor import ArgMaxPredictor
 import sqlite3
+import random
 
 # params
 db_fname = 'macys.db'
@@ -14,20 +16,36 @@ dict_fname = 'tokens.dict'
 model_fname = 'lda.pickle'
 NUM_INIT_ITEMS = 10
 NUM_ITEMS_PER_ROUND = 10
-#CATEGORY = 'T-Shirts'            # ~0.3 accurancy
-#CATEGORY = 'Sunglasses'          # ~0.9
-#CATEGORY = 'Calvin Klein'        # ~0.3
-#CATEGORY = 'Swimwear'            # ~0.03 huh?
-#CATEGORY = 'Lucky Brand Jeans'   # ~0.2
-CATEGORY = 'Dress Shirts'        # ~0.3
-ROUNDS = 100
+ROUNDS = 25
+                                  # argmax  bayes
+#CATEGORY = 'T-Shirts'            # 0.096   0.356
+#CATEGORY = 'Sunglasses'          # 0.996   0.996
+CATEGORY = 'Calvin Klein'        # 0.5     0.576
+#CATEGORY = 'Swimwear'            # 0.31    0.207
+#CATEGORY = 'Lucky Brand Jeans'   # 0.23    0.254
+#CATEGORY = 'Dress Shirts'        # 0.98    0.884
 
-selectItemsStmt = 'SELECT Id FROM Categories WHERE Category = :category LIMIT :limit'
+selectItemsStmt = 'SELECT Id FROM Categories WHERE Category = :category'
 selectCatsStmt = 'SELECT Category FROM Categories WHERE Id = :id'
 
+def randomPop(data):
+    if data != []:
+        pos = random.randrange( len(data) )
+        elem = data[pos]
+        data[pos] = data[-1]
+        del data[-1]
+        return elem
+    else:
+        return None
+
 def getInitItems(cursor, category):
-    cursor.execute(selectItemsStmt, (category, NUM_INIT_ITEMS))
-    return [row[0] for row in cursor.fetchall()]
+    """Randomly choose and return NUM_INIT_ITEMS from `category'."""
+    cursor.execute(selectItemsStmt, (category,))
+    all_items = [row[0] for row in cursor.fetchall()]
+    init_items = []
+    for i in range(NUM_INIT_ITEMS):
[…]

### 2014-02-16 bdales/bdales.py
Initialize the HaveRecs column of Products within macys.py.

@@ -37,8 +37,8 @@ selectCategoryStmt =\
      'ParentCategory = :ParentCategory AND Category = :Category')
 insertProductStmt =\
     ('INSERT INTO Products(Id, Url, ImgFile, Description, Prices, Available, '
-     'Time) VALUES (:Id, :Url, :ImgFile, :Description, :Prices, 1, '
-     'CURRENT_TIMESTAMP)')
+     'HaveRecs, Time) VALUES (:Id, :Url, :ImgFile, :Description, :Prices, 1, '
+     '0, CURRENT_TIMESTAMP)')
 insertCategoryStmt =\
     ('INSERT INTO Categories(Id, ParentCategory, Category) '
      'VALUES (:Id, :ParentCategory, :Category)')

### 2014-02-16 bdales/urlGet.py
No image downloads from urlGet.py.

@@ -4,6 +4,7 @@ import sys
 import os
 from optparse import OptionParser
 from selenium import webdriver
+from ghost import Ghost
 
 def getParser(usage=None):
     parser = OptionParser(usage=usage)
@@ -26,10 +27,12 @@ def main():
         outfile = sys.stdout
 
     # Attempt to open url
-    browser = webdriver.Firefox()
-    browser.get(url)
-    content = browser.page_source.encode('utf-8')
-    browser.close()
+    profile = webdriver.FirefoxProfile();
+    profile.set_preference('permissions.default.image', 2);
+    driver = webdriver.Firefox(firefox_profile=profile);
+    driver.get(url)
+    content = driver.page_source.encode('utf-8')
+    driver.close()
 
     outfile.write(content)
     outfile.close()

### 2014-02-16 bdales/bdales_recs.py
Add bdales_rec.py, which uses selenium.

@@ -1,0 +1,101 @@
+#!/usr/local/bin/python
+
+import urllib2
+from bs4 import BeautifulSoup, NavigableString
+import sqlite3
+import sys
+from selenium import webdriver
+
+# params
+productUrlTemplate = 'http://www1.bloomingdales.com/shop/product/?ID=%d'
+cycleDriverInterval = 100
+loadTimeout = 30
+
+# db params
+db_fname = 'bdales.db'
+createRecommendsTableStmt =\
+    ('CREATE TABLE IF NOT EXISTS Recommends(Id1 INT, Id2 INT, '
+     'PRIMARY KEY (Id1, Id2), FOREIGN KEY(Id1) REFERENCES Products(Id), '
+     'FOREIGN KEY(Id2) REFERENCES Products(Id))')
+selectRecommendStmt = 'SELECT * FROM Recommends WHERE Id1 = :Id1 AND Id2 = :Id2'
+insertRecommendStmt = 'INSERT INTO Recommends(Id1, Id2) VALUES (:Id1, :Id2)'
+selectProductsStmt =\
+     'SELECT Id FROM Products WHERE Available = 1 AND HaveRecs = 0'
+selectProductStmt = 'SELECT * FROM Products WHERE Id = :Id'
+updateProductStmt = 'UPDATE Products SET HaveRecs = 1 WHERE Id = :Id'
+
+def getWebDriver():
+    profile = webdriver.FirefoxProfile()
+    profile.set_preference('permissions.default.image', 2)
+    driver = webdriver.Firefox(firefox_profile=profile)
+    driver.set_page_load_timeout(loadTimeout)
+    return driver
+
+def processRecommend(id1, productTag, db_curs):
+    # process productTag
+    try:
+        id2 = int(productTag['id'])
+    except ValueError:
+        print >> sys.stderr, 'WARNING: Failed to parse productId.'
+        return 0
+    # check for pre-existing recommendation
+    db_curs.execute(selectRecommendStmt, (id1, id2))
+    if db_curs.fetchone():
+        return 0
+    # check db for product
+    print 'Id2 = %d' % id2
+    db_curs.execute(selectProductStmt, (id2,))
+    if not db_curs.fetchone():
+        print >> sys.stderr, 'WARNING: Product not in DB.'
+        return 0
+    # insert new recommendation
+    db_curs.execute(insertRecommendStmt, (id1, id2))
+    return 1
+
+def getRecommends(driver, productId, db_curs):
[…]

### 2014-02-16 bdales/bdales_recs.py
Add multiprocessing to bdales_recs.py.

@@ -1,14 +1,17 @@
 #!/usr/local/bin/python
 
+import multiprocessing as mp
+import time
 import urllib2
-from bs4 import BeautifulSoup, NavigableString
+from bs4 import BeautifulSoup
 import sqlite3
 import sys
+import selenium
 from selenium import webdriver
 
 # params
+numWorkers = 4
 productUrlTemplate = 'http://www1.bloomingdales.com/shop/product/?ID=%d'
-cycleDriverInterval = 100
 loadTimeout = 30
 
 # db params
@@ -24,12 +27,14 @@ selectProductsStmt =\
 selectProductStmt = 'SELECT * FROM Products WHERE Id = :Id'
 updateProductStmt = 'UPDATE Products SET HaveRecs = 1 WHERE Id = :Id'
 
+
 def getWebDriver():
     profile = webdriver.FirefoxProfile()
     profile.set_preference('permissions.default.image', 2)
     driver = webdriver.Firefox(firefox_profile=profile)
     driver.set_page_load_timeout(loadTimeout)
     return driver
+
 
 def processRecommend(id1, productTag, db_curs):
     # process productTag
@@ -43,38 +48,72 @@ def processRecommend(id1, productTag, db_curs):
     if db_curs.fetchone():
         return 0
     # check db for product
-    print 'Id2 = %d' % id2
+    #print 'Id2 = %d' % id2
     db_curs.execute(selectProductStmt, (id2,))
     if not db_curs.fetchone():
-        print >> sys.stderr, 'WARNING: Product not in DB.'
+        #print >> sys.stderr, 'WARNING: Product not in DB.'
         return 0
     # insert new recommendation
     db_curs.execute(insertRecommendStmt, (id1, id2))
     return 1
 
+
 def getRecommends(driver, productId, db_curs):
-    print 'productId = %d' % productId
+    #print 'productId = %d' % productId
+    insertCnt = 0
     productUrl = productUrlTemplate % productId
     try:
         # fetch html
-        print 'Opening url: %s' % productUrl
+        #print 'Opening url: %s' % productUrl
         driver.get(productUrl)
         html = driver.page_source
         soup = BeautifulSoup(html)
-        insertCnt = 0
         for productTag in\
             soup.findAll('div', {'class': 'productThumbnail showQuickView'}):
             insertCnt +=\
                 processRecommend(productId, productTag, db_curs)
         if insertCnt == 0:
-            print 'No recommends found.'
+            #print 'No recommends found.'
+            pass
[…]

### 2014-02-16 bdales/bdales_recs.py
Change bdales_recs.py to use Queues instead of Pipes.

@@ -83,9 +83,7 @@ def getRecommends(driver, productId, db_curs):
     return insertCnt
 
 
-def worker(master_pipe, worker_pipe):
-    master_pipe.close()
-
+def worker(q):
     # connect to db
     db_conn = sqlite3.connect(db_fname)
     with db_conn:
@@ -96,24 +94,18 @@ def worker(master_pipe, worker_pipe):
         insertCnt = 0
         while True:
             try:
-                productId = worker_pipe.recv()
+                productId = q.get(block=False)
                 insertCnt += getRecommends(driver, productId, db_curs)
                 db_conn.commit()
-            except EOFError:
+            except Queue.Full:
                 break
 
         driver.quit()
 
     #print 'Inserted %d new recommendations.' % insertCnt
 
-    worker_pipe.close()
-        
 
-def master(pipes):
-    # close worker ends of pipes
-    for w in range(numWorkers):
-        pipes[w][1].close()
-
+def master(queues):
     # connect to db
     db_conn = sqlite3.connect(db_fname)
     with db_conn:
@@ -128,33 +120,31 @@ def master(pipes):
         for row in db_curs.fetchall():
             productId = row[0]
             workerIdx = iter % numWorkers
-            pipes[workerIdx][0].send(productId)
+            queues[workerIdx].put(productId)
             iter += 1
 
-    # close master ends of pipes
+    # close queues
     for w in range(numWorkers):
-        pipes[w][0].close()
+        queues[w].close()
 
 
 def main():
-    # create pipes
-    pipes = []
+    # create queues
+    queues = []
     for w in range(numWorkers):
-        master_pipe, worker_pipe = mp.Pipe()
-        pipes.append((master_pipe, worker_pipe))
+        queues.append(mp.Queue())
 
     # create worker processes
     workers = []
     for w in range(numWorkers):
-        w = mp.Process(target=worker, args=(pipes[w][0], pipes[w][1]))
-        workers.append(w)
+        workers.append(mp.Process(target=worker, args=(queues[w],)))
 
     # start worker processes
     for w in range(numWorkers):
         workers[w].start()
 
     # delegate work
-    master(pipes)
+    master(queues)
 
 
 if __name__ == '__main__':

### 2014-02-17 bdales/bdales_recs.py
Improve bdales_recs.py robustness.

@@ -1,6 +1,7 @@
 #!/usr/local/bin/python
 
 import multiprocessing as mp
+import Queue
 import time
 import urllib2
 from bs4 import BeautifulSoup
@@ -10,9 +11,11 @@ import selenium
 from selenium import webdriver
 
 # params
-numWorkers = 4
+numWorkers = 6
 productUrlTemplate = 'http://www1.bloomingdales.com/shop/product/?ID=%d'
-loadTimeout = 30
+workerTimeout = 30
+pageLoadTimeout = 20
+workerQueueSize = 10
 
 # db params
 db_fname = 'bdales.db'
@@ -32,7 +35,7 @@ def getWebDriver():
     profile = webdriver.FirefoxProfile()
     profile.set_preference('permissions.default.image', 2)
     driver = webdriver.Firefox(firefox_profile=profile)
-    driver.set_page_load_timeout(loadTimeout)
+    driver.set_page_load_timeout(pageLoadTimeout)
     return driver
 
 
@@ -92,20 +95,18 @@ def worker(q):
         driver = getWebDriver()
 
         insertCnt = 0
+        # TODO: Need a terminal condition here.
         while True:
-            try:
-                productId = q.get(block=False)
-                insertCnt += getRecommends(driver, productId, db_curs)
-                db_conn.commit()
-            except Queue.Full:
-                break
+            productId = q.get()
+            insertCnt += getRecommends(driver, productId, db_curs)
+            db_conn.commit()
 
         driver.quit()
 
     #print 'Inserted %d new recommendations.' % insertCnt
 
 
-def master(queues):
+def master(workers, queues):
     # connect to db
     db_conn = sqlite3.connect(db_fname)
     with db_conn:
@@ -120,7 +121,15 @@ def master(queues):
         for row in db_curs.fetchall():
             productId = row[0]
             workerIdx = iter % numWorkers
-            queues[workerIdx].put(productId)
+            try:
+                print 'Sending productId %d to worker %d' % (productId, workerIdx)
+                queues[workerIdx].put(productId, timeout=workerTimeout)
+            except Queue.Full:
+                # replace worker
+                print 'WARNING: Replacing worker %d' % workerIdx
+                workers[workerIdx].terminate()
+                workers[workerIdx] = mp.Process(target=worker, args=(queues[workerIdx],))
+                queues[workerIdx].put(productId, timeout=workerTimeout)
             iter += 1
 
     # close queues
[…]

### 2014-02-17 bdales/bdales_recs.py
Fix worker restart.

@@ -129,6 +129,8 @@ def master(workers, queues):
                 print 'WARNING: Replacing worker %d' % workerIdx
                 workers[workerIdx].terminate()
                 workers[workerIdx] = mp.Process(target=worker, args=(queues[workerIdx],))
+                workers[workerIdx].start()
+                time.sleep(5)
                 queues[workerIdx].put(productId, timeout=workerTimeout)
             iter += 1
 

### 2014-02-17 bdales/bdales_recs.py
Minor tweaks.

@@ -13,8 +13,8 @@ from selenium import webdriver
 # params
 numWorkers = 6
 productUrlTemplate = 'http://www1.bloomingdales.com/shop/product/?ID=%d'
-workerTimeout = 30
-pageLoadTimeout = 20
+workerTimeout = 15
+pageLoadTimeout = 15
 workerQueueSize = 10
 
 # db params
@@ -93,15 +93,16 @@ def worker(q):
         db_curs = db_conn.cursor()
 
         driver = getWebDriver()
+        try:
+            insertCnt = 0
+            # TODO: Need a terminal condition here.
+            while True:
+                productId = q.get()
+                insertCnt += getRecommends(driver, productId, db_curs)
+                db_conn.commit()
 
-        insertCnt = 0
-        # TODO: Need a terminal condition here.
-        while True:
-            productId = q.get()
-            insertCnt += getRecommends(driver, productId, db_curs)
-            db_conn.commit()
-
-        driver.quit()
+        finally:
+            driver.quit()
 
     #print 'Inserted %d new recommendations.' % insertCnt
 
@@ -130,7 +131,7 @@ def master(workers, queues):
                 workers[workerIdx].terminate()
                 workers[workerIdx] = mp.Process(target=worker, args=(queues[workerIdx],))
                 workers[workerIdx].start()
-                time.sleep(5)
+                time.sleep(2) # a couple of seconds to start 
                 queues[workerIdx].put(productId, timeout=workerTimeout)
             iter += 1
 
@@ -155,7 +156,7 @@ def main():
         workers[w].start()
 
     # allow time for browsers to start
-    time.sleep(2*numWorkers)
+    time.sleep(numWorkers)
 
     # delegate work
     master(workers, queues)

### 2014-02-17 bdales/bdales_recs.py
Improve bdales_recs.py robustness.

@@ -2,6 +2,8 @@
 
 import multiprocessing as mp
 import Queue
+import eventlet
+from eventlet.timeout import Timeout
 import time
 import urllib2
 from bs4 import BeautifulSoup
@@ -13,8 +15,8 @@ from selenium import webdriver
 # params
 numWorkers = 6
 productUrlTemplate = 'http://www1.bloomingdales.com/shop/product/?ID=%d'
-workerTimeout = 15
-pageLoadTimeout = 15
+workerTimeout = 30
+webDriverTimeout = 30
 workerQueueSize = 10
 
 # db params
@@ -35,7 +37,10 @@ def getWebDriver():
     profile = webdriver.FirefoxProfile()
     profile.set_preference('permissions.default.image', 2)
     driver = webdriver.Firefox(firefox_profile=profile)
-    driver.set_page_load_timeout(pageLoadTimeout)
+    # TODO: Verify that these timeouts are optimal.
+    driver.implicitly_wait(webDriverTimeout)
+    driver.set_page_load_timeout(webDriverTimeout)
+    driver.set_script_timeout(webDriverTimeout)
     return driver
 
 
@@ -86,28 +91,25 @@ def getRecommends(driver, productId, db_curs):
     return insertCnt
 
 
-def worker(q):
+def worker(driver, q):
     # connect to db
     db_conn = sqlite3.connect(db_fname)
     with db_conn:
         db_curs = db_conn.cursor()
 
-        driver = getWebDriver()
-        try:
-            insertCnt = 0
-            # TODO: Need a terminal condition here.
-            while True:
-                productId = q.get()
-                insertCnt += getRecommends(driver, productId, db_curs)
-                db_conn.commit()
-
-        finally:
-            driver.quit()
+        insertCnt = 0
+        # TODO: Need a terminal condition here.
+        while True:
+            productId = q.get()
+            # TODO: Verify that the next line is useful.
+            driver.switch_to_window(driver.current_window_handle)
+            insertCnt += getRecommends(driver, productId, db_curs)
+            db_conn.commit()
 
     #print 'Inserted %d new recommendations.' % insertCnt
 
 
-def master(workers, queues):
+def master(drivers, queues, workers):
     # connect to db
     db_conn = sqlite3.connect(db_fname)
     with db_conn:
@@ -122,17 +124,23 @@ def master(workers, queues):
         for row in db_curs.fetchall():
             productId = row[0]
             workerIdx = iter % numWorkers
-            try:
[…]

### 2014-02-17 cgi-bin/products.py
Updating

@@ -1,6 +1,7 @@
 #!/usr/bin/env python
 
 import sys; sys.path.insert(0, 'lib') # this line is necessary for the rest
+sys.path.append('/Users/jirvine/Desktop/cs399/cs399/macys')
 import os                             # of the imports to work!
 
 import json
@@ -21,12 +22,6 @@ import pprint
 # helper method to convert times from database (which will return a string)
 # into datetime objects. This will allow you to compare times correctly (using
 # ==, !=, <, >, etc.) instead of lexicographically as strings.
-
-# Sample use:
-# current_time = string_to_time(sqlitedb.getTime())
-
-def string_to_time(date_str):
-    return datetime.strptime(date_str, '%Y-%m-%d %H:%M:%S')
 
 # helper method to render a template in the templates/ directory
 #

### 2014-02-17 cgi-bin/sqlitedb.py

@@ -29,6 +29,7 @@ def transaction():
 #     t.commit()
 #
 # check out http://webpy.org/cookbook/transactions for examples
+
 def getProductID():
     query_string = 'select min(Id) as x from Products'
     results = query(query_string)
@@ -41,14 +42,17 @@ def getProduct(productId):
 
 def getNextProduct(productId, liked):
     productId = int(productId)
-    if productId == 0:
-        productId = 1082639 # Magic productId for testing
+    if (liked == "first"):
+        global sessionId 
+        sessionId += 1
+        wardrobe.initSession(sessionId)
+        return getProduct(productId)
     likes = []
     dislikes = []
     if liked == 'disliked':
-        dislikes.append(int(productId))
+        dislikes.append(productId)
     else:
-        likes.append(int(productId))
+        likes.append(productId)
     wardrobe.feedback(sessionId, likes, dislikes)
     items = wardrobe.nextItems(sessionId, 1)
     result = getProduct(items[0])

### 2014-02-17 cgi-bin/static/js/index.js

@@ -1,18 +1,17 @@
 var currentProduct=new Object();
-// Set currentProduct initially to have smalled id
-$.post("/leftorright", 
-        {
-            'productId':0,
-            'liked':"neither"
-        },
-        function(data,status){
-            var product = JSON.parse(data);
-            document.getElementById("test").innerHTML = "Id = " + product.Id +
-                ", Description = " + product.Description;
-            currentProduct = product;
-            changeProduct(product);
-        });
+var defaultId = 1082639;
 
+// Set currentProduct initially to have smallest id
+var firstId = getIdFromURL();
+getNextProduct(firstId, "first");
+
+function getIdFromURL() {
+    var name = "firstId";
+    name = name.replace(/[\[]/, "\\[").replace(/[\]]/, "\\]");
+    var regex = new RegExp("[\\?&]" + name + "=([^&#]*)"),
+        results = regex.exec(location.search);
+    return results == null ? defaultId : decodeURIComponent(results[1].replace(/\+/g, " "));
+}
 
 function changeProduct(product) {
     image = document.getElementById("productImage");
@@ -40,24 +39,24 @@ function writeToTest() {
     test.innerHTML = "This is just a test";
 }
 
-function JQAjax(list) {
+function getNextProduct(productId, liked) {
     $.post("/leftorright", 
             {
-                'productId':currentProduct.Id,
-                'liked':list
+                'productId':productId,
+    'liked':liked
             },
             function(data,status){
                 var product = JSON.parse(data);
                 document.getElementById("test").innerHTML = "Id = " + product.Id +
-                    ", Description = " + product.Description;
-                changeProduct(product);
-                currentProduct = product;
+        ", Description = " + product.Description;
+    changeProduct(product);
+    currentProduct = product;
             });
 }
 
 function move(list) {
     addToList(list);
-    JQAjax(list);
+    getNextProduct(currentProduct.Id, list);
 }
 
 $(document).keyup(function(e) {

### 2014-02-17 macys/MagicWardrobe.py

@@ -1,6 +1,6 @@
 #!/usr/local/bin/python
 
-from RRBFSPredictor import RRBFSPredictor
+from WalkPredictor import WalkPredictor
 import pickle
 
 # params
@@ -14,7 +14,7 @@ class MagicWardrobe:
     def __init__(self):
         """Initializes predictor/feedback engine."""
         graph = self._loadGraph()
-        self.predictor = RRBFSPredictor(graph)
+        self.predictor = WalkPredictor(graph)
         pass
 
     def initSession(self, sessionId):

### 2014-02-17 macys/RRBFSPredictor.py

@@ -21,6 +21,7 @@ class RRBFSPredictor(ItemPredictor):
         self.sessions = {}
 
     def initSession(self, sessionId):
+        self.sessions = {}
         self.sessions[sessionId] = self.Session()
 
     def feedback(self, sessionId, likes, dislikes):

### 2014-02-17 proposal/new_proposal.tex
Added new proposal

\documentclass[10pt]{article}

% =========================================================================
% document style changes
% =========================================================================

\usepackage{amsmath}                    % AMS math packages
\usepackage{amssymb}                    %
\usepackage[]{graphpap}
\usepackage[T1]{fontenc}                % for \mathrm{}
\usepackage{courier}                    % for \texttt{}
\usepackage{bbm}                        % for \mathbbm{1} (indicator function)
    \usepackage{booktabs}
    \usepackage{graphicx}
    \usepackage{caption}
    \usepackage{subcaption}

    %\setlength{\parskip}{\baselineskip}     % skip line following paragraphs
    \setlength{\parskip}{0.0in}
    \setlength{\parindent}{0.3in}             % Control margins and amount of text
    \setlength{\topmargin}{-0.8in}
    \setlength{\oddsidemargin}{-.15in}       % changed from {-.15in}
    \setlength{\textheight}{9.5in}
    \setlength{\textwidth}{6.8in}
    \pagestyle{empty}                       % No page numbers

    \newcommand{\spc}{\vspace{0.25in}}      % Shortcut commands
    \newcommand{\ds}{\displaystyle}         %\newcommand{\ds}[1]{\displaystyle{#1}}
    \newcommand{\ra}{\rightarrow}

    \begin{document}                        % This is where the document begins

{\LARGE\bf
    \begin{tabbing}
    \hspace{2.8in} \= \hspace{1.3in} \= \hspace{1.2in} \= \\

        %=========    Heading   ==================================================
        CS 399 \> PA4 \> Bill Chickering (bchick)\\
        \normalsize Jan 18, 2014 \> \> Jamie Irvine (jirvine)
        % =========================================================================
        \end{tabbing}
} \vspace{.4in}

\section*{Abstract}
\emph{Will write this when the paper is done}

\section*{Introduction}
For any large retailer, it is valuable to know the relationship between
products they sell or could sell. This information can be used to build
recommendations for similar products, understand subsets of products that
appeal to certain users, or expand a catalogue to include new products that
will be liked by the existing user-base. Currently, many retailers already
leverage user traffic and content information to learn these relationships
within their catalogue. However, in spaces where competing retailers sell
many of the same products, such as department stores or movie vendors, more
information can be gleaned by examining the inter-product relationships that
competing retailers have built for their own catalogues.

In this paper, we focus on department stores. This space is prime for our
analysis since there are a number of competing stores that have overlapping
catalogues. We model the relationships between products with a
\emph{recommendation graph}, which is directed graph between the products in a
catalogue. A directed edge from product A to product B means that if a user
likes product A, he/she is likely to like product B. First we construct the
recommendation graphs of different department stores using the
recommendations given for each product on their websites. With these
original graphs, we aim to add meaningful edges within one retailers graph
(representing a new recommendation) and between different retailer graphs
(representing a candidate to add to a catalogue).

We compare three approaches to extending the recommendation graphs. The
first approach is a baseline that uses a content-driven similarity function
to add edges between items. It ignores any information from the original
recommendation graphs. The second approach focuses exclusively on the
recommendation graphs. This method uses an identity function to determine
which items in different graphs are the same and thus links the graphs
together. Using a random walk algorithm, we determine which unlinked items
should be linked. The third approach is uses a combination of the first two
to leverage both kinds of information.

\section*{Data}
We collect the data by scraping the websites of major departments stores for
products with their descriptions, prices, photos, etc. From each product
page, we also gather the website's top recommended items for that product.
For this paper, we scraped the catalogues of Macy's, Bloomingdale's, etc.
Each of these websites requires their own scraper, but the same data is
accessible from all of them, including at least four recommended products
for each product in the catalogue. This allows us to build a directed
recommendation graph for each department store.

It is worth noting that in reality, a company would have more data about
their own products than what they post on their website. However, the
recommendation graph is still quite informative and is accessible to anyone.
It is realistic that a company could gather graphs like this from each of
their competitors, as we have here.

\section*{Methods}
\subsection*{Method 1}
create similarity scores between items. If items are similar, we add edges
in both directions to the recommendation graph. To determine the threshold
of how high the score must be before adding an edge, we calculate the
average similarity between items that are connected in the original graph.
These edges can connect two items in the same catalogue or items across
catalogues.

\emph{More about the specifics of the function here}

\subsection*{Method 2}
The second method uses the recommendation graphs themselves to determine new
recommendation edges. Since there is a significant overlap of products
between department stores (x\% between Macy's and Bloomingdales for example),
        recognizing which products are the same connects the otherwise disjoint
        graphs of different companies. To identify which products are identical, we
        construct an identity function. \emph{This function works like this and was
            tested like this, so we know it works.}

            With connected graphs, we can apply a random walk algorithm to determine
            which new edges would be most appropriate to add. In the random walk, each
            product node is a source of a 'walker', which has equal probability of
            moving down each outgoing edge. We add edges based on the probability that a
            walker goes from one product to another. Since the bench-mark for an
            appropriate threshold is less obvious in this method, we use the top X
            scores, the number of edges above the threshold found in method 1. While
            this is somewhat arbitrary, it allows for a fair comparison between the
            methods.

            \emph{More about the specifics of the random walk algorithm here}

            \subsection*{Method 3}
            The final method leverages both content-similarity and the structure of the
            recommendation graphs. First, we combine the recommendation graphs using an
            identity function and run a random-walk algorithm, as in method 2. This
            determines how 'far,' in a sense, any product is from any other product. To
            determine if there should be an edge added from product A to product B, we
            use the content-similarity function of method 1 to compare A to every other
            product in the catalogue and sum these scores up with the weights from the
            random-walk originating at product B. Thus content-similarity from product A
            to a product that is near product B in the recommendation graph increases
            the likelihood that product A should be connected to product B. \emph{An equation
                would probably be good in here}. For this method, we also add the top X
                edges based on the threshold in method 1.

                \emph{More about the specifics of the implementation here}

                \section*{Evaluation}
                Empirically evaluating the methods is somewhat difficult with the data we
                have access to. With all of the user-traffic data, one could see if an added
                recommendation edge connects two products that are in fact co-clicked on
                co-purchased. Without this, the question of whether one product should be
                recommended for another can be answered with more subjective human
                judgement.

                For this experiment, we collect the top edges to be added for each of the
                three methods. We show a subset of them to a number of people who rate them
                in exchange for small sexual favors. The tester is asked 'How true is the
                following statement: If someone likes product 1, they will probably like
                product 2', followed by the pictures and descriptions of the two items of
                the directed edge. The tester answers on a 5 point scale and the results are
                aggregated. Each tester is shown edges from each of three methods, to
                decrease bias from a biased tester.

                \emph{More about how many testers we use}

                \section*{Results}
                \emph{We'll see}

                \section*{Conclusion}
                \emph{Will write this when the paper is done}

                \end{document}

### 2014-02-17 proposal/new_proposal.tex
Minor tweaks and formatting fixes.

@@
 leverage user traffic and content information to learn these relationships
-within their catalogue. However, in spaces where competing retailers sell
-many of the same products, such as department stores or movie vendors, more
+within their catalogue. However, in spaces where competing retailers sell many
+of the same products, such as department stores or movie vendors, more
 information can be gleaned by examining the inter-product relationships that
@@
 
-In this paper, we focus on department stores. This space is prime for our
-analysis since there are a number of competing stores that have overlapping
+In this paper, we focus on department stores. This space is well suited for our
+analysis since there are a number of competing stores with overlapping
 catalogues. We model the relationships between products with a
-\emph{recommendation graph}, which is directed graph between the products in a
-catalogue. A directed edge from product A to product B means that if a user
+\emph{recommendation graph}, which is a directed graph between the products in
+a catalogue. A directed edge from product A to product B means that if a user
 likes product A, he/she is likely to like product B. First we construct the
-recommendation graphs of different department stores using the
-recommendations given for each product on their websites. With these
-original graphs, we aim to add meaningful edges within one retailers graph
-(representing a new recommendation) and between different retailer graphs
-(representing a candidate to add to a catalogue).
+recommendation graphs of different department stores using the recommendations
+given for each product on their websites. With these original graphs, we aim to
+add meaningful edges within one retailer's graph (representing a new
+recommendation) and between different retailer graphs (representing a candidate
+to add to a catalogue).
 
-We compare three approaches to extending the recommendation graphs. The
-first approach is a baseline that uses a content-driven similarity function
-to add edges between items. It ignores any information from the original
+We compare three approaches to extending the recommendation graphs. The first
+approach is a baseline that uses a content-driven similarity function to add
+edges between items. It ignores any information from the original
 recommendation graphs. The second approach focuses exclusively on the
-recommendation graphs. This method uses an identity function to determine
-which items in different graphs are the same and thus links the graphs
-together. Using a random walk algorithm, we determine which unlinked items
-should be linked. The third approach is uses a combination of the first two
-to leverage both kinds of information.
+recommendation graphs. This method uses an identity function to determine which
+items in different graphs are the same and thus links the graphs together.
+Using a random walk algorithm, we determine which unlinked items should be
+linked. The third approach is uses a combination of the first two to leverage
+both kinds of information.
 
@@
 We collect the data by scraping the websites of major departments stores for
-products with their descriptions, prices, photos, etc. From each product
-page, we also gather the website's top recommended items for that product.
-For this paper, we scraped the catalogues of Macy's, Bloomingdale's, etc.
-Each of these websites requires their own scraper, but the same data is
-accessible from all of them, including at least four recommended products
-for each product in the catalogue. This allows us to build a directed
-recommendation graph for each department store.
+products with their descriptions, prices, photos, etc. From each product page,
+we also gather the website's top recommended items for that product.  For this
+paper, we scraped the catalogues of Macy's, Bloomingdale's, etc.  Each of these
+websites requires their own scraper, but the same data is accessible from all
+of them, including at least four recommended products for each product in the
+catalogue. This allows us to build a directed recommendation graph for each
+department store.
 
-It is worth noting that in reality, a company would have more data about
-their own products than what they post on their website. However, the
-recommendation graph is still quite informative and is accessible to anyone.
-It is realistic that a company could gather graphs like this from each of
-their competitors, as we have here.
+It is worth noting that in reality, a company would have more data about their
+own products than what they post on their website. However, the recommendation
+graph is still quite informative and is accessible to anyone.  It is realistic
+that a company could gather graphs like this from each of their competitors, as
+we have here.
 
@@
 \subsection*{Method 1}
-create similarity scores between items. If items are similar, we add edges
-in both directions to the recommendation graph. To determine the threshold
-of how high the score must be before adding an edge, we calculate the
-average similarity between items that are connected in the original graph.
-These edges can connect two items in the same catalogue or items across
-catalogues.
+Create similarity scores between items. If items are similar, we add edges in
+both directions to the recommendation graph. To determine the threshold of how
+high the score must be before adding an edge, we calculate the average
+similarity between items that are connected in the original graph.  These edges
+can connect two items in the same catalogue or items across catalogues.
 
@@
 The second method uses the recommendation graphs themselves to determine new
-recommendation edges. Since there is a significant overlap of products
-between department stores (x\% between Macy's and Bloomingdales for example),
-        recognizing which products are the same connects the otherwise disjoint
-        graphs of different companies. To identify which products are identical, we
-        construct an identity function. \emph{This function works like this and was
-            tested like this, so we know it works.}
+recommendation edges. Since there is a significant overlap of products between
+department stores (x\% between Macy's and Bloomingdales for example),
+recognizing which products are the same connects the otherwise disjoint graphs
+of different companies. To identify which products are identical, we construct
+an identity function. \emph{This function works like this and was tested like
+this, so we know it works.}
 
-            With connected graphs, we can apply a random walk algorithm to determine
-            which new edges would be most appropriate to add. In the random walk, each
-            product node is a source of a 'walker', which has equal probability of
-            moving down each outgoing edge. We add edges based on the probability that a
-            walker goes from one product to another. Since the bench-mark for an
-            appropriate threshold is less obvious in this method, we use the top X
-            scores, the number of edges above the threshold found in method 1. While
-            this is somewhat arbitrary, it allows for a fair comparison between the
-            methods.
+With connected graphs, we can apply a random walk algorithm to determine which
+new edges would be most appropriate to add. In the random walk, each product
+node is a source of a 'walker', which has equal probability of moving down each
+outgoing edge. We add edges based on the probability that a walker goes from
+one product to another. Since the bench-mark for an appropriate threshold is
+less obvious in this method, we use the top X scores, the number of edges above
+the threshold found in method 1. While this is somewhat arbitrary, it allows
+for a fair comparison between the methods.
 
@@
 determines how 'far,' in a sense, any product is from any other product. To
-            determine if there should be an edge added from product A to product B, we
-            use the content-similarity function of method 1 to compare A to every other
-            product in the catalogue and sum these scores up with the weights from the
-            random-walk originating at product B. Thus content-similarity from product A
-            to a product that is near product B in the recommendation graph increases
-            the likelihood that product A should be connected to product B. \emph{An equation
-                would probably be good in here}. For this method, we also add the top X
-                edges based on the threshold in method 1.
+determine if there should be an edge added from product A to product B, we use
+the content-similarity function of method 1 to compare A to every other product
+in the catalogue and sum these scores up with the weights from the random-walk
+originating at product B. Thus content-similarity from product A to a product
+that is near product B in the recommendation graph increases the likelihood
+that product A should be connected to product B. \emph{An equation would
+probably be good in here}. For this method, we also add the top X edges based
+on the threshold in method 1.
 
@@
 \section*{Evaluation}
-                Empirically evaluating the methods is somewhat difficult with the data we
-                have access to. With all of the user-traffic data, one could see if an added
+Empirically evaluating the methods is somewhat difficult with the data we have
+access to. With all of the user-traffic data, one could see if an added
 recommendation edge connects two products that are in fact co-clicked on
 co-purchased. Without this, the question of whether one product should be
-                recommended for another can be answered with more subjective human
-                judgement.
+recommended for another can be answered with more subjective human judgement.
 
-                For this experiment, we collect the top edges to be added for each of the
-                three methods. We show a subset of them to a number of people who rate them
-                in exchange for small sexual favors. The tester is asked 'How true is the
+For this experiment, we collect the top edges to be added for each of the three
+methods. We show a subset of them to a number of people who rate them in
+exchange for small sexual favors. The tester is asked 'How true is the
 following statement: If someone likes product 1, they will probably like
-                product 2', followed by the pictures and descriptions of the two items of
-                the directed edge. The tester answers on a 5 point scale and the results are
-                aggregated. Each tester is shown edges from each of three methods, to
-                decrease bias from a biased tester.
+product 2', followed by the pictures and descriptions of the two items of the
+directed edge. The tester answers on a 5 point scale and the results are
+aggregated. Each tester is shown edges from each of three methods, to decrease
+bias from a biased tester.
 

### 2014-02-17 proposal/new_proposal.tex
New introduction.

@@
 \section*{Introduction}
-For any large retailer, it is valuable to know the relationship between
-products they sell or could sell. This information can be used to build
-recommendations for similar products, understand subsets of products that
-appeal to certain users, or expand a catalogue to include new products that
-will be liked by the existing user-base. Currently, many retailers already
-leverage user traffic and content information to learn these relationships
-within their catalogue. However, in spaces where competing retailers sell many
-of the same products, such as department stores or movie vendors, more
-information can be gleaned by examining the inter-product relationships that
-competing retailers have built for their own catalogues.
+A key challenge in retail is choosing which items to incorporate into one's
+collection of offered goods and services. These decisions can significantly
+impact a retailer's revenue in the short run as well as effect customer
+experience, and hence, their relationship with the retailer. Several factors
+must be considered including the number of product lines, the vareity of
+products in each line, as well as the consistency and relationships between
+products. In this study, we focus on the latter and explore techniques that
+leverage publicly available product recommendation information for the purpose
+of improving product assortment decisions.
 
-In this paper, we focus on department stores. This space is well suited for our
-analysis since there are a number of competing stores with overlapping
-catalogues. We model the relationships between products with a
-\emph{recommendation graph}, which is a directed graph between the products in
-a catalogue. A directed edge from product A to product B means that if a user
-likes product A, he/she is likely to like product B. First we construct the
-recommendation graphs of different department stores using the recommendations
-given for each product on their websites. With these original graphs, we aim to
-add meaningful edges within one retailer's graph (representing a new
-recommendation) and between different retailer graphs (representing a candidate
-to add to a catalogue).
+It has become standard practice for online retailers to recommend one or more
+products to prospective customers who have viewed or purchased an item. These
+recommendations are generally derived from collaborative filtering or other
+data-mining techniques employed by the retailer or a third-party service
+provider. Importantly, these recommendations provide a source of publicly
+available business intelligence by relating the items in their catalogue.
+Specifically, these online recommendations logically form a directed graph in
+with the vertices are products and an edge pointing from item A to item B
+indicates that customers who view or purchase item A are recommended item B.
+Given the recommendation graphs of two or more distinct retailers, our goal is
+to determine new, meaningful edges that connect these graphs in a way that would
+allow one retailer to relate their products to those of another retailer.
 
+In this study we choose to focus on department stores, which typically have
+overlapping product catalogues. We intend to exploit this overlap by identifying
+identical inter-retailer products in order to achieve some initial connectivity
+between their recommendation graphs. From there, we will explore and compare a
+vareity of techniques for choosing additional edges to further connect these
+originally disconnected graphs. In addition to relating {\em inter-retailer} items, we
+are also interested in leveraging other retailer's recommendation graphs in
+order to determine new {\em intra-retailer} edges, which can be valuable for
+improving customer recommendations.
+
+\section*{Data}
+We collect the data by scraping the websites of major departments stores for
+products with their descriptions, prices, photos, etc. From each product page,
+we also gather the website's top recommended items for that product.  For this
+paper, we scraped the catalogues of Bloomingdale's, Neiman Marcus, Nordstrom,
+etc.. Each of these websites requires their own scraper, but the same data is
+accessible from all of them, including at least four recommended products for
+each product in the catalogue. This allows us to build a directed recommendation
+graph for each department store.
+
+It is worth noting that in reality, a company would have more data about their
+own products than what they post on their website. However, the recommendation
+graph is still quite informative and is accessible to anyone.  It is realistic
+that a company could gather graphs like this from each of their competitors, as
+we have here.
+
+\section*{Methods}
 We compare three approaches to extending the recommendation graphs. The first
@@
 
-\section*{Data}
-We collect the data by scraping the websites of major departments stores for
-products with their descriptions, prices, photos, etc. From each product page,
-we also gather the website's top recommended items for that product.  For this
-paper, we scraped the catalogues of Macy's, Bloomingdale's, etc.  Each of these
-websites requires their own scraper, but the same data is accessible from all
-of them, including at least four recommended products for each product in the
-catalogue. This allows us to build a directed recommendation graph for each
-department store.
-
-It is worth noting that in reality, a company would have more data about their
-own products than what they post on their website. However, the recommendation
-graph is still quite informative and is accessible to anyone.  It is realistic
-that a company could gather graphs like this from each of their competitors, as
-we have here.
-
-\section*{Methods}
 \subsection*{Method 1}

### 2014-02-18 proposal/new_proposal.tex
small edits to proposal

@@
 Specifically, these online recommendations logically form a directed graph in
-with the vertices are products and an edge pointing from item A to item B
+which the vertices are products and an edge pointing from item A to item B
 indicates that customers who view or purchase item A are recommended item B.
@@
 \subsection*{Method 1}
-Create similarity scores between items. If items are similar, we add edges in
+For the baseline method, we create similarity scores between items based on the
+given content of each item. If items are similar, we add edges in
 both directions to the recommendation graph. To determine the threshold of how
@@
 recommendation graphs. First, we combine the recommendation graphs using an
+
 identity function and run a random-walk algorithm, as in method 2. This
@@
 For this experiment, we collect the top edges to be added for each of the three
-methods. We show a subset of them to a number of people who rate them in
-exchange for small sexual favors. The tester is asked 'How true is the
-following statement: If someone likes product 1, they will probably like
-product 2', followed by the pictures and descriptions of the two items of the
-directed edge. The tester answers on a 5 point scale and the results are
-aggregated. Each tester is shown edges from each of three methods, to decrease
-bias from a biased tester.
+methods. We show a subset of them to a number of people who rate them subjectively. 
+The tester is asked 'How true is the following statement: If someone likes
+product 1, they will probably like product 2', followed by the pictures and
+descriptions of the two items of the directed edge. The tester answers on a 5
+point scale and the results are aggregated. Each tester is shown edges from
+each of three methods, to decrease bias from a biased tester.
 

### 2014-02-18 macys/buildDirectedRecGraph.py
Add buildDirectedRecGraph.py

@@ -1,0 +1,48 @@
+#!/usr/local/bin/python
+
+"""
+Create and save an undirected graph from Recommends relations.
+"""
+import sqlite3
+import pickle
+
+# params
+db_fname = 'macys.db'
+save_fname = 'recDirectedGraph.pickle'
+
+selectRecommendsStmt = 'SELECT Id1, Id2 FROM Recommends'
+
+class Node(object):
+    def __init__(self):
+        self.inbound = []
+        self.outbound = []
+
+def makeGraph(db_curs):
+    graph = {}
+    node_cnt = 0
+    edge_cnt = 0
+    db_curs.execute(selectRecommendsStmt)
+    for row in db_curs.fetchall():
+        id1 = row[0]
+        id2 = row[1]
+        if id1 not in graph:
+            node_cnt += 1
+            graph[id1] = Node()
+        graph[id1].outbound.append(id2)
+        if id2 not in graph:
+            node_cnt += 1
+            graph[id2] = Node()
+        graph[id2].inbound.append(id1)
+        edge_cnt += 1
+    print '%d nodes, %d edges' % (node_cnt, edge_cnt)
+    return graph
+
+def main():
+    db_conn = sqlite3.connect(db_fname)
+    with db_conn:
+        db_curs = db_conn.cursor()
+        graph = makeGraph(db_curs)
+        pickle.dump(graph, open(save_fname, 'w'))
+
+if __name__ == '__main__':
+    main()

### 2014-02-18 common/buildDirectedRecGraph.py
Generalize buildRecGraph.py and buildDirectedRecGraph.py.

@@ -5,10 +5,7 @@ Create and save an undirected graph from Recommends relations.
 """
 import sqlite3
 import pickle
-
-# params
-db_fname = 'macys.db'
-save_fname = 'recDirectedGraph.pickle'
+from optparse import OptionParser
 
 selectRecommendsStmt = 'SELECT Id1, Id2 FROM Recommends'
 
@@ -16,6 +13,15 @@ class Node(object):
     def __init__(self):
         self.inbound = []
         self.outbound = []
+
+def getParser(usage=None):
+    parser = OptionParser(usage=usage)
+    parser.add_option('-d', '--database', dest='dbname', default=None,
+        help='sqlite3 database file containing Recommends table.', metavar='FILE')
+    parser.add_option('-o', '--output', dest='outfilename',
+        default='recDirectedGraph.pickle',
+        help='Output pickle file containing recommendation graph.', metavar='OUT_FILE')
+    return parser
 
 def makeGraph(db_curs):
     graph = {}
@@ -38,11 +44,18 @@ def makeGraph(db_curs):
     return graph
 
 def main():
-    db_conn = sqlite3.connect(db_fname)
+    # Parse options
+    usage = 'Usage: %prog [options] -d database'
+    parser = getParser(usage=usage)
+    (options, args) = parser.parse_args()
+    if not options.dbname:
+        parser.error('Database required')
+
+    db_conn = sqlite3.connect(options.dbname)
     with db_conn:
         db_curs = db_conn.cursor()
         graph = makeGraph(db_curs)
-        pickle.dump(graph, open(save_fname, 'w'))
+        pickle.dump(graph, open(options.outfilename, 'w'))
 
 if __name__ == '__main__':
     main()

### 2014-02-18 common/buildRecGraph.py

@@ -5,12 +5,17 @@ Create and save an undirected graph from Recommends relations.
 """
 import sqlite3
 import pickle
-
-# params
-db_fname = 'macys.db'
-save_fname = 'recGraph.pickle'
+from optparse import OptionParser
 
 selectRecommendsStmt = 'SELECT Id1, Id2 FROM Recommends'
+
+def getParser(usage=None):
+    parser = OptionParser(usage=usage)
+    parser.add_option('-d', '--database', dest='dbname', default=None,
+        help='sqlite3 database file containing Recommends table.', metavar='FILE')
+    parser.add_option('-o', '--output', dest='outfilename', default='recGraph.pickle',
+        help='Output pickle file containing recommendation graph.', metavar='OUT_FILE')
+    return parser
 
 def makeGraph(db_curs):
     graph = {}
@@ -35,11 +40,18 @@ def makeGraph(db_curs):
     return graph
 
 def main():
-    db_conn = sqlite3.connect(db_fname)
+    # Parse options
+    usage = 'Usage: %prog [options] -d database'
+    parser = getParser(usage=usage)
+    (options, args) = parser.parse_args()
+    if not options.dbname:
+        parser.error('Database required')
+
+    db_conn = sqlite3.connect(options.dbname)
     with db_conn:
         db_curs = db_conn.cursor()
         graph = makeGraph(db_curs)
-        pickle.dump(graph, open(save_fname, 'w'))
+        pickle.dump(graph, open(options.outfilename, 'w'))
 
 if __name__ == '__main__':
     main()

### 2014-02-18 common/Factory.py
Add Product.py and Factory.py.

@@ -1,0 +1,51 @@
+#!/usr/local/bin/python
+
+import sqlite3
+from Product import Product
+
+class Factory(object):
+    """Generates Product objects from various retailer databases.
+    """
+
+    def __init__(self):
+        self.retailers = {}
+
+    def registerRetailer(self, db_conn, name):
+        self.retailers[name] = db_conn
+
+    def getRetailer(self, name):
+        return self.retailers[name]
+
+    def getProduct(self, retailer, productId):
+        if retailer not in self.retailers:
+            raise ValueError('Unrecognized retailer.')
+        return Product(self.retailers[retailer], productId, retailer=retailer)
+
+
+##########################
+# ROUGH UNIT TEST
+##########################
+
+# params
+db_name = 'macys/macys.db'
+
+def main():
+    db_conn = sqlite3.connect(db_name)
+    factory = Factory()
+    retailer = 'Macys'
+    factory.registerRetailer(db_conn, retailer)
+    product = factory.getProduct(retailer, 827236)
+    product.printProduct()
+    print '====== Outgoing Edges ======'
+    for productId in product.outgoing:
+        p = factory.getProduct(product.retailer, productId)
+        p.printProduct()
+    print '############################'
+    print '====== Incoming Edges ======'
+    for productId in product.incoming:
+        p = factory.getProduct(product.retailer, productId)
+        p.printProduct()
+    print '############################'
+
+if __name__ == '__main__':
+    main()

### 2014-02-18 common/Product.py

@@ -1,0 +1,73 @@
+#!/usr/local/bin/python
+
+import sqlite3
+
+# params
+selectProductStmt =\
+    ('SELECT Url, ImgFile, Description, Prices, Available, Time, HaveRecs '
+     'FROM Products WHERE Id = :Id')
+selectCategoriesStmt =\
+     'SELECT ParentCategory, Category FROM Categories WHERE Id = :Id'
+selectOutgoingEdgesStmt = 'SELECT Id2 FROM Recommends WHERE Id1 = :Id'
+selectIncomingEdgesStmt = 'SELECT Id1 FROM Recommends WHERE Id2 = :Id'
+
+class Product(object):
+    """Represents a Product from a retailer database.
+    """
+
+    def __init__(self, db_conn, productId, retailer=None):
+        self.id = productId
+        self.retailer = retailer
+        db_curs = db_conn.cursor()
+        # Get Product relation
+        db_curs.execute(selectProductStmt, {'Id': productId})
+        if db_curs.rowcount == 0:
+            raise ValueError('productId not found in database')
+        row = db_curs.fetchone()
+        self.url = row[0]
+        self.imgfile = row[1]
+        self.description = row[2]
+        self.prices = row[3]
+        self.available = row[4]
+        self.time = row[5]
+        self.haverecs = row[6]
+        # Get Categories relations
+        db_curs.execute(selectCategoriesStmt, {'Id': productId})
+        self.categories = [(row[0], row[1]) for row in db_curs.fetchall()]
+        # Get outgoing edges
+        db_curs.execute(selectOutgoingEdgesStmt, {'Id': productId})
+        self.outgoing = [row[0] for row in db_curs.fetchall()]
+        # Get incoming edges
+        db_curs.execute(selectIncomingEdgesStmt, {'Id': productId})
+        self.incoming = [row[0] for row in db_curs.fetchall()]
+
+    def printProduct(self):
+        print '  productId = %d' % self.id
+        print '   retailer = %s' % self.retailer
+        print '        url = %s' % self.url
+        print '    imgfile = %s' % self.imgfile
+        print 'description = %s' % self.description
+        print '     prices = %s' % self.prices
+        print '  available = %d' % self.available
+        print '       time = %s' % self.time
+        print '   haverecs = %d' % self.haverecs
+        for cat in self.categories:
+            print '  parentCat = %s, cat = %s' % (cat[0], cat[1])
[…]

### 2014-02-18 common/buildDirectedRecGraph.py

@@ -11,8 +11,8 @@ selectRecommendsStmt = 'SELECT Id1, Id2 FROM Recommends'
 
 class Node(object):
     def __init__(self):
-        self.inbound = []
-        self.outbound = []
+        self.incoming = []
+        self.outgoing = []
 
 def getParser(usage=None):
     parser = OptionParser(usage=usage)
@@ -34,11 +34,11 @@ def makeGraph(db_curs):
         if id1 not in graph:
             node_cnt += 1
             graph[id1] = Node()
-        graph[id1].outbound.append(id2)
+        graph[id1].outoing.append(id2)
         if id2 not in graph:
             node_cnt += 1
             graph[id2] = Node()
-        graph[id2].inbound.append(id1)
+        graph[id2].incoming.append(id1)
         edge_cnt += 1
     print '%d nodes, %d edges' % (node_cnt, edge_cnt)
     return graph

### 2014-02-19 common/buildDirectedRecGraph.py
Add RRBFSDirectoredPredictor and update MagicWardrobe to use it.

@@ -9,11 +9,6 @@ from optparse import OptionParser
 
 selectRecommendsStmt = 'SELECT Id1, Id2 FROM Recommends'
 
-class Node(object):
-    def __init__(self):
-        self.incoming = []
-        self.outgoing = []
-
 def getParser(usage=None):
     parser = OptionParser(usage=usage)
     parser.add_option('-d', '--database', dest='dbname', default=None,
@@ -33,12 +28,12 @@ def makeGraph(db_curs):
         id2 = row[1]
         if id1 not in graph:
             node_cnt += 1
-            graph[id1] = Node()
-        graph[id1].outoing.append(id2)
+            graph[id1] = ([],[])
+        graph[id1][0].append(id2)
         if id2 not in graph:
             node_cnt += 1
-            graph[id2] = Node()
-        graph[id2].incoming.append(id1)
+            graph[id2] = ([],[])
+        graph[id2][1].append(id1)
         edge_cnt += 1
     print '%d nodes, %d edges' % (node_cnt, edge_cnt)
     return graph

### 2014-02-19 macys/MagicWardrobe.py

@@ -1,10 +1,10 @@
 #!/usr/local/bin/python
 
-from RRBFSPredictor import RRBFSPredictor
+from RRBFSDirectedPredictor import RRBFSDirectedPredictor
 import pickle
 
 # params
-graph_fname = 'recGraph.pickle'
+graph_fname = 'recDirectedGraph.pickle'
 
 class MagicWardrobe:
     """
@@ -14,7 +14,7 @@ class MagicWardrobe:
     def __init__(self):
         """Initializes predictor/feedback engine."""
         graph = self._loadGraph()
-        self.predictor = RRBFSPredictor(graph)
+        self.predictor = RRBFSDirectedPredictor(graph)
         pass
 
     def initSession(self, sessionId):

### 2014-02-19 macys/RRBFSDirectedPredictor.py

@@ -1,0 +1,85 @@
+#!/usr/local/bin/python
+
+from ItemPredictor import ItemPredictor
+from RoundRobinBFS_directed import roundRobinBFS
+from collections import deque
+
+class RRBFSDirectedPredictor(ItemPredictor):
+    """An ItemPredictor that is implemented using a round robin BFS algorithm on
+    a recommendation graph.
+    """
+
+    class Session(object):
+        """Inner class that defines a user session."""
+        def __init__(self):
+            self.likes = set()
+            self.dislikes = set()
+            self.sources = deque()
+
+    def __init__(self, graph):
+        self.graph = graph
+        self.sessions = {}
+
+    def initSession(self, sessionId):
+        self.sessions[sessionId] = self.Session()
+
+    def feedback(self, sessionId, likes, dislikes):
+        assert sessionId in self.sessions
+        likes = set(likes)
+        dislikes = set(dislikes)
+        assert likes.isdisjoint(dislikes)
+        assert likes.isdisjoint(self.sessions[sessionId].likes)
+        assert dislikes.isdisjoint(self.sessions[sessionId].dislikes)
+        self.sessions[sessionId].likes =\
+            self.sessions[sessionId].likes.union(likes)
+        self.sessions[sessionId].dislikes =\
+            self.sessions[sessionId].dislikes.union(dislikes)
+        [self.sessions[sessionId].sources.appendleft(item) for item in likes]
+
+    def nextItems(self, sessionId, num):
+        assert sessionId in self.sessions
+        items = []
+        firstTry = True
+        while len(items) < num:
+            if not firstTry:
+                # if necessary, moves dislikes to likes to find items
+                item = self.sessions[sessionId].dislikes.pop()
+                self.sessions[sessionId].likes.add(item)
+                self.sessions[sessionId].sources.appendleft(item)
+            items += roundRobinBFS(self.graph,
+                                   self.sessions[sessionId].sources,
+                                   self.sessions[sessionId].dislikes,
+                                   num - len(items))
+            firstTry = False
+        return items
+        
+        
+
+##########################
+# ROUGH UNIT TEST
+##########################
+
+import pickle
+
+# params
+graph_fname = 'recDirectedGraph.pickle'
+
+def loadGraph():
+    with open(graph_fname, 'r') as f:
+        graph = pickle.load(f)
+    return graph
+
+def main():
+    graph = loadGraph()
[…]

### 2014-02-19 macys/RoundRobinBFS_directed.py

@@ -1,0 +1,66 @@
+#!/usr/local/bin/python
+
+"""
+Performs a breadth first search (BFS) originating from multiple sources. Returns
+the first K nodes visited.
+"""
+
+from collections import deque
+
+def roundRobinBFS(graph, sources, ignores, k):
+
+    # create circular queue of queues
+    # and initialize each queue
+    queues = deque(maxlen=len(sources))
+    for s in sources:
+        q = deque()
+        for neighbor in graph[s][0]:
+            q.appendleft(neighbor)
+        queues.appendleft(q)
+   
+    found = [] 
+    visited = set(sources).union(set(ignores))
+    while queues:
+        q = queues.pop()
+        while q:
+            n = q.pop()
+            if n not in visited:
+                break
+        if n in visited:
+            continue
+        visited.add(n)
+        found.append(n)
+        if len(found) >= k:
+            break
+        for neighbor in graph[n][0]:
+            if neighbor not in visited:
+                q.appendleft(neighbor)
+        if q:
+            queues.appendleft(q)
+    return found
+
+
+##########################
+# ROUGH UNIT TEST
+##########################
+
+import pickle
+
+# params
+graph_fname = 'recDirectedGraph.pickle'
+
+def loadGraph():
+    with open(graph_fname, 'r') as f:
+        graph = pickle.load(f)
+    return graph
+
+def main():
+    graph = loadGraph()
+    sources = set([45, 74])
+    ignores = set([888757, 826391])
+    k = 10
+    found = roundRobinBFS(graph, sources, ignores, k)
+    print 'found =', found
+
+if __name__ == '__main__':
+    main()

### 2014-03-18 amazon/experiment.py
Implement predSim.

@@ -4,11 +4,13 @@ import multiprocessing as mp
 import Queue
 from optparse import OptionParser
 import sqlite3
+import numpy as np
 import csv
 import os
 import sys
 
 import similarity
+import predictions as pred
 
 # params
 stepSize = 10
@@ -46,13 +48,19 @@ def getParser(usage=None):
     parser.add_option('-c', '--cosineFunc', dest='cosineFunc',
         default='prefSim',
         help=('Similarity function to use: "prefSim" (default), "randSim", '
-              '"prefSimAlt1", or "randSimAlt1"'), metavar='FUNCNAME')
+              '"prefSimAlt1", "randSimAlt1", or "predSim"'), metavar='FUNCNAME')
     parser.add_option('-s', '--stepSize', dest='stepSize', type='int',
         default=None, help='Step size prefSim or prefSimAlt1.', metavar='NUM')
     parser.add_option('-K', dest='K', type='int', default=None,
         help='Parameter K for prefSimAlt1 or randSimAlt1.', metavar='NUM')
     parser.add_option('--sigma', dest='sigma', type='float', default=None,
         help='Parameter sigma for prefSimAlt1 or randSimAlt1.', metavar='FLOAT')
+    parser.add_option('--errorFileName', dest='errorFileName',
+        default='data/userErrors.csv', help='User prediction errors file.',
+        metavar='FILE')
+    parser.add_option('--epsilon', dest='epsilon', type='float', default=0.5,
+        help='Parameter epsilon for weightedRandSim or wightedPrefSim.',
+        metavar='FLOAT')
     return parser
 
 def getReviewPairTimes(reviewsA, reviewsB):
@@ -79,6 +87,9 @@ def getReviewPairTimes(reviewsA, reviewsB):
     return reviewPairTimes
 
 def expt1(db_conn, writer, cosineFunc, productId1, productId2):
+    # create covariance matrix (only used by predSim)
+    sigma = np.array([[similarity.sigmaXX, similarity.sigmaXY],
+                      [similarity.sigmaXY, similarity.sigmaXX]])
     db_curs = db_conn.cursor()
     # fetch product reviews
     db_curs.execute(selectReviewsOrderByTimeStmt, (productId1,))
@@ -110,20 +121,30 @@ def expt1(db_conn, writer, cosineFunc, productId1, productId2):
             stepReviews2 = sorted(reviews2[stepBegin_j:j], key=lambda x: x[1])
             stepBegin_i = i
             stepBegin_j = j
-
[…]

### 2014-03-18 amazon/errorHist.py
Add error and prediction histogram scripts.

@@ -1,0 +1,66 @@
+#!/usr/local/bin/python
+
+"""
+Generate histogram data for prediction errors aggregated by user via
+aggregatePredictions.py.
+"""
+
+from optparse import OptionParser
+import csv
+import os
+import sys
+
+# params
+userIdIdx = 0
+countIdx = 1
+averageIdx = 2
+varianceIdx = 3
+absAverageIdx = 4
+uncertaintyIdx = 5
+
+def getParser(usage=None):
+    parser = OptionParser(usage=usage)
+    parser.add_option('-b', '--bins', type='int', dest='numBins', default=100,
+        help='Number of bins.', metavar='NUM')
+    return parser
+
+def incrementBin(bins, error):
+    bins[int(round((error)*(len(bins) - 1)/2.0))] += 1
+
+def makeHistogram(reader, bins):
+    for row in reader:
+        absAverage = float(row[absAverageIdx])
+        incrementBin(bins, absAverage)
+    return bins
+
+def printHistogram(writer, bins):
+    for i in range(len(bins)):
+        writer.writerow([round((i + 0.5)*2.0/len(bins), 3), bins[i]])
+
+def main():
+    # Parse options
+    usage = 'Usage: %prog [options] <csvfile>'
+    parser = getParser(usage=usage)
+    (options, args) = parser.parse_args()
+    if len(args) != 1:
+        parser.error('Wrong number of arguments')
+    inputFileName = args[0]
+    if not os.path.isfile(inputFileName):
+        print >> sys.stderr, 'Cannot find input file: %s' % inputFileName
+        return
+
+    # Open input file
+    print >> sys.stderr, 'Reading from %s . . .' % inputFileName
+    inputfile = open(inputFileName, 'r')
+    reader = csv.reader(inputfile)
+
+    # Generate histogram data
+    bins = [0]*options.numBins
+    bins = makeHistogram(reader, bins)
+
+    # Print histogram data
+    writer = csv.writer(sys.stdout)
+    printHistogram(writer, bins)
+
+if __name__ == '__main__':
+    main()

### 2014-03-18 amazon/predictionHist.py

@@ -1,0 +1,85 @@
+#!/usr/local/bin/python
+
+"""
+Generate histogram data for predictions (aka "implied" similarities) produced by
+predictions.py.
+"""
+
+from optparse import OptionParser
+import csv
+import os
+import sys
+import math
+from fnmatch import fnmatch
+
+# params
+userIdIdx = 0
+rating1Idx = 1
+rating2Idx = 2
+predictionIdx = 3
+errorIdx = 4
+
+def getParser(usage=None):
+    parser = OptionParser(usage=usage)
+    parser.add_option('-b', '--bins', type='int', dest='numBins', default=100,
+        help='Number of bins.', metavar='NUM')
+    parser.add_option('-p', '--pattern', dest='pattern', default=None,
+        help='Input file pattern.', metavar='PATTERN')
+    parser.add_option('-c', '--cosineFunc', dest='cosineFunc',
+        default='prefSim',
+        help=('Similarity function to use: "prefSim" (default), "randSim", '
+              '"prefSimAlt1", or "randSimAlt1".'), metavar='FUNCNAME')
+    return parser
+
+def incrementBin(bins, cosineSim):
+    bins[int(round((1.0 + cosineSim)*(len(bins) - 1)/2.0))] += 1
+
+def makeHistogram(inputDir, pattern, bins):
+    for filename in os.listdir(inputDir):
+        fullpath = os.path.join(inputDir, filename)
+        if not os.path.isfile(fullpath):
+            continue
+        if not fnmatch(filename, pattern):
+            continue
+        print >> sys.stderr, 'Processing %s . . .' % filename
+        with open(fullpath, 'rb') as csvfile:
+            reader = csv.reader(csvfile)
+            for row in reader:
+                #userId = row[userIdIdx]
+                #rating1 = float(row[rating1Idx])
+                #rating2 = float(row[rating2Idx])
+                prediction = float(row[predictionIdx])
+                #error = float(row[errorIdx])
+                incrementBin(bins, prediction)
+    return bins
+
+def printHistogram(writer, bins):
+    for i in range(len(bins)):
+        writer.writerow([round((i + 0.5)*2.0/len(bins) - 1.0, 3), bins[i]])
+
+def main():
+    # Parse options
+    usage = 'Usage: %prog [options] <inputDir>'
+    parser = getParser(usage=usage)
+    (options, args) = parser.parse_args()
+    if len(args) != 1:
[…]

### 2014-03-18 amazon/regressionEval.py
adding regressionEval.py

@@ -1,0 +1,156 @@
+#!/usr/local/bin/python
+
+"""
+Evaluates the linear regression predictor
+"""
+
+from optparse import OptionParser
+import sqlite3
+import math
+import random
+import os
+import csv
+import numpy as np
+import mlpy
+
+def getParser(usage=None):
+    parser = OptionParser(usage=usage)
+    parser.add_option('-p', '--paramsFile', dest='paramsFile',
+            default='regressionParams.csv', help='File with regression' +
+            ' parameters', metavar='FILE')
+    parser.add_option('-o', '--outputFile', dest='outputFile', default='regResults.csv', 
+            help='File to output results of evaluation', metavar='FILE')
+    return parser
+
+def predict(params, n, x):
+    if not n in params:
+        return x
+    b = float(params[n][1])
+    m = float(params[n][2])
+    return m*x + b
+
+def addToDict(dictionary, index, val):
+    if not index in dictionary:
+        dictionary[index] = val
+    dictionary[index] += val
+
+def avgDict(total_dict, time):
+    avg_dict = {}
+    for num_in_common in time:
+        avg_dict[num_in_common] = total_dict[num_in_common]/time[num_in_common]
+    return avg_dict
+
+def writeTimeDicts(err, abs_err, sqr_err, reg_err, reg_abs_err, reg_sqr_err, outfile):
+    with open(outfile, 'wb') as out:
+        writer = csv.writer(out)
+        # header
+        header = [];
+        header.append("num_in_common")
+        header.append("avg_error")
+        header.append("avg_abs_error")
+        header.append("avg_sqr_error")
+        header.append("reg_avg_error")
+        header.append("reg_avg_abs_error")
+        header.append("reg_avg_sqr_error")
+        writer.writerow(header)
+        for num_in_common in err:
+            output = [];
+            output.append(num_in_common)
+            output.append(err[num_in_common])
+            output.append(abs_err[num_in_common])
+            output.append(sqr_err[num_in_common])
+            output.append(reg_err[num_in_common])
+            output.append(reg_abs_err[num_in_common])
+            output.append(reg_sqr_err[num_in_common])
+            writer.writerow(output)
+
+def main():
+    # Parse options
+    usage = 'Usage: %prog [options] testFile'
+    parser = getParser(usage=usage)
+    (options, args) = parser.parse_args()
+    if len(args) != 1:
+        parser.error('Wrong number of arguments')
+    testFile = args[0]
[…]

### 2014-03-18 amazon/regression.py
adding modified regression.py (?)

@@ -14,17 +14,11 @@ import csv
 import numpy as np
 import mlpy
 
-# params
-trainFileTemplate = 'Train_%s_in_common.csv'
-testFileName = 'Testset.csv'
-
 def getParser(usage=None):
     parser = OptionParser(usage=usage)
     parser.add_option('-o', '--outputFile', dest='outFile',
             default='regressionParams.csv', help='Output file for regression' +
             ' parameters', metavar='FILE')
-    parser.add_option('-s', '--step', dest='step', type='int', default=1,
-        help='Step size for K.', metavar='NUM')
     return parser
 
 def linearRegression(x, y):

### 2014-03-18 amazon/cosineSimSim.py
Change cosineSimSim to dynamically adjust covariance.

@@ -16,14 +16,18 @@ def getParser(usage=None):
     parser.add_option('--sigmaXX', type='float', dest='sigmaXX', default=1.0,
        help=('Diagonal component of covariance matrix for multivariate '
              'Gaussian distribution prior for ratings.'), metavar='FLOAT')
-    parser.add_option('--sigmaXY', type='float', dest='sigmaXY', default=0.3,
-       help=('Off-diagonal component of covariance matrix for multivariate '
-             'Gaussian distribution prior for ratings.'), metavar='FLOAT')
+    parser.add_option('--rho', type='float', dest='rho', default=0.1,
+       help=('Scalar constant for off-diagonal component of covariance matrix '
+             'for multivariate Gaussian distribution prior for ratings.'),
+             metavar='FLOAT')
     return parser
 
-def cosineSimSim(dimensions, mu, sigma):
-    assert(sigma.shape == (2, 2))
-    assert(mu.shape == (2,))
+def cosineSimSim(dimensions, sigmaXX, rho, rating1, rating2):
+    # Form mu vector and sigma matrix
+    mu = np.array([rating1, rating2])
+    sigma = np.array([[sigmaXX, rho*rating1*rating2],
+                      [rho*rating1*rating2, sigmaXX]])
+
     innerProd = 0
     var1 = 0
     var2 = 0
@@ -49,13 +53,9 @@ def main():
     # Seed random module
     random.seed(options.randomSeed)
 
-    # Form mu vector and sigma matrix
-    mu = np.array([rating1, rating2])
-    sigma = np.array([[options.sigmaXX, options.sigmaXY],
-                      [options.sigmaXY, options.sigmaXX]])
-
     # Run simulation
-    simsim = cosineSimSim(options.dimensions, mu, sigma)
+    simsim = cosineSimSim(options.dimensions, options.sigmaXX, options.rho,
+                          rating1, rating2)
 
     # Output
     print simsim

### 2014-03-18 amazon/predictions.py

@@ -46,13 +46,13 @@ def getParser(usage=None):
     parser.add_option('--sigmaXX', type='float', dest='sigmaXX', default=1.0,
        help=('Diagonal component of covariance matrix for multivariate '
              'Gaussian distribution prior for ratings.'), metavar='FLOAT')
-    parser.add_option('--sigmaXY', type='float', dest='sigmaXY', default=0.3,
-       help=('Off-diagonal component of covariance matrix for multivariate '
-             'Gaussian distribution prior for ratings.'), metavar='FLOAT')
+    parser.add_option('--rho', type='float', dest='rho', default=0.1,
+       help=('Scalar constant for off-diagonal component of covariance matrix '
+             'for multivariate Gaussian distribution prior for ratings.'),
+             metavar='FLOAT')
     return parser
 
-def getPredictions(dimensions, sigma, reviews1, reviews2):
-    assert(sigma.shape == (2, 2))
+def getPredictions(dimensions, sigmaXX, rho, reviews1, reviews2):
     predictions = []
     # compute product biases
     bias1 = np.mean([review[2] for review in reviews1])
@@ -76,18 +76,16 @@ def getPredictions(dimensions, sigma, reviews1, reviews2):
                 continue # ignore duplicate reviews
             rating1 = reviews1[i][2] - bias1
             rating2 = reviews2[j][2] - bias2
-            # make mu vector
-            mu = np.array([rating1, rating2])
             # make prediction
             predictions.append((userId1, rating1, rating2,
-                               cosineSimSim(dimensions, mu, sigma)))
+                               cosineSimSim(dimensions, sigmaXX, rho,
+                                            rating1, rating2)))
             i += 1
             j += 1
     return predictions
 
-def processPair(db_conn, writer, dimensions, sigma, cosineFunc,
+def processPair(db_conn, writer, dimensions, sigmaXX, rho, cosineFunc,
                 productId1, productId2):
-    assert(sigma.shape == (2, 2))
     db_curs = db_conn.cursor()
     # fetch product reviews
     db_curs.execute(selectReviewsStmt, (productId1,))
@@ -97,15 +95,14 @@ def processPair(db_conn, writer, dimensions, sigma, cosineFunc,
     # compute cosine similarity using the provided function.
     cosineSim, numUserCommon = cosineFunc(reviews1, reviews2)
     # get predictions
-    predictions = getPredictions(dimensions, sigma, reviews1, reviews2)
[…]

### 2014-03-18 amazon/similarity.py

@@ -6,10 +6,10 @@ import csv
 import sys
 from collections import defaultdict
 
-import cosineSimSim
 import predictions as pred
 
 # params
+constSimScore = 0.33
 K = 20
 sigma = 1.0
 writer = csv.writer(sys.stdout)
@@ -22,6 +22,13 @@ sigmaXY = 0.3
 userIdIdx = 0
 averageIdx = 2
 varianceIdx = 3
+absAverageIdx = 4
+
+def constSim(reviewsA, reviewsB):
+    """Provides a baseline similarity function by always returning the
+       average item-item similarity.
+    """
+    return (constSimScore, countUsersCommon(reviewsA, reviewsB))
 
 def prefSim(reviewsA, reviewsB):
     """Compute the "preferential" similarity between items A and B.
@@ -74,12 +81,27 @@ def predSim(reviewsA, reviewsB):
     """Compute the average "implicitly" predicted similarity between
        items A and B.
     """
-    print 'DBG: EXECUTING predSim()'
     sigma = np.array([[sigmaXX, sigmaXY],
                       [sigmaXY, sigmaXX]])
     predictions = pred.getPredictions(dimensions, sigma, reviewsA, reviewsB)
     avgPrediction = np.mean([p[3] for p in predictions])
     return (avgPrediction, len(predictions))
+
+def weightedPredSim(reviewsA, reviewsB):
+    """Compute the average "implicitly" predicted similarity between
+       items A and B.
+    """
+    sigma = np.array([[sigmaXX, sigmaXY],
+                      [sigmaXY, sigmaXX]])
+    predictions = pred.getPredictions(dimensions, sigma, reviewsA, reviewsB)
+    printPredictions(predictions, weights)
+    avgPrediction = np.mean([p[3]*weights[p[0]] for p in predictions])
+    return (avgPrediction, len(predictions))
+
+def printPredictions(predictions, weights):
+    for p in predictions:
+        print 'W: %0.3f, R1: % 0.3f, R2: % 0.3f, P: % 0.3f' %\
+            (weights[p[0]], p[1], p[2], p[3])
 
 def cosineSim(reviewsA, reviewsB, onlyCommon=False, fudgeFactor=None,
               weights=None):
@@ -158,6 +180,29 @@ def cosineSim(reviewsA, reviewsB, onlyCommon=False, fudgeFactor=None,
             varB += (K - numUsersCommon)*sigma**2
[…]

### 2014-03-19 amazon/experiment.py
Incorporate regSim into experiment.

@@ -49,6 +49,16 @@ def getParser(usage=None):
         default='prefSim',
         help=('Similarity function to use: "prefSim" (default), "randSim", '
               '"prefSimAlt1", "randSimAlt1", or "predSim"'), metavar='FUNCNAME')
+    parser.add_option('--regSimRawFunc', dest='regSimRawFunc',
+        default='randSim',
+        help=('Similarity function to used for raw similarity by regSim: '
+              '"prefSim", or "randSim" (default)'), metavar='FUNCNAME')
+    parser.add_option('--regSimParamsFile', dest='regSimParamsFile',
+        default='output/regSim_params.csv', help='Parameter file for regSim.',
+        metavar='FILE')
+    parser.add_option('--max-common-reviewers', dest='maxUsersCommon',
+        type='int', default=100,
+        help='Maximum number of common reviewers for regSim.', metavar='NUM')
     parser.add_option('-s', '--stepSize', dest='stepSize', type='int',
         default=None, help='Step size prefSim or prefSimAlt1.', metavar='NUM')
     parser.add_option('-K', dest='K', type='int', default=None,
@@ -275,6 +285,12 @@ def main():
         print >> sys.stderr,\
             'Invalid Similarity function: %s' % options.cosineFunc
         return
+    try:
+        regSimRawFunc = getattr(similarity, options.regSimRawFunc)
+    except KeyError:
+        print >> sys.stderr,\
+            'Invalid Raw Similarity function: %s' % options.regSimRawFunc
+        return
     if options.outputDir:
         outputDir = options.outputDir
     else:
@@ -293,6 +309,11 @@ def main():
         similarity.K = options.K
     if options.sigma:
         similarity.sigma = options.sigma
+
+    # regSim
+    if options.cosineFunc == 'regSim':
+        similarity.initRegSim(options.regSimParamsFile, regSimRawFunc,
+                   options.maxUsersCommon)
 
     # weighted similarity
     if options.cosineFunc == 'weightedRandSim':

### 2014-03-19 amazon/regressionTrain.py

@@ -1,0 +1,83 @@
+#!/usr/local/bin/python
+
+"""
+Perform linear regressions on similarity scores.
+Take as input the output of regressionTrainData.py and produce parameters file
+for use by regSim.
+"""
+
+from optparse import OptionParser
+import csv
+import os
+import sys
+import mlpy
+
+# params
+scoreIdx = 0
+trueScoreIdx = 1
+
+def getParser(usage=None):
+    parser = OptionParser(usage=usage)
+    parser.add_option('-m', '--max-common-reviewers', dest='maxUsersCommon',
+        type='int', default=100, help='Maximum number of common reviewers.',
+        metavar='NUM')
+    return parser
+
+def linearRegression(x, y):
+    x2d = []
+    for i in x:
+        x2d.append([i, 0])
+    ols = mlpy.OLS()
+    ols.learn(x2d, y)
+    params = []
+    params.append(ols.beta0())
+    params.append(ols.beta()[0])
+    return params
+
+def getSqrError(params, x, y):
+    totalSqrError = 0
+    b = params[0]
+    m = params[1]
+    for i in range(len(x)):
+        prediction = m*float(x[i]) + b
+        totalSqrError += (prediction - float(y[i]))*(prediction - float(y[i]))
+    return totalSqrError
+
+def main():
+    # Parse options
+    usage = 'Usage: %prog [options] inputDir'
+    parser = getParser(usage=usage)
+    (options, args) = parser.parse_args()
+    if len(args) != 1:
+        parser.error('Wrong number of arguments')
+    inputDir = args[0]
+    if not os.path.isdir(inputDir):
+        print >> sys.stderr, 'Cannot find: %s' % inputDir
+        return
+
+    # setup output writer
+    writer = csv.writer(sys.stdout)
+
+    # process data
+    inputfileTemplate = '%d.csv'
+    for numUsersCommon in range(1, options.maxUsersCommon+1):
+        inputfileName = inputfileTemplate % numUsersCommon
+        #print 'Processing %s . . .' % inputfileName
+        fullpath = os.path.join(inputDir, inputfileName)
+        if not os.path.isfile(fullpath):
+            print >> sys.stderr, 'ERROR: Cannot find %s' % fullpath
+            break
+        with open(fullpath, 'rb') as csvfile:
+            reader = csv.reader(csvfile)
+            # read data
+            data = [row for row in reader]
+            scores = [row[scoreIdx] for row in data]
+            trueScores = [row[trueScoreIdx] for row in data]
[…]

### 2014-03-19 amazon/regressionTrainData.py

@@ -1,0 +1,74 @@
+#!/usr/local/bin/python
+
+"""
+Takes output files from simVsK.py and generates input files for regression.py.
+"""
+
+from optparse import OptionParser
+import sqlite3
+import csv
+import os
+import sys
+
+# params
+numUsersCommonIdx = 0
+scoreIdx = 1
+
+def getParser(usage=None):
+    parser = OptionParser(usage=usage)
+    parser.add_option('-o', '--output-dir', dest='outputDir',
+        default='regressionTrainOutput', help='Output directory.', metavar='DIR')
+    parser.add_option('-m', '--max-common-reviewers', dest='maxUsersCommon',
+        type='int', default=100, help='Maximum number of common reviewers.',
+        metavar='NUM')
+    return parser
+
+def getTrueScore(fullpath):
+    """Retrieves last score in file."""
+    trueScore = None
+    with open(fullpath, 'rb') as csvfile:
+        reader = csv.reader(csvfile)
+        for row in reader:
+            trueScore = row[scoreIdx]
+    return trueScore
+
+def main():
+    # Parse options
+    usage = 'Usage: %prog [options] inputDir'
+    parser = getParser(usage=usage)
+    (options, args) = parser.parse_args()
+    if len(args) != 1:
+        parser.error('Wrong number of arguments')
+    inputDir = args[0]
+    if not os.path.isdir(inputDir):
+        print >> sys.stderr, 'Cannot find: %s' % inputDir
+        return
+
+    # Create output writers
+    outfileTemplate = '%d.csv'
+    writers = [None] # create empty entry in order to index from 1
+    for i in range(options.maxUsersCommon):
+        outfileName = os.path.join(options.outputDir, outfileTemplate % (i+1))
+        outfile = open(outfileName, 'wb')
+        writers.append(csv.writer(outfile))
+
+    print 'Reading files in %s . . .' % inputDir
+    for filename in os.listdir(inputDir):
+        fullpath = os.path.join(inputDir, filename)
+        if not os.path.isfile(fullpath):
+            continue
+        print 'Processing %s . . .' % filename
+        trueScore = getTrueScore(fullpath)
+        with open(fullpath, 'rb') as csvfile:
+            reader = csv.reader(csvfile)
+            for row in reader:
+                numUsersCommon = int(row[numUsersCommonIdx])
+                if numUsersCommon > options.maxUsersCommon:
+                    break
+                score = float(row[scoreIdx])
+                # write output
[…]

### 2014-03-19 amazon/simVsK.py

@@ -33,7 +33,7 @@ def getParser(usage=None):
         default='prefSim',
         help=('Similarity function to use: "prefSim" (default), "randSim", '
               '"prefSimAlt1", or "randSimAlt1"'), metavar='FUNCNAME')
-    parser.add_option('-s', '--step', dest='step', type='int', default=10,
+    parser.add_option('-s', '--step', dest='step', type='int', default=1,
         help='Step size for K.', metavar='NUM')
     return parser
 

### 2014-03-19 amazon/similarity.py

@@ -9,7 +9,7 @@ from collections import defaultdict
 import predictions as pred
 
 # params
-constSimScore = 0.33
+constSimScore = 0.27
 K = 20
 sigma = 1.0
 writer = csv.writer(sys.stdout)
@@ -23,6 +23,17 @@ userIdIdx = 0
 averageIdx = 2
 varianceIdx = 3
 absAverageIdx = 4
+
+def regSim(reviewsA, reviewsB):
+    """Uses params from linear regressions on training data to translate
+       raw similarity to an improved esimate.
+    """
+    rawSim, numUsersCommon = regSim_rawFunc(reviewsA, reviewsB)
+    if numUsersCommon <= regSim_maxCommonUsers:
+        intercept, slope = regSimParams[numUsersCommon]
+        return (intercept + slope*rawSim, numUsersCommon)
+    else:
+        return (rawSim, numUsersCommon)
 
 def constSim(reviewsA, reviewsB):
     """Provides a baseline similarity function by always returning the
@@ -238,6 +249,35 @@ def initWeightedSim(errorFileName, epsilon):
     averageW = totalW/len(w)
     weights = defaultdict(lambda x: averageW, w)
 
+#########################
+#
+#  Linear Regression
+#
+##############################
+
+regSimParams = {}
+regSim_rawFunc = None
+regSim_maxCommonUsers = 0
+
+numUsersCommonIdx = 0
+interceptIdx = 1
+slopeIdx = 2
+
+def initRegSim(paramFileName, cosineFunc, maxCommonUsers):
+    global regSimParams
+    global regSim_rawFunc
+    global regSim_maxCommonUsers
+    regSim_rawFunc = cosineFunc
+    regSim_maxCommonUsers = maxCommonUsers
+    print 'Reading %s . . .' % paramFileName
+    with open(paramFileName, 'rb') as csvfile:
+        reader = csv.reader(csvfile)
+        for row in reader:
+            numUsersCommon = int(row[numUsersCommonIdx])
+            intercept = float(row[interceptIdx])
+            slope = float(row[slopeIdx])
+            regSimParams[numUsersCommon] = (intercept, slope)
+
 ############
 #
 #  Stand-alone similarity
@@ -263,6 +303,16 @@ def getParser(usage=None):
         default='prefSim',
         help=('Similarity function to use: "prefSim" (default), "randSim", '
               'or "predSim"'), metavar='FUNCNAME')
+    parser.add_option('--regSimRawFunc', dest='regSimRawFunc',
+        default='randSim',
[…]

### 2014-03-19 amazon/partitionPairs.py
Add partitionPairs.py.

@@ -1,0 +1,74 @@
+#!/usr/local/bin/python
+
+"""
+Partition product pairs into a training set and testing test.
+"""
+
+from optparse import OptionParser
+import os
+import sys
+import csv
+import random
+
+def getParser(usage=None):
+    parser = OptionParser(usage=usage)
+    parser.add_option('-f', '--fraction', type='float', dest='fraction',
+        default=0.3, help='Fraction of pairs in test set.', metavar='FLOAT')
+    parser.add_option('--trainOut', dest='trainOut', default=None,
+        help='Training pairs output file.', metavar='FILE')
+    parser.add_option('--testOut', dest='testOut', default=None,
+        help='Testing pairs output file.', metavar='FILE')
+    return parser
+
+def main():
+    # Parse options
+    usage = 'Usage: %prog [options] <csvfile>'
+    parser = getParser(usage=usage)
+    (options, args) = parser.parse_args()
+    if len(args) != 1:
+        parser.error('Wrong number of arguments')
+    inputFileName = args[0]
+    if not os.path.isfile(inputFileName):
+        print >> sys.stderr, 'Cannot find input file: %s' % inputFileName
+        return
+    progName = os.path.basename(sys.argv[0]).split('.')[0]
+    if options.trainOut:
+        trainOut = options.trainOut
+    else:
+        trainOut = 'data/%s.train.csv' % progName
+    if options.testOut:
+        testOut = options.testOut
+    else:
+        testOut = 'data/%s.test.csv' % progName
+
+    # open input file
+    print 'Reading from %s. . .' % inputFileName
+    inFile = open(inputFileName, 'rb') 
+    reader = csv.reader(inFile)
+
+    # open output files
+    print 'Writing training data to %s . . .' % trainOut
+    trainFile = open(trainOut, 'wb')
+    trainWriter = csv.writer(trainFile)
+    print 'Writing testing data to %s . . .' % testOut
+    testFile = open(testOut, 'wb')
+    testWriter = csv.writer(testFile)
+
+    # read and shuffle data
+    pairs = [(row[0], row[1]) for row in reader]
+    random.shuffle(pairs)
+
+    # write data
+    trainCount = 0
+    testCount = 0
+    for pair in pairs:
+       if random.random() >= options.fraction:
+           trainCount += 1
+           trainWriter.writerow(pair)
+       else:
+           testCount += 1
+           testWriter.writerow(pair)
[…]

### 2014-03-19 amazon/SimilarityGrid.py
Implement predSim.

@@ -1,0 +1,86 @@
+import collections
+import csv
+import sys
+
+class SimilarityGrid(collections.MutableMapping):
+    """A dictionary of similarity scores by rating pairs."""
+    def __init__(self, minRating, maxRating, stepRating):
+        self.minRating = minRating
+        self.maxRating = maxRating
+        self.stepRating = stepRating
+        self.store = {}
+        # initialize grid
+        for key in self:
+            self[key] = [0, 0]
+        # include null entry for bad keys
+        self[(self.maxRating, self.maxRating)] = None
+
+    def __keytransform__(self, key):
+        rating1 = min(key[0], key[1])
+        rating2 = max(key[0], key[1])
+        if rating1 < self.minRating or\
+           rating2 > self.maxRating - self.stepRating:
+            return -1
+        idx = 0
+        p = self.minRating
+        while p < self.maxRating:
+            q = p
+            while q < self.maxRating:
+                if rating1 >= p and rating1 < p + self.stepRating and\
+                   rating2 >= q and rating2 < q + self.stepRating:
+                    return idx
+                idx += 1
+                q += self.stepRating
+            p += self.stepRating
+        raise Exception("Unreachable code.")
+
+    def __getitem__(self, key):
+        return self.store[self.__keytransform__(key)]
+
+    def __setitem__(self, key, value):
+        self.store[self.__keytransform__(key)] = value
+
+    def __delitem__(self, key):
+        del self.store[self.__keytransform__(key)]
+
+    def __len__(self):
+        return len(self.store)
+
+    def __iter__(self):
+        self.current = [self.minRating, self.minRating]
+        return self
+
+    def next(self):
+        if self.current[0] >= self.maxRating and\
+           self.current[1] >= self.maxRating:
+            raise StopIteration
+        elif self.current[1] >= self.maxRating:
+           self.current[0] += self.stepRating
+           self.current[1] = self.current[0] + self.stepRating
+        else:
+           self.current[1] += self.stepRating
+        return (round(self.current[0], 3),
+                round(self.current[1] - self.stepRating, 3))
+
+    def writeToFile(self, outputFile):
+        writer = csv.writer(outputFile)
+        for (rating1, rating2) in self:
+            entry = self[(rating1, rating2)]
+            if entry:
+                writer.writerow([round(rating1, 3), round(rating2, 3),
+                                 entry[0], entry[1]])
+
+    def readFromFile(self, inputFile):
[…]

### 2014-03-19 amazon/experiment.py

@@ -48,7 +48,8 @@ def getParser(usage=None):
     parser.add_option('-c', '--cosineFunc', dest='cosineFunc',
         default='prefSim',
         help=('Similarity function to use: "prefSim" (default), "randSim", '
-              '"prefSimAlt1", "randSimAlt1", or "predSim"'), metavar='FUNCNAME')
+              '"prefSimAlt1", "randSimAlt1", "regSim" or "predSim"'),
+        metavar='FUNCNAME')
     parser.add_option('--regSimRawFunc', dest='regSimRawFunc',
         default='randSim',
         help=('Similarity function to used for raw similarity by regSim: '
@@ -59,6 +60,15 @@ def getParser(usage=None):
     parser.add_option('--max-common-reviewers', dest='maxUsersCommon',
         type='int', default=100,
         help='Maximum number of common reviewers for regSim.', metavar='NUM')
+    parser.add_option('--simGridFile', dest='simGridFile',
+        default='output/simGrid_randSim.csv',
+        help='CSV containing simGrid data.', metavar='FILE')
+    parser.add_option('--minRating', type='float', dest='minRating',
+        default=-2.0, help='Minimum rating.', metavar='FLOAT')
+    parser.add_option('--maxRating', type='float', dest='maxRating',
+        default=2.0, help='Maximum rating.', metavar='FLOAT')
+    parser.add_option('--stepRating', type='float', dest='stepRating',
+        default=0.1, help='Step rating.', metavar='FLOAT')
     parser.add_option('-s', '--stepSize', dest='stepSize', type='int',
         default=None, help='Step size prefSim or prefSimAlt1.', metavar='NUM')
     parser.add_option('-K', dest='K', type='int', default=None,
@@ -97,9 +107,6 @@ def getReviewPairTimes(reviewsA, reviewsB):
     return reviewPairTimes
 
 def expt1(db_conn, writer, cosineFunc, productId1, productId2):
-    # create covariance matrix (only used by predSim)
-    sigma = np.array([[similarity.sigmaXX, similarity.sigmaXY],
-                      [similarity.sigmaXY, similarity.sigmaXX]])
     db_curs = db_conn.cursor()
     # fetch product reviews
     db_curs.execute(selectReviewsOrderByTimeStmt, (productId1,))
@@ -131,30 +138,20 @@ def expt1(db_conn, writer, cosineFunc, productId1, productId2):
             stepReviews2 = sorted(reviews2[stepBegin_j:j], key=lambda x: x[1])
[…]

### 2014-03-19 amazon/makeSimGrid.py

@@ -1,0 +1,129 @@
+#!/usr/local/bin/python
+
+"""
+Compute the average similarity scores associated with rating pairs.
+"""
+
+from optparse import OptionParser
+import sqlite3
+import numpy
+import csv
+import os
+import sys
+import collections
+
+import similarity
+from SimilarityGrid import SimilarityGrid
+
+# db params
+dbTimeout = 5
+selectReviewsStmt =\
+    ('SELECT Time, UserId, AdjustedScore '
+     'FROM Reviews '
+     'WHERE ProductId = :ProductId '
+     'ORDER BY UserId')
+
+def getParser(usage=None):
+    parser = OptionParser(usage=usage)
+    parser.add_option('-d', '--database', dest='db_fname',
+        default='data/amazon.db', help='sqlite3 database file.', metavar='FILE')
+    parser.add_option('-c', '--cosineFunc', dest='cosineFunc',
+        default='randSim',
+        help='Similarity function to use: "prefSim" or "randSim" (default)',
+        metavar='FUNCNAME')
+    parser.add_option('--minRating', type='float', dest='minRating',
+        default=-2.0, help='Minimum rating.', metavar='FLOAT')
+    parser.add_option('--maxRating', type='float', dest='maxRating',
+        default=2.0, help='Maximum rating.', metavar='FLOAT')
+    parser.add_option('--stepRating', type='float', dest='stepRating',
+        default=0.1, help='Step rating.', metavar='FLOAT')
+    return parser
+
+def fillGrid(db_conn, inputfile, cosineFunc, simGrid):
+    count = 0
+    bad_keys = 0
+    db_curs = db_conn.cursor()
+    for line in inputfile:
+        tokens = line.split(',')
+        productId1 = tokens[0].strip()
+        productId2 = tokens[1].strip()
+        print >> sys.stderr, '%s, %s' % (productId1, productId2)
+        # fetch product reviews
+        db_curs.execute(selectReviewsStmt, (productId1,))
+        reviews1 = [row for row in db_curs.fetchall()]
+        db_curs.execute(selectReviewsStmt, (productId2,))
+        reviews2 = [row for row in db_curs.fetchall()]
+        # compute cosine similarity using the provided function.
+        cosineSim, numUserCommon = cosineFunc(reviews1, reviews2)
+        # compute product biases
+        bias1 = numpy.mean([review[2] for review in reviews1])
+        bias2 = numpy.mean([review[2] for review in reviews2])
+        # update grid
+        i = 0
+        j = 0
[…]

### 2014-03-19 amazon/predictSim.py

@@ -1,0 +1,48 @@
+#!/usr/local/bin/python
+
+from optparse import OptionParser
+import os
+import sys
+
+from SimilarityGrid import SimilarityGrid
+
+def getParser(usage=None):
+    parser = OptionParser(usage=usage)
+    parser.add_option('-f', '--simGridFile', dest='simGridFile',
+        default='output/simGrid_randSim.csv',
+        help='CSV containing simGrid data.', metavar='FILE')
+    parser.add_option('--minRating', type='float', dest='minRating',
+        default=-2.0, help='Minimum rating.', metavar='FLOAT')
+    parser.add_option('--maxRating', type='float', dest='maxRating',
+        default=2.0, help='Maximum rating.', metavar='FLOAT')
+    parser.add_option('--stepRating', type='float', dest='stepRating',
+        default=0.1, help='Step rating.', metavar='FLOAT')
+    return parser
+
+def main():
+    # Parse options
+    usage = 'Usage: %prog [options] rating1 rating2'
+    parser = getParser(usage=usage)
+    (options, args) = parser.parse_args()
+    if len(args) != 2:
+        parser.error('Wrong number of arguments')
+    rating1 = float(args[0])
+    rating2 = float(args[1])
+
+    # Initialize similarity grid
+    simGrid = SimilarityGrid(options.minRating, options.maxRating,
+                             options.stepRating)
+    inputFile = open(options.simGridFile, 'rb')
+    simGrid.readFromFile(inputFile)
+
+    # Get entry
+    scoreTotal = simGrid[(rating1, rating2)][0]
+    count = simGrid[(rating1, rating2)][1]
+    if count > 0:
+        estimate = scoreTotal/count
+    else:
+        estimate = '?'
+    print ('(%0.3f/%d) =' % (scoreTotal, count)), estimate
+ 
+if __name__ == '__main__':
+    main()

### 2014-03-19 amazon/similarity.py

@@ -5,8 +5,10 @@ import numpy as np
 import csv
 import sys
 from collections import defaultdict
+import sympy 
 
 import predictions as pred
+from SimilarityGrid import SimilarityGrid
 
 # params
 constSimScore = 0.27
@@ -17,6 +19,11 @@ step = sys.maxint
 dimensions = 100
 sigmaXX = 1.0
 sigmaXY = 0.3
+
+# modelSim params
+mu_s = constSimScore
+sigma_s = 0.2
+sigma_r = 0.1
 
 # errorFile
 userIdIdx = 0
@@ -92,12 +99,110 @@ def predSim(reviewsA, reviewsB):
     """Compute the average "implicitly" predicted similarity between
        items A and B.
     """
-    sigma = np.array([[sigmaXX, sigmaXY],
-                      [sigmaXY, sigmaXX]])
-    predictions = pred.getPredictions(dimensions, sigma, reviewsA, reviewsB)
-    avgPrediction = np.mean([p[3] for p in predictions])
-    return (avgPrediction, len(predictions))
+    #sigma = np.array([[sigmaXX, sigmaXY],
+    #                  [sigmaXY, sigmaXX]])
+    #predictions = pred.getPredictions(dimensions, sigma, reviewsA, reviewsB)
+    #avgPrediction = np.mean([p[3] for p in predictions])
+    #return (avgPrediction, len(predictions))
 
+    # compute product biases:
+    if reviewsA:
+        biasA = np.mean([review[2] for review in reviewsA])
+    else:
+        biasA = 0.0
+    if reviewsB:
+        biasB = np.mean([review[2] for review in reviewsB])
+    else:
+        biasB = 0.0
+    totalScore = 0
+    totalCount = 0
+    numUsersCommon = 0
+    i = 0
+    j = 0
+    while i < len(reviewsA) and j < len(reviewsB):
+        userIdA = reviewsA[i][1]
+        userIdB = reviewsB[j][1]
+        if userIdA < userIdB:
+            i += 1
+        elif userIdA > userIdB:
+            j += 1
+        else:
+            timeA = reviewsA[i][0]
+            timeB = reviewsB[j][0]
+            if timeA == timeB:
+                i += 1
+                j += 1
+                continue # ignore duplicate reviews
+            numUsersCommon += 1
+            scoreA = reviewsA[i][2] - biasA
+            scoreB = reviewsB[j][2] - biasB
+            entry = simGrid[(scoreA, scoreB)]
+            if entry:
[…]

### 2014-03-20 amazon/makeModelGrid.py
Implement modelSim.

@@ -1,0 +1,70 @@
+#!/usr/local/bin/python
+
+"""
+Compute the similarity score using a simple mode for rating pairs.
+"""
+
+from optparse import OptionParser
+import csv
+import os
+import sys
+import sympy
+
+from SimilarityGrid import SimilarityGrid
+
+def getParser(usage=None):
+    parser = OptionParser(usage=usage)
+    parser.add_option('--mu_s', type='float', dest='mu_s',
+        default=0.27, help='Mean of score distribution prior.', metavar='FLOAT')
+    parser.add_option('--sigma_s', type='float', dest='sigma_s',
+        default=0.2, help='Standard deviation of score distribution prior.',
+        metavar='FLOAT')
+    parser.add_option('--sigma_r', type='float', dest='sigma_r',
+        default=0.1, help='Standard deviation of rating distribution prior.',
+        metavar='FLOAT')
+    parser.add_option('--minRating', type='float', dest='minRating',
+        default=-2.0, help='Minimum rating.', metavar='FLOAT')
+    parser.add_option('--maxRating', type='float', dest='maxRating',
+        default=2.0, help='Maximum rating.', metavar='FLOAT')
+    parser.add_option('--stepRating', type='float', dest='stepRating',
+        default=0.1, help='Step rating.', metavar='FLOAT')
+    return parser
+
+def fillGrid(mu_s, sigma_s, sigma_r, simGrid):
+    for (rating1, rating2) in simGrid:
+        p = round(rating1*rating2, 3)
+        q = round(rating1**2 + rating2**2, 3)
+        s = sympy.Symbol('s')
+        solutions =\
+            sympy.solve(-p*s**2 + q*s - p +\
+                        (sigma_r/sigma_s)**2*(1 - s**2)**2*(s - mu_s), s)
+        if solutions:
+            score = solutions[0]
+        else:
+            print >> sys.stderr, 'WARNING: No solution to eq found.'
+            score = mu_s
+        print >> sys.stderr,\
+            'rating1 = % 0.3f, rating2 = % 0.3f, score = % 0.3f' %\
+            (rating1, rating2, score)
+        if simGrid[(rating1, rating2)]:
+            simGrid[(rating1, rating2)][0] = score
+        
+def main():
+    # Parse options
+    usage = 'Usage: %prog [options]'
+    parser = getParser(usage=usage)
+    (options, args) = parser.parse_args()
+
+    # Initialize similarity grid
[…]

### 2014-03-20 amazon/modelPredictions.py

@@ -1,0 +1,234 @@
+#!/usr/local/bin/python
+
+import multiprocessing as mp
+import Queue
+from optparse import OptionParser
+import sqlite3
+import csv
+import os
+import sys
+import numpy as np
+from sympy import Symbol, solve
+from sympy.solvers.solvers import check_assumptions
+
+import similarity
+
+# params
+outputTemplateTemplate = '%s_%%s_%%s.csv'
+workerTimeout = 60
+workerQueueSize = 100
+END_OF_QUEUE = -1
+
+# db params
+dbTimeout = 5
+selectReviewsStmt =\
+    ('SELECT Time, UserId, AdjustedScore '
+     'FROM Reviews '
+     'WHERE ProductId = :ProductId '
+     'ORDER BY UserId')
+
+def getParser(usage=None):
+    parser = OptionParser(usage=usage)
+    parser.add_option('-w', '--numWorkers', dest='numWorkers', type='int',
+        default=1, help='Number of worker processes.', metavar='NUM')
+    parser.add_option('-d', '--database', dest='db_fname',
+        default='data/amazon.db', help='sqlite3 database file.', metavar='FILE')
+    parser.add_option('-o', '--output-dir', dest='outputDir',
+        default='modelPredictions', help='Output directory.', metavar='DIR')
+    parser.add_option('-c', '--cosineFunc', dest='cosineFunc',
+        default='prefSim',
+        help='Similarity function to use: "prefSim" (default) or "randSim"',
+              metavar='FUNCNAME')
+    parser.add_option('--modelGridFile', dest='modelGridFile',
+        default='output/simGrid_modelSim.csv',
+        help='CSV containing simGrid data for modelSim.', metavar='FILE')
+    parser.add_option('--sigma_r', type='float', dest='sigma_r',
+        default=0.2, help='Standard deviation of rating distribution.',
+        metavar='FLOAT')
+    parser.add_option('--sigma_s', type='float', dest='sigma_s',
+        default=0.2, help='Standard deviation of score distribution.',
+        metavar='FLOAT')
+    parser.add_option('--minRating', type='float', dest='minRating',
+        default=-2.0, help='Minimum rating.', metavar='FLOAT')
+    parser.add_option('--maxRating', type='float', dest='maxRating',
+        default=2.0, help='Maximum rating.', metavar='FLOAT')
+    parser.add_option('--stepRating', type='float', dest='stepRating',
[…]

### 2014-03-20 amazon/similarity.py

@@ -147,6 +147,52 @@ def predSim(reviewsA, reviewsB):
     else:
         return (constSimScore, numUsersCommon)
 
+#def modelSim(reviewsA, reviewsB):
+#    """Compute the average "implicitly" predicted similarity between
+#       items A and B according to a simple model.
+#    """
+#    # compute product biases:
+#    if reviewsA:
+#        biasA = np.mean([review[2] for review in reviewsA])
+#    else:
+#        biasA = 0.0
+#    if reviewsB:
+#        biasB = np.mean([review[2] for review in reviewsB])
+#    else:
+#        biasB = 0.0
+#    totalScore = 0
+#    totalCount = 0
+#    numUsersCommon = 0
+#    i = 0
+#    j = 0
+#    while i < len(reviewsA) and j < len(reviewsB):
+#        userIdA = reviewsA[i][1]
+#        userIdB = reviewsB[j][1]
+#        if userIdA < userIdB:
+#            i += 1
+#        elif userIdA > userIdB:
+#            j += 1
+#        else:
+#            timeA = reviewsA[i][0]
+#            timeB = reviewsB[j][0]
+#            if timeA == timeB:
+#                i += 1
+#                j += 1
+#                continue # ignore duplicate reviews
+#            numUsersCommon += 1
+#            scoreA = reviewsA[i][2] - biasA
+#            scoreB = reviewsB[j][2] - biasB
+#            entry = modelGrid[(scoreA, scoreB)]
+#            if entry and entry[0] >= -1 and entry[0] <= 1:
+#                totalScore += entry[0]
+#                totalCount += math.sqrt(scoreA**2 + scoreB**2)
+#            i += 1
+#            j += 1
+#    if totalCount > 0:
+#        return (totalScore/totalCount, numUsersCommon)
+#    else:
+#        return (constSimScore, numUsersCommon)
+
 def modelSim(reviewsA, reviewsB):
     """Compute the MLE similarity based upon a simple model."""
     if reviewsA:
@@ -202,7 +248,7 @@ def modelSim(reviewsA, reviewsB):
 
     #print solutions
     return (solutionTotal/weightTotal, numUsersCommon)
-    
+   
 def weightedPredSim(reviewsA, reviewsB):
     """Compute the average "implicitly" predicted similarity between
        items A and B.
@@ -384,6 +430,20 @@ def initRegSim(paramFileName, cosineFunc, maxCommonUsers):
 
 #################
 #
+#  modelSim
+#
+####################
+
+modelGrid = None
+
[…]

### 2014-03-20 paper/paper.tex
Started paper

\documentclass[10pt]{article}

% =========================================================================
% document style changes
% =========================================================================

\usepackage{amsmath}                    % AMS math packages
\usepackage{amssymb}                    %
\usepackage[]{graphpap}
\usepackage[T1]{fontenc}                % for \mathrm{}
\usepackage{courier}                    % for \texttt{}
\usepackage{bbm}                        % for \mathbbm{1} (indicator function)
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}

%\setlength{\parskip}{\baselineskip}     % skip line following paragraphs
\setlength{\parskip}{0.0in}
\setlength{\parindent}{0.3in}           % Control margins and amount of text
\setlength{\topmargin}{-0.8in}
\setlength{\oddsidemargin}{-.15in}      % changed from {-.15in}
\setlength{\textheight}{9.5in}
\setlength{\textwidth}{6.8in}
\pagestyle{empty}                       % No page numbers

\newcommand{\spc}{\vspace{0.25in}}      % Shortcut commands
\newcommand{\ds}{\displaystyle}         %\newcommand{\ds}[1]{\displaystyle{#1}}
\newcommand{\ra}{\rightarrow}
\DeclareMathOperator*{\argmax}{arg\!\max}
\DeclareMathOperator*{\argmin}{arg\!\min}

\begin{document}                        % This is where the document begins

{\LARGE\bf
\begin{tabbing}
\hspace{2.8in} \= \hspace{1.3in} \= \hspace{1.2in} \= \\

%=========    Heading   ==================================================
CS 399 \> Paper \> Bill Chickering (bchick)\\
\normalsize Mar 21, 2014 \> \> Jamie Irvine (jirvine)
% =========================================================================
\end{tabbing}
} \vspace{.4in}

\section*{Abstract}
\emph{Will write this when the paper is done}

\section*{Introduction}
Determining the similarity of two items is super important because... With large
amounts of rating data, collaborative filtering is a good method for
calculating similarity. However, with sparse rating data, this
technique gives noisy and unreliable approximations for similarity. Even worse, 
low confidence measurements of similarity are indistinguishable from high
confidence ones. In this paper, we develop a novel method which incorporates
confidence into the similarity score to produce more accurate estimates of
similarity.

Item-based collaborative filtering is a common technique for measuring the similarity 
of two items. There are different forms of collaborative filtering. For
this paper, we use a traditional approach. Each item is represented as a vector
of ratings. The similarity score of two items is computed by measuring the similarity
of the two rating vectors.

There are a few ways to compute the similarity of two vectors. One approach is
to use Cosine-similarity, defined as the cosine of the angle between the two vectors:
\begin{align}
PearsSim(A, B) = ...
\end{align}
where this = that ... Another popular approach is to use Pearson correlation:
\begin{align}
CosSim(A, B) = ...
\end{align}
where this = that.  Other similarity functions exist, such as one-sided similarities, but Pearson
correlation and Cosine-similarity are the most popular and will be the two
measurements used in this paper. \footnotemark

\footnotetext{ Note that the difference between the two similarities is how they handle
unpaired ratings; that is ratings from a user who has not rated the other
item. Cosine-similarity considers the unknown rating from an unpaired rating to
be $0$ and then calculates the cosine of the two vectors. Pearson correlation simply
throws away all unpaired ratings and calculates the cosine of the two modified vectors.}

Both similarity measurements primarily leverage information from common users,
that is users who have rated both items. Because of this, they perform
well with a large number common users, but give unreliable results when the
items have few common users. In an extreme case, if only one user has rated item
$A$ and item $B$, $PearsSim(A, B) = 1$ or $-1$. Not only is this unlikely to be an accurate
assessment of the true similarity of $A$ and $B$, it also gives the most extreme
results possible, without giving any indication that this is a low-confidence
calculation.

In this paper, we construct a more accurate similarity score by incorporating the
number of common users, $n$, into the similarity function.

\section*{Problem}
The goal is to calculate a modified similiary function that leverages the number of
commonn users to better estimate the true similarity. Formally, we design
$ModSim$ such that for two items $A$ and $B$ with $n$ users in common
\begin{align}
ModSim(A, B, n) \approx TrueSim(A, B)
\end{align}
where $TrueSim$ is the true similarity of the two items. Of course, there is no
way to actually know the true similarity of two items, but for a good enough
similarity function, such as $PearsSim$ or $CosSim$ and a large enough $n$, we
can get a good estimate. Thus, we model $TrueSim$ as 
\begin{align}
TrueSim(A, B) = lim_{n\to\infty}Sim(A, B)
\end{align}
where $Sim$ is the most appropriate generic similarity function for the dataset.

Since $TrueSim$ dependends on the choice of $Sim$, we abstract away the details
of $Sim$ and use its output directly. In all, the goal is to construct $ModSim$ such that
\begin{align}
ModSim(Sim(A, B), n) \approx lim_{n\to\infty}Sim(A, B)
\end{align}

\section*{Model}
We model the problem probabilistically. Let $Y$ be a random variable
representing the $TrueSim$ of a randomly chosen pair of items. Let $X_n$ be a
random variable representing the $Sim$ of a pair of items with $n$ common users.

We model $Y$ as a Normal distribution 
\begin{align}
Y \sim N(\mu, \sigma_{1}^2)
\end{align}
where $\mu$ and $\sigma_{1}^2$ are the average similarity score and the variance of similarity scores
of all pairs of items, respectively. Since $X_n$ represents a noisy reading of
the true similarity $Y$, we model $X_n | Y$ as a Normal distribution around $Y$:
\begin{align}
(X_n | Y=y) \sim N(y, \sigma_{2, n}^2)
\end{align}
Note that $\sigma_{2, n}^2$ represents how noisy the estimation of $Sim$ is
when there are $n$ common users. This should decrease as $n$ increases.

\begin{itemize}
    \item illustration
\end{itemize}

In probabilistic terms, for a calculated value $X_n$, we want to find the value 
of $Y$ that most likely produced $X_n$. Thus we desire the Maximum Likelihood
Estimate of $Y | X_n$
\begin{align}
\argmax_yP(Y=y|X_n=x) &= \argmax_y\left[\frac{P(X_n=x |
Y=y)P(Y=y)}{P(X_n=x)}\right]
\\&= \argmax_y\left[P(X_n=x|Y=y)P(Y=y)\right]
\\&= 
\argmax_y\left[\frac{1}{\sigma_{2,n}\sqrt{2\pi}}\exp{\left(\frac{-(x-y)^2}{2\sigma_{2,n}^2}\right)}
\frac{1}{\sigma_{1}\sqrt{2\pi}}\exp{\left(\frac{-(y-\mu)^2}{2\sigma_{1}^2}\right)}\right]
\\&= \argmin_y\left[\frac{(x-y)^2}{2\sigma_{2,n}^2} +
\frac{(y-\mu)^2}{2\sigma_{1}^2}\right]
\\&= \argmin_y\left[\left(\sigma_{1}^2+\sigma_{2,n}^2\right)y^2 - 
2\left(\sigma_{1}^2x+\sigma_{2,n}^2\mu\right)y\right]
\end{align}
We find the exact minimum by taking the derivative with respect to $y$ and
setting it to zero:
\begin{align}
\frac{d}{dy}\left[\left(\sigma_{1}^2+\sigma_{2,n}^2\right)y^2 - 
2\left(\sigma_{1}^2x+\sigma_{2,n}^2\mu\right)y\right]
&= 2\left(\sigma_{1}^2+\sigma_{2,n}^2\right)y - 
2\left(\sigma_{1}^2x+\sigma_{2,n}^2\mu\right) 
\\&= 0
\end{align}
Solving for y:
\begin{align}
y &= \frac{\sigma_{1}^2x+\sigma_{2,n}^2\mu}{\sigma_{1}^2+\sigma_{2,n}^2}
\end{align}
which can be rewritten as:
\begin{align}
\left(y - \mu\right) &= \frac{\sigma_{1}^2}{\sigma_{1}^2+\sigma_{2,n}^2}\left(x-\mu\right)
\end{align}

\begin{itemize}
    \item MLE with bias
    \item Features - linear, mu is fixed point
    \item Slope - between 0 and 1. As sigma2 changes. 
    \item Model sigma2 as a function of users in common
    \item Show final parameterized model - discuss params, speculation in model (?)
\end{itemize}

\section*{Technique}

\begin{itemize}
    \item Discussion of abandonment of parameterization in exchange for linear regression
    \item Pick data that's big enough to have gold standard
    \item Slice data and group by n
    \item Run linear regression on each group of n
    \item For predicting true similarity, take a given point and project on appropriate
    \item regression.
\end{itemize}

\section*{Experiment}

\section*{Results}

\section*{Conclusion}

\end{document}

### 2014-03-20 amazon/aggregateExpt2.py
Fix up modelSim.

@@ -105,7 +105,10 @@ def main():
         print >> sys.stderr, 'Processing %s . . .' % filename
         filenames.append(filename)
         # record all data
-        data.append(getData(fullpath))
+        filedata = getData(fullpath)
+        if not filedata:
+            continue
+        data.append(filedata)
         # parse productIds from filename
         productId1 = filename.split('_')[1]
         productId2 = filename.split('_')[2].split('.')[0]

### 2014-03-20 amazon/experiment.py

@@ -8,9 +8,10 @@ import numpy as np
 import csv
 import os
 import sys
+import math
 
 import similarity
-import predictions as pred
+import modelPredictions as modpred
 
 # params
 stepSize = 10
@@ -157,6 +158,9 @@ def expt2(db_conn, writer, cosineFunc, productId1, productId2):
     reviews1 = [(row[0], row[1], row[2]) for row in db_curs.fetchall()]
     db_curs.execute(selectReviewsOrderByUserStmt, (productId2,))
     reviews2 = [(row[0], row[1], row[2]) for row in db_curs.fetchall()]
+    # compute product biases
+    bias1 = np.mean([review[2] for review in reviews1])
+    bias2 = np.mean([review[2] for review in reviews2])
     # determine review pair times
     reviewPairTimes = getReviewPairTimes(reviews1, reviews2)
     # sort everything by time
@@ -188,16 +192,40 @@ def expt2(db_conn, writer, cosineFunc, productId1, productId2):
         stepReviews2 = sorted(reviews2[stepBegin_j:j], key=lambda x: x[1])
         stepBegin_i = i
         stepBegin_j = j
-        # append to past reviews and sort by userId
-        pastReviews1 = sorted(pastReviews1 + stepReviews1,
-                              key=lambda x: x[1])
-        assert(pastReviews1)
-        pastReviews2 = sorted(pastReviews2 + stepReviews2,
-                              key=lambda x: x[1])
-        assert(pastReviews2)
-        # compute cosine similarity at present time slice
-        # using the provided function.
-        cosineSim, numUserCommon = cosineFunc(pastReviews1, pastReviews2)
+        # special handling for modelSim
+        if cosineFunc == similarity.modelSim:
+            predReviews1 =\
+                sorted(predReviews1 + stepReviews1, key=lambda x: x[1])
+            predReviews2 =\
+                sorted(predReviews2 + stepReviews2, key=lambda x: x[1])
+            predictions += modpred.getPredictions(predReviews1, predReviews2,
+                                                  bias1, bias2)
+            # remove predictors from predReviews to avoid double counting
+            predReviews1 = [r for r in predReviews1\
+                            if r[1] not in [p[0] for p in predictions]]
+            predReviews2 = [r for r in predReviews2\
+                            if r[1] not in [p[0] for p in predictions]]
+            # compute weighted mean
+            totalScore = 0
+            totalWeight = 0
+            for p in predictions:
+                #weight = abs(p[1]*p[2])
+                weight = 1
[…]

### 2014-03-20 amazon/modelPredictions.py

@@ -37,35 +37,25 @@ def getParser(usage=None):
         default='modelPredictions', help='Output directory.', metavar='DIR')
     parser.add_option('-c', '--cosineFunc', dest='cosineFunc',
         default='prefSim',
-        help='Similarity function to use: "prefSim" (default) or "randSim"',
-              metavar='FUNCNAME')
-    parser.add_option('--modelGridFile', dest='modelGridFile',
-        default='output/simGrid_modelSim.csv',
-        help='CSV containing simGrid data for modelSim.', metavar='FILE')
+        help=('Similarity function to use as reference: '
+             '"prefSim" (default) or "randSim"'),
+        metavar='FUNCNAME')
     parser.add_option('--sigma_r', type='float', dest='sigma_r',
         default=0.2, help='Standard deviation of rating distribution.',
         metavar='FLOAT')
     parser.add_option('--sigma_s', type='float', dest='sigma_s',
         default=0.2, help='Standard deviation of score distribution.',
         metavar='FLOAT')
-    parser.add_option('--minRating', type='float', dest='minRating',
-        default=-2.0, help='Minimum rating.', metavar='FLOAT')
-    parser.add_option('--maxRating', type='float', dest='maxRating',
-        default=2.0, help='Maximum rating.', metavar='FLOAT')
-    parser.add_option('--stepRating', type='float', dest='stepRating',
-        default=0.1, help='Step rating.', metavar='FLOAT')
     return parser
 
-def getPredictions(reviews1, reviews2, sigma_r=None, sigma_s=None):
+def getPredictions(reviews1, reviews2, bias1, bias2,
+                   sigma_r=None, sigma_s=None):
     mu_s = similarity.mu_s
     if not sigma_s:
         sigma_s = similarity.sigma_s
     if not sigma_r:
         sigma_r = similarity.sigma_r
     predictions = []
-    # compute product biases
-    bias1 = np.mean([review[2] for review in reviews1])
-    bias2 = np.mean([review[2] for review in reviews2])
     # update histograms
     i = 0
     j = 0
@@ -90,8 +80,12 @@ def getPredictions(reviews1, reviews2, sigma_r=None, sigma_s=None):
             q = round(rating1**2 + rating2**2, 3)
             sigRatio = round((sigma_r/sigma_s)**2, 3)
             s = Symbol('s')
-            solutions =\
[…]

### 2014-03-20 amazon/similarity.py

@@ -23,7 +23,7 @@ sigmaXY = 0.3
 # modelSim params
 mu_s = constSimScore
 sigma_s = 0.2
-sigma_r = 0.1
+sigma_r = 0.3
 
 # errorFile
 userIdIdx = 0

### 2014-03-20 amazon/modelPredictions.py
Implement modelSim.py and modelSimVsRating.py.

@@ -12,6 +12,7 @@ from sympy import Symbol, solve
 from sympy.solvers.solvers import check_assumptions
 
 import similarity
+import modelSim
 
 # params
 outputTemplateTemplate = '%s_%%s_%%s.csv'
@@ -36,27 +37,32 @@ def getParser(usage=None):
     parser.add_option('-o', '--output-dir', dest='outputDir',
         default='modelPredictions', help='Output directory.', metavar='DIR')
     parser.add_option('-c', '--cosineFunc', dest='cosineFunc',
-        default='prefSim',
+        default='randSim',
         help=('Similarity function to use as reference: '
-             '"prefSim" (default) or "randSim"'),
+             '"randSim" (default) or "prefSim"'),
         metavar='FUNCNAME')
-    parser.add_option('--sigma_r', type='float', dest='sigma_r',
-        default=0.2, help='Standard deviation of rating distribution.',
+    parser.add_option('--mu_s', type='float', dest='mu_s',
+        default=None,  help='Mean of score distribution.',
         metavar='FLOAT')
     parser.add_option('--sigma_s', type='float', dest='sigma_s',
-        default=0.2, help='Standard deviation of score distribution.',
+        default=None, help='Standard deviation of score distribution.',
+        metavar='FLOAT')
+    parser.add_option('--sigma_r', type='float', dest='sigma_r',
+        default=None, help='Standard deviation of rating distribution.',
         metavar='FLOAT')
     return parser
 
 def getPredictions(reviews1, reviews2, bias1, bias2,
-                   sigma_r=None, sigma_s=None):
-    mu_s = similarity.mu_s
-    if not sigma_s:
-        sigma_s = similarity.sigma_s
-    if not sigma_r:
-        sigma_r = similarity.sigma_r
+                   mu_s=None, sigma_r=None, sigma_s=None):
+    # override params
+    if mu_s:
+        modelSim.mu_s = mu_s
+    if sigma_s:
+        modelSim.sigma_s = sigma_s
+    if sigma_r:
+        modelSim.sigma_r = sigma_r
     predictions = []
-    # update histograms
+    # iterate over common reviewers
     i = 0
     j = 0
     while i < len(reviews1) and j < len(reviews2):
@@ -75,39 +81,13 @@ def getPredictions(reviews1, reviews2, bias1, bias2,
                 continue # ignore duplicate reviews
             rating1 = reviews1[i][2] - bias1
[…]

### 2014-03-20 amazon/modelSim.py

@@ -1,0 +1,87 @@
+#!/usr/local/bin/python
+
+from optparse import OptionParser
+import os
+import sys
+import math
+import numpy as np
+from sympy import Symbol, Eq, solve
+from sympy.solvers.solvers import check_assumptions
+
+"""
+Determine a estimate for item-item similarity given two product ratings by
+a single reviewer.
+"""
+
+# "official" modelSim params
+mu_s = 0.27
+sigma_s = 0.2
+sigma_r = 0.2
+
+def getParser(usage=None):
+    parser = OptionParser(usage=usage)
+    parser.add_option('--mu_s', type='float', dest='mu_s',
+        default=None, help='Mean of score distribution.',
+        metavar='FLOAT')
+    parser.add_option('--sigma_s', type='float', dest='sigma_s',
+        default=None, help='Standard deviation of score distribution.',
+        metavar='FLOAT')
+    parser.add_option('--sigma_r', type='float', dest='sigma_r',
+        default=None, help='Standard deviation of rating distribution.',
+        metavar='FLOAT')
+    return parser
+
+def modelSim(rating1, rating2):
+    # rounding seems to help sympy
+    mu = round(mu_s, 3)
+    sigRatio = round((sigma_r/sigma_s)**2, 3)
+    p = round(rating1*rating2, 3)
+    q = round(rating1**2 + rating2**2, 3)
+    s = Symbol('s')
+    try:
+        solutions = solve(-p*s**2 + q*s - p +\
+                          sigRatio*(1 - s**2)**2*(s - mu), s)
+    except:
+        print >> sys.stderr, 'WARNING: sympy.solve raised error.'
+        solutions = []
+    prediction = None
+    for candidate in solutions:
+        if (check_assumptions(candidate, real=True) and\
+            candidate > -1 and candidate < 1) or\
+             candidate == 0:
+            prediction = candidate
+    if not prediction:
+        if solutions == [0.0]:
+            prediction = 0.0
+        elif solutions and solutions[0] == 1.0:
+            prediction = 1.0
+        elif solutions and solutions[0] == -1.0:
+            prediction = -1.0
+        else:
+            print >> sys.stderr, 'WARNING: No solution found:', solutions
+            prediction = mu_s
+    return prediction
+
+def main():
+    # Parse options
+    usage = 'Usage: %prog [options] rating1 rating2'
+    parser = getParser(usage=usage)
+    (options, args) = parser.parse_args()
[…]

### 2014-03-20 amazon/modelSimVsRating.py

@@ -1,0 +1,61 @@
+#!/usr/local/bin/python
+
+"""
+Generate modelSim scores as a function of a single product rating
+while holding the other product rating held fixed.
+"""
+
+from optparse import OptionParser
+import numpy as np
+import csv
+import os
+import sys
+
+import modelSim
+
+def getParser(usage=None):
+    parser = OptionParser(usage=usage)
+    parser.add_option('--mu_s', type='float', dest='mu_s',
+        default=None,  help='Mean of score distribution.',
+        metavar='FLOAT')
+    parser.add_option('--sigma_s', type='float', dest='sigma_s',
+        default=None, help='Standard deviation of score distribution.',
+        metavar='FLOAT')
+    parser.add_option('--sigma_r', type='float', dest='sigma_r',
+        default=None, help='Standard deviation of rating distribution.',
+        metavar='FLOAT')
+    parser.add_option('--minRating', type='float', dest='minRating',
+       default=-1.0, help='Minimum rating value.', metavar='FLOAT')
+    parser.add_option('--maxRating', type='float', dest='maxRating',
+       default=2.0, help='Maximum rating value.', metavar='FLOAT')
+    parser.add_option('--stepRating', type='float', dest='stepRating',
+       default=0.1, help='Step rating value.', metavar='FLOAT')
+    return parser
+
+def main():
+    # Parse options
+    usage = 'Usage: %prog [options] rating1'
+    parser = getParser(usage=usage)
+    (options, args) = parser.parse_args()
+    if len(args) != 1:
+        parser.error('Wrong number of arguments')
+    rating1 = float(args[0])
+
+    # modelSim param overrides
+    if options.mu_s:
+        modelSim.mu_s = options.mu_s
+    if options.sigma_s:
+        modelSim.sigma_s = options.sigma_s
+    if options.sigma_r:
+        modelSim.sigma_r = options.sigma_r
+
+    # Generate curve
+    writer = csv.writer(sys.stdout)
+    rating2 = options.minRating
+    while rating2 < options.maxRating:
+        sim = modelSim.modelSim(rating1, rating2)
+        writer.writerow([round(rating2, 3), sim])
+        rating2 += options.stepRating
+
+if __name__ == '__main__':
+    main()

### 2014-03-20 amazon/modelGenRating.py
Add modelGenRating.py.

@@ -1,0 +1,62 @@
+#!/usr/local/bin/python
+
+"""
+Generates rating pairs given their cosine similarity according to a simple
+model.
+"""
+
+from optparse import OptionParser
+import os
+import sys
+import numpy as np
+
+import modelSim
+
+def getParser(usage=None):
+    parser = OptionParser(usage=usage)
+    parser.add_option('-n', '--numRatings', dest='numRatings', type='int',
+        default=10, help='Number of rating pairs to generate.', metavar='NUM')
+    parser.add_option('--sigma_s', type='float', dest='sigma_s',
+        default=None, help='Standard deviation of score distribution.',
+        metavar='FLOAT')
+    parser.add_option('--sigma_r', type='float', dest='sigma_r',
+        default=None, help='Standard deviation of rating distribution.',
+        metavar='FLOAT')
+    return parser
+
+def genRatings(sigma_s, sigma_r, score):
+    mean = np.array([0, 0])
+    cov = sigma_r*np.array([[1.0, score],
+                            [score, 1.0]])
+    return np.random.multivariate_normal(mean, cov)
+
+def main():
+    # Parse options
+    usage = 'Usage: %prog [options] similarityScore'
+    parser = getParser(usage=usage)
+    (options, args) = parser.parse_args()
+    if len(args) != 1:
+        parser.error('Wrong number of arguments')
+    try:
+        score = float(args[0])
+    except:
+        print >> sys.stderr, 'ERROR: Invalid similarity score.'
+        return
+
+    # modelSim param overrides
+    if options.sigma_s:
+        sigma_s = options.sigma_s
+    else:
+        sigma_s = modelSim.sigma_s
+    if options.sigma_r:
+        sigma_r = options.sigma_r
+    else:
+        sigma_r = modelSim.sigma_r
+
+    # generate ratings
+    for i in range(options.numRatings):
+        rating1, rating2 = genRatings(sigma_s, sigma_r, score)
+        print 'rating1 = % 0.3f, rating2 = % 0.3f' % (rating1, rating2)
+
+if __name__ == '__main__':
+    main()

### 2014-03-20 amazon/experiment.py
Further develop modelSim.

@@ -77,8 +77,8 @@ def getParser(usage=None):
     parser.add_option('--sigma', dest='sigma', type='float', default=None,
         help='Parameter sigma for prefSimAlt1 or randSimAlt1.', metavar='FLOAT')
     parser.add_option('--errorFileName', dest='errorFileName',
-        default='data/userErrors.csv', help='User prediction errors file.',
-        metavar='FILE')
+        default='output/aggregatePredictions_modelSim.csv',
+        help='User prediction errors file for modelSim.', metavar='FILE')
     parser.add_option('--epsilon', dest='epsilon', type='float', default=0.5,
         help='Parameter epsilon for weightedRandSim or wightedPrefSim.',
         metavar='FLOAT')
@@ -193,7 +193,8 @@ def expt2(db_conn, writer, cosineFunc, productId1, productId2):
         stepBegin_i = i
         stepBegin_j = j
         # special handling for modelSim
-        if cosineFunc == similarity.modelSim:
+        if cosineFunc == similarity.modelSim or\
+           cosineFunc == similarity.weightedModelSim:
             predReviews1 =\
                 sorted(predReviews1 + stepReviews1, key=lambda x: x[1])
             predReviews2 =\
@@ -209,8 +210,13 @@ def expt2(db_conn, writer, cosineFunc, productId1, productId2):
             totalScore = 0
             totalWeight = 0
             for p in predictions:
-                #weight = abs(p[1]*p[2])
-                weight = 1
+                if cosineFunc == similarity.weightedModelSim:
+                    try:
+                        weight = similarity.weights[p[0]]
+                    except:
+                        weight = similarity.avgWeight
+                else:
+                    weight = 1
                 totalScore += p[3]*weight
                 totalWeight += weight
             cosineSim = totalScore/totalWeight
@@ -330,7 +336,7 @@ def main():
                                options.stepRating, options.simGridFile)
 
     # weighted similarity
-    if options.cosineFunc == 'weightedRandSim':
+    if options.cosineFunc == 'weightedModelSim':
         similarity.initWeightedSim(options.errorFileName, options.epsilon)
 
     # create queues

### 2014-03-20 amazon/modelGenRating.py

@@ -16,18 +16,18 @@ def getParser(usage=None):
     parser = OptionParser(usage=usage)
     parser.add_option('-n', '--numRatings', dest='numRatings', type='int',
         default=10, help='Number of rating pairs to generate.', metavar='NUM')
-    parser.add_option('--sigma_s', type='float', dest='sigma_s',
-        default=None, help='Standard deviation of score distribution.',
-        metavar='FLOAT')
     parser.add_option('--sigma_r', type='float', dest='sigma_r',
         default=None, help='Standard deviation of rating distribution.',
         metavar='FLOAT')
+    parser.add_option('--rho', type='float', dest='rho',
+        default=None, help='Similarity multiplicative constant.',
+        metavar='FLOAT')
     return parser
 
-def genRatings(sigma_s, sigma_r, score):
+def genRatings(sigma_r, rho, score):
     mean = np.array([0, 0])
-    cov = sigma_r*np.array([[1.0, score],
-                            [score, 1.0]])
+    cov = sigma_r*np.array([[1.0, rho*score],
+                            [rho*score, 1.0]])
     return np.random.multivariate_normal(mean, cov)
 
 def main():
@@ -44,18 +44,18 @@ def main():
         return
 
     # modelSim param overrides
-    if options.sigma_s:
-        sigma_s = options.sigma_s
-    else:
-        sigma_s = modelSim.sigma_s
     if options.sigma_r:
         sigma_r = options.sigma_r
     else:
         sigma_r = modelSim.sigma_r
+    if options.rho:
+        rho = options.rho
+    else:
+        rho = modelSim.rho
 
     # generate ratings
     for i in range(options.numRatings):
-        rating1, rating2 = genRatings(sigma_s, sigma_r, score)
+        rating1, rating2 = genRatings(sigma_r, rho, score)
         print 'rating1 = % 0.3f, rating2 = % 0.3f' % (rating1, rating2)
 
 if __name__ == '__main__':

### 2014-03-20 amazon/modelSim.py

@@ -15,8 +15,9 @@ a single reviewer.
 
 # "official" modelSim params
 mu_s = 0.27
-sigma_s = 0.2
-sigma_r = 0.2
+sigma_s = 0.3
+sigma_r = 0.5
+rho = 2.2
 
 def getParser(usage=None):
     parser = OptionParser(usage=usage)
@@ -29,37 +30,45 @@ def getParser(usage=None):
     parser.add_option('--sigma_r', type='float', dest='sigma_r',
         default=None, help='Standard deviation of rating distribution.',
         metavar='FLOAT')
+    parser.add_option('--rho', type='float', dest='rho',
+        default=None, help='Similarity multiplicative constant.',
+        metavar='FLOAT')
     return parser
 
 def modelSim(rating1, rating2):
     # rounding seems to help sympy
     mu = round(mu_s, 3)
     sigRatio = round((sigma_r/sigma_s)**2, 3)
+    r = round(rho, 3)
+    r_sq = round(rho**2, 3)
     p = round(rating1*rating2, 3)
     q = round(rating1**2 + rating2**2, 3)
     s = Symbol('s')
+    solutions = None
     try:
-        solutions = solve(-p*s**2 + q*s - p +\
-                          sigRatio*(1 - s**2)**2*(s - mu), s)
+        solutions = solve(-p*r_sq*s**2 + q*r*s - p +\
+                          sigRatio*(1 - r_sq*s**2)**2*(s - mu), s)
     except:
         print >> sys.stderr, 'WARNING: sympy.solve raised error.'
-        solutions = []
     prediction = None
-    for candidate in solutions:
-        if (check_assumptions(candidate, real=True) and\
-            candidate > -1 and candidate < 1) or\
-             candidate == 0:
-            prediction = candidate
-    if not prediction:
-        if solutions == [0.0]:
-            prediction = 0.0
-        elif solutions and solutions[0] == 1.0:
-            prediction = 1.0
-        elif solutions and solutions[0] == -1.0:
-            prediction = -1.0
-        else:
-            print >> sys.stderr, 'WARNING: No solution found:', solutions
-            prediction = mu_s
+    if not solutions:
+        prediction = mu_s
+    else:
+        for candidate in solutions:
+            if (check_assumptions(candidate, real=True) and\
+                candidate > -1 and candidate < 1) or\
+                 candidate == 0:
+                prediction = candidate
+        if not prediction:
+            if solutions == [0.0]:
+                prediction = 0.0
[…]

### 2014-03-20 amazon/similarity.py

@@ -147,11 +147,16 @@ def predSim(reviewsA, reviewsB):
     else:
         return (constSimScore, numUsersCommon)
 
+def modelSim(reviewsA, reviewsB):
+    """this is just a place holder."""
+    raise NotImplemented
+
+def weightedModelSim(reviewsA, reviewsB):
+    """this is just a place holder."""
+    raise NotImplemented
+
 #def modelSim(reviewsA, reviewsB):
-#    """Compute the average "implicitly" predicted similarity between
-#       items A and B according to a simple model.
-#    """
-#    # compute product biases:
+#    """Compute the MLE similarity based upon a simple model."""
 #    if reviewsA:
 #        biasA = np.mean([review[2] for review in reviewsA])
 #    else:
@@ -160,9 +165,9 @@ def predSim(reviewsA, reviewsB):
 #        biasB = np.mean([review[2] for review in reviewsB])
 #    else:
 #        biasB = 0.0
-#    totalScore = 0
-#    totalCount = 0
 #    numUsersCommon = 0
+#    solutionTotal = 0
+#    weightTotal = 0
 #    i = 0
 #    j = 0
 #    while i < len(reviewsA) and j < len(reviewsB):
@@ -182,69 +187,18 @@ def predSim(reviewsA, reviewsB):
 #            numUsersCommon += 1
 #            scoreA = reviewsA[i][2] - biasA
 #            scoreB = reviewsB[j][2] - biasB
-#            entry = modelGrid[(scoreA, scoreB)]
-#            if entry and entry[0] >= -1 and entry[0] <= 1:
-#                totalScore += entry[0]
-#                totalCount += math.sqrt(scoreA**2 + scoreB**2)
+#            p = round(scoreA*scoreB, 3)
+#            q = round(scoreA**2 + scoreB**2, 3)
+#            s = sympy.Symbol('s')
+#            solutions =\
+#                sympy.solve(-p*s**2 + q*s - p +\
+#                            (sigma_r/sigma_s)**2*(1 - s**2)**2*(s - mu_s), s,
+#                            minimal=True, quick=True)
+#            weight = math.sqrt(q)
+#            solutionTotal += weight*solutions[0]
+#            weightTotal += weight
 #            i += 1
 #            j += 1
-#    if totalCount > 0:
-#        return (totalScore/totalCount, numUsersCommon)
-#    else:
-#        return (constSimScore, numUsersCommon)
-
-def modelSim(reviewsA, reviewsB):
[…]

### 2014-03-20 amazon/make3DRatingData.py
Add make3DRatingData.py.

@@ -1,0 +1,130 @@
+#!/usr/local/bin/python
+
+"""
+Generate 3D histogram of rating pair data.
+"""
+
+from optparse import OptionParser
+import sqlite3
+import csv
+import os
+import sys
+import numpy as np
+
+# db params
+dbTimeout = 5
+selectReviewsStmt =\
+    ('SELECT Time, UserId, AdjustedScore '
+     'FROM Reviews '
+     'WHERE ProductId = :ProductId '
+     'ORDER BY UserId')
+
+def getParser(usage=None):
+    parser = OptionParser(usage=usage)
+    parser.add_option('-d', '--database', dest='db_fname',
+        default='data/amazon.db', help='sqlite3 database file.', metavar='FILE')
+    parser.add_option('--minRating', type='float', dest='minRating',
+        default=-1.5, help='Minimum rating.', metavar='NUM')
+    parser.add_option('--stepRating', type='float', dest='stepRating',
+        default=0.075, help='Step rating.', metavar='NUM')
+    parser.add_option('--numBins', type='int', dest='numBins',
+        default=40, help='Number of bins on a side.', metavar='NUM')
+    return parser
+
+def incrementHistogram(histogram, minRating, stepRating, numBins,
+                       rating1, rating2):
+    if rating1 >= minRating and rating1 < minRating + numBins*stepRating and\
+       rating2 >= minRating and rating2 < minRating + numBins*stepRating:
+        idx1 = int(round((rating1 - minRating)/stepRating))
+        idx2 = int(round((rating2 - minRating)/stepRating))
+        if idx1 >= numBins:
+            idx1 = numBins - 1
+        if idx2 >= numBins:
+            idx2 = numBins - 1
+        histogram[idx1][idx2] += 1
+
+def fillHistogram(db_conn, inputfile, histogram,
+                  minRating, stepRating, numBins):
+    db_curs = db_conn.cursor()
+    for line in inputfile:
+        tokens = line.split(',')
+        productId1 = tokens[0].strip()
+        productId2 = tokens[1].strip()
+        print >> sys.stderr, 'Processing %s, %s . . .' %\
+                             (productId1, productId2)
+        # fetch product reviews
+        db_curs.execute(selectReviewsStmt, (productId1,))
+        reviews1 = [(row[0], row[1], row[2]) for row in db_curs.fetchall()]
+        db_curs.execute(selectReviewsStmt, (productId2,))
[…]

### 2014-03-22 amazon/experiment.py
Update default errorFileName param of experiment.py.

@@ -77,7 +77,7 @@ def getParser(usage=None):
     parser.add_option('--sigma', dest='sigma', type='float', default=None,
         help='Parameter sigma for prefSimAlt1 or randSimAlt1.', metavar='FLOAT')
     parser.add_option('--errorFileName', dest='errorFileName',
-        default='output/aggregatePredictions_modelSim.csv',
+        default='output/aggregatePredictions_modelSim_2.2_0.5_0.3.csv',
         help='User prediction errors file for modelSim.', metavar='FILE')
     parser.add_option('--epsilon', dest='epsilon', type='float', default=0.5,
         help='Parameter epsilon for weightedRandSim or wightedPrefSim.',

### 2014-03-22 paper/paper.tex
Paper has twonormals image and goes until Technique section of linear stuff

@@
 \section*{Introduction}
-Determining the similarity of two items is super important because... With large
+The similarity of two items is a valuable measurement for a number of
+data-mining applications, such as recommender systems. With large
 amounts of rating data, collaborative filtering is a good method for
@@
 \begin{align}
-PearsSim(A, B) = ...
+CosSim(A, B) = \frac{\sum\limits_{u\in U_{AB}}
+r_{u,A}r_{u,B}}{\sqrt{\sum\limits_{u\in U_{A}} r_{u,A}^2}
+\sqrt{\sum\limits_{u\in U_{B}} r_{u,B}^2}}
 \end{align}
-where this = that ... Another popular approach is to use Pearson correlation:
+where $U_{I}$ is the set of all users who rated item $I$, $U_{AB}$ is the set of all 
+users who rated both item $A$ and item $B$ and $r_{u,I}$ is the rating user $u$ gave 
+to item $I$. Another popular approach is to use Pearson correlation:
 \begin{align}
-CosSim(A, B) = ...
+PearsSim(A, B) = \frac{\sum\limits_{u\in U_{AB}}
+r_{u,A}r_{u,B}}{\sqrt{\sum\limits_{u\in U_{AB}} r_{u,A}^2}
+\sqrt{\sum\limits_{u\in U_{AB}} r_{u,B}^2}}
 \end{align}
-where this = that.  Other similarity functions exist, such as one-sided similarities, but Pearson
+Other similarity functions exist, such as one-sided similarities, but Pearson
 correlation and Cosine-similarity are the most popular and will be the two
@@
 representing the $TrueSim$ of a randomly chosen pair of items. Let $X_n$ be a
-random variable representing the $Sim$ of a pair of items with $n$ common users.
+random variable representing the calculated $Sim$ of a pair of items with $n$ common users.
 
@@
 
-\begin{itemize}
-    \item illustration
-\end{itemize}
+\begin{figure}[!htbp]
+    \centering
+    \includegraphics[width=0.5\textwidth]{twonormals.png}
+    \caption{The blue curve represents $Y$, the distribution of true similarities
+    between a random pair of items. The red curve represents $X|Y$, the
+    estimated similarity when a small number of common users exist.}
+    \label{fig:two_normals}
+\end{figure}
 
@@
 
-\begin{itemize}
-    \item MLE with bias
-    \item Features - linear, mu is fixed point
-    \item Slope - between 0 and 1. As sigma2 changes. 
-    \item Model sigma2 as a function of users in common
-    \item Show final parameterized model - discuss params, speculation in model (?)
-\end{itemize}
+The model suggests that there is a linear correlation between $X_n$, the
+similarity calculated with $n$ common users, and $Y$, the true similarity. Since 
+$\sigma_{2,n}^2\ge0$, the slope is always $\le1$, meaning that the calculated
+distance from $\mu$ (the average similarity) overestimates the true distance. This makes
+sense, since there is a prior distribution of true similarity weighted around $\mu$.
+
+The parameters $\mu$, $\sigma_{1}^2$, and $\sigma_{2,n}^2$ all can be approximated 
+using the method of moments over training data where there are enough common
+users that $TrueSim$ can be approximated. This method requires approximating
+$\sigma_{2,n}^2$ for every value of $n$ separately. Alternatively, we can model
+$\sigma_{2,n}^2$ as a function of $n$. Recall that $\sigma_{2,n}^2$ is the
+noisiness of the measurement $Sim$ about $TrueSim$ and $TrueSim = lim_{n \to
+\infty}Sim$. Thus $Sim$ can be thought of as a sampling $n$ users from the
+infinite set used to calculate $TrueSim$. Although $Sim$ may not be linear, we
+the Central Limit Theorem motivates the intuition that the variance of $Sim$
+would decrease as the $sqrt(n)$. Thus we model $\sigma_{2,n}^2$ as:
+
+\begin{align}
+\sigma_{2,n}^2 = \frac{\alpha}{\sqrt{n}}
+\end{align}
+Yielding the single parameter model:
+\begin{align}
+\left(Y - \mu\right) = \frac{\sigma_{1}^2}{\sigma_{1}^2+\frac{\alpha}{\sqrt{n}}}
+\left(X_n-\mu\right)
+\end{align}
+In terms of a modified similarity function, as desired earlier, we now have:
+\begin{align}
+ModSim(Sim, n) = \frac{\sigma_{1}^2}{\sigma_{1}^2+\frac{\alpha}{\sqrt{n}}}
+\left(Sim-\mu\right) + \mu
+\end{align}
 

### 2014-03-22 amazon/modelPredictions.py
Implement better model.

@@ -41,26 +41,31 @@ def getParser(usage=None):
         help=('Similarity function to use as reference: '
              '"randSim" (default) or "prefSim"'),
         metavar='FUNCNAME')
-    parser.add_option('--mu_s', type='float', dest='mu_s',
-        default=None,  help='Mean of score distribution.',
-        metavar='FLOAT')
-    parser.add_option('--sigma_s', type='float', dest='sigma_s',
-        default=None, help='Standard deviation of score distribution.',
-        metavar='FLOAT')
-    parser.add_option('--sigma_r', type='float', dest='sigma_r',
-        default=None, help='Standard deviation of rating distribution.',
-        metavar='FLOAT')
+    parser.add_option('--mu_s', type='float', dest='mu_s', default=None,
+        help='Mean of score distribution.', metavar='FLOAT')
+    parser.add_option('--sigma_s', type='float', dest='sigma_s', default=None,
+        help='Standard deviation of score distribution.', metavar='FLOAT')
+    parser.add_option('--mu_r', type='float', dest='mu_r', default=None,
+        help='Mean of rating distribution.', metavar='FLOAT')
+    parser.add_option('--sigma_r', type='float', dest='sigma_r', default=None,
+        help='Standard deviation of rating distribution.', metavar='FLOAT')
+    parser.add_option('--alpha', type='int', dest='alpha', default=None,
+        help='Similarity root.', metavar='INT')
     return parser
 
-def getPredictions(reviews1, reviews2, bias1, bias2,
-                   mu_s=None, sigma_r=None, sigma_s=None):
+def getPredictions(reviews1, reviews2, bias1, bias2, mu_s=None, sigma_s=None,
+                   mu_r=None, sigma_r=None, alpha=None):
     # override params
     if mu_s:
         modelSim.mu_s = mu_s
     if sigma_s:
         modelSim.sigma_s = sigma_s
+    if mu_r:
+        modelSim.mu_r = mu_r
     if sigma_r:
         modelSim.sigma_r = sigma_r
+    if alpha:
+        modelSim.alpha = alpha
     predictions = []
     # iterate over common reviewers
     i = 0
@@ -87,8 +92,8 @@ def getPredictions(reviews1, reviews2, bias1, bias2,
             j += 1
     return predictions
 
-def processPair(db_conn, writer, cosineFunc, mu_s, sigma_r, sigma_s,
-                productId1, productId2):
[…]

### 2014-03-22 amazon/modelSim.py

@@ -16,38 +16,39 @@ a single reviewer.
 # "official" modelSim params
 mu_s = 0.27
 sigma_s = 0.3
-sigma_r = 0.5
-rho = 2.2
+mu_r = 0.2
+sigma_r = 0.2
+alpha = 2
 
 def getParser(usage=None):
     parser = OptionParser(usage=usage)
-    parser.add_option('--mu_s', type='float', dest='mu_s',
-        default=None, help='Mean of score distribution.',
-        metavar='FLOAT')
-    parser.add_option('--sigma_s', type='float', dest='sigma_s',
-        default=None, help='Standard deviation of score distribution.',
-        metavar='FLOAT')
-    parser.add_option('--sigma_r', type='float', dest='sigma_r',
-        default=None, help='Standard deviation of rating distribution.',
-        metavar='FLOAT')
-    parser.add_option('--rho', type='float', dest='rho',
-        default=None, help='Similarity multiplicative constant.',
-        metavar='FLOAT')
+    parser.add_option('--mu_s', type='float', dest='mu_s', default=None,
+        help='Mean of score distribution.', metavar='FLOAT')
+    parser.add_option('--sigma_s', type='float', dest='sigma_s', default=None,
+        help='Standard deviation of score distribution.', metavar='FLOAT')
+    parser.add_option('--mu_r', type='float', dest='mu_r', default=None,
+        help='Mean of rating distribution.', metavar='FLOAT')
+    parser.add_option('--sigma_r', type='float', dest='sigma_r', default=None,
+        help='Standard deviation of rating distribution.', metavar='FLOAT')
+    parser.add_option('--alpha', type='int', dest='alpha', default=None,
+        help='Similarity root.', metavar='INT')
     return parser
 
 def modelSim(rating1, rating2):
     # rounding seems to help sympy
-    mu = round(mu_s, 3)
-    sigRatio = round((sigma_r/sigma_s)**2, 3)
-    r = round(rho, 3)
-    r_sq = round(rho**2, 3)
-    p = round(rating1*rating2, 3)
-    q = round(rating1**2 + rating2**2, 3)
+    p = round((rating1 - mu_r)*(rating2 - mu_r), 3)
+    q = round((rating1 - mu_r)**2 + (rating2 - mu_r)**2, 3)
+    mu_s_ = round(mu_s, 3)
+    sigRatio = round(alpha*(sigma_r/sigma_s)**2, 3)
[…]

### 2014-03-23 amazon/experiment.py
Re-implement modelSim.

@@ -45,7 +45,7 @@ def getParser(usage=None):
         help='Output file prefix.', metavar='STR')
     parser.add_option('-e', '--exptFunc', dest='exptFunc', default='expt1',
         help='Experiment function to use: "expt1" (default) or "expt2"',
-        metavar='FUNCNAME')
+        metavar='EXPTNAME')
     parser.add_option('-c', '--cosineFunc', dest='cosineFunc',
         default='prefSim',
         help=('Similarity function to use: "prefSim" (default), "randSim", '
@@ -77,9 +77,9 @@ def getParser(usage=None):
     parser.add_option('--sigma', dest='sigma', type='float', default=None,
         help='Parameter sigma for prefSimAlt1 or randSimAlt1.', metavar='FLOAT')
     parser.add_option('--errorFileName', dest='errorFileName',
-        default='output/aggregatePredictions_modelSim_2.2_0.5_0.3.csv',
+        default='output/aggregatePredictions_modelSim.csv',
         help='User prediction errors file for modelSim.', metavar='FILE')
-    parser.add_option('--epsilon', dest='epsilon', type='float', default=0.5,
+    parser.add_option('--epsilon', dest='epsilon', type='float', default=0.01,
         help='Parameter epsilon for weightedRandSim or wightedPrefSim.',
         metavar='FLOAT')
     return parser

### 2014-03-23 amazon/modelPredictions.py

@@ -1,5 +1,10 @@
 #!/usr/local/bin/python
 
+"""
+Processes a list of product pairs and for each produces a list of users, their
+implied similarity estimates based on a model, and their error.
+"""
+
 import multiprocessing as mp
 import Queue
 from optparse import OptionParser
@@ -8,16 +13,14 @@ import csv
 import os
 import sys
 import numpy as np
-from sympy import Symbol, solve
-from sympy.solvers.solvers import check_assumptions
 
 import similarity
 import modelSim
 
 # params
 outputTemplateTemplate = '%s_%%s_%%s.csv'
-workerTimeout = 60
-workerQueueSize = 100
+workerTimeout = 20
+workerQueueSize = 20
 END_OF_QUEUE = -1
 
 # db params
@@ -39,8 +42,7 @@ def getParser(usage=None):
     parser.add_option('-c', '--cosineFunc', dest='cosineFunc',
         default='randSim',
         help=('Similarity function to use as reference: '
-             '"randSim" (default) or "prefSim"'),
-        metavar='FUNCNAME')
+             '"randSim" (default) or "prefSim"'), metavar='FUNCNAME')
     parser.add_option('--mu_s', type='float', dest='mu_s', default=None,
         help='Mean of score distribution.', metavar='FLOAT')
     parser.add_option('--sigma_s', type='float', dest='sigma_s', default=None,
@@ -49,12 +51,10 @@ def getParser(usage=None):
         help='Mean of rating distribution.', metavar='FLOAT')
     parser.add_option('--sigma_r', type='float', dest='sigma_r', default=None,
         help='Standard deviation of rating distribution.', metavar='FLOAT')
-    parser.add_option('--alpha', type='int', dest='alpha', default=None,
-        help='Similarity root.', metavar='INT')
     return parser
 
 def getPredictions(reviews1, reviews2, bias1, bias2, mu_s=None, sigma_s=None,
-                   mu_r=None, sigma_r=None, alpha=None):
+                   mu_r=None, sigma_r=None):
     # override params
     if mu_s:
         modelSim.mu_s = mu_s
@@ -64,8 +64,6 @@ def getPredictions(reviews1, reviews2, bias1, bias2, mu_s=None, sigma_s=None,
         modelSim.mu_r = mu_r
     if sigma_r:
         modelSim.sigma_r = sigma_r
-    if alpha:
-        modelSim.alpha = alpha
     predictions = []
     # iterate over common reviewers
     i = 0
[…]

### 2014-03-23 amazon/modelSim.py

@@ -5,8 +5,7 @@ import os
 import sys
 import math
 import numpy as np
-from sympy import Symbol, Eq, solve
-from sympy.solvers.solvers import check_assumptions
+from scipy.optimize import brentq
 
 """
 Determine a estimate for item-item similarity given two product ratings by
@@ -17,8 +16,7 @@ a single reviewer.
 mu_s = 0.27
 sigma_s = 0.3
 mu_r = 0.2
-sigma_r = 0.2
-alpha = 2
+sigma_r = 0.3
 
 def getParser(usage=None):
     parser = OptionParser(usage=usage)
@@ -30,51 +28,20 @@ def getParser(usage=None):
         help='Mean of rating distribution.', metavar='FLOAT')
     parser.add_option('--sigma_r', type='float', dest='sigma_r', default=None,
         help='Standard deviation of rating distribution.', metavar='FLOAT')
-    parser.add_option('--alpha', type='int', dest='alpha', default=None,
-        help='Similarity root.', metavar='INT')
     return parser
 
 def modelSim(rating1, rating2):
-    # rounding seems to help sympy
-    p = round((rating1 - mu_r)*(rating2 - mu_r), 3)
-    q = round((rating1 - mu_r)**2 + (rating2 - mu_r)**2, 3)
-    mu_s_ = round(mu_s, 3)
-    sigRatio = round(alpha*(sigma_r/sigma_s)**2, 3)
-    exp1 = round((1 + alpha)/float(alpha), 3)
-    exp2 = round((1 - alpha)/float(alpha), 3)
-    exp3 = round((3 - alpha)/float(alpha), 3)
-    exp4 = round(2/alpha, 3)
-    s = Symbol('s')
-    solutions = None
-    try:
-        solutions = solve(q*s - p*s**exp1 - p*s**exp2 + p*s**exp3 +\
-                          sigRatio*(1 - s**exp4)**2*(s - mu_s_), s)
-    except:
-        print >> sys.stderr, 'WARNING: sympy.solve raised error.'
-    prediction = None
-    if not solutions:
-        prediction = mu_s
+    p = (rating1 - mu_r)*(rating2 - mu_r)
+    q = (rating1 - mu_r)**2 + (rating2 - mu_r)**2
+    extremum = brentq(lambda s: -p*s**2 + q*s - p +\
+                  (sigma_r/sigma_s)**2*(1 - s**2)**2*(s - mu_s),
+                  -100, 100, xtol=0.0001)
[…]

### 2014-03-23 amazon/partitionProducts.py
Reformat code in partitionProducts.py.

@@ -12,28 +12,32 @@ import random
 # db params
 dropStoresTableStmt = 'DROP TABLE IF EXISTS Stores'
 dropStoreProductsTableStmt = 'DROP TABLE IF EXISTS StoreProducts'
-createStoresTableStmt = 'CREATE TABLE Stores(StoreId INT PRIMARY KEY, Name TEXT)'
+createStoresTableStmt =\
+    'CREATE TABLE Stores(StoreId INT PRIMARY KEY, Name TEXT)'
 createStoreProductsTableStmt =\
     ('CREATE TABLE StoreProducts(StoreId INT, ProductId INT, PRIMARY '
-     'KEY(StoreId, ProductId), FOREIGN KEY(StoreId) REFERENCES Stores(StoreId), '
+     'KEY(StoreId, ProductId), '
+     'FOREIGN KEY(StoreId) REFERENCES Stores(StoreId), '
      'FOREIGN KEY(ProductId) REFERENCES Products(ProductId))')
 insertStoreStmt = 'INSERT INTO Stores (StoreId) VALUES (:StoreId)'
 selectProductsStmt = 'SELECT ProductId FROM Products'
 insertStoreProductStmt =\
-    'INSERT INTO StoreProducts (StoreId, ProductId) VALUES (:StoreId, :ProductId)'
+    ('INSERT INTO StoreProducts (StoreId, ProductId) '
+     'VALUES (:StoreId, :ProductId)')
 
 def getParser(usage=None):
     parser = OptionParser(usage=usage)
-    parser.add_option('-d', '--database', dest='db_fname', default='data/amazon.db',
-        help='sqlite3 database file.', metavar='FILE')
-    parser.add_option('-n', '--numStores', dest='numStores', type='int', default=2,
-        help='Number of StoreId values.', metavar='NUM')
+    parser.add_option('-d', '--database', dest='db_fname',
+        default='data/amazon.db', help='sqlite3 database file.', metavar='FILE')
+    parser.add_option('-n', '--numStores', dest='numStores', type='int',
+        default=2, help='Number of StoreId values.', metavar='NUM')
     parser.add_option('-r', '--ratios', dest='ratios', default='1:1',
         help=('Ratios of products per store. For example, "2:1" indicates that '
-              'store 1 has a product catalog that is twice the size of store 2\'s.'),
-        metavar='RATIOS')
-    parser.add_option('-o', '--overlap', dest='overlap', type='float', default=0.0,
[…]

### 2014-03-23 amazon/similarity.py
Remove reference to predictions.py from similarity.py.

@@ -12,7 +12,6 @@ import sys
 from collections import defaultdict
 import sympy 
 
-import predictions as pred
 from SimilarityGrid import SimilarityGrid
 
 # params

### 2014-03-23 common/rmBelowLength.py
Add rmBelowLength.py.

@@ -1,0 +1,51 @@
+#!/usr/local/bin/python
+
+"""
+Delete all files in a directory that have fewer than n lines.
+"""
+
+from optparse import OptionParser
+import os
+import sys
+
+def getParser(usage=None):
+    parser = OptionParser(usage=usage)
+    return parser
+
+def linesLessThan(n, fullpath):
+    count = 0
+    with open(fullpath, 'r') as f:
+        for i, l in enumerate(f):
+            if i + 1 >= n:
+                return False
+    return True
+
+def main():
+    # Parse options
+    usage = 'Usage: %prog [options] N <dir>'
+    parser = getParser(usage=usage)
+    (options, args) = parser.parse_args()
+    if len(args) != 2:
+        parser.error('Wrong number of arguments')
+    try:
+        n = int(args[0])
+    except ValueError:
+        print >> sys.stderr, 'Invalid number of lines: %s' % args[0]
+        return
+    if not os.path.isdir(args[1]):
+        print >> sys.stderr, 'Cannot find directory: %s' % args[1]
+        return
+    inputDir = args[1]
+
+    num_removes = 0
+    for filename in os.listdir(inputDir):
+        fullpath = os.path.join(inputDir, filename)
+        if not os.path.isfile(fullpath):
+            continue
+        if linesLessThan(n, fullpath):
+            os.remove(fullpath)
+            num_removes += 1
+    print 'Removed %d files.' % num_removes
+
+if __name__ == '__main__':
+    main()

### 2014-03-24 amazon/aggregatePredictions.py
Tweak program options.

@@ -20,12 +20,8 @@ errorIdx = 4
 
 def getParser(usage=None):
     parser = OptionParser(usage=usage)
-    parser.add_option('-p', '--pattern', dest='pattern', default=None,
+    parser.add_option('-p', '--pattern', dest='pattern', default='*.csv',
         help='Input file pattern.', metavar='PATTERN')
-    parser.add_option('-c', '--cosineFunc', dest='cosineFunc',
-        default='prefSim',
-        help=('Similarity function to use: "prefSim" (default), "randSim", '
-              '"prefSimAlt1", or "randSimAlt1".'), metavar='FUNCNAME')
     return parser
 
 def main():
@@ -39,10 +35,6 @@ def main():
     if not os.path.isdir(inputDir):
         print >> sys.stderr, 'Cannot find: %s' % inputDir
         return
-    if options.pattern:
-        pattern = options.pattern
-    else:
-        pattern = '%s_*_*.csv' % options.cosineFunc
 
     # compile user predictions
     userErrors = {}
@@ -50,7 +42,7 @@ def main():
         fullpath = os.path.join(inputDir, filename)
         if not os.path.isfile(fullpath):
             continue
-        if not fnmatch(filename, pattern):
+        if not fnmatch(filename, options.pattern):
             continue
         print >> sys.stderr, 'Processing %s . . .' % filename
         with open(fullpath, 'rb') as csvfile:

### 2014-03-24 amazon/experiment.py

@@ -51,6 +51,9 @@ def getParser(usage=None):
         help=('Similarity function to use: "prefSim" (default), "randSim", '
               '"prefSimAlt1", "randSimAlt1", "regSim" or "predSim"'),
         metavar='FUNCNAME')
+    parser.add_option('--constSimScore', type='float', dest='constSimScore',
+        default=None, help='Constant similarity estimate used by constSim.',
+        metavar='FLOAT')
     parser.add_option('--regSimRawFunc', dest='regSimRawFunc',
         default='randSim',
         help=('Similarity function to used for raw similarity by regSim: '
@@ -324,6 +327,8 @@ def main():
         similarity.K = options.K
     if options.sigma:
         similarity.sigma = options.sigma
+    if options.constSimScore:
+        similarity.constSimScore = options.constSimScore
 
     # regSim
     if options.cosineFunc == 'regSim':

### 2014-03-24 amazon/modelSim.py

@@ -13,10 +13,15 @@ a single reviewer.
 """
 
 # "official" modelSim params
-mu_s = 0.27
-sigma_s = 0.3
-mu_r = 0.2
-sigma_r = 0.3
+#mu_s = 0.27 # topEdges_80_randSim
+#mu_s = 0.16 # topEdges_40_randSim
+mu_s = 0.050 # topEdges_80_prefSim
+#sigma_s = 0.3 # randSim
+sigma_s = 0.050 # prefSim
+mu_r = 0.085 # topEdges_80
+#mu_r = 0.026 # topEdges_40
+#sigma_r = 0.3 # randSim
+sigma_r = 0.15 # prefSim
 
 def getParser(usage=None):
     parser = OptionParser(usage=usage)

### 2014-03-24 amazon/similarity.py

@@ -15,16 +15,13 @@ import sympy
 from SimilarityGrid import SimilarityGrid
 
 # params
-constSimScore = 0.27
+#constSimScore = 0.16   # topEdges_40_randSim
+#constSimScore = 0.27    # topEdges_80_randSim
+constSimScore = 0.050    # topEdges_80_prefSim
 K = 20
 sigma = 1.0
 writer = csv.writer(sys.stdout)
 step = sys.maxint
-
-# modelSim params
-mu_s = constSimScore
-sigma_s = 0.2
-sigma_r = 0.3
 
 # errorFile
 userIdIdx = 0

### 2014-03-24 paper/paper.tex
Clean up paper paper formatting.

@@
-\documentclass[10pt]{article}
+\documentclass[11pt]{article}
 
@@
 %\setlength{\parskip}{\baselineskip}     % skip line following paragraphs
-\setlength{\parskip}{0.0in}
-\setlength{\parindent}{0.3in}           % Control margins and amount of text
-\setlength{\topmargin}{-0.8in}
-\setlength{\oddsidemargin}{-.15in}      % changed from {-.15in}
-\setlength{\textheight}{9.5in}
-\setlength{\textwidth}{6.8in}
-\pagestyle{empty}                       % No page numbers
+%\pagestyle{empty}                       % No page numbers
+\setlength{\topmargin}{-.5in}
+\setlength{\textheight}{9in}
+\setlength{\oddsidemargin}{.125in}
+\setlength{\textwidth}{6.25in}
 
@@
 
-{\LARGE\bf
-\begin{tabbing}
-\hspace{2.8in} \= \hspace{1.3in} \= \hspace{1.2in} \= \\
-
-%=========    Heading   ==================================================
-CS 399 \> Paper \> Bill Chickering (bchick)\\
-\normalsize Mar 21, 2014 \> \> Jamie Irvine (jirvine)
-% =========================================================================
-\end{tabbing}
-} \vspace{.4in}
+\title{Improving Item-Item Similarity Estimation during Collaborative Filtering}
+\author{Bill Chickering and Jamie Irvine\\
+CS 399 with Anand Rajaraman\\
+Stanford University}
+\renewcommand{\today}{March 25, 2014}
+\maketitle
 
@@
 The similarity of two items is a valuable measurement for a number of
-data-mining applications, such as recommender systems. With large
-amounts of rating data, collaborative filtering is a good method for
-calculating similarity. However, with sparse rating data, this
-technique gives noisy and unreliable approximations for similarity. Even worse, 
-low confidence measurements of similarity are indistinguishable from high
-confidence ones. In this paper, we develop a novel method which incorporates
-confidence into the similarity score to produce more accurate estimates of
-similarity.
+data-mining applications, such as recommender systems. With large amounts of
+rating data, collaborative filtering is a good method for calculating
+similarity. However, with sparse rating data, this technique gives noisy and
+unreliable approximations for similarity. Even worse, low confidence
+measurements of similarity are indistinguishable from high confidence ones. In
+this paper, we develop a novel method which incorporates confidence into the
+similarity score to produce more accurate estimates of similarity.
 
-Item-based collaborative filtering is a common technique for measuring the similarity 
-of two items. There are different forms of collaborative filtering. For
-this paper, we use a traditional approach. Each item is represented as a vector
-of ratings. The similarity score of two items is computed by measuring the similarity
-of the two rating vectors.
+Item-based collaborative filtering is a common technique for measuring the
+similarity of two items. There are different forms of collaborative filtering.
+For this paper, we use a traditional approach. Each item is represented as a
+vector of ratings. The similarity score of two items is computed by measuring
+the similarity of the two rating vectors.
 
 There are a few ways to compute the similarity of two vectors. One approach is
-to use Cosine-similarity, defined as the cosine of the angle between the two vectors:
+to use Cosine-similarity, defined as the cosine of the angle between the two
+vectors:
 \begin{align}
@@
 \end{align}
-where $U_{I}$ is the set of all users who rated item $I$, $U_{AB}$ is the set of all 
-users who rated both item $A$ and item $B$ and $r_{u,I}$ is the rating user $u$ gave 
-to item $I$. Another popular approach is to use Pearson correlation:
+where $U_{I}$ is the set of all users who rated item $I$, $U_{AB}$ is the set of
+all users who rated both item $A$ and item $B$ and $r_{u,I}$ is the rating user
+$u$ gave to item $I$. Another popular approach is to use Pearson correlation:
 \begin{align}
@@
 \end{align}
+
 Other similarity functions exist, such as one-sided similarities, but Pearson
@@
 
-\footnotetext{ Note that the difference between the two similarities is how they handle
-unpaired ratings; that is ratings from a user who has not rated the other
+\footnotetext{ Note that the difference between the two similarities is how they
+handle unpaired ratings; that is ratings from a user who has not rated the other
 item. Cosine-similarity considers the unknown rating from an unpaired rating to
-be $0$ and then calculates the cosine of the two vectors. Pearson correlation simply
-throws away all unpaired ratings and calculates the cosine of the two modified vectors.}
+be $0$ and then calculates the cosine of the two vectors. Pearson correlation
+simply throws away all unpaired ratings and calculates the cosine of the two
+modified vectors.}
 
 Both similarity measurements primarily leverage information from common users,
-that is users who have rated both items. Because of this, they perform
-well with a large number common users, but give unreliable results when the
-items have few common users. In an extreme case, if only one user has rated item
-$A$ and item $B$, $PearsSim(A, B) = 1$ or $-1$. Not only is this unlikely to be an accurate
+that is users who have rated both items. Because of this, they perform well with
+a large number common users, but give unreliable results when the items have few
+common users. In an extreme case, if only one user has rated item $A$ and item
+$B$, $PearsSim(A, B) = 1$ or $-1$. Not only is this unlikely to be an accurate
 assessment of the true similarity of $A$ and $B$, it also gives the most extreme
@@
 
-In this paper, we construct a more accurate similarity score by incorporating the
-number of common users, $n$, into the similarity function.
+In this paper, we construct a more accurate similarity score by incorporating
+the number of common users, $n$, into the similarity function.
 
 \section*{Problem}
-The goal is to calculate a modified similiary function that leverages the number of
-commonn users to better estimate the true similarity. Formally, we design
+
+The goal is to calculate a modified similiary function that leverages the number
+of commonn users to better estimate the true similarity. Formally, we design
 $ModSim$ such that for two items $A$ and $B$ with $n$ users in common
+
 \begin{align}
@@
 Since $TrueSim$ dependends on the choice of $Sim$, we abstract away the details
-of $Sim$ and use its output directly. In all, the goal is to construct $ModSim$ such that
+of $Sim$ and use its output directly. In all, the goal is to construct $ModSim$
+such that
+
 \begin{align}
@@
 \section*{Model}
+
 We model the problem probabilistically. Let $Y$ be a random variable
 representing the $TrueSim$ of a randomly chosen pair of items. Let $X_n$ be a
-random variable representing the calculated $Sim$ of a pair of items with $n$ common users.
+random variable representing the calculated $Sim$ of a pair of items with $n$
+common users.
 
@@
 \end{align}
-where $\mu$ and $\sigma_{1}^2$ are the average similarity score and the variance of similarity scores
-of all pairs of items, respectively. Since $X_n$ represents a noisy reading of
-the true similarity $Y$, we model $X_n | Y$ as a Normal distribution around $Y$:
+where $\mu$ and $\sigma_{1}^2$ are the average similarity score and the variance
+of similarity scores of all pairs of items, respectively. Since $X_n$ represents
+a noisy reading of the true similarity $Y$, we model $X_n | Y$ as a Normal
+distribution around $Y$:
 \begin{align}
@@
 \end{align}
-Note that $\sigma_{2, n}^2$ represents how noisy the estimation of $Sim$ is
-when there are $n$ common users. This should decrease as $n$ increases.
+
+Note that $\sigma_{2, n}^2$ represents how noisy the estimation of $Sim$ is when
+there are $n$ common users. This should decrease as $n$ increases.
 
@@
     \includegraphics[width=0.5\textwidth]{twonormals.png}
-    \caption{The blue curve represents $Y$, the distribution of true similarities
-    between a random pair of items. The red curve represents $X|Y$, the
-    estimated similarity when a small number of common users exist.}
+	\caption{The blue curve represents $Y$, the distribution of true
+    similarities between a random pair of items. The red curve represents $X|Y$,
+    the estimated similarity when a small number of common users exist.}
     \label{fig:two_normals}
@@
 \\&= 
-\argmax_y\left[\frac{1}{\sigma_{2,n}\sqrt{2\pi}}\exp{\left(\frac{-(x-y)^2}{2\sigma_{2,n}^2}\right)}
-\frac{1}{\sigma_{1}\sqrt{2\pi}}\exp{\left(\frac{-(y-\mu)^2}{2\sigma_{1}^2}\right)}\right]
+\argmax_y\left[\frac{1}{\sigma_{2,n}\sqrt{2\pi}}\exp{\left(\frac{-(x-y)^2}
+{2\sigma_{2,n}^2}\right)}
+\frac{1}{\sigma_{1}\sqrt{2\pi}}\exp{\left(\frac{-(y-\mu)^2}
+{2\sigma_{1}^2}\right)}\right]
 \\&= \argmin_y\left[\frac{(x-y)^2}{2\sigma_{2,n}^2} +
@@
 \begin{align}
-\left(y - \mu\right) &= \frac{\sigma_{1}^2}{\sigma_{1}^2+\sigma_{2,n}^2}\left(x-\mu\right)
+\left(y - \mu\right) &= \frac{\sigma_{1}^2}{\sigma_{1}^2+\sigma_{2,n}^2}
+\left(x-\mu\right)
 \end{align}
@@
 $\sigma_{2,n}^2\ge0$, the slope is always $\le1$, meaning that the calculated
-distance from $\mu$ (the average similarity) overestimates the true distance. This makes
-sense, since there is a prior distribution of true similarity weighted around $\mu$.
+distance from $\mu$ (the average similarity) overestimates the true distance.
+This makes sense, since there is a prior distribution of true similarity
+weighted around $\mu$.
 
-The parameters $\mu$, $\sigma_{1}^2$, and $\sigma_{2,n}^2$ all can be approximated 
-using the method of moments over training data where there are enough common
-users that $TrueSim$ can be approximated. This method requires approximating
-$\sigma_{2,n}^2$ for every value of $n$ separately. Alternatively, we can model
-$\sigma_{2,n}^2$ as a function of $n$. Recall that $\sigma_{2,n}^2$ is the
-noisiness of the measurement $Sim$ about $TrueSim$ and $TrueSim = lim_{n \to
-\infty}Sim$. Thus $Sim$ can be thought of as a sampling $n$ users from the
+The parameters $\mu$, $\sigma_{1}^2$, and $\sigma_{2,n}^2$ all can be
+approximated using the method of moments over training data where there are
+enough common users that $TrueSim$ can be approximated. This method requires
+approximating $\sigma_{2,n}^2$ for every value of $n$ separately. Alternatively,
+we can model $\sigma_{2,n}^2$ as a function of $n$. Recall that $\sigma_{2,n}^2$
+is the noisiness of the measurement $Sim$ about $TrueSim$ and $TrueSim = lim_{n
+\to \infty}Sim$. Thus $Sim$ can be thought of as a sampling $n$ users from the
 infinite set used to calculate $TrueSim$. Although $Sim$ may not be linear, we
@@
 would decrease as the $sqrt(n)$. Thus we model $\sigma_{2,n}^2$ as:
-
 \begin{align}
-\sigma_{2,n}^2 = \frac{\alpha}{\sqrt{n}}
+\sigma_{2,n}^2 = \frac{\alpha}{\sqrt{n}},
 \end{align}
-Yielding the single parameter model:
+yielding the single parameter model:
 \begin{align}

### 2014-03-24 paper/paper.tex
Added method of moments and alpha techniques to similarity.py and experiment.py. Added momentsTrain.py and alphaParams.py to get parameters for these techniques. Added technique descriptions to paper

--- 2014-03-22 paper/paper.tex
@@
 The model suggests that there is a linear correlation between $X_n$, the
-similarity calculated with $n$ common users, and $Y$, the true similarity. Since 
-$\sigma_{2,n}^2\ge0$, the slope is always $\le1$, meaning that the calculated
-distance from $\mu$ (the average similarity) overestimates the true distance. This makes
-sense, since there is a prior distribution of true similarity weighted around $\mu$.
+similarity calculated with $n$ common users, and $Y$, the true similarity.
+Moreover, the best predicted $y$ is a linear combination of the mean and the
+observed similiatiry. Since $\sigma_{2,n}^2\ge0$, the slope is always $\le1$, 
+meaning that the calculated distance from $\mu$ (the average similarity) 
+overestimates the true distance. This makes sense, since there is a prior 
+distribution of true similarity weighted around $\mu$.
 
@@
 \section*{Technique}
+Based on this maodel, we try three different techniques to predict the true similarity from a
+measured similarity. The techniques vary in their fidelity to the model. For
+each, we limit our data to all pairs of items that have enough common users that
+we can approximate $TrueSim \approx Sim$. Then we pick subsets of the common
+users of various sizes $n$ and calculate $Sim_{n}$. With examples of $Sim$, $n$
+and the resulting $TrueSim$, we can implement supervised learning.
 
-\begin{itemize}
-    \item Discussion of abandonment of parameterization in exchange for linear regression
-    \item Pick data that's big enough to have gold standard
-    \item Slice data and group by n
-    \item Run linear regression on each group of n
-    \item For predicting true similarity, take a given point and project on appropriate
-    \item regression.
-\end{itemize}
+The first technique is the least faithful to the model. It simply runs a number
+of linear regressions between $Sim$ and $TrueSim$ for each $n$. This ignores the 
+data-specific parameters $\mu$, $\sigma_{1}$ and $\sigma_{2,n}$ to find the best 
+linear fit. It also treats each $n$ completely independently. This technique 
+necessarilty has the lowest training error of all linear techniques, potentially 
+at the risk of overfitting.
+
+The second technique is more general. It approximates $\mu$ and $\sigma_{1}$
+using the method of moments on the training data of all $TrueSim$ scores. Then,
+$\sigma_{2,n}$ is also approximated using a method of moments on the training
+data of all $Sim_n$ for each $n$ separately. Here $\sigma_{2,n}$'s are also
+considered independently of each other.
+
+The third and most general technique models $\sigma_{2,n}$ as a function of
+$n$ using equation 19. As in the previous technique, $\mu$, $\sigma_{1}$ and each
+$\sigma_{2,n}$ are approximated using the method of moments over the training
+data. Then $\alpha$ is calculated that minimizes the sum of squared errors
+between $\sigma_{2,n}^2$ and $\frac{\alpha}{\sqrt{n}}$. This $\alpha$ is used to
+construct the three-parameter prediction function of equation 19.
 

### 2014-03-24 amazon/alphaParams.py

@@ -1,0 +1,66 @@
+#!/usr/local/bin/python
+
+"""
+Fit sigma2_n to the function alpha / sqrt(n).
+Take as input the parameter file output from momemtsTrain.py and produce 
+alpha parameters file for use by momSim.
+"""
+
+from optparse import OptionParser
+import csv
+import os
+import sys
+import mlpy
+import math
+
+def getParser(usage=None):
+    parser = OptionParser(usage=usage)
+    return parser
+
+def harmonic(limit):
+    total = 0.0
+    for n in range(1, limit+1):
+        total += 1.0 / n
+    return total
+
+def learnAlpha(sigma2_n):
+    harmonicSum = harmonic(len(sigma2_n))
+    total = 0.0
+    for i in range(len(sigma2_n)):
+        total += float(sigma2_n[i]) / math.sqrt(i + 1)
+    return total / harmonicSum
+
+def main():
+    # parse options
+    usage = 'Usage: %prog momParamFile'
+    parser = getParser(usage=usage)
+    (options, args) = parser.parse_args()
+    if len(args) != 1:
+        parser.error('Wrong number of arguments')
+    inputFileName = args[0]
+    if not os.path.isfile(inputFileName):
+        print >> sys.stderr, 'Cannot find: %s' % inputDir
+        return
+
+    # setup output writer
+    writer = csv.writer(sys.stdout)
+
+    variableIdx = 0
+    valueIdx = 1
+    # process data
+    with open(inputFileName, 'rb') as csvfile:
+        reader = csv.reader(csvfile)
+        # read data
+        data = [row for row in reader]
+        variables = [row[variableIdx] for row in data]
+        values = [row[valueIdx] for row in data]
+        # first two rows of input are mu and sigma1
+        mu = values[0]
+        sigma1 = values[1]
+        sigma2_n = values[2:]
+        alpha = learnAlpha(sigma2_n)
+        # write sigma2s
+        writer.writerow([mu, sigma1, alpha])
+
+if __name__ == '__main__':
+    main()

### 2014-03-24 amazon/experiment.py

@@ -49,18 +49,32 @@ def getParser(usage=None):
     parser.add_option('-c', '--cosineFunc', dest='cosineFunc',
         default='prefSim',
         help=('Similarity function to use: "prefSim" (default), "randSim", '
-              '"prefSimAlt1", "randSimAlt1", "regSim" or "predSim"'),
+              '"prefSimAlt1", "randSimAlt1", "regSim", "momSim", "alphaSim" or "predSim"'),
         metavar='FUNCNAME')
     parser.add_option('--regSimRawFunc', dest='regSimRawFunc',
         default='randSim',
         help=('Similarity function to used for raw similarity by regSim: '
               '"prefSim", or "randSim" (default)'), metavar='FUNCNAME')
+    parser.add_option('--momSimRawFunc', dest='momSimRawFunc',
+        default='randSim',
+        help=('Similarity function to used for raw similarity by momSim: '
+              '"prefSim", or "randSim" (default)'), metavar='FUNCNAME')
+    parser.add_option('--alphaSimRawFunc', dest='alphaSimRawFunc',
+        default='randSim',
+        help=('Similarity function to used for raw similarity by alphaSim: '
+              '"prefSim", or "randSim" (default)'), metavar='FUNCNAME')
     parser.add_option('--regSimParamsFile', dest='regSimParamsFile',
         default='output/regSim_params.csv', help='Parameter file for regSim.',
         metavar='FILE')
+    parser.add_option('--momSimParamsFile', dest='momSimParamsFile',
+        default='output/momSim_params.csv', help='Parameter file for momSim.',
+        metavar='FILE')
+    parser.add_option('--alphaSimParamsFile', dest='alphaSimParamsFile',
+        default='output/alphaSim_params.csv', help='Parameter file for alphaSim.',
+        metavar='FILE')
     parser.add_option('--max-common-reviewers', dest='maxUsersCommon',
         type='int', default=100,
-        help='Maximum number of common reviewers for regSim.', metavar='NUM')
+        help='Maximum number of common reviewers for regSim, momSim, or alphaSim.', metavar='NUM')
     parser.add_option('--simGridFile', dest='simGridFile',
         default='output/simGrid_randSim.csv',
         help='CSV containing simGrid data.', metavar='FILE')
@@ -306,6 +320,18 @@ def main():
         print >> sys.stderr,\
             'Invalid Raw Similarity function: %s' % options.regSimRawFunc
         return
[…]

### 2014-03-24 amazon/momentsTrain.py

@@ -1,0 +1,81 @@
+#!/usr/local/bin/python
+
+"""
+Use method of moments to learn mu, sigma1 and sigma2_n for each n.
+Take as input the output of regressionTrainData.py and produce parameters file
+for use by momSim.
+"""
+
+from optparse import OptionParser
+import csv
+import os
+import sys
+import mlpy
+
+# params
+scoreIdx = 0
+trueScoreIdx = 1
+
+def getParser(usage=None):
+    parser = OptionParser(usage=usage)
+    parser.add_option('-m', '--max-common-reviewers', dest='maxUsersCommon',
+        type='int', default=100, help='Maximum number of common reviewers.',
+        metavar='NUM')
+    return parser
+
+def mean(scores):
+    total = 0.0
+    for score in scores:
+        total += float(score)
+    return total / len(scores)
+
+def variance(scores):
+    total = 0.0
+    sqrTotal = 0.0
+    for score in scores:
+        score = float(score)
+        total += score
+        sqrTotal += score*score
+    mean = total / len(scores)
+    secondMoment = sqrTotal / len(scores)
+    return secondMoment - mean*mean
+
+def main():
+    # Parse options
+    usage = 'Usage: %prog [options] inputDir'
+    parser = getParser(usage=usage)
+    (options, args) = parser.parse_args()
+    if len(args) != 1:
+        parser.error('Wrong number of arguments')
+    inputDir = args[0]
+    if not os.path.isdir(inputDir):
+        print >> sys.stderr, 'Cannot find: %s' % inputDir
+        return
+
+    # setup output writer
+    writer = csv.writer(sys.stdout)
+
+    # process data
+    inputfileTemplate = '%d.csv'
+    for numUsersCommon in range(1, options.maxUsersCommon+1):
+        inputfileName = inputfileTemplate % numUsersCommon
+        #print 'Processing %s . . .' % inputfileName
+        fullpath = os.path.join(inputDir, inputfileName)
+        if not os.path.isfile(fullpath):
+            print >> sys.stderr, 'ERROR: Cannot find %s' % fullpath
+            break
+        with open(fullpath, 'rb') as csvfile:
+            reader = csv.reader(csvfile)
+            # read data
+            data = [row for row in reader]
+            scores = [row[scoreIdx] for row in data]
+            trueScores = [row[trueScoreIdx] for row in data]
+            # first two rows of output are mu and sigma1
+            if numUsersCommon == 1:
[…]

### 2014-03-24 amazon/similarity.py

@@ -42,6 +42,35 @@ def regSim(reviewsA, reviewsB):
     else:
         return (rawSim, numUsersCommon)
 
+def momSim(reviewsA, reviewsB):
+    """Uses params from method of moments on training data to translate
+       raw similarity to an improved esimate.
+    """
+    rawSim, numUsersCommon = momSim_rawFunc(reviewsA, reviewsB)
+    if numUsersCommon <= momSim_maxCommonUsers:
+        mu = momSimParams['mu']
+        sigma1 = momSimParams['sigma1']
+        sigma2_n = momSimParams['sigma2_n'][numUsersCommon]
+        predictSim = (sigma1 / (sigma1 + sigma2_n)) * (rawSim - mu) + mu
+        return (predictSim, numUsersCommon)
+    else:
+        return (rawSim, numUsersCommon)
+
+def alphaSim(reviewsA, reviewsB):
+    """Uses params from method of moments and alpha fit for sigma2s to translate
+       raw similarity to an improved esimate.
+    """
+    rawSim, numUsersCommon = alphaSim_rawFunc(reviewsA, reviewsB)
+    if numUsersCommon <= alphaSim_maxCommonUsers:
+        mu = float(alphaSimParams['mu'])
+        sigma1 = float(alphaSimParams['sigma1'])
+        alpha = float(alphaSimParams['alpha'])
+        sigma2 = alpha / math.sqrt(int(numUsersCommon))
+        predictSim = (sigma1 / (sigma1 + sigma2)) * (rawSim - mu) + mu
+        return (predictSim, numUsersCommon)
+    else:
+        return (rawSim, numUsersCommon)
+
 def constSim(reviewsA, reviewsB):
     """Provides a baseline similarity function by always returning the
        average item-item similarity.
@@ -385,6 +414,68 @@ def initRegSim(paramFileName, cosineFunc, maxCommonUsers):
             slope = float(row[slopeIdx])
             regSimParams[numUsersCommon] = (intercept, slope)
 
+#########################
+#
+#  Method of Moments
+#
+##############################
+
+momSimParams = {}
+momSim_rawFunc = None
+momSim_maxCommonUsers = 0
+
+variableIdx = 0
+valueIdx = 1
+
+def initMomSim(paramFileName, cosineFunc, maxCommonUsers):
+    global momSimParams
+    global momSim_rawFunc
+    global momSim_maxCommonUsers
[…]

### 2014-03-24 paper/paper.tex

--- 2014-03-24 paper/paper.tex
@@
 measurements of similarity are indistinguishable from high confidence ones. In
-this paper, we develop a novel method which incorporates confidence into the
-similarity score to produce more accurate estimates of similarity.
+this paper, we develop and compare a number of methods to produce more accurate
+estimates of similarity by incorporating confidence into the similarity score.
 
@@
 The model suggests that there is a linear correlation between $X_n$, the
-similarity calculated with $n$ common users, and $Y$, the true similarity. Since
-$\sigma_{2,n}^2\ge0$, the slope is always $\le1$, meaning that the calculated
-distance from $\mu$ (the average similarity) overestimates the true distance.
-This makes sense, since there is a prior distribution of true similarity
-weighted around $\mu$.
+similarity calculated with $n$ common users, and $Y$, the true similarity.
+Moreover, the best predicted $y$ is a linear combination of the mean and the
+observed similiatiry. Since $\sigma_{2,n}^2\ge0$, the slope is always $\le1$, 
+meaning that the calculated distance from $\mu$ (the average similarity) 
+overestimates the true distance. This makes sense, since there is a prior 
+distribution of true similarity weighted around $\mu$.
 
-The parameters $\mu$, $\sigma_{1}^2$, and $\sigma_{2,n}^2$ all can be
-approximated using the method of moments over training data where there are
-enough common users that $TrueSim$ can be approximated. This method requires
-approximating $\sigma_{2,n}^2$ for every value of $n$ separately. Alternatively,
-we can model $\sigma_{2,n}^2$ as a function of $n$. Recall that $\sigma_{2,n}^2$
-is the noisiness of the measurement $Sim$ about $TrueSim$ and $TrueSim = lim_{n
-\to \infty}Sim$. Thus $Sim$ can be thought of as a sampling $n$ users from the
+The parameters $\mu$, $\sigma_{1}^2$, and $\sigma_{2,n}^2$ all can be approximated 
+using the method of moments over training data where there are enough common
+users that $TrueSim$ can be approximated. This method requires approximating
+$\sigma_{2,n}^2$ for every value of $n$ separately. Alternatively, we can model
+$\sigma_{2,n}^2$ as a function of $n$. Recall that $\sigma_{2,n}^2$ is the
+noisiness of the measurement $Sim$ about $TrueSim$ and $TrueSim = lim_{n \to
+\infty}Sim$. Thus $Sim$ can be thought of as a sampling $n$ users from the
 infinite set used to calculate $TrueSim$. Although $Sim$ may not be linear, we

### 2014-03-24 paper/paper.tex
Add figures to paper.

@@
 \usepackage{graphicx}
-\usepackage{caption}
+\usepackage[font={small}]{caption}
 \usepackage{subcaption}
@@
 
+\begin{figure}[!htbp]
+    \centering
+    \includegraphics[width=0.7\textwidth]{RatingDist_40.png}
+    \includegraphics[width=0.7\textwidth]{RatingDist_80.png}
+	\caption{Distribution of rating pairs by common reviewers for item pairs.
+Top figure shows distribution for item pairs having at least 40 common
+reviewers. Bottom figure shows distribution for item pairs having at least 80
+common reviewers.}
+    \label{fig:RatingDist}
+\end{figure}
+
+\begin{figure}[!htbp]
+    \centering
+    \includegraphics[width=1.0\textwidth]{Histograms.png}
+	\caption{Histograms of {\em TrueSimR} (top) and {\em TrueSimP} (bottom) for
+training datasets containing at least 40 common reviewers (left) and 80 common
+reviewers (right).}
+    \label{fig:Histograms}
+\end{figure}
+
+\begin{figure}[!htbp]
+    \centering
+    \includegraphics[width=0.9\textwidth]{MSE_SimR_40.png}
+    \includegraphics[width=0.9\textwidth]{MSE_SimR_80.png}
+	\caption{Mean square error {\em vs} number of common reviewers, $n$, for
+estimates of {\em TrueSimR} for item pairs ultimately having at least 40 common
+reviewers (top) and 80 common reviewers (bottom). The dashed curve represents
+{\em SimR} computed using only the ratings from the first $n$ reviewers. The
+flat red line indicate the MSE that results from always guessing the average
+{\em TrueSimR} from the training set. The green curve is {\em RegSimR} and the
+blue curve is {\em ModelSimR}. }
+    \label{fig:MSE_SimR}
+\end{figure}
+
+\begin{figure}[!htbp]
+    \centering
+    \includegraphics[width=0.9\textwidth]{MSE_SimP_80.png}
+	\caption{Mean square error {\em vs} number of common reviewers, $n$, for
+estimates of {\em TrueSimP} for item pairs ultimately having at least 80 common
+reviewers. The dashed curve represents
+{\em SimP} computed using only the ratings from the first $n$ reviewers. The
+flat red line indicate the MSE that results from always guessing the average
+{\em TrueSimP} from the training set. The green curve is {\em RegSimP} and the
+blue curve is {\em ModelSimP}. }
+    \label{fig:MSE_SimP}
+\end{figure}
+
 \section*{Conclusion}

### 2014-03-24 paper/paper.tex
small changes to paper

--- 2014-03-24 paper/paper.tex
@@
 of similarity scores of all pairs of items, respectively. Since $X_n$ represents
-a noisy reading of the true similarity $Y$, we model $X_n | Y$ as a Normal
-distribution around $Y$:
+a noisy reading of the true similarity $Y$, we model $X_n | Y$ as a Gaussian
+error around $Y$:
 \begin{align}
@@
     similarities between a random pair of items. The red curve represents $X|Y$,
-    the estimated similarity when a small number of common users exist.}
+    the observed similarity when a small number of common users exist.}
     \label{fig:two_normals}

### 2014-03-24 paper/paper.tex
Add opening paragraph.

@@
 \section*{Introduction}
-The similarity of two items is a valuable measurement for a number of
-data-mining applications, such as recommender systems. With large amounts of
-rating data, collaborative filtering is a good method for calculating
-similarity. However, with sparse rating data, this technique gives noisy and
-unreliable approximations for similarity. Even worse, low confidence
-measurements of similarity are indistinguishable from high confidence ones. In
-this paper, we develop and compare a number of methods to produce more accurate
-estimates of similarity by incorporating confidence into the similarity score.
+Accurate estimations of the similarity of two users or items is valuable for a
+number of data-mining applications including recommender systems, product
+assortment and differentiation, as well as branding. Collaborative filtering is
+often used to exploit user-item rating data to make such estimates typically by
+computing the Pearson correlation or cosine similarity of vector representations
+of the users or items \cite{Breese98}. Given sufficient user-item rating data,
+this approach can be very effective. On the otherhand, when very little rating
+data for one or both users/items is available, unmodified, these metrics can
+yield dramatic discrepancies compared to values computed later once sufficient
+data has been acquired. In this report, we describe two novel techniques for
+estimating the true similarity of two users or items given limited common rating
+data.
+
+.....
+
+Even worse, low confidence measurements of similarity are indistinguishable from
+high confidence ones. In this paper, we develop and compare a number of methods
+to produce more accurate estimates of similarity by incorporating confidence
+into the similarity score.
 

### 2014-03-24 paper/paper.tex
small changes to paper

--- 2014-03-24 paper/paper.tex
@@
 representing the $TrueSim$ of a randomly chosen pair of items. Let $X_n$ be a
-random variable representing the calculated $Sim$ of a pair of items with $n$
+random variable representing the observed $Sim$ of a pair of items with $n$
 common users.
@@
     \centering
-    \includegraphics[width=0.5\textwidth]{twonormals.png}
+    \includegraphics[width=0.7\textwidth]{twonormals.png}
 	\caption{The blue curve represents $Y$, the distribution of true
@@
 
-In probabilistic terms, for a calculated value $X_n$, we want to find the value
+In probabilistic terms, for an observed value $X_n$, we want to find the value
 of $Y$ that most likely produced $X_n$. Thus we desire the Maximum Likelihood
@@
 The model suggests that there is a linear correlation between $X_n$, the
-similarity calculated with $n$ common users, and $Y$, the true similarity.
+observed similarity with $n$ common users, and $Y$, the true similarity.
 Moreover, the best predicted $y$ is a linear combination of the mean and the
-observed similiatiry. Since $\sigma_{2,n}^2\ge0$, the slope is always $\le1$, 
-meaning that the calculated distance from $\mu$ (the average similarity) 
-overestimates the true distance. This makes sense, since there is a prior 
-distribution of true similarity weighted around $\mu$.
+observed similiarity. Since $\sigma_{2,n}^2\ge0$, the slope is always $\le1$, 
+meaning that the observed distance from $\mu$ (the average similarity) 
+overestimates the true distance. This is reasonable, since there is a prior 
+that expects the true similarity to be around $\mu$.
 
-The parameters $\mu$, $\sigma_{1}^2$, and $\sigma_{2,n}^2$ all can be approximated 
+The parameters $\mu$, $\sigma_{1}^2$, and $\sigma_{2,n}^2$ can be approximated 
 using the method of moments over training data where there are enough common
 users that $TrueSim$ can be approximated. This method requires approximating
-$\sigma_{2,n}^2$ for every value of $n$ separately. Alternatively, we can model
-$\sigma_{2,n}^2$ as a function of $n$. Recall that $\sigma_{2,n}^2$ is the
-noisiness of the measurement $Sim$ about $TrueSim$ and $TrueSim = lim_{n \to
+$\sigma_{2,n}^2$ for every value of $n$ separately. To gain a more general 
+model, we can model $\sigma_{2,n}^2$ as a function of $n$. Recall that 
+$\sigma_{2,n}^2$ is the noisiness of the measurement $Sim$ about $TrueSim$ and $TrueSim = lim_{n \to
 \infty}Sim$. Thus $Sim$ can be thought of as a sampling $n$ users from the

### 2014-03-24 paper/paper.tex
Edited first paragraph

@@
 \section*{Introduction}
+Determining the similarity of two users or items is valuable for a
+number of data-mining applications including recommender systems, product
+assortment and differentiation, and branding. Collaborative filtering is
+often used to exploit user-item rating data to make such estimates. This is 
+typically done by computing the Pearson correlation or cosine similarity of 
+vector representations of the users or items \cite{Breese98}. Given sufficient 
+user-item rating data, this approach can be very effective. However, when the
+rating data is sparse for one or both users/items, these metrics can
+yield dramatic discrepancies compared to values computed with sufficient
+data at a later point in time. In this report, we describe two novel techniques 
+for estimating the true similarity of two users or items given limited common 
+rating data.
+
+...
 Accurate estimations of the similarity of two users or items is valuable for a

### 2014-03-24 paper/paper.tex
Add second paragraph.

@@
 computing the Pearson correlation or cosine similarity of vector representations
-of the users or items \cite{Breese98}. Given sufficient user-item rating data,
+of the users or items \cite{Su2009}. Given sufficient user-item rating data,
 this approach can be very effective. On the otherhand, when very little rating
@@
 data.
+
+To keep the present work focused, we confine ourselves to memory-based
+collaborative filtering (CF) in which we work with a logical user-item matrix
+with elements corresponding to ratings given by a particular user to a
+particular item. Further, we choose to limit our analysis to item-item
+similarity but note that these techniques should be equally effective at
+improving user-user similarity estimation. An item is now represented as vector
+of the ratings they have received from all users. We now formalize the notion of
+item-item similarity. Two alternative models for the generation of user-item
+ratings are useful here in arriving at two different definitions of the {\em
+true} item-item similarity.
+
+In the first model, we imagine users are randomly presented with items to rate.
+Given a sufficient period of time, all users will eventually rate all items such
+that all elements of an item's vector correspond to a provided rating. The true
+similarity in this model of {\em random} rating, which we denote $TrueSimR$, is
+then the vector product of two item vectors. That is, 
+\begin{align}
+TrueSimR(A, B) = \frac{\sum\limits_{u\in U}
+r_{u,A}r_{u,B}}{\sqrt{\sum\limits_{u\in U} r_{u,A}^2}
+\sqrt{\sum\limits_{u\in U} r_{u,B}^2}}
+\end{align}
+where $U$ is the set of all users and $r_{u,I}$ is the rating given by user $u$
+to item $I$.
+
+In the second model, we imagine an unlimited supply of users each of whom
+chooses to rate a subset of items. At any point in time, we may estimate the
+similarity in this model of {\em preferential} rating, by computing the cosine
+similarity of the two item vectors. Only in the limit of inifinite time does
+this estimate converge with the true similarity, which we denote $TrueSimP$.
+That is,
+\begin{align}
+TrueSimP(A, B) = lim_{t\to\infty}\frac{\sum\limits_{u\in U_{AB}}
+r_{u,A}r_{u,B}}{\sqrt{\sum\limits_{u\in U_A} r_{u,A}^2}
+\sqrt{\sum\limits_{u\in U_B} r_{u,B}^2}}
+\end{align}
+where $U_A$ is the set of users who have rated $A$, $U_B$ is the set of users
+who have rated $B$, and $U_{AB}$ is the set of users who have rated both $A$ and
+$B$.
 

### 2014-03-24 paper/paper.tex
Remove old first paragraph.

@@
 number of data-mining applications including recommender systems, product
-assortment and differentiation, and branding. Collaborative filtering is
+assortment and differentiation, as well as branding. Collaborative filtering is
 often used to exploit user-item rating data to make such estimates. This is 
 typically done by computing the Pearson correlation or cosine similarity of 
-vector representations of the users or items \cite{Breese98}. Given sufficient 
+vector representations of the users or items \cite{Su2009}. Given sufficient 
 user-item rating data, this approach can be very effective. However, when the
@@
 rating data.
-
-...
-Accurate estimations of the similarity of two users or items is valuable for a
-number of data-mining applications including recommender systems, product
-assortment and differentiation, as well as branding. Collaborative filtering is
-often used to exploit user-item rating data to make such estimates typically by
-computing the Pearson correlation or cosine similarity of vector representations
-of the users or items \cite{Su2009}. Given sufficient user-item rating data,
-this approach can be very effective. On the otherhand, when very little rating
-data for one or both users/items is available, unmodified, these metrics can
-yield dramatic discrepancies compared to values computed later once sufficient
-data has been acquired. In this report, we describe two novel techniques for
-estimating the true similarity of two users or items given limited common rating
-data.
 

### 2014-03-24 paper/paper.tex
merging
Tweak definition of TrueSimP.

@@
 In the second model, we imagine an unlimited supply of users each of whom
-chooses to rate a subset of items. At any point in time, we may estimate the
-similarity in this model of {\em preferential} rating, by computing the cosine
-similarity of the two item vectors. Only in the limit of inifinite time does
-this estimate converge with the true similarity, which we denote $TrueSimP$.
-That is,
+chooses to rate a subset of items. The lack of a rating is now meaningful since
+it indicates a lack of interest of a particular user for a particular item. This
+is captured by the model through a default rating of zero \footnote{We address
+the issue of rating bias and the precise meaning of a zero valued rating later
+in this report.}. At any point in time, we may estimate the true similarity in
+this model of {\em preferential} rating, by computing the cosine similarity of
+the two item vectors. Only in the limit of inifinite time does this estimate
+converge to the true similarity, which we denote $TrueSimP$.  That is,
 \begin{align}

### 2014-03-25 paper/paper.tex
edited paragraphs on TrueSim. Changed names to Rand, Pref

@@
 To keep the present work focused, we confine ourselves to memory-based
-collaborative filtering (CF) in which we work with a logical user-item matrix
-with elements corresponding to ratings given by a particular user to a
+collaborative filtering (CF) in which we work with a logical user-item matrix.
+Each element of the matrix corresponds to a rating given by a particular user to a
 particular item. Further, we choose to limit our analysis to item-item
 similarity but note that these techniques should be equally effective at
-improving user-user similarity estimation. An item is now represented as vector
-of the ratings they have received from all users. We now formalize the notion of
-item-item similarity. Two alternative models for the generation of user-item
-ratings are useful here in arriving at two different definitions of the {\em
-true} item-item similarity.
+improving user-user similarity estimation. In this representaion, an item is
+a vector of the ratings it has received from all users. We now formalize the 
+notion of {\em true} item-item similarity. Two alternative models for the 
+generation of user-item ratings are useful in arriving at two different 
+definitions of the {\em true} item-item similarity.
 
-In the first model, we imagine users are randomly presented with items to rate.
-Given a sufficient period of time, all users will eventually rate all items such
-that all elements of an item's vector correspond to a provided rating. The true
-similarity in this model of {\em random} rating, which we denote $TrueSimR$, is
-then the vector product of two item vectors. That is, 
+In the first model, we imagine users are randomly presented items to rate. Given 
+a sufficient period of time, all users will eventually rate all items, creating
+item vectors that contain provided ratings from all users. The true similarity 
+in this model of {\em random} rating, denoted $TrueSimRand$, is then the vector 
+product of two complete item vectors. That is, 
 \begin{align}
-TrueSimR(A, B) = \frac{\sum\limits_{u\in U}
+TrueSimRand(A, B) = \frac{\sum\limits_{u\in U}
 r_{u,A}r_{u,B}}{\sqrt{\sum\limits_{u\in U} r_{u,A}^2}
@@
 \end{align}
-where $U$ is the set of all users and $r_{u,I}$ is the rating given by user $u$
-to item $I$.
+where $A$ and $B$ are items, $U$ is the set of all users and $r_{u,I}$ is the 
+rating given by user $u$ to item $I$.
 
@@
 the two item vectors. Only in the limit of inifinite time does this estimate
-converge to the true similarity, which we denote $TrueSimP$.  That is,
+converge to the true similarity, which we denote $TrueSimPref$.  That is,
 \begin{align}
-TrueSimP(A, B) = lim_{t\to\infty}\frac{\sum\limits_{u\in U_{AB}}
+TrueSimPref(A, B) = lim_{t\to\infty}\frac{\sum\limits_{u\in U_{AB}}
+
 r_{u,A}r_{u,B}}{\sqrt{\sum\limits_{u\in U_A} r_{u,A}^2}

### 2014-03-25 paper/paper.tex
Add first paragraph of Estimating... section.

--- 2014-03-24 paper/paper.tex
@@
 
+\section*{True Item-Item Similarity}
+
 In the first model, we imagine users are randomly presented with items to rate.
@@
 that all elements of an item's vector correspond to a provided rating. The true
-similarity in this model of {\em random} rating, which we denote $TrueSimR$, is
-then the vector product of two item vectors. That is, 
+similarity in this model of {\em random} rating, which we denote $TrueSimRand$,
+is then the vector product of two item vectors. That is,
 \begin{align}
-TrueSimR(A, B) = \frac{\sum\limits_{u\in U}
+TrueSimRand(A, B) = \frac{\sum\limits_{u\in U}
 r_{u,A}r_{u,B}}{\sqrt{\sum\limits_{u\in U} r_{u,A}^2}
-\sqrt{\sum\limits_{u\in U} r_{u,B}^2}}
+\sqrt{\sum\limits_{u\in U} r_{u,B}^2}},
 \end{align}
@@
 is captured by the model through a default rating of zero \footnote{We address
-the issue of rating bias and the precise meaning of a zero valued rating later
+the issue of rating bias and the precise meaning of a zero valued ratings later
 in this report.}. At any point in time, we may estimate the true similarity in
@@
 the two item vectors. Only in the limit of inifinite time does this estimate
-converge to the true similarity, which we denote $TrueSimP$.  That is,
+converge to the true similarity, which we denote $TrueSimPref$.  That is,
 \begin{align}
-TrueSimP(A, B) = lim_{t\to\infty}\frac{\sum\limits_{u\in U_{AB}}
+TrueSimPref(A, B) = lim_{t\to\infty}\frac{\sum\limits_{u\in U_{AB}}
 r_{u,A}r_{u,B}}{\sqrt{\sum\limits_{u\in U_A} r_{u,A}^2}
-\sqrt{\sum\limits_{u\in U_B} r_{u,B}^2}}
+\sqrt{\sum\limits_{u\in U_B} r_{u,B}^2}},
 \end{align}
@@
 $B$.
+
+\section*{Estimating Item-Item Similarity}
+
+Of course, most systems will never acquire ratings from all users for all items.
+Nor are we able to work in the limit of infinite time. We must therefore make
+estimates of the true item-item similarity using only partial information. In
+the case of $TrueSimPref$, the cosine similarity computed at any point in time
+serves as a simple and obvious estimate. In using this estimate, we are treating
+all missing ratings as intentionally omitted. When estimating $TrueSimRand$, the
+choice of how to handle missing ratings is perhaps less obvious. Following the
+convention of Breese et al. \cite{Breese1998}, we choose to use the Pearson
+correlation \footnote{The Pearson correlation is traditionally characterized by
+subtracting an expectation value from each dimension. This is only approximately
+true in our case. We discuss the how we subtract rating biases later in this
+report.} for this estimate, ignoring all ratings by users who have not rated
+both items. To summarize, our naive estimates are
+\begin{align}
+PearSim(A, B) = \frac{\sum\limits_{u\in U_{AB}}
+r_{u,A}r_{u,B}}{\sqrt{\sum\limits_{u\in U_{AB}} r_{u,A}^2}
+\sqrt{\sum\limits_{u\in U_{AB}} r_{u,B}^2}}
+\stackrel{?}{\approx} TrueSimRand(A, B)
+\end{align}
+
+\begin{align}
+CosSim(A, B) = \frac{\sum\limits_{u\in U_{AB}}
+r_{u,A}r_{u,B}}{\sqrt{\sum\limits_{u\in U_A} r_{u,A}^2}
+\sqrt{\sum\limits_{u\in U_B} r_{u,B}^2}}
+\stackrel{?}{\approx} TrueSimPref(A, B).
+\end{align}
 

### 2014-03-25 paper/paper.tex

@@
 To keep the present work focused, we confine ourselves to memory-based
-collaborative filtering (CF) in which we work with a logical user-item matrix
-with elements corresponding to ratings given by a particular user to a
+collaborative filtering (CF) in which we work with a logical user-item matrix.
+Each element of the matrix corresponds to a rating given by a particular user to a
 particular item. Further, we choose to limit our analysis to item-item
 similarity but note that these techniques should be equally effective at
-improving user-user similarity estimation. An item is now represented as vector
-of the ratings they have received from all users. We now formalize the notion of
-item-item similarity. Two alternative models for the generation of user-item
-ratings are useful here in arriving at two different definitions of the {\em
-true} item-item similarity.
+improving user-user similarity estimation. In this representaion, an item is
+a vector of the ratings it has received from all users. We now formalize the 
+notion of {\em true} item-item similarity. Two alternative models for the 
+generation of user-item ratings are useful in arriving at two different 
+definitions of the {\em true} item-item similarity.
 
@@
 
-In the first model, we imagine users are randomly presented with items to rate.
-Given a sufficient period of time, all users will eventually rate all items such
-that all elements of an item's vector correspond to a provided rating. The true
-similarity in this model of {\em random} rating, which we denote $TrueSimRand$,
-is then the vector product of two item vectors. That is,
+In the first model, we imagine users are randomly presented items to rate. Given 
+a sufficient period of time, all users will eventually rate all items, creating
+item vectors that contain provided ratings from all users. The true similarity 
+in this model of {\em random} rating, denoted $TrueSimRand$, is then the vector 
+product of two complete item vectors. That is, 
 \begin{align}

### 2014-03-25 paper/paper.tex
Adding to bib
Edited probabilistic model section for linear model

@@
 a vector of the ratings it has received from all users. We now formalize the 
-notion of {\em true} item-item similarity. Two alternative models for the 
-generation of user-item ratings are useful in arriving at two different 
-definitions of the {\em true} item-item similarity.
+notion of item-item similarity. Two alternative models for the generation of 
+user-item ratings are useful in arriving at two different definitions of the 
+{\em true} item-item similarity.
 
@@
 
-In the second model, we imagine an unlimited supply of users each of whom
+In the second model, we imagine an unlimited supply of users, each of whom
 chooses to rate a subset of items. The lack of a rating is now meaningful since
@@
 
-\section*{Model}
+\section*{Linear Approach}
+\subsection*{Probabilistic Model}
 
-We model the problem probabilistically. Let $Y$ be a random variable
-representing the $TrueSim$ of a randomly chosen pair of items. Let $X_n$ be a
-random variable representing the observed $Sim$ of a pair of items with $n$
-common users.
+In our first approach, we model true and observed similarity probabilistically. For
+this analysis, we can consider an abstact $TrueSim$ which ignores the details of 
+$TrueSimRand$ or $TrueSimPref$. Similarly, we consider $Sim_{n}$ and $GoldSim$ to be
+the generic observed similarity scores. Let $Y$ be a random variable representing 
+the $TrueSim$ of a randomly chosen pair of items. Let $X_{n}$ be a random variable 
+representing the $Sim_{n}$ of a pair of items when $n$ common users are observed.
 
@@
 of similarity scores of all pairs of items, respectively. Since $X_n$ represents
-a noisy reading of the true similarity $Y$, we model $X_n | Y$ as a Gaussian
-error around $Y$:
+a noisy reading of the true similarity $Y$, we model $X_n | Y=y$ as a Gaussian
+error around $y$:
 \begin{align}
@@
 
-Note that $\sigma_{2, n}^2$ represents how noisy the estimation of $Sim$ is when
-there are $n$ common users. This should decrease as $n$ increases.
+Note that $\sigma_{2, n}^2$ represents how noisy $Sim_n$ is as an estimation of
+$TrueSim$ and therefore should decrease as $n$ increases. The probability
+distribution of $X_n$ can be visualized in Figure 1.
 
@@
 
-In probabilistic terms, for an observed value $X_n$, we want to find the value
-of $Y$ that most likely produced $X_n$. Thus we desire the Maximum Likelihood
-Estimate of $Y | X_n$
+The problem of approximating the $TrueSim$ from $Sim_n$ can now be seen as
+finding the value of $Y$ that most likely produced $X_n$. In other words we 
+desire the Maximum Likelihood Estimate of $Y | X_n$:
 \begin{align}
-\argmax_yP(Y=y|X_n=x) &= \argmax_y\left[\frac{P(X_n=x |
-Y=y)P(Y=y)}{P(X_n=x)}\right]
+\hat{y} &= \argmax_yP(Y=y|X_n=x) 
+\\&= \argmax_y\left[\frac{P(X_n=x | Y=y)P(Y=y)}{P(X_n=x)}\right]
 \\&= \argmax_y\left[P(X_n=x|Y=y)P(Y=y)\right]
@@
 \end{align}
-Solving for y:
+This produces the following linear equation for \hat{y} in terms of x:
 \begin{align}
-y &= \frac{\sigma_{1}^2x+\sigma_{2,n}^2\mu}{\sigma_{1}^2+\sigma_{2,n}^2}
+\hat{y} &= \frac{\sigma_{1}^2x+\sigma_{2,n}^2\mu}{\sigma_{1}^2+\sigma_{2,n}^2}
 \end{align}
@@
 \begin{align}
-\left(y - \mu\right) &= \frac{\sigma_{1}^2}{\sigma_{1}^2+\sigma_{2,n}^2}
+\left(\hat{y} - \mu\right) &= \frac{\sigma_{1}^2}{\sigma_{1}^2+\sigma_{2,n}^2}
 \left(x-\mu\right)
@@
 
-The model suggests that there is a linear correlation between $X_n$, the
-observed similarity with $n$ common users, and $Y$, the true similarity.
-Moreover, the best predicted $y$ is a linear combination of the mean and the
-observed similiarity. Since $\sigma_{2,n}^2\ge0$, the slope is always $\le1$, 
-meaning that the observed distance from $\mu$ (the average similarity) 
-overestimates the true distance. This is reasonable, since there is a prior 
-that expects the true similarity to be around $\mu$.
+The model suggests that there is a linear correlation between $X_n$ ($Sim_n$)
+and $Y$ ($TrueSim$). Moreover, the best predicted $TrueSim$ is a linear 
+combination of the mean and the observed similiarity. Since $\sigma_{2,n}^2$ is 
+nonnegative, the slope is always less than or equal to one, meaning that the 
+observed distance from $\mu$ (the average true similarity) overestimates the 
+true distance. This is reasonable, since there is a prior that expects the true 
+similarity to be around $\mu$.
 
@@
 \section*{Conclusion}
+\begin{thebibliography}{9}
+
+\bibitem{Su2009}
+    Xiaoyuan Su and Taghi M. Khoshgoftaar.
+    ``A Survey of Collaborative Filtering Techniques.''
+    \emph{Advances in Artificial Intelligence.} vol. 2009,
+    Article ID 421425, 19 pages, 2009. doi:10.1155/2009/421425
+\bibitem{Breese1998}
+    Breese, John S., David Heckerman, and Carl Kadie, 
+    ``Empirical analysis of predictive algorithms for collaborative filtering.'' 
+    \emph{Proceedings of the Fourteenth conference on Uncertainty in artificial intelligence.}
+    Morgan Kaufmann Publishers Inc., 1998.
+\bibitem{Lemire1998}
+    Daniel Lemire and Anna Maclachlan.
+    ``Slope One Predictors for Online Rating-Based Collaborative Filtering.''
+    \emph{SDM.} Vol. 5, 2005.
+
+\end{thebibliography}
 

### 2014-03-25 paper/paper.tex
Add a couple of paragraphs to Model-based Technique section.

@@
 \section*{Introduction}
-Determining the similarity of two users or items is valuable for a
-number of data-mining applications including recommender systems, product
-assortment and differentiation, as well as branding. Collaborative filtering is
-often used to exploit user-item rating data to make such estimates. This is 
-typically done by computing the Pearson correlation or cosine similarity of 
-vector representations of the users or items \cite{Su2009}. Given sufficient 
-user-item rating data, this approach can be very effective. However, when the
-rating data is sparse for one or both users/items, these metrics can
-yield dramatic discrepancies compared to values computed with sufficient
-data at a later point in time. In this report, we describe two novel techniques 
+Determining the similarity of two users or items is valuable for a number of
+data-mining applications including recommender systems, product assortment and
+differentiation, as well as branding. Collaborative filtering is often used to
+exploit user-item rating data to make such estimates. This is typically done by
+computing the Pearson correlation or cosine similarity of vector representations
+of the users or items \cite{Su2009}. Given sufficient user-item rating data,
+this approach can be very effective. However, when the rating data is sparse for
+one or both users/items, these metrics can yield dramatic discrepancies compared
+to values computed with sufficient data at a later point in time. Moreover,
+these early, low-confidence measurements can be indistinguishable from later,
+high-confidence measurements.  In this report, we describe two novel techniques
 for estimating the true similarity of two users or items given limited common
-rating data.
+rating data by incorporating confidence into the measurement.
 
@@
 collaborative filtering (CF) in which we work with a logical user-item matrix.
-Each element of the matrix corresponds to a rating given by a particular user to a
-particular item. Further, we choose to limit our analysis to item-item
+Each element of the matrix corresponds to a rating given by a particular user to
+a particular item. Further, we choose to limit our analysis to item-item
 similarity but note that these techniques should be equally effective at
-improving user-user similarity estimation. In this representaion, an item is
-a vector of the ratings it has received from all users. We now formalize the 
+improving user-user similarity estimation. In this representation, an item is a
+vector of the ratings it has received from all users. We now formalize the
 notion of item-item similarity. Two alternative models for the generation of
@@
 
-.....
-
-Even worse, low confidence measurements of similarity are indistinguishable from
-high confidence ones. In this paper, we develop and compare a number of methods
-to produce more accurate estimates of similarity by incorporating confidence
-into the similarity score.
-
-Item-based collaborative filtering is a common technique for measuring the
-similarity of two items. There are different forms of collaborative filtering.
-For this paper, we use a traditional approach. Each item is represented as a
-vector of ratings. The similarity score of two items is computed by measuring
-the similarity of the two rating vectors.
-
-There are a few ways to compute the similarity of two vectors. One approach is
-to use Cosine-similarity, defined as the cosine of the angle between the two
-vectors:
-\begin{align}
-CosSim(A, B) = \frac{\sum\limits_{u\in U_{AB}}
-r_{u,A}r_{u,B}}{\sqrt{\sum\limits_{u\in U_{A}} r_{u,A}^2}
-\sqrt{\sum\limits_{u\in U_{B}} r_{u,B}^2}}
-\end{align}
-where $U_{I}$ is the set of all users who rated item $I$, $U_{AB}$ is the set of
-all users who rated both item $A$ and item $B$ and $r_{u,I}$ is the rating user
-$u$ gave to item $I$. Another popular approach is to use Pearson correlation:
-\begin{align}
-PearsSim(A, B) = \frac{\sum\limits_{u\in U_{AB}}
-r_{u,A}r_{u,B}}{\sqrt{\sum\limits_{u\in U_{AB}} r_{u,A}^2}
-\sqrt{\sum\limits_{u\in U_{AB}} r_{u,B}^2}}
-\end{align}
-
-Other similarity functions exist, such as one-sided similarities, but Pearson
-correlation and Cosine-similarity are the most popular and will be the two
-measurements used in this paper. \footnotemark
-
-\footnotetext{ Note that the difference between the two similarities is how they
-handle unpaired ratings; that is ratings from a user who has not rated the other
-item. Cosine-similarity considers the unknown rating from an unpaired rating to
-be $0$ and then calculates the cosine of the two vectors. Pearson correlation
-simply throws away all unpaired ratings and calculates the cosine of the two
-modified vectors.}
-
-Both similarity measurements primarily leverage information from common users,
-that is users who have rated both items. Because of this, they perform well with
-a large number common users, but give unreliable results when the items have few
-common users. In an extreme case, if only one user has rated item $A$ and item
-$B$, $PearsSim(A, B) = 1$ or $-1$. Not only is this unlikely to be an accurate
-assessment of the true similarity of $A$ and $B$, it also gives the most extreme
-results possible, without giving any indication that this is a low-confidence
-calculation.
-
-In this paper, we construct a more accurate similarity score by incorporating
-the number of common users, $n$, into the similarity function.
+Both similarity estimates primarily leverage information from common users, that
+is users who have rated both items. Because of this, they perform well with a
+large number common users, but give unreliable, and dramatically varying,
+results when the items have few common users. In an extreme case, if only one
+user has rated item $A$ and item $B$, $PearsSim(A, B) = 1$ or $-1$. Not only is
+this unlikely to be an accurate assessment of the true similarity of $A$ and
+$B$, it also gives the most extreme results possible, without giving any
+indication that this is a low-confidence calculation.
 
@@
 
+\section*{A Model-based Technique}
+
+We now construct a relatively simple probabilistic model that is readily
+exploited for making estimates of $TrueSimRand$ and, to a lesser extent,
+$TrueSimPref$. In addition, this approach has the added benefit of allowing us
+to infer implied predictions of $TrueSimRand$ made by individual users each time
+they rate two items. This latter feature allows us to define a new metric for
+users that we call {\em user-predictivity}, which allows us to ascribe varying
+levels of confidence to their contributions to a similarity estimate.
+
+Let $Y$ be a random variable representing the $TrueSimRand$ of a randomly chosen
+pair of items. As before, we model $Y$ as a Normal distribution
+\begin{align}
+Y \sim N(\mu_s, \sigma_s^2),
+\end{align}
+where $\mu_s$ and $\sigma_s$ are the average and standard deviation of all
+similarity scores, respectively. Now, let $\vec{R} = \left(R_A~~R_B\right)$ be
+a random two element vector representing the ratings given to items $A$ and $B$.
+We model the conditional distribution of $\vec{R} | Y$ as a multivariate Normal
+distribution
+\begin{align}
+\vec{R} | Y=y \sim N(\vec{\mu_r}, \mathbf{\Sigma_y}),
+\end{align}
+where the elements of $\vec{\mu_r}$ are the average ratings and
+$\mathbf{\Sigma_y}$ is the $y$-dependent $2\times2$ covariance matrix.
+
 \section*{Experiment}

### 2014-03-25 paper/paper.tex
Fix latex error (\hat only allowed in math mode).
Change section titles for Model-based Approach.
Develop Probabilistic Model subsection of Model-based Approach.

@@
 \end{align}
-This produces the following linear equation for \hat{y} in terms of x:
+This produces the following linear equation for $\hat{y}$ in terms of $x$:
 \begin{align}
@@
 
-\section*{A Model-based Technique}
+\section*{Model-based Approach}
+\subsection*{Probabilistic Model}
 
@@
 \end{align}
-where $\mu_s$ and $\sigma_s$ are the average and standard deviation of all
-similarity scores, respectively. Now, let $\vec{R} = \left(R_A~~R_B\right)$ be
-a random two element vector representing the ratings given to items $A$ and $B$.
-We model the conditional distribution of $\vec{R} | Y$ as a multivariate Normal
+where $\mu_s$ and $\sigma_s^2$ are the average and variance of all similarity
+scores, respectively. Now, let $\vec{R} = \left(R_A~~R_B\right)$ be a random two
+element vector representing the ratings given to items $A$ and $B$.  We model
+the conditional distribution of $\vec{R} | Y$ as a multivariate Normal
 distribution
@@
 where the elements of $\vec{\mu_r}$ are the average ratings and
-$\mathbf{\Sigma_y}$ is the $y$-dependent $2\times2$ covariance matrix.
+$\mathbf{\Sigma_y}$ is the $y$-dependent $2\times2$ covariance matrix. Now,
+being that $Y=y$ is the Pearson correlation of the ratings $R_A | y$ and $R_B |
+y$ generated by this distribution, if we assume the variance of individual
+ratings $\sigma_r^2$ is independent of $y$, we have that
+\begin{align}
+y = \frac{\mathrm{COV}(R_A | y, R_B | y)}{\sigma_r^2},
+\end{align}
+where $\mathrm{COV}(R_A | y, R_B | y)$ is the covariance of the ratings $R_A |
+y$ and $R_B | y$.  It therefore follows that if the ratings are indeed normally
+distributed that the covariance matrix is related to the similarity $y$ by
+\begin{align}
+\mathbf{\Sigma_y} = \sigma_r^2 \cdot
+\left[ \begin{array}{cc}  1 & y \\ y & 1 \end{array} \right]
+\end{align}
 

### 2014-03-25 amazon/similarity.py
Add missing term to modelSim equation.

@@ -16,8 +16,8 @@ from SimilarityGrid import SimilarityGrid
 
 # params
 #constSimScore = 0.16   # topEdges_40_randSim
-#constSimScore = 0.27    # topEdges_80_randSim
-constSimScore = 0.050    # topEdges_80_prefSim
+constSimScore = 0.27    # topEdges_80_randSim
+#constSimScore = 0.050    # topEdges_80_prefSim
 K = 20
 sigma = 1.0
 writer = csv.writer(sys.stdout)

### 2014-03-25 paper/paper.tex
pulling
Extend model-based section.
Develop model-based section.
Changed names of sections. Edited probabilstic model section of ratings model

--- 2014-03-25 paper/paper.tex
@@
 \newcommand{\ra}{\rightarrow}
+\newcommand{\Cov}{\mathrm{Cov}}
 \DeclareMathOperator*{\argmax}{arg\!\max}
@@
 
-\section*{True Item-Item Similarity}
+\section*{Similarity}
+\subsection*{True Item-Item Similarity}
 
@@
 
-\section*{Estimating Item-Item Similarity}
+\subsection*{Estimating Item-Item Similarity}
 
@@
 
-\section*{Linear Approach}
+\section*{Score-based Approach}
 \subsection*{Probabilistic Model}
@@
 \begin{align}
-Y \sim N(\mu, \sigma_{1}^2)
+Y \sim N(\mu, \sigma_{s}^2)
 \end{align}
-where $\mu$ and $\sigma_{1}^2$ are the average similarity score and the variance 
+where $\mu$ and $\sigma_{s}^2$ are the average similarity score and the variance 
 of similarity scores of all pairs of items, respectively. Since $X_n$ represents
@@
 \begin{align}
-(X_n | Y=y) \sim N(y, \sigma_{2, n}^2)
+(X_n | Y=y) \sim N(y, \sigma_{err, n}^2)
 \end{align}
 
-Note that $\sigma_{2, n}^2$ represents how noisy $Sim_n$ is as an estimation of
+Note that $\sigma_{err, n}^2$ represents how noisy $Sim_n$ is as an estimation of
 $TrueSim$ and therefore should decrease as $n$ increases. The probability
@@
 \\&= 
-\argmax_y\left[\frac{1}{\sigma_{2,n}\sqrt{2\pi}}\exp{\left(\frac{-(x-y)^2}
-{2\sigma_{2,n}^2}\right)}
-\frac{1}{\sigma_{1}\sqrt{2\pi}}\exp{\left(\frac{-(y-\mu)^2}
-{2\sigma_{1}^2}\right)}\right]
-\\&= \argmin_y\left[\frac{(x-y)^2}{2\sigma_{2,n}^2} +
-\frac{(y-\mu)^2}{2\sigma_{1}^2}\right]
-\\&= \argmin_y\left[\left(\sigma_{1}^2+\sigma_{2,n}^2\right)y^2 - 
-2\left(\sigma_{1}^2x+\sigma_{2,n}^2\mu\right)y\right]
+\argmax_y\left[\frac{1}{\sigma_{err,n}\sqrt{2\pi}}\exp{\left(\frac{-(x-y)^2}
+{2\sigma_{err,n}^2}\right)}
+\frac{1}{\sigma_{s}\sqrt{2\pi}}\exp{\left(\frac{-(y-\mu)^2}
+{2\sigma_{s}^2}\right)}\right]
+\\&= \argmin_y\left[\frac{(x-y)^2}{2\sigma_{err,n}^2} +
+\frac{(y-\mu)^2}{2\sigma_{s}^2}\right]
+\\&= \argmin_y\left[\left(\sigma_{s}^2+\sigma_{err,n}^2\right)y^2 - 
+2\left(\sigma_{s}^2x+\sigma_{err,n}^2\mu\right)y\right]
 \end{align}
@@
 \begin{align}
-\frac{d}{dy}\left[\left(\sigma_{1}^2+\sigma_{2,n}^2\right)y^2 - 
-2\left(\sigma_{1}^2x+\sigma_{2,n}^2\mu\right)y\right]
-&= 2\left(\sigma_{1}^2+\sigma_{2,n}^2\right)y - 
-2\left(\sigma_{1}^2x+\sigma_{2,n}^2\mu\right) 
+\frac{d}{dy}\left[\left(\sigma_{s}^2+\sigma_{err,n}^2\right)y^2 - 
+2\left(\sigma_{s}^2x+\sigma_{err,n}^2\mu\right)y\right]
+&= 2\left(\sigma_{s}^2+\sigma_{err,n}^2\right)y - 
+2\left(\sigma_{s}^2x+\sigma_{err,n}^2\mu\right) 
 \\&= 0
 \end{align}
-This produces the following linear equation for \hat{y} in terms of x:
+This produces the following linear equation for $\hat{y}$ in terms of $x$:
 \begin{align}
-\hat{y} &= \frac{\sigma_{1}^2x+\sigma_{2,n}^2\mu}{\sigma_{1}^2+\sigma_{2,n}^2}
+\hat{y} &= \frac{\sigma_{s}^2x+\sigma_{err,n}^2\mu}{\sigma_{s}^2+\sigma_{err,n}^2}
 \end{align}
@@
 \begin{align}
-\left(\hat{y} - \mu\right) &= \frac{\sigma_{1}^2}{\sigma_{1}^2+\sigma_{2,n}^2}
+\left(\hat{y} - \mu\right) &= \frac{\sigma_{s}^2}{\sigma_{s}^2+\sigma_{err,n}^2}
 \left(x-\mu\right)
@@
 and $Y$ ($TrueSim$). Moreover, the best predicted $TrueSim$ is a linear 
-combination of the mean and the observed similiarity. Since $\sigma_{2,n}^2$ is 
+combination of the mean and the observed similiarity. Since $\sigma_{err,n}^2$ is 
 nonnegative, the slope is always less than or equal to one, meaning that the 
@@
 
-The parameters $\mu$, $\sigma_{1}^2$, and $\sigma_{2,n}^2$ can be approximated 
+The parameters $\mu$, $\sigma_{s}^2$, and $\sigma_{err,n}^2$ can be approximated 
 using the method of moments over training data where there are enough common
 users that $TrueSim$ can be approximated. This method requires approximating
-$\sigma_{2,n}^2$ for every value of $n$ separately. To gain a more general 
-model, we can model $\sigma_{2,n}^2$ as a function of $n$. Recall that 
-$\sigma_{2,n}^2$ is the noisiness of the measurement $Sim$ about $TrueSim$ and $TrueSim = lim_{n \to
+$\sigma_{err,n}^2$ for every value of $n$ separately. To gain a more general 
+model, we can model $\sigma_{err,n}^2$ as a function of $n$. Recall that 
+$\sigma_{err,n}^2$ is the noisiness of the measurement $Sim$ about $TrueSim$ and $TrueSim = lim_{n \to
 \infty}Sim$. Thus $Sim$ can be thought of as a sampling $n$ users from the
@@
 the Central Limit Theorem motivates the intuition that the variance of $Sim$
-would decrease as the $sqrt(n)$. Thus we model $\sigma_{2,n}^2$ as:
+would decrease as the $sqrt(n)$. Thus we model $\sigma_{err,n}^2$ as:
 \begin{align}
-\sigma_{2,n}^2 = \frac{\alpha}{\sqrt{n}},
+\sigma_{err,n}^2 = \frac{\alpha}{\sqrt{n}},
 \end{align}
@@
 \begin{align}
-\left(Y - \mu\right) = \frac{\sigma_{1}^2}{\sigma_{1}^2+\frac{\alpha}{\sqrt{n}}}
+\left(Y - \mu\right) = \frac{\sigma_{s}^2}{\sigma_{s}^2+\frac{\alpha}{\sqrt{n}}}
 \left(X_n-\mu\right)
@@
 \begin{align}
-ModSim(Sim, n) = \frac{\sigma_{1}^2}{\sigma_{1}^2+\frac{\alpha}{\sqrt{n}}}
+ModSim(Sim, n) = \frac{\sigma_{s}^2}{\sigma_{s}^2+\frac{\alpha}{\sqrt{n}}}
 \left(Sim-\mu\right) + \mu
@@
 
-\section*{Technique}
+\subsection*{Technique}
 Based on this maodel, we try three different techniques to predict the true similarity from a
@@
 of linear regressions between $Sim$ and $TrueSim$ for each $n$. This ignores the 
-data-specific parameters $\mu$, $\sigma_{1}$ and $\sigma_{2,n}$ to find the best 
+data-specific parameters $\mu$, $\sigma_{s}$ and $\sigma_{err,n}$ to find the best 
 linear fit. It also treats each $n$ completely independently. This technique 
@@
 
-The second technique is more general. It approximates $\mu$ and $\sigma_{1}$
+The second technique is more general. It approximates $\mu$ and $\sigma_{s}$
 using the method of moments on the training data of all $TrueSim$ scores. Then,
-$\sigma_{2,n}$ is also approximated using a method of moments on the training
-data of all $Sim_n$ for each $n$ separately. Here $\sigma_{2,n}$'s are also
+$\sigma_{err,n}$ is also approximated using a method of moments on the training
+data of all $Sim_n$ for each $n$ separately. Here $\sigma_{err,n}$'s are also
 considered independently of each other.
 
-The third and most general technique models $\sigma_{2,n}$ as a function of
-$n$ using equation 19. As in the previous technique, $\mu$, $\sigma_{1}$ and each
-$\sigma_{2,n}$ are approximated using the method of moments over the training
+The third and most general technique models $\sigma_{err,n}$ as a function of
+$n$ using equation 19. As in the previous technique, $\mu$, $\sigma_{s}$ and each
+$\sigma_{err,n}$ are approximated using the method of moments over the training
 data. Then $\alpha$ is calculated that minimizes the sum of squared errors
-between $\sigma_{2,n}^2$ and $\frac{\alpha}{\sqrt{n}}$. This $\alpha$ is used to
+between $\sigma_{err,n}^2$ and $\frac{\alpha}{\sqrt{n}}$. This $\alpha$ is used to
 construct the three-parameter prediction function of equation 19.
 
-\section*{A Model-based Technique}
+\section*{Ratings-based Approach}
+\subsection*{Probabilistic Model}
 
-We now construct a relatively simple probabilistic model that is readily
-exploited for making estimates of $TrueSimRand$ and, to a lesser extent,
-$TrueSimPref$. In addition, this approach has the added benefit of allowing us
-to infer implied predictions of $TrueSimRand$ made by individual users each time
-they rate two items. This latter feature allows us to define a new metric for
-users that we call {\em user-predictivity}, which allows us to ascribe varying
-levels of confidence to their contributions to a similarity estimate.
+We now construct a different probabilistic model based on the pairs of ratings
+themselves. This model is readily exploited for making estimates of 
+$TrueSimRand$ and, to a lesser extent, $TrueSimPref$. In addition, this approach 
+has the added benefit of allowing us to infer implied predictions of 
+$TrueSimRand$ made by individual users each time they rate two items. This 
+latter feature allows us to define a new metric for users that we call {\em 
+user-predictivity}, which allows us to ascribe varying levels of confidence to 
+their contributions to a similarity estimate.
 
@@
 \end{align}
-where $\mu_s$ and $\sigma_s$ are the average and standard deviation of all
-similarity scores, respectively. Now, let $\vec{R} = \left(R_A~~R_B\right)$ be
-a random two element vector representing the ratings given to items $A$ and $B$.
-We model the conditional distribution of $\vec{R} | Y$ as a multivariate Normal
+where $\mu_s$ and $\sigma_s^2$ are the average and variance of all true similarity
+scores, respectively. Now, let $\vec{R} = \left(R_A~~R_B\right)$ be a random two
+element vector representing the ratings given to items $A$ and $B$.  We model
+the conditional distribution of $\vec{R} | Y$ as a multivariate Normal
 distribution
 \begin{align}
-\vec{R} | Y=y \sim N(\vec{\mu_r}, \mathbf{\Sigma_y}),
+(\vec{R} | Y=y) \sim N(\vec{\mu_r}, \mathbf{\Sigma_y}),
 \end{align}
@@
 $\mathbf{\Sigma_y}$ is the $y$-dependent $2\times2$ covariance matrix. 
+Since $y$ is the $TrueSimRand$ score, it can be thought of as the Pearson 
+correlation over the distribution of ratings $R_A | Y$ and $R_B | Y$. Assuming
+that the variance of individual ratings $\sigma_r^2$ is independent of $y$, we have 
+\begin{align}
+y = \frac{\Cov{\left[R_A | Y=y , R_B | Y=y\right]}}{\sigma_r^2},
+\end{align}
+It therefore follows that if the ratings are indeed normally distributed 
+then the covariance matrix is related to the true similarity $y$ by
+\begin{align}
+\mathbf{\Sigma_y} = \sigma_r^2 \cdot
+\left[ \begin{array}{cc}  1 & y \\ y & 1 \end{array} \right]
+\end{align}
 

### 2014-03-25 paper/paper.tex
Develop model-based section further.

@@
 \mathbf{\Sigma_y} = \sigma_r^2 \cdot
-\left[ \begin{array}{cc}  1 & y \\ y & 1 \end{array} \right]
+\left[ \begin{array}{cc}  1 & y \\ y & 1 \end{array} \right].
 \end{align}
+
+Using this model, we may view the problem of estimating the similarity $Y$ of
+items $A$ and $B$ that most likely produced ratings $R_A$ and $_B$ as
+that of finding the Maximum Likelihood Estimate of $Y | \vec{R}$:
+\begin{align}
+P(Y | \vec{R}) = \frac{P(\vec{R} | Y)P(Y)}{P(\vec{R})}.
+\end{align}
+
+\begin{align}
+\hat{y} &= \argmax_yP(Y=y|\vec{R}=\vec{r}) 
+\\&= \argmax_y\left[\frac{P(\vec{R}=\vec{r} |
+Y=y)P(Y=y)}{P(\vec{R}=\vec{r})}\right]
+\\&= \argmax_y\left[P(\vec{R}=\vec{r}|Y=y)P(Y=y)\right]
+\\&=
+\argmax_y\left[\frac{1}{2\pi\sqrt{\left|\mathbf{\Sigma_y}\right|}}\exp{\left(-\frac{1}{2}
+(\vec{r} - \vec{\mu_r})^\mathrm{T}\mathbf{\Sigma_y}^{-1}(\vec{r} - \vec{\mu_r})\right)}
+\frac{1}{\sigma_{s}\sqrt{2\pi}}\exp{\left(\frac{-(y-\mu_s)^2}
+{2\sigma_{s}^2}\right)}\right]
+\\&= \argmin_y\left[(\vec{r} -
+\vec{\mu_r})^\mathrm{T}\mathbf{\Sigma_y}^{-1}(\vec{r} - \vec{\mu_r}) + \frac{(y -
+\mu_s)^2}{\sigma_{s}^2} + \ln{\left|\mathbf{\Sigma_y}\right|} \right]
+\\&=
+\argmin_y\left[\vec{r^\prime}^\mathrm{T}\mathbf{\Sigma_y}^{-1}\vec{r^\prime}
+ + \frac{(y - \mu_s)^2}{\sigma_{s}^2} + \ln{\left|\mathbf{\Sigma_y}\right|} \right]
+\\&= \argmin_y\left[\frac{1}{\sigma_r^2}\frac{1}{1 -
+y^2}\left[r_A^\prime(r_A^\prime - yr_B^\prime) + r_B^\prime(r_B^\prime -
+yr_A^\prime)\right] + \frac{(y - \mu_s)^2}{\sigma_{s}^2} +
+\ln{\left[\sigma_r^2(1 - y^2)\right]} \right],
+\end{align}
+where we used the fact that
+\begin{align}
+\mathbf{\Sigma_y}^{-1} = \frac{1}{\sigma_r^2}
+\left[ \begin{array}{cc}  1 & y \\ y & 1 \end{array} \right]^{-1} =
+\frac{1}{\sigma_r^2}\frac{1}{1 - y^2}\left[ \begin{array}{cc}  1 & -y \\ -y & 1
+\end{array} \right].
+\end{align}
+and that
+\begin{align}
+\left|\mathbf{\Sigma}\right| = \sigma_r^2(1 - y^2)
+\end{align}
+along with the substitution $\vec{r^\prime} = (r_A^\prime~~r_B^\prime) = \vec{r}
+- \vec{\mu_r}$. Next, we find the minimum by taking the derivative with respect
+to $y$ and setting it to zero:
+\begin{align}
+\frac{d}{dy}\left[ \frac{1}{\sigma_r^2}\frac{1}{1 -
+y^2}\left[r_A^\prime(r_A^\prime - yr_B^\prime) + r_B^\prime(r_B^\prime -
+yr_A^\prime)\right] + \frac{(y - \mu_s)^2}{\sigma_{s}^2} +
+\ln{\left[\sigma_r^2(1 - y^2)\right]} \right] &= 0 \\
+-r_A^\prime r_B^\prime y^2 + ({r_A^\prime}^2 + {r_B^\prime}^2)y - r_A^\prime
+r_B^\prime + \frac{\sigma_r^2}{\sigma_s^2}(1 - y^2)^2(y - \mu_s) - \sigma_r^2(1
+- y^2)y &= 0.
+\end{align}
+This final expression implicitly yields the MLE of $y$ as a function of the
+ratings $r_A$ and $r_B$. That is, constraining the parameters $\mu_r$,
+$\sigma_r$, $\mu_s$, and $\sigma_s$ to realistic values, this expression defines
+a surface that can be to efficiently estimate $y$ given a pair of ratings.
 

### 2014-03-25 paper/paper.tex
Fix typo and add comment.

@@
 \end{align}
-where $\mu_s$ and $\sigma_s^2$ are the average and variance of all true similarity
-scores, respectively. Now, let $\vec{R} = \left(R_A~~R_B\right)$ be a random two
-element vector representing the ratings given to items $A$ and $B$.  We model
-the conditional distribution of $\vec{R} | Y$ as a multivariate Normal
+where $\mu_s$ and $\sigma_s^2$ are the average and variance of all true
+similarity scores, respectively. Now, let $\vec{R} = \left(R_A~~R_B\right)$ be a
+random two element vector representing the ratings given to items $A$ and $B$.
+We model the conditional distribution of $\vec{R} | Y$ as a multivariate Normal
 distribution
@@
 where the elements of $\vec{\mu_r}$ are the average ratings and
-$\mathbf{\Sigma_y}$ is the $y$-dependent $2\times2$ covariance matrix. 
-Since $y$ is the $TrueSimRand$ score, it can be thought of as the Pearson 
-correlation over the distribution of ratings $R_A | Y$ and $R_B | Y$. Assuming
-that the variance of individual ratings $\sigma_r^2$ is independent of $y$, we have 
+$\mathbf{\Sigma_y}$ is the $y$-dependent $2\times2$ covariance matrix.  Since
+$y$ is the $TrueSimRand$ score, it can be thought of as the Pearson correlation
+over the distribution of ratings $R_A | Y$ and $R_B | Y$. Assuming that the
+variance of individual ratings $\sigma_r^2$ is independent of $y$, we have 
 \begin{align}
@@
 Using this model, we may view the problem of estimating the similarity $Y$ of
-items $A$ and $B$ that most likely produced ratings $R_A$ and $_B$ as
+items $A$ and $B$ that most likely produced ratings $R_A$ and $R_B$ as
 that of finding the Maximum Likelihood Estimate of $Y | \vec{R}$:
@@
 ratings $r_A$ and $r_B$. That is, constraining the parameters $\mu_r$,
-$\sigma_r$, $\mu_s$, and $\sigma_s$ to realistic values, this expression defines
-a surface that can be to efficiently estimate $y$ given a pair of ratings.
+$\sigma_r$, $\mu_s$, and $\sigma_s$ to realistic values (e.g. $\sigma_r,
+\sigma_s > 0$), this expression defines a surface that can be to efficiently
+estimate $y$ given a pair of ratings.
 

### 2014-03-25 paper/paper.tex
Fix typo.
Begin User Predictivity subsection.

@@
 it indicates a lack of interest of a particular user for a particular item. This
-is captured by the model through a default rating of zero \footnote{We address
+is captured by the model through a default rating of zero\footnote{We address
 the issue of rating bias and the precise meaning of a zero valued ratings later
@@
 convention of Breese et al. \cite{Breese1998}, we choose to use the Pearson
-correlation \footnote{The Pearson correlation is traditionally characterized by
+correlation\footnote{The Pearson correlation is traditionally characterized by
 subtracting an expectation value from each dimension. This is only approximately
@@
 $\sigma_r$, $\mu_s$, and $\sigma_s$ to realistic values (e.g. $\sigma_r,
-\sigma_s > 0$), this expression defines a surface that can be to efficiently
-estimate $y$ given a pair of ratings.
+\sigma_s > 0$), this expression defines a surface that can be used to
+efficiently estimate $y$ given a pair of ratings\footnote{Note that this
+equation neglects the constraint that $-1 \leq y \leq 1$. As a consequence, some
+combinations of ratings and parameters can yield MLEs for $y$ that exceed $1$ in
+magnitude. This is easily fixed, however, by applying this additional
+constraint.}.
+
+\subsection*{User Predictivity}
+
+As already mentioned, the estimator $PearSim$ is unable to say much about the
+similarity of two items when only provided ratings from a single common user. In
+this case, $PearSim$ yields $\pm1$ (or the indeterminate value $0/0$ in the case
+that both ratings are zero) depending on whether the ratings have the same sign
+or not. The MLE provided by the above model rectifies this by generating
+similarity estimates in the range $[-1, +1]$ given a single pair of ratings. We
+therefore consider each user who rates a pair of items as implicitly predicting
+the similarity of the items. A key discovery of the present work is that the
+accuracy of users' predictions in a training set is correlated with the accuracy
+in a test set. We exploit this finding by ascribing a quantity we call {\em user
+predictivity} to each user defined as
+\begin{align}
+p_u = e^{-\epsilon_u/\lambda},
+\end{align}
+where $\lambda$ is a parameter to be learned and
+\begin{align}
+\epsilon_u = \frac{1}{\left|I_u\right|}
+\sum_{A,B \in I_u}{\left|ModelSimRand(A, B) - GoldSimRand(A, B)\right|}
+\end{align}
+with $I_u$ the set of item pairs rated by user $u$ such that $\epsilon_u$ is
+the average absolute error of user $u$'s similarity predictions.
 

### 2014-03-25 paper/paper.tex
working on technique section

--- 2014-03-25 paper/paper.tex
@@
 
-\section*{Problem}
-
-The goal is to calculate a modified similiary function that leverages the number
-of commonn users to better estimate the true similarity. Formally, we design
-$ModSim$ such that for two items $A$ and $B$ with $n$ users in common
-
-\begin{align}
-ModSim(A, B, n) \approx TrueSim(A, B)
-\end{align}
-where $TrueSim$ is the true similarity of the two items. Of course, there is no
-way to actually know the true similarity of two items, but for a good enough
-similarity function, such as $PearsSim$ or $CosSim$ and a large enough $n$, we
-can get a good estimate. Thus, we model $TrueSim$ as 
-\begin{align}
-TrueSim(A, B) = lim_{n\to\infty}Sim(A, B)
-\end{align}
-where $Sim$ is the most appropriate generic similarity function for the dataset.
-
-Since $TrueSim$ dependends on the choice of $Sim$, we abstract away the details
-of $Sim$ and use its output directly. In all, the goal is to construct $ModSim$
-such that
-
 \begin{align}
@@
 
-In our first approach, we model true and observed similarity probabilistically. For
-this analysis, we can consider an abstact $TrueSim$ which ignores the details of 
+In our first approach, we model true and observed similarity probabilistically. The
+goal is to use the observed similarity score directly to approximate the true similarity.
+In this way, we can consider an abstact $TrueSim$ which ignores the details of 
 $TrueSimRand$ or $TrueSimPref$. Similarly, we consider $Sim_{n}$ and $GoldSim$ to be
@@
 combination of the mean and the observed similiarity. Since $\sigma_{err,n}^2$ is 
-nonnegative, the slope is always less than or equal to one, meaning that the 
+nonnegative, the slope is always less than or equal to one, meaningthat the 
 observed distance from $\mu$ (the average true similarity) overestimates the 
@@
 
-The parameters $\mu$, $\sigma_{s}^2$, and $\sigma_{err,n}^2$ can be approximated 
-using the method of moments over training data where there are enough common
-users that $TrueSim$ can be approximated. This method requires approximating
-$\sigma_{err,n}^2$ for every value of $n$ separately. To gain a more general 
-model, we can model $\sigma_{err,n}^2$ as a function of $n$. Recall that 
-$\sigma_{err,n}^2$ is the noisiness of the measurement $Sim$ about $TrueSim$ and $TrueSim = lim_{n \to
-\infty}Sim$. Thus $Sim$ can be thought of as a sampling $n$ users from the
-infinite set used to calculate $TrueSim$. Although $Sim$ may not be linear, we
-the Central Limit Theorem motivates the intuition that the variance of $Sim$
-would decrease as the $sqrt(n)$. Thus we model $\sigma_{err,n}^2$ as:
+This model requires calculating $\sigma_{err,n}^2$ for every value of $n$ 
+separately. To gain a more general model, which relates the $\sigma_{err, n}$'s, 
+we model $\sigma_{err,n}^2$ as a function of $n$. Recall that $\sigma_{err,n}^2$ 
+is the noisiness of the observed $Sim_n$ about $TrueSim$. Since the $TrueSim$ is
+the $Sim$ when all users that will rate the items have rated them, $Sim_n$ can be
+thought of as sampling $n$ users from the set used to calculate $TrueSim$. 
+Although $Sim$ is likely more complicated than an average, the Central Limit 
+Theorem motivates the intuition that the variance of $Sim$ would decrease as the 
+$\sqrt{n}$. Thus we model $\sigma_{err,n}^2$ as:
 \begin{align}
@@
 \end{align}
-In terms of a modified similarity function, as desired earlier, we now have:
-\begin{align}
-ModSim(Sim, n) = \frac{\sigma_{s}^2}{\sigma_{s}^2+\frac{\alpha}{\sqrt{n}}}
-\left(Sim-\mu\right) + \mu
-\end{align}
 
 \subsection*{Technique}
+The parameters $\mu$, $\sigma_{s}^2$, and $\sigma_{err,n}^2$ can be approximated 
+using the method of moments over training data. $\mu$ and $\sigma_{s}$ are
+calculated
+
 Based on this maodel, we try three different techniques to predict the true similarity from a
@@
 \begin{align}
-P(Y | \vec{R}) = \frac{P(\vec{R} | Y)P(Y)}{P(\vec{R})}.
-\end{align}
-
-\begin{align}
 \hat{y} &= \argmax_yP(Y=y|\vec{R}=\vec{r}) 
@@
 This final expression implicitly yields the MLE of $y$ as a function of the
-ratings $r_A$ and $r_B$. That is, constraining the parameters $\mu_r$,
-$\sigma_r$, $\mu_s$, and $\sigma_s$ to realistic values, this expression defines
-a surface that can be to efficiently estimate $y$ given a pair of ratings.
+ratings $r_A$ and $r_B$. That is, by constraining the parameters $\mu_r$,
+$\sigma_r$, $\mu_s$, and $\sigma_s$ to realistic values, the equation
+can be used to efficiently estimate $y$ given a pair of ratings.
 

### 2014-03-25 paper/paper.tex

@@
 ratings $r_A$ and $r_B$. That is, by constraining the parameters $\mu_r$,
-$\sigma_r$, $\mu_s$, and $\sigma_s$ to realistic values, the equation
-can be used to efficiently estimate $y$ given a pair of ratings.
+$\sigma_r$, $\mu_s$, and $\sigma_s$ to realistic values (e.g. $\sigma_r,
+\sigma_s > 0$), this expression defines a surface that can be used to
+efficiently estimate $y$ given a pair of ratings\footnote{Note that this
+equation neglects the constraint that $-1 \leq y \leq 1$. As a consequence, some
+combinations of ratings and parameters can yield MLEs for $y$ that exceed $1$ in
+magnitude. This is easily fixed, however, by applying this additional
+constraint.}.
+
+\subsection*{User Predictivity}
+
+As already mentioned, the estimator $PearSim$ is unable to say much about the
+similarity of two items when only provided ratings from a single common user. In
+this case, $PearSim$ yields $\pm1$ (or the indeterminate value $0/0$ in the case
+that both ratings are zero) depending on whether the ratings have the same sign
+or not. The MLE provided by the above model rectifies this by generating
+similarity estimates in the range $[-1, +1]$ given a single pair of ratings. We
+therefore consider each user who rates a pair of items as implicitly predicting
+the similarity of the items. A key discovery of the present work is that the
+accuracy of users' predictions in a training set is correlated with the accuracy
+in a test set. We exploit this finding by ascribing a quantity we call {\em user
+predictivity} to each user defined as
+\begin{align}
+p_u = e^{-\epsilon_u/\lambda},
+\end{align}
+where $\lambda$ is a parameter to be learned and
+\begin{align}
+\epsilon_u = \frac{1}{\left|I_u\right|}
+\sum_{A,B \in I_u}{\left|ModelSimRand(A, B) - GoldSimRand(A, B)\right|}
+\end{align}
+with $I_u$ the set of item pairs rated by user $u$ such that $\epsilon_u$ is
+the average absolute error of user $u$'s similarity predictions.
 
@@
 \section*{Conclusion}
+\section*{Future Work}
+
 \begin{thebibliography}{9}

### 2014-03-25 paper/paper.tex
Finished linear techniques section. Added graph. Going to bed

@@
 
-\begin{align}
-ModSim(Sim(A, B), n) \approx lim_{n\to\infty}Sim(A, B)
-\end{align}
+In the following sections, we present two approaches to better approximate the
+true similarity given a limited number of common reviewers. Both approaches are
+motivated by probabilistic models with distributions grounded in the data. The
+first approach leverages the observed similarity score and the number of common
+users to calculate the true similarity. The second approach leverages the
+rating-vectors themselves.
 
@@
 
-In our first approach, we model true and observed similarity probabilistically. The
-goal is to use the observed similarity score directly to approximate the true similarity.
-In this way, we can consider an abstact $TrueSim$ which ignores the details of 
-$TrueSimRand$ or $TrueSimPref$. Similarly, we consider $Sim_{n}$ and $GoldSim$ to be
-the generic observed similarity scores. Let $Y$ be a random variable representing 
-the $TrueSim$ of a randomly chosen pair of items. Let $X_{n}$ be a random variable 
-representing the $Sim_{n}$ of a pair of items when $n$ common users are observed.
+In our first approach, we model the probability distributions of true and 
+observed similarity. The goal is to use the observed similarity score directly 
+to approximate the true similarity. In this way, we can consider an abstact 
+$TrueSim$ which ignores the details of $TrueSimRand$ or $TrueSimPref$. 
+Similarly, we consider $Sim_n$ to be a generic observed similarity score when
+the two itmes have $n$ common reviewers. Now, let $Y$ be a random variable 
+representing the $TrueSim$ of a randomly chosen pair of items. Let $X_{n}$ be a
+random variable representing the $Sim_{n}$ of a pair of items when $n$ common 
+users are observed.
 
@@
 and $Y$ ($TrueSim$). Moreover, the best predicted $TrueSim$ is a linear 
-combination of the mean and the observed similiarity. Since $\sigma_{err,n}^2$ is 
-nonnegative, the slope is always less than or equal to one, meaningthat the 
+combination of the average true similarity and the observed similiarity. Since $\sigma_{err,n}^2$ is 
+nonnegative, the slope is always less than or equal to one, meaning that the 
 observed distance from $\mu$ (the average true similarity) overestimates the 
-true distance. This is reasonable, since there is a prior that expects the true 
-similarity to be around $\mu$.
+true distance. This is reasonable, since prior suggests that the true 
+similarity is likely to be near $\mu$.\footnote{It is worth noting that despite
+the resemblance to Slope One Predictors \cite{Lemire1998}, this is a fundamentally 
+different technique. Here, we have a linear relationship between observed 
+scores and true scores, whereas the Slope One model seeks a linear relationship 
+between the ratings a user gives to one items versus another.}
 
-This model requires calculating $\sigma_{err,n}^2$ for every value of $n$ 
-separately. To gain a more general model, which relates the $\sigma_{err, n}$'s, 
-we model $\sigma_{err,n}^2$ as a function of $n$. Recall that $\sigma_{err,n}^2$ 
-is the noisiness of the observed $Sim_n$ about $TrueSim$. Since the $TrueSim$ is
-the $Sim$ when all users that will rate the items have rated them, $Sim_n$ can be
-thought of as sampling $n$ users from the set used to calculate $TrueSim$. 
-Although $Sim$ is likely more complicated than an average, the Central Limit 
-Theorem motivates the intuition that the variance of $Sim$ would decrease as the 
-$\sqrt{n}$. Thus we model $\sigma_{err,n}^2$ as:
+This model requires a different parameter $\sigma_{err,n}^2$ for each $n$.
+To gain a more generality, we model $\sigma_{err,n}^2$ as a function of $n$
+instead of considering them independent of each other. Recall that 
+$\sigma_{err,n}^2$ is the noisiness of the observed $Sim_n$ about $TrueSim$. 
+Since the $TrueSim$ is the $Sim$ when all users that will rate the items 
+have rated them, $Sim_n$ can be thought of as sampling $n$ users from the 
+set used to calculate $TrueSim$. Although $Sim$ is likely more complicated 
+than an average, the Central Limit Theorem motivates the intuition that the 
+variance of $Sim$ would decrease as the sqaure root of $n$. Thus we model 
+$\sigma_{err,n}^2$ as:
 \begin{align}
@@
 \subsection*{Technique}
-The parameters $\mu$, $\sigma_{s}^2$, and $\sigma_{err,n}^2$ can be approximated 
-using the method of moments over training data. $\mu$ and $\sigma_{s}$ are
-calculated
+While it may be impossible to get $TrueSim$ for any pair, we
+can get close by calculating $Sim_n$ when a large number of users have rated
+both items. Formally, we define $GoldSim$ = $Sim_N$ where $N$ is a number 
+large enough that $GoldSim \approx TrueSim$.
 
-Based on this maodel, we try three different techniques to predict the true similarity from a
-measured similarity. The techniques vary in their fidelity to the model. For
-each, we limit our data to all pairs of items that have enough common users that
-we can approximate $TrueSim \approx Sim$. Then we pick subsets of the common
-users of various sizes $n$ and calculate $Sim_{n}$. With examples of $Sim$, $n$
-and the resulting $TrueSim$, we can implement supervised learning.
+With a measurable $GoldSim$ as our goal, we can apply supervised learning 
+techniques. First, we find pairs of items that have at least $N$ common users 
+and calculate their $GoldSim$. Next we withold some of the users to produce 
+estimated $Sim_n$ for $n=1,...,N$. By doing this over many pairs of items, we 
+produce pairs of ($Sim_n$, $TrueSim$) that can be used for training and testing.
 
-The first technique is the least faithful to the model. It simply runs a number
-of linear regressions between $Sim$ and $TrueSim$ for each $n$. This ignores the 
-data-specific parameters $\mu$, $\sigma_{s}$ and $\sigma_{err,n}$ to find the best 
-linear fit. It also treats each $n$ completely independently. This technique 
-necessarilty has the lowest training error of all linear techniques, potentially 
-at the risk of overfitting.
+Now, the parameters $\mu$, $\sigma_{s}^2$, and $\sigma_{err,n}^2$ can be 
+approximated using the method of moments over the training data. The values of 
+$\sigma_{err,n}^2$ are fit to the function $\frac{\alpha}{\sqrt{n}}$ and the $\alpha$ 
+that minimizes the sum of squared error is computed.
 
-The second technique is more general. It approximates $\mu$ and $\sigma_{s}$
-using the method of moments on the training data of all $TrueSim$ scores. Then,
-$\sigma_{err,n}$ is also approximated using a method of moments on the training
-data of all $Sim_n$ for each $n$ separately. Here $\sigma_{err,n}$'s are also
-considered independently of each other.
+The following graph compares the results of three linear fits, with varying
+degrees of fidelity to the probabilistic model. The most faithful fit uses
+$\mu$, $\sigma_{s}^2$ and $\alpha$ as calculated above to predict the $GoldSim$
+using Eqn (18). The next fit uses the calculated values of $\sigma_{err,n}^2$ directly,
+following Eqn (16) for predictions. The third fit simply uses a series of linear
+regressions, one for each $n$. This model only assumes that for each $n$,
+$Sim_n$ and $GoldSim$ are linearly correlated.
 
-The third and most general technique models $\sigma_{err,n}$ as a function of
-$n$ using equation 19. As in the previous technique, $\mu$, $\sigma_{s}$ and each
-$\sigma_{err,n}$ are approximated using the method of moments over the training
-data. Then $\alpha$ is calculated that minimizes the sum of squared errors
-between $\sigma_{err,n}^2$ and $\frac{\alpha}{\sqrt{n}}$. This $\alpha$ is used to
-construct the three-parameter prediction function of equation 19.
+\begin{figure}[!htbp]
+    \centering
+    \includegraphics[width=0.7\textwidth]{LinearTechniques.png}
+	\caption{Comparison of different linear models. For each number of reviewers
+    in common $n$, the graphs represent the average squared error of predicted
+    $GoldSim$ calculated from the observed $Sim_n$. As a baseline, the technique
+    of guessing the average $GoldSim$ is included.}
+    \label{fig:Linear}
+\end{figure}
+
+The linear regression technique trades the generality of the model for a better
+fit on the training data. As the results show, the trade-off is worth it.
 

### 2014-03-25 paper/paper.tex
Touched up intro and score-based section

@@
 Similarly, we consider $Sim_n$ to be a generic observed similarity score when
-the two itmes have $n$ common reviewers. Now, let $Y$ be a random variable 
+the two items have $n$ common reviewers. Now, let $Y$ be a random variable 
 representing the $TrueSim$ of a randomly chosen pair of items. Let $X_{n}$ be a
@@
 The model suggests that there is a linear correlation between $X_n$ ($Sim_n$)
-and $Y$ ($TrueSim$). Moreover, the best predicted $TrueSim$ is a linear 
-combination of the average true similarity and the observed similiarity. Since $\sigma_{err,n}^2$ is 
-nonnegative, the slope is always less than or equal to one, meaning that the 
-observed distance from $\mu$ (the average true similarity) overestimates the 
-true distance. This is reasonable, since prior suggests that the true 
-similarity is likely to be near $\mu$.\footnote{It is worth noting that despite
+and $Y$ ($TrueSim$).\footnote{It is worth noting that despite
 the resemblance to Slope One Predictors \cite{Lemire1998}, this is a fundamentally 
@@
 between the ratings a user gives to one items versus another.} 
+Moreover, the best predicted $TrueSim$ is a linear 
+combination of the average true similarity and the observed similiarity. Since $\sigma_{err,n}^2$ is 
+nonnegative, the slope is always less than or equal to one, meaning that the 
+observed distance from $\mu$ (the average true similarity) overestimates the 
+true distance. This is reasonable, since the prior suggests that the true 
+similarity is likely to be near $\mu$.
 
 This model requires a different parameter $\sigma_{err,n}^2$ for each $n$.
-To gain a more generality, we model $\sigma_{err,n}^2$ as a function of $n$
+To gain more generality, we model $\sigma_{err,n}^2$ as a function of $n$
 instead of considering them independent of each other. Recall that 
 $\sigma_{err,n}^2$ is the noisiness of the observed $Sim_n$ about $TrueSim$. 
-Since the $TrueSim$ is the $Sim$ when all users that will rate the items 
+Since $TrueSim$ is $Sim$ when all users that will rate the items 
 have rated them, $Sim_n$ can be thought of as sampling $n$ users from the 
-set used to calculate $TrueSim$. Although $Sim$ is likely more complicated 
+set used to calculate $TrueSim$. Although $Sim$ is usually more complicated 
 than an average, the Central Limit Theorem motivates the intuition that the 
@@
 
-With a measurable $GoldSim$ as our goal, we can apply supervised learning 
-techniques. First, we find pairs of items that have at least $N$ common users 
-and calculate their $GoldSim$. Next we withold some of the users to produce 
-estimated $Sim_n$ for $n=1,...,N$. By doing this over many pairs of items, we 
-produce pairs of ($Sim_n$, $TrueSim$) that can be used for training and testing.
-
-Now, the parameters $\mu$, $\sigma_{s}^2$, and $\sigma_{err,n}^2$ can be 
-approximated using the method of moments over the training data. The values of 
-$\sigma_{err,n}^2$ are fit to the function $\frac{\alpha}{\sqrt{n}}$ and the $\alpha$ 
-that minimizes the sum of squared error is computed.
+With a measurable $GoldSim$ as our predictive goal, we can apply supervised 
+learning techniques. First, we find pairs of items that have at least $N$ common
+users and calculate their $GoldSim$. Next we withold some of the users to 
+produce estimated $Sim_n$ for $n=1,...,N$. By doing this over many pairs of 
+items, we produce pairs of ($Sim_n$, $TrueSim$) across different values of $n$ 
+that can be used for training and testing. Now the parameters $\mu$, 
+$\sigma_{s}^2$, and $\sigma_{err,n}^2$ can be approximated using the method of 
+moments over the training data. The values of $\sigma_{err,n}^2$ are fit to the 
+function $\frac{\alpha}{\sqrt{n}}$ and the $\alpha$ that minimizes the sum of 
+squared error is computed.
 
@@
     \includegraphics[width=0.7\textwidth]{LinearTechniques.png}
-	\caption{Comparison of different linear models. For each number of reviewers
-    in common $n$, the graphs represent the average squared error of predicted
+	\caption{Comparison of different linear techniques. For each number of reviewers
+    in common, $n$, the graphs represent the average squared error of predicted
     $GoldSim$ calculated from the observed $Sim_n$. As a baseline, the technique
-    of guessing the average $GoldSim$ is included.}
+    of guessing the average $GoldSim$ is included. For these experiments,
+    $N=100$ for determining $GoldSim$ and $70\%$ of the data was used for
+    training.}
     \label{fig:Linear}

### 2014-03-25 paper/paper.tex
Working on Problem section

@@
 
+\section*{Problem}
 Both similarity estimates primarily leverage information from common users, that
 is users who have rated both items. Because of this, they perform well with a
-large number common users, but give unreliable, and dramatically varying,
+large number of common users, but give unreliable, and dramatically varying,
 results when the items have few common users. In an extreme case, if only one
@@
 rating-vectors themselves.
+
+\subsection*{The Experiment}
+We construct an experiment to test the efficacy of a predictive model. 
+For either similarity metric, there is not enough data to actually calculate
+$TrueSim$ for any pair. Instead, we can get close by calculating the
+corresponding $Sim_n$ metric when a large number of users have rated both 
+items. Formally, we define $GoldSim$ = $Sim_N$ where $N$ is a number large 
+enough that $GoldSim \approx TrueSim$.
+
+With a measurable $GoldSim$ as our predictive goal, we can implement a simple
+predictivity experiment. First, we find pairs of items that have at least 
+$N$ common users and calculate their $GoldSim$. Next, for each of these pairs
+of items, we withhold user ratings, leaving $n$ common reviewers. This 
+simulates a pair of items at times when there is a sparse amount of rating data. 
+We repeat theis for $n=1,...,N$ to get varying degrees of sparsity. With the
+sparse ratings alone, the model in question generates a prediction of the 
+$GoldSim$, which is compared to the actual $GoldSim$. This is repeated over 
+a large number of test pairs and the average squared error of the prediction 
+is calculated for each $n$.
+
+For this experimetn, we use Amazon's book data. Books are rated by users on a 1
+to 5 star-scale. As in 
+Some stuff about the experiment details: the data we use, how we subtract the
+biases, the biasedness of the experimental dataset, training vs testing.
 
@@
 both items. Formally, we define $GoldSim$ = $Sim_N$ where $N$ is a number 
-large enough that $GoldSim \approx TrueSim$.
+large enough that $GoldSim \approx TrueSim$. Using this approach to approximate
+$TrueSim$, we can look at the distribution of $GoldSim$ scores (Fig. 2) and see 
+that the Normal distribution of $TrueSim$ is a reasonable model.
+
+\begin{figure}[!htbp]
+    \centering
+    \includegraphics[width=0.8\textwidth]{Histograms.png}
+	\caption{Histograms of {\em TrueSimR} (top) and {\em TrueSimP} (bottom) for
+training datasets containing at least 40 common reviewers (left) and 80 common
+reviewers (right).}
+    \label{fig:Histograms}
+\end{figure}
 
@@
 
-The following graph compares the results of three linear fits, with varying
+The graph of Fig. 3 compares the results of three linear fits, with varying
 degrees of fidelity to the probabilistic model. The most faithful fit uses
@@
 regressions, one for each $n$. This model only assumes that for each $n$,
-$Sim_n$ and $GoldSim$ are linearly correlated.
+$Sim_n$ and $GoldSim$ are linearly correlated. The linear regression technique 
+trades the generality of the model for a better fit on the training data. 
+As the results show, the trade-off is worth it.
 
@@
     $GoldSim$ calculated from the observed $Sim_n$. As a baseline, the technique
-    of guessing the average $GoldSim$ is included. For these experiments,
-    $N=100$ for determining $GoldSim$ and $70\%$ of the data was used for
-    training.}
+    of guessing the average $GoldSim$ is included. For these experiments, we use
+    $Sim$ = $PearSim$ and $N=100$ for determining $GoldSim$.}
     \label{fig:Linear}
 \end{figure}
-
-The linear regression technique trades the generality of the model for a better
-fit on the training data. As the results show, the trade-off is worth it.
 
@@
     \centering
-    \includegraphics[width=1.0\textwidth]{Histograms.png}
-	\caption{Histograms of {\em TrueSimR} (top) and {\em TrueSimP} (bottom) for
-training datasets containing at least 40 common reviewers (left) and 80 common
-reviewers (right).}
-    \label{fig:Histograms}
-\end{figure}
-
-\begin{figure}[!htbp]
-    \centering
     \includegraphics[width=0.9\textwidth]{MSE_SimR_40.png}
@@
     \emph{SDM.} Vol. 5, 2005.
+\bibitem{Koren2009}
+    Koren, Yehuda, Robert Bell, and Chris Volinsky. 
+    ``Matrix factorization techniques for recommender systems.''
+    \emph{Computer} 42.8, 2009. 30-37.
 

### 2014-03-25 paper/paper.tex
Tweak Score-based section.

@@
 motivated by probabilistic models with distributions grounded in the data. The
-first approach leverages the observed similarity score and the number of common
-users to calculate the true similarity. The second approach leverages the
-rating-vectors themselves.
+first method leverages the observed naive similarity estimate and the number of
+common users to better estimate the true similarity. The second method leverages
+the observed ratings themselves for the same ends.
 
@@
 \begin{align}
-Y \sim N(\mu, \sigma_{s}^2)
+Y \sim N(\mu, \sigma_{s}^2),
 \end{align}
@@
 \end{align}
-
-Note that $\sigma_{err, n}^2$ represents how noisy $Sim_n$ is as an estimation of
+Note that $\sigma_{err, n}^2$ represents how noisy $Sim_n$ is as an estimate of
 $TrueSim$ and therefore should decrease as $n$ increases. The probability
-distribution of $X_n$ can be visualized in Figure 1.
+distribution of $X_n$ is illustrated in Fig. 1.
 
@@
 
-The problem of approximating the $TrueSim$ from $Sim_n$ can now be seen as
-finding the value of $Y$ that most likely produced $X_n$. In other words we 
-desire the Maximum Likelihood Estimate of $Y | X_n$:
+The problem of approximating $TrueSim$ given $Sim_n$ can now be seen as finding
+the value of $Y$ that most likely produced $X_n$. In other words we desire the
+Maximum Likelihood Estimate of $Y | X_n$:
 \begin{align}
@@
 \end{align}
-This produces the following linear equation for $\hat{y}$ in terms of $x$:
+This produces the following linear equation for $\hat{y}$ in terms of $x$
 \begin{align}
-\hat{y} &= \frac{\sigma_{s}^2x+\sigma_{err,n}^2\mu}{\sigma_{s}^2+\sigma_{err,n}^2}
+\hat{y} &=
+\frac{\sigma_{s}^2x+\sigma_{err,n}^2\mu}{\sigma_{s}^2+\sigma_{err,n}^2},
 \end{align}
-which can be rewritten as:
+which can be rewritten as
 \begin{align}
 \left(\hat{y} - \mu\right) &= \frac{\sigma_{s}^2}{\sigma_{s}^2+\sigma_{err,n}^2}
-\left(x-\mu\right)
+\left(x-\mu\right).
 \end{align}
@@
 The model suggests that there is a linear correlation between $X_n$ ($Sim_n$)
-and $Y$ ($TrueSim$).\footnote{It is worth noting that despite
-the resemblance to Slope One Predictors \cite{Lemire1998}, this is a fundamentally 
-different technique. Here, we have a linear relationship between observed 
-scores and true scores, whereas the Slope One model seeks a linear relationship 
-between the ratings a user gives to one items versus another.} 
-Moreover, the best predicted $TrueSim$ is a linear 
-combination of the average true similarity and the observed similiarity. Since $\sigma_{err,n}^2$ is 
-nonnegative, the slope is always less than or equal to one, meaning that the 
-observed distance from $\mu$ (the average true similarity) overestimates the 
-true distance. This is reasonable, since the prior suggests that the true 
-similarity is likely to be near $\mu$.
+and $Y$ ($TrueSim$).\footnote{It is worth noting that despite the resemblance to
+Slope One Predictors \cite{Lemire1998}, this is a fundamentally different
+technique. Here, we have a linear relationship between observed scores and true
+scores, whereas the Slope One model seeks a linear relationship between the
+ratings a user gives to different items.} Moreover, the best predicted $TrueSim$
+is a linear combination of the average true similarity and the observed
+similiarity. Since $\sigma_{err,n}^2$ is nonnegative, the slope is always less
+than or equal to one, meaning that, on average, the observed distance from $\mu$
+overestimates the true distance from $\mu$. This is reasonable, since the prior
+distribution suggests that the true similarity is likely to be near $\mu$.
 
-This model requires a different parameter $\sigma_{err,n}^2$ for each $n$.
-To gain more generality, we model $\sigma_{err,n}^2$ as a function of $n$
-instead of considering them independent of each other. Recall that 
-$\sigma_{err,n}^2$ is the noisiness of the observed $Sim_n$ about $TrueSim$. 
-Since $TrueSim$ is $Sim$ when all users that will rate the items 
-have rated them, $Sim_n$ can be thought of as sampling $n$ users from the 
-set used to calculate $TrueSim$. Although $Sim$ is usually more complicated 
-than an average, the Central Limit Theorem motivates the intuition that the 
-variance of $Sim$ would decrease as the sqaure root of $n$. Thus we model 
-$\sigma_{err,n}^2$ as:
+This model requires a different parameter $\sigma_{err,n}^2$ for each $n$.  To
+gain more generality, we model $\sigma_{err,n}^2$ as a function of $n$ instead
+of considering them independent of each other. Recall that $\sigma_{err,n}^2$ is
+the noisiness of the observed $Sim_n$ about $TrueSim$.  $Sim$ converges to
+$TrueSim$ as the number of users $n$ increases.  The Central Limit Theorem
+therefore motivates the intuition that the variance of $Sim$ would decrease as
+the square root of $n$. Thus we model $\sigma_{err,n}^2$ as
 \begin{align}
@@
 \end{align}
-yielding the single parameter model:
+yielding the single parameter model
 \begin{align}
 \left(Y - \mu\right) = \frac{\sigma_{s}^2}{\sigma_{s}^2+\frac{\alpha}{\sqrt{n}}}
-\left(X_n-\mu\right)
+\left(X_n-\mu\right).
 \end{align}
@@
 \subsection*{Technique}
-While it may be impossible to get $TrueSim$ for any pair, we
-can get close by calculating $Sim_n$ when a large number of users have rated
-both items. Formally, we define $GoldSim$ = $Sim_N$ where $N$ is a number 
-large enough that $GoldSim \approx TrueSim$. Using this approach to approximate
-$TrueSim$, we can look at the distribution of $GoldSim$ scores (Fig. 2) and see 
-that the Normal distribution of $TrueSim$ is a reasonable model.
+To evaluate this method of improving similarity estimates, we would like to know
+{\em a priori} the true similarity for a set of item pairs. This way, we could
+compute the error in estimates made based on only partial rating data. Of
+course, the metric $TrueSim$ is a theoretical construct and not something we
+have access to in the real world. Fortunately, we can achieve good
+approximations of $TrueSim$ given sufficient data. We therefore define $GoldSim$
+= $Sim_N$ where $N$ is a number large enough that $GoldSim \approx TrueSim$.
+Using this approach to approximate $TrueSim$, we can look at the distribution of
+$GoldSim$ scores (Fig. 2) and see that the Normal distribution of $TrueSim$ is a
+reasonable model.
 

### 2014-03-25 paper/paper.tex
Finished first draft of problem section

--- 2014-03-25 paper/paper.tex
@@
 
-For this experimetn, we use Amazon's book data. Books are rated by users on a 1
-to 5 star-scale. As in 
-Some stuff about the experiment details: the data we use, how we subtract the
-biases, the biasedness of the experimental dataset, training vs testing.
+For this experiment, we use Amazon's book data. Books are rated by users on a 1
+to 5 star-scale. Before any analysis of the data, we remove the biases from each
+rating, as in \cite{Koren2009}:
+\begin{align}
+\hat{r_{ui}} = r_{ui} - \mu_r - b_i - b_u
+\end{align}
+where $r_{ui}$ is the raw rating user $u$ gave item $i$, $\mu_r$ is the average
+rating over the entire dataset, $b_i$ is the item bias, calculated as the
+average rating for item $i$ after $\mu_r$ has been subtracted from each, and 
+$b_u$ is the user bias, calculated as the average rating from user $u$ after
+$\mu_r$ and $b_i$ have been subtracted from each.
 

### 2014-03-25 paper/paper.tex
Fix histograms fig. caption.

--- 2014-03-25 paper/paper.tex
@@
     \includegraphics[width=0.8\textwidth]{Histograms.png}
-	\caption{Histograms of {\em TrueSimR} (top) and {\em TrueSimP} (bottom) for
-training datasets containing at least 40 common reviewers (left) and 80 common
-reviewers (right).}
+	\caption{Histograms of {\em TrueSimRand} (top) and {\em TrueSimPref}
+(bottom) for training datasets containing at least 40 common reviewers (left)
+and 80 common reviewers (right).}
     \label{fig:Histograms}
@@
 $\mu$, $\sigma_{s}^2$ and $\alpha$ as calculated above to predict the $GoldSim$
-using Eqn (18). The next fit uses the calculated values of $\sigma_{err,n}^2$ directly,
-following Eqn (16) for predictions. The third fit simply uses a series of linear
-regressions, one for each $n$. This model only assumes that for each $n$,
-$Sim_n$ and $GoldSim$ are linearly correlated. The linear regression technique 
-trades the generality of the model for a better fit on the training data. 
-As the results show, the trade-off is worth it.
+using Eqn (18). The next fit uses the calculated values of $\sigma_{err,n}^2$
+directly, following Eqn (16) for predictions. The third fit simply uses a series
+of linear regressions, one for each $n$. This model only assumes that for each
+$n$, $Sim_n$ and $GoldSim$ are linearly correlated. The linear regression
+technique trades the generality of the model for a better fit on the training
+data.  As the results show, the trade-off is worth it.
 
@@
     \includegraphics[width=0.7\textwidth]{LinearTechniques.png}
-	\caption{Comparison of different linear techniques. For each number of reviewers
-    in common, $n$, the graphs represent the average squared error of predicted
-    $GoldSim$ calculated from the observed $Sim_n$. As a baseline, the technique
-    of guessing the average $GoldSim$ is included. For these experiments, we use
-    $Sim$ = $PearSim$ and $N=100$ for determining $GoldSim$.}
+	\caption{Comparison of different linear techniques. For each number of
+reviewers in common, $n$, the graphs represent the average squared error of
+predicted $GoldSim$ calculated from the observed $Sim_n$. As a baseline, the
+technique of guessing the average $GoldSim$ is included. For these experiments,
+we use $Sim$ = $PearSim$ and $N=100$ for determining $GoldSim$.}
     \label{fig:Linear}

### 2014-03-25 paper/paper.tex
Added subscripts to mu and some other small changes

--- 2014-03-25 paper/paper.tex
@@
 \begin{align}
-Y \sim N(\mu, \sigma_{s}^2),
+Y \sim N(\mu_s, \sigma_{s}^2),
 \end{align}
-where $\mu$ and $\sigma_{s}^2$ are the average similarity score and the variance
+where $\mu_s$ and $\sigma_{s}^2$ are the average similarity score and the variance
 of similarity scores of all pairs of items, respectively. Since $X_n$ represents
@@
 {2\sigma_{err,n}^2}\right)}
-\frac{1}{\sigma_{s}\sqrt{2\pi}}\exp{\left(\frac{-(y-\mu)^2}
+\frac{1}{\sigma_{s}\sqrt{2\pi}}\exp{\left(\frac{-(y-\mu_s)^2}
 {2\sigma_{s}^2}\right)}\right]
 \\&= \argmin_y\left[\frac{(x-y)^2}{2\sigma_{err,n}^2} +
-\frac{(y-\mu)^2}{2\sigma_{s}^2}\right]
+\frac{(y-\mu_s)^2}{2\sigma_{s}^2}\right]
 \\&= \argmin_y\left[\left(\sigma_{s}^2+\sigma_{err,n}^2\right)y^2 - 
-2\left(\sigma_{s}^2x+\sigma_{err,n}^2\mu\right)y\right]
+2\left(\sigma_{s}^2x+\sigma_{err,n}^2\mu_s\right)y\right]
 \end{align}
@@
 \frac{d}{dy}\left[\left(\sigma_{s}^2+\sigma_{err,n}^2\right)y^2 - 
-2\left(\sigma_{s}^2x+\sigma_{err,n}^2\mu\right)y\right]
+2\left(\sigma_{s}^2x+\sigma_{err,n}^2\mu_s\right)y\right]
 &= 2\left(\sigma_{s}^2+\sigma_{err,n}^2\right)y - 
-2\left(\sigma_{s}^2x+\sigma_{err,n}^2\mu\right) 
+2\left(\sigma_{s}^2x+\sigma_{err,n}^2\mu_s\right) 
 \\&= 0
@@
 \hat{y} &=
-\frac{\sigma_{s}^2x+\sigma_{err,n}^2\mu}{\sigma_{s}^2+\sigma_{err,n}^2},
+\frac{\sigma_{s}^2x+\sigma_{err,n}^2\mu_s}{\sigma_{s}^2+\sigma_{err,n}^2},
 \end{align}
@@
 \begin{align}
-\left(\hat{y} - \mu\right) &= \frac{\sigma_{s}^2}{\sigma_{s}^2+\sigma_{err,n}^2}
-\left(x-\mu\right).
+\left(\hat{y} - \mu_s\right) &= \frac{\sigma_{s}^2}{\sigma_{s}^2+\sigma_{err,n}^2}
+\left(x-\mu_s\right).
 \end{align}
@@
 similiarity. Since $\sigma_{err,n}^2$ is nonnegative, the slope is always less
-than or equal to one, meaning that, on average, the observed distance from $\mu$
-overestimates the true distance from $\mu$. This is reasonable, since the prior
-distribution suggests that the true similarity is likely to be near $\mu$.
+than or equal to one. This means that, on average, the observed distance from $\mu_s$
+overestimates the true distance from $\mu_s$. This is reasonable, since the prior
+distribution suggests that the true similarity is likely to be near $\mu_s$.
 
@@
 \begin{align}
-\left(Y - \mu\right) = \frac{\sigma_{s}^2}{\sigma_{s}^2+\frac{\alpha}{\sqrt{n}}}
-\left(X_n-\mu\right).
+\left(Y - \mu_s\right) = \frac{\sigma_{s}^2}{\sigma_{s}^2+\frac{\alpha}{\sqrt{n}}}
+\left(X_n-\mu_s\right).
 \end{align}
@@
 items, we produce pairs of ($Sim_n$, $TrueSim$) across different values of $n$ 
-that can be used for training and testing. Now the parameters $\mu$, 
+that can be used for training and testing. Now the parameters $\mu_s$, 
 $\sigma_{s}^2$, and $\sigma_{err,n}^2$ can be approximated using the method of 
@@
 degrees of fidelity to the probabilistic model. The most faithful fit uses
-$\mu$, $\sigma_{s}^2$ and $\alpha$ as calculated above to predict the $GoldSim$
+$\mu_s$, $\sigma_{s}^2$ and $\alpha$ as calculated above to predict the $GoldSim$
 using Eqn (18). The next fit uses the calculated values of $\sigma_{err,n}^2$ directly,
@@
     of guessing the average $GoldSim$ is included. For these experiments, we use
-    $Sim$ = $PearSim$ and $N=100$ for determining $GoldSim$.}
+    $Sim$ = $PearSim$ and $N=70$ for determining $GoldSim$.}
     \label{fig:Linear}

### 2014-03-25 paper/paper.tex

@@
 $\mu_s$, $\sigma_{s}^2$ and $\alpha$ as calculated above to predict the $GoldSim$
-using Eqn (18). The next fit uses the calculated values of $\sigma_{err,n}^2$ directly,
-following Eqn (16) for predictions. The third fit simply uses a series of linear
-regressions, one for each $n$. This model only assumes that for each $n$,
-$Sim_n$ and $GoldSim$ are linearly correlated. The linear regression technique 
-trades the generality of the model for a better fit on the training data. 
-As the results show, the trade-off is worth it.
+using Eqn (18). The next fit uses the calculated values of $\sigma_{err,n}^2$
+directly, following Eqn (16) for predictions. The third fit simply uses a series
+of linear regressions, one for each $n$. This model only assumes that for each
+$n$, $Sim_n$ and $GoldSim$ are linearly correlated. The linear regression
+technique trades the generality of the model for a better fit on the training
+data.  As the results show, the trade-off is worth it.
 
@@
     \includegraphics[width=0.7\textwidth]{LinearTechniques.png}
-	\caption{Comparison of different linear techniques. For each number of reviewers
-    in common, $n$, the graphs represent the average squared error of predicted
-    $GoldSim$ calculated from the observed $Sim_n$. As a baseline, the technique
-    of guessing the average $GoldSim$ is included. For these experiments, we use
-    $Sim$ = $PearSim$ and $N=70$ for determining $GoldSim$.}
+	\caption{Comparison of different linear techniques. For each number of
+reviewers in common, $n$, the graphs represent the average squared error of
+predicted $GoldSim$ calculated from the observed $Sim_n$. As a baseline, the
+technique of guessing the average $GoldSim$ is included. For these experiments,
+we use $Sim$ = $PearSim$ and $N=70$ for determining $GoldSim$.}
     \label{fig:Linear}

### 2014-03-25 paper/paper.tex
Cleaned up technique section, moved to experiment section

--- 2014-03-25 paper/paper.tex
@@
 We construct an experiment to test the efficacy of a predictive model. 
-For either similarity metric, there is not enough data to actually calculate
-$TrueSim$ for any pair. Instead, we can get close by calculating the
-corresponding $Sim_n$ metric when a large number of users have rated both 
-items. Formally, we define $GoldSim$ = $Sim_N$ where $N$ is a number large 
-enough that $GoldSim \approx TrueSim$.
+We would like to know {\em a priori} the true similarity for a pair of items.
+For either true similarity metric, $TrueSim$ is a theoretical construct and not 
+something we have access to in the real world. Fortunately, we can achieve good
+approximations of $TrueSim$ given sufficient data. We therefore define $GoldSim$
+= $Sim_N$ where $N$ is a number large enough that $GoldSim \approx TrueSim$.
 
-With a measurable $GoldSim$ as our predictive goal, we can implement a simple
+With a measurable $GoldSim$ as our goal, we can implement a simple
 predictivity experiment. First, we find pairs of items that have at least 
@@
 of items, we withhold user ratings, leaving $n$ common reviewers. This 
-simulates a pair of items at times when there is a sparse amount of rating data. 
-We repeat theis for $n=1,...,N$ to get varying degrees of sparsity. With the
+simulates a pair of items at a time when there is sparser rating data. 
+We repeat this for $n=1,...,N$ to get varying degrees of sparsity. With the
 sparse ratings alone, the model in question generates a prediction of the 
@@
 where $\mu_s$ and $\sigma_{s}^2$ are the average similarity score and the variance
-of similarity scores of all pairs of items, respectively. Since $X_n$ represents
+of similarity scores of all pairs of items, respectively. Figure 1 shows the
+distributions of $GoldSim$, which motivates a Gaussina model for $TrueSim$.
+
+\begin{figure}[!htbp]
+    \centering
+    \includegraphics[width=0.8\textwidth]{Histograms.png}
+	\caption{Histograms of {\em GoldPearSim} (top) and {\em GoldCosSim}
+(bottom) for training datasets containing at least 40 common reviewers (left)
+and 80 common reviewers (right).}
+    \label{fig:Histograms}
+\end{figure}
+
+Since $X_n$ represents
 a noisy reading of the true similarity $Y$, we model $X_n | Y=y$ as a Gaussian
@@
 $TrueSim$ and therefore should decrease as $n$ increases. The probability
-distribution of $X_n$ is illustrated in Fig. 1.
+distribution of $X_n$ is illustrated in Figure 2.
 
@@
 
-\subsection*{Technique}
-To evaluate this method of improving similarity estimates, we would like to know
-{\em a priori} the true similarity for a set of item pairs. This way, we could
-compute the error in estimates made based on only partial rating data. Of
-course, the metric $TrueSim$ is a theoretical construct and not something we
-have access to in the real world. Fortunately, we can achieve good
-approximations of $TrueSim$ given sufficient data. We therefore define $GoldSim$
-= $Sim_N$ where $N$ is a number large enough that $GoldSim \approx TrueSim$.
-Using this approach to approximate $TrueSim$, we can look at the distribution of
-$GoldSim$ scores (Fig. 2) and see that the Normal distribution of $TrueSim$ is a
-reasonable model.
-
-\begin{figure}[!htbp]
-    \centering
-    \includegraphics[width=0.8\textwidth]{Histograms.png}
-	\caption{Histograms of {\em TrueSimR} (top) and {\em TrueSimP} (bottom) for
-training datasets containing at least 40 common reviewers (left) and 80 common
-reviewers (right).}
-    \label{fig:Histograms}
-\end{figure}
-
+\subsection*{Technique and Results}
 With a measurable $GoldSim$ as our predictive goal, we can apply supervised 
@@
 items, we produce pairs of ($Sim_n$, $TrueSim$) across different values of $n$ 
-that can be used for training and testing. Now the parameters $\mu_s$, 
-$\sigma_{s}^2$, and $\sigma_{err,n}^2$ can be approximated using the method of 
-moments over the training data. The values of $\sigma_{err,n}^2$ are fit to the 
-function $\frac{\alpha}{\sqrt{n}}$ and the $\alpha$ that minimizes the sum of 
-squared error is computed.
+that can be used for training. Now the parameters $\mu_s$, $\sigma_{s}^2$, and 
+$\sigma_{err,n}^2$ can be approximated using the method of moments over the 
+training data. The values of $\sigma_{err,n}^2$ are fit to the function 
+$\frac{\alpha}{\sqrt{n}}$ and the $\alpha$ that minimizes the sum of squared 
+error is computed.
 

### 2014-03-25 paper/paper.tex
Tweak Rating-based section.

@@
 \end{align}
-where the elements of $\vec{\mu_r}$ are the average ratings and
+where $\vec{\mu_r} = (\mu_r~~\mu_r)$ are the average ratings and
 $\mathbf{\Sigma_y}$ is the $y$-dependent $2\times2$ covariance matrix.  Since
@@
 \ln{\left[\sigma_r^2(1 - y^2)\right]} \right] &= 0 \\
--r_A^\prime r_B^\prime y^2 + ({r_A^\prime}^2 + {r_B^\prime}^2)y - r_A^\prime
-r_B^\prime + \frac{\sigma_r^2}{\sigma_s^2}(1 - y^2)^2(y - \mu_s) - \sigma_r^2(1
-- y^2)y &= 0.
+-r_A^\prime r_B^\prime \hat{y}^2 + ({r_A^\prime}^2 + {r_B^\prime}^2)\hat{y} - r_A^\prime
+r_B^\prime + \frac{\sigma_r^2}{\sigma_s^2}(1 - \hat{y}^2)^2(\hat{y} - \mu_s) - \sigma_r^2(1
+- \hat{y}^2)\hat{y} &= 0.
 \end{align}
@@
 that both ratings are zero) depending on whether the ratings have the same sign
-or not. The MLE provided by the above model rectifies this by generating
-similarity estimates in the range $[-1, +1]$ given a single pair of ratings. We
-therefore consider each user who rates a pair of items as implicitly predicting
-the similarity of the items. A key discovery of the present work is that the
-accuracy of users' predictions in a training set is correlated with the accuracy
-in a test set. We exploit this finding by ascribing a quantity we call {\em user
-predictivity} to each user defined as
+or not. The value of $\hat{y}$ provided by the above model rectifies this by
+generating similarity estimates in the range $[-1, +1]$ given a single pair of
+ratings. We therefore consider each user who rates a pair of items as implicitly
+predicting the similarity of the items. A key discovery of the present work is
+that the accuracy of users' predictions in a training set is correlated with the
+accuracy in a test set. We exploit this finding by ascribing a quantity we call
+{\em user predictivity} to each user defined as
 \begin{align}
@@
 \end{align}
-with $I_u$ the set of item pairs rated by user $u$ such that $\epsilon_u$ is
-the average absolute error of user $u$'s similarity predictions.
+with $I_u$ the set of item pairs rated by user $u$. We are once again using
+$GoldSim$ as a proxy for $TrueSim$ (See the {\bf The Experiment} section) such
+that $\epsilon_u$ is the average absolute error of user $u$'s similarity
+predictions.
 
-\section*{Experiment}
-
-\section*{Results}
+\subsection*{The Technique}
 

### 2014-03-25 paper/paper.tex
Deleted extra sections at bottom
Add Data section.

@@
 
-\subsection*{The Experiment}
-We construct an experiment to test the efficacy of a predictive model. 
-We would like to know {\em a priori} the true similarity for a pair of items.
-For either true similarity metric, $TrueSim$ is a theoretical construct and not 
-something we have access to in the real world. Fortunately, we can achieve good
-approximations of $TrueSim$ given sufficient data. We therefore define $GoldSim$
-= $Sim_N$ where $N$ is a number large enough that $GoldSim \approx TrueSim$.
+\subsection*{Data}
+In the following section, we will describe an off-line experiment used to
+evaluate our similarity estimators. But first, we provide some details about the
+data used for this work. The data, originally from Amazon.com, was acquired from
+the Stanford Network Analysis Project\footnote{http://snap.stanford.edu}. It
+consists of ~5.6M ratings by ~1.2M users of ~600k music items (e.g. albums and
+songs). Each rating is from 1 to 5 stars and is associated with a particular
+user and item at a particular time.  For practical reasons, we focus our work on
+two subsets of this data. The first consists of item pairs having at least 40
+common reviewers (i.e. users that have provided a rating for the item). The
+second consists of item pairs having at least 80 common reviewers. As discussed
+below, we must confine ourselves to item pairs that have many common reviewers
+in order to determine a good approximation of the their true similarity.
+
+\subsection*{Experiment}
+Ideally, we would know {\em a priori} the true similarity for a set of item
+pairs. But both $TrueSimRand$ and $TrueSimPref$ are theoretical constructs and
+not metrics we have access to in the real world.  Fortunately, we can achieve
+good approximations of either $TrueSim$ given sufficient data. We therefore
+define $GoldSim$ = $Sim_N$ where $N$ is a number large enough that $GoldSim
+\approx TrueSim$.
 
@@
 
-\section*{Score-based Approach}
+\section*{Score-based Method}
 \subsection*{Probabilistic Model}
@@
 
-\section*{Ratings-based Approach}
+\section*{Ratings-based Method}
 \subsection*{Probabilistic Model}
@@
 
-\subsection*{The Technique}
+\subsection*{Technique and Results}
 

### 2014-03-25 paper/paper.tex
Modify Experiment section.

@@
 user and item at a particular time.  For practical reasons, we focus our work on
-two subsets of this data. The first consists of item pairs having at least 40
-common reviewers (i.e. users that have provided a rating for the item). The
-second consists of item pairs having at least 80 common reviewers. As discussed
-below, we must confine ourselves to item pairs that have many common reviewers
-in order to determine a good approximation of the their true similarity.
+two subsets of this data. The first consists of the 13,752 item pairs having at
+least 40 common users (i.e. users that have provided a rating for the item). The
+second consists of the 1,312 item pairs having at least 80 common reviewers. As
+discussed below, we must confine ourselves to item pairs that have many common
+reviewers in order to determine a good approximation of the their true
+similarity.
 
@@
 Ideally, we would know {\em a priori} the true similarity for a set of item
-pairs. But both $TrueSimRand$ and $TrueSimPref$ are theoretical constructs and
-not metrics we have access to in the real world.  Fortunately, we can achieve
-good approximations of either $TrueSim$ given sufficient data. We therefore
-define $GoldSim$ = $Sim_N$ where $N$ is a number large enough that $GoldSim
-\approx TrueSim$.
+pairs. We could then generate similarity estimates and readily determine the
+true error in these estimates. But both $TrueSimRand$ and $TrueSimPref$ are
+theoretical constructs and not metrics we have access to in the real world.
+Fortunately, we can achieve good approximations of either $TrueSim$ given
+sufficient data. We therefore define $GoldSim = Sim_N$ where $N$ is a number
+large enough that $GoldSim \approx TrueSim$. With a measurable $GoldSim$ as our
+goal, we can implement a simple predictivity experiment.
 
-With a measurable $GoldSim$ as our goal, we can implement a simple
-predictivity experiment. First, we find pairs of items that have at least 
+To obtain good $GoldSim$ values, we work with subsets consisting of item pairs
+with at least either $N=40$ and $80$ common users. These choices are a
+compromise between having enough data for statistically significant results and
+having a sufficiently large $N$ that $GoldSim \approx TrueSim$. Note that even
+with $N=80$ the average absolute error in $PearSim$ (i.e. $\left<\left| PearSim
+- GoldSim \right|\right>$) is $\approx 0.1$\footnote{See Fig.
+  \ref{fig:MSE_SimR}}.
+
+...
+
+First, we find pairs of items that have at least 
 $N$ common users and calculate their $GoldSim$. Next, for each of these pairs

### 2014-03-25 paper/paper.tex
Minor changes

--- 2014-03-25 paper/paper.tex
@@
 
-In the following sections, we present two approaches to better approximate the
+In this paper, we present two approaches to better approximate the
 true similarity given a limited number of common reviewers. Both approaches are
@@
 technique of guessing the average $GoldSim$ is included. For these experiments,
-we use $Sim$ = $PearSim$ and $N=70$ for determining $GoldSim$.}
+we use $Sim$ = $PearSim$ and $N=80$ for determining $GoldSim$.}
     \label{fig:Linear}

### 2014-03-25 paper/paper.tex
Modify Data and Experiment sections.

--- 2014-03-25 paper/paper.tex
@@
 
+Before any analysis of the data, we remove the biases from each rating, as in
+\cite{Koren2009}:
+\begin{align}
+\hat{r_{ui}} = r_{ui} - \mu_r - b_u - b_i
+\end{align}
+where $r_{ui}$ is the raw rating user $u$ gave item $i$, $\mu_r$ is the average
+rating over the entire dataset, $b_u$ is the user bias, calculated as the
+average rating from user $u$ after $\mu_r$ has been subtracted from each, and
+$b_i$ is the item bias, calculated as the average rating of item $i$ after
+$\mu_r$ and $b_u$ have been subtracted from each.
+
 \subsection*{Experiment}
@@
 with $N=80$ the average absolute error in $PearSim$ (i.e. $\left<\left| PearSim
-- GoldSim \right|\right>$) is $\approx 0.1$\footnote{See Fig.
-  \ref{fig:MSE_SimR}}.
+- GoldSim \right|\right>$) is $\approx 0.1$ (see Fig.  \ref{fig:MSE_SimR}). So
+  while $PearSim$ and $CosSim$ are relatively stable after this many common
+users, there does remain a significant level of variance.
 
-...
-
-First, we find pairs of items that have at least 
-$N$ common users and calculate their $GoldSim$. Next, for each of these pairs
-of items, we withhold user ratings, leaving $n$ common reviewers. This 
-simulates a pair of items at a time when there is sparser rating data. 
-We repeat this for $n=1,...,N$ to get varying degrees of sparsity. With the
-sparse ratings alone, the model in question generates a prediction of the 
-$GoldSim$, which is compared to the actual $GoldSim$. This is repeated over 
-a large number of test pairs and the average squared error of the prediction 
-is calculated for each $n$.
-
-For this experiment, we use Amazon's book data. Books are rated by users on a 1
-to 5 star-scale. Before any analysis of the data, we remove the biases from each
-rating, as in \cite{Koren2009}:
-\begin{align}
-\hat{r_{ui}} = r_{ui} - \mu_r - b_i - b_u
-\end{align}
-where $r_{ui}$ is the raw rating user $u$ gave item $i$, $\mu_r$ is the average
-rating over the entire dataset, $b_i$ is the item bias, calculated as the
-average rating for item $i$ after $\mu_r$ has been subtracted from each, and 
-$b_u$ is the user bias, calculated as the average rating from user $u$ after
-$\mu_r$ and $b_i$ have been subtracted from each.
+Our off-line experiment simulates the passing of time, applying ratings to items
+in the order they actually occured. The similarity of item pairs is then
+estimated each time a user has rated both items such that $N$ is incremented by
+one. In this way, we can compare various estimators by comparing their outputs
+to each other and to the known $GoldSim$ value for each pair. We proceed like
+this for each pair in a test set for $n=1,...,N$ to get varying degrees of
+sparsity. We then compute the average squared error of the estimate for each
+$n$.
 

### 2014-03-25 paper/paper.tex
edited last two paragraphs of experiment section

@@
 Fortunately, we can achieve good approximations of either $TrueSim$ given
-sufficient data. We therefore define $GoldSim = Sim_N$ where $N$ is a number
-large enough that $GoldSim \approx TrueSim$. With a measurable $GoldSim$ as our
-goal, we can implement a simple predictivity experiment.
+sufficient data. We therefore define $GoldPearSim = PearSim_N$ and $GoldCosSim =
+CosSim_N$ where $N$ is a large enough number of common users that $GoldSim 
+\approx TrueSim$.\footnote{To calculate $GoldSim$ in practice, we use $Sim$ with
+all rating data, as long as the number of common users is greater than or equal
+to $N$} With a measurable $GoldSim$ as our goal, we can implement 
+a simple predictivity experiment.
 
-To obtain good $GoldSim$ values, we work with subsets consisting of item pairs
-with at least either $N=40$ and $80$ common users. These choices are a
-compromise between having enough data for statistically significant results and
-having a sufficiently large $N$ that $GoldSim \approx TrueSim$. Note that even
-with $N=80$ the average absolute error in $PearSim$ (i.e. $\left<\left| PearSim
-- GoldSim \right|\right>$) is $\approx 0.1$ (see Fig.  \ref{fig:MSE_SimR}). So
-  while $PearSim$ and $CosSim$ are relatively stable after this many common
-users, there does remain a significant level of variance.
+To obtain good $GoldSim$ values, we work only with the item pairs that have at 
+least either $N=40$ and $80$ common users. These choices are a compromise 
+between having enough data for statistically significant results and having 
+a sufficiently large $N$ that $GoldSim \approx TrueSim$.
 
-Our off-line experiment simulates the passing of time, applying ratings to items
-in the order they actually occured. The similarity of item pairs is then
-estimated each time a user has rated both items such that $N$ is incremented by
-one. In this way, we can compare various estimators by comparing their outputs
-to each other and to the known $GoldSim$ value for each pair. We proceed like
-this for each pair in a test set for $n=1,...,N$ to get varying degrees of
-sparsity. We then compute the average squared error of the estimate for each
-$n$.
+In order to test predicted $TrueSim$, we construct an off-line experiment to
+simulate observing sparse amounts of rating data. Our experiment reenacts  
+the passing of time. Starting with all ratings withheld, we apply ratings to 
+items in the order they actually occured. The similarity of item pairs is then
+estimated each time the number of common users $n$ is incremented. In this way,
+we can quanitatively compare various estimators by calculating their mean
+squared error when approximating $GoldSim$ across pairs of items. Since 
+different estimators may perform better for different degrees of sparsity, 
+we compute the mean squared error for $n=1,...,N$ separately.
 

### 2014-03-25 paper/paper.tex
Begin Technique and Results subsection under Rating-based Method.

@@
 with $I_u$ the set of item pairs rated by user $u$. We are once again using
-$GoldSim$ as a proxy for $TrueSim$ (See the {\bf The Experiment} section) such
-that $\epsilon_u$ is the average absolute error of user $u$'s similarity
-predictions.
+$GoldSim$ as a proxy for $TrueSim$ (See the {\bf Experiment} section) such that
+$\epsilon_u$ is the average absolute error of user $u$'s similarity predictions.
 
 \subsection*{Technique and Results}
+
+The parameter values $\mu_s$ and $\sigma_s$ are obtained by fitting the
+distribution of $GoldSimRand$ values in the training set (see Fig.
+\ref{fig:Histograms}). For example, the $N \geq 80$ dataset we have $\mu_s =
+0.27$ and $\sigma_s = 0.25$. The choice of $\mu_r$ and $\sigma_r$ is less clear.
+Some insight can be gained, however, by examining the prior distribution of
+ratings for our datasets, which are shown in Fig. \ref{fig:RatingDist}.
 
@@
 \end{figure}
+
+The upper panel of Fig. \ref{fig:RatingDist} shows data for the $N \geq 40$
+training set while the lower panel shows the same for the $N \geq 80$ training
+set. In these density plots, for each rating pair, we randomly choose which is
+Rating1 and which is Rating2. The two density plots present similar
+distributions. These data, together with the similarity score histograms of Fig.
+\ref{fig:Histograms}, suggest that the data for our two data sets are similar
+and that large biases are not introduced by varying the number of common users,
+$N$. At the same time, both rating distributions reveal significant correlations
+between Rating1 and Rating2 as evidenced by the elongation in the $45^\circ$
+direction. This is consistent with the data of Fig. \ref{fig:Histograms}, which
+shows the mean $PearSim > 0$ for both sets. Further, these data indicate a
+bias toward positive values among rating pairs compared to those of ratings in
+general, which have an average value of zero by construction\footnote{See
+discussion of bias subtraction in {\bf Data} section.}. This latter feature
+indicates that items with large numbers of common users tend to be rated higher
+by their common users than by their other users since we have already subtracted
+the item's average rating thereby avoiding any simple popularity effects.
+
+We observe that while the data of Fig. \ref{fig:RatingDist} clearly do not
+conform precisely to a multivariate Gaussian distribution, this approximation is
+not an unreasonable one.
 

### 2014-03-25 amazon/modelSim.py
Add mu and sigma for various datasets to modelSim.py.

@@ -15,9 +15,12 @@ a single reviewer.
 # "official" modelSim params
 #mu_s = 0.27 # topEdges_80_randSim
 mu_s = 0.16 # topEdges_40_randSim
-#mu_s = 0.050 # topEdges_80_prefSim
-sigma_s = 0.3 # randSim
-#sigma_s = 0.050 # prefSim
+mu_s = 0.014 # topEdges_40_prefSim
+#mu_s = 0.054 # topEdges_80_prefSim
+sigma_s = 0.25 # topEdges_80_randSim
+#sigma_s = 0.36 # topEdges_40_randSim
+#sigma_s = 0.034 # topEdges_40_prefSim
+#sigma_s = 0.049 # topEdges_80_prefSim
 #mu_r = 0.085 # topEdges_80
 mu_r = 0.25 # fix
 #mu_r = 0.026 # topEdges_40

### 2014-03-25 paper/paper.tex
Add to Technique and Results subsection of rating-based method.

@@
 \epsilon_u = \frac{1}{\left|I_u\right|}
-\sum_{A,B \in I_u}{\left|ModelSimRand(A, B) - GoldSimRand(A, B)\right|}
+\sum_{A,B \in I_u}{\left|\hat{y}(r_{u,A},r_{u,B}) - GoldSim(A, B)\right|}
 \end{align}
 with $I_u$ the set of item pairs rated by user $u$. We are once again using
-$GoldSim$ as a proxy for $TrueSim$ (See the {\bf Experiment} section) such that
+$GoldSim$ as a proxy for $TrueSim$ (See {\bf Experiment} section) such that
 $\epsilon_u$ is the average absolute error of user $u$'s similarity predictions.
@@
 The parameter values $\mu_s$ and $\sigma_s$ are obtained by fitting the
-distribution of $GoldSimRand$ values in the training set (see Fig.
+distribution of $GoldSim$ values in the training set (see Fig.
 \ref{fig:Histograms}). For example, the $N \geq 80$ dataset we have $\mu_s =
@@
 conform precisely to a multivariate Gaussian distribution, this approximation is
-not an unreasonable one.
+not an unreasonable one. In passing, we remark on the faint clusters that flank
+the central cluster in the figures. We understand these to be artifacts of the
+bias subtraction. Moreover, a more sophisticated Rating-based model could
+account for these by considering a $P(\vec{R} | Y)$ that is a mixture of
+Gaussians. Motivated partly by these density plots and partly by estimator
+performance against the training set, we choose $\mu_r = 0.25$ and $\sigma_r =
+0.3$ when estimating $TrueSimRand$.
+
+Having chosen values for $\mu_s$, $\sigma_s$, $\mu_r$, and $\sigma_r$ we are now
+able to estimate the similarity of item pairs. We do this for each common user
+$u$ in the training set in order to determine their average absolute error
+$\epsilon_u$. Next, we apply supervised learning techniques to optimize the
+value of the parameter $\lambda$ that appears in the definition of the {\em user
+predictivity} $p_u$. We do this by minimizing the error in the similarity
+estimates made against the training set where each estimate is the weighted
+average of the common user's implied prediction. That is, our new estimate for
+the similarity of items $A$ and $B$ is
+\begin{align}
+UserPredSim(A,B) = \frac{\sum_{u \in U_{AB}}{p_u\cdot\hat{y}(r_{u,A}, r_{u,B})}}
+{\sum_{u \in U_{AB}}{p_u}},
+\end{align}
+where, as before, $U_{AB}$ is the set of common users for items $A$ and $B$.
 
@@
 flat red line indicate the MSE that results from always guessing the average
-{\em TrueSimR} from the training set. The green curve is {\em RegSimR} and the
-blue curve is {\em ModelSimR}. }
+{\em TrueSimR} from the training set. The green curve is {\em LinRegSim} and the
+blue curve is {\em UserPredSim}. }
     \label{fig:MSE_SimR}
@@
 estimates of {\em TrueSimP} for item pairs ultimately having at least 80 common
-reviewers. The dashed curve represents
-{\em SimP} computed using only the ratings from the first $n$ reviewers. The
-flat red line indicate the MSE that results from always guessing the average
-{\em TrueSimP} from the training set. The green curve is {\em RegSimP} and the
-blue curve is {\em ModelSimP}. }
+reviewers. The dashed curve represents {\em CosSim} computed using only the
+ratings from the first $n$ reviewers. The flat red line indicate the MSE that
+results from always guessing the average {\em TrueSimPref} from the training
+set. The green curve is {\em LinRegSim} and the blue curve is {\em UserPredSim}.
+}
     \label{fig:MSE_SimP}

### 2014-03-25 paper/paper.tex
Tweaking techniques of predictivity model

--- 2014-03-25 paper/paper.tex
@@
 distribution of $GoldSimRand$ values in the training set (see Fig.
-\ref{fig:Histograms}). For example, the $N \geq 80$ dataset we have $\mu_s =
+\ref{fig:Histograms}). For example, in the $N \geq 80$ dataset we have $\mu_s =
 0.27$ and $\sigma_s = 0.25$. The choice of $\mu_r$ and $\sigma_r$ is less clear.
@@
 reviewers. Bottom figure shows distribution for item pairs having at least 80
-common reviewers.}
+common reviewers. Rating1 and Rating2 are chosen in arbitrary order.}
     \label{fig:RatingDist}
@@
 
-The upper panel of Fig. \ref{fig:RatingDist} shows data for the $N \geq 40$
-training set while the lower panel shows the same for the $N \geq 80$ training
-set. In these density plots, for each rating pair, we randomly choose which is
-Rating1 and which is Rating2. The two density plots present similar
-distributions. These data, together with the similarity score histograms of Fig.
+Fig. \ref{fig:RetingDist} shows density plots of rating pairs, for $N \geq 40$
+and $N \geq 80$. These data, together with the similarity score histograms of Fig.
 \ref{fig:Histograms}, suggest that the data for our two data sets are similar
@@
 \bibitem{Breese1998}
-    Breese, John S., David Heckerman, and Carl Kadie, 
+    John S. Breese, David Heckerman, and Carl Kadie, 
     ``Empirical analysis of predictive algorithms for collaborative filtering.'' 
@@
 \bibitem{Koren2009}
-    Koren, Yehuda, Robert Bell, and Chris Volinsky. 
+    Yehuda Koren, Robert Bell, and Chris Volinsky. 
     ``Matrix factorization techniques for recommender systems.''

### 2014-03-25 paper/paper.tex

--- 2014-03-25 paper/paper.tex
@@
 distribution of $GoldSim$ values in the training set (see Fig.
-\ref{fig:Histograms}). For example, the $N \geq 80$ dataset we have $\mu_s =
+\ref{fig:Histograms}). For example, in the $N \geq 80$ dataset we have $\mu_s =
 0.27$ and $\sigma_s = 0.25$. The choice of $\mu_r$ and $\sigma_r$ is less clear.

### 2014-03-25 paper/paper.tex
Added choice of parameters_r to earlier paragraph

@@
 0.27$ and $\sigma_s = 0.25$. The choice of $\mu_r$ and $\sigma_r$ is less clear.
-Some insight can be gained, however, by examining the prior distribution of
-ratings for our datasets, which are shown in Fig. \ref{fig:RatingDist}.
+Some insight can be gained by examining the prior distribution of ratings for 
+our datasets, which are shown in Fig. \ref{fig:RatingDist}. From there, we
+experiment with different values of these parameters to minimize the resulting
+error over the training set. We find $\mu_r = 0.25$ and $\sigma_r = 0.3$ to be
+good parameters and use them when estimating $TrueSimRand$.
 
@@
 account for these by considering a $P(\vec{R} | Y)$ that is a mixture of
-Gaussians. Motivated partly by these density plots and partly by estimator
-performance against the training set, we choose $\mu_r = 0.25$ and $\sigma_r =
-0.3$ when estimating $TrueSimRand$.
+Gaussians.
 

### 2014-03-25 paper/paper.tex
Further develop Technique and Results section under rating-based method.

@@
 yr_A^\prime)\right] + \frac{(y - \mu_s)^2}{\sigma_{s}^2} +
-\ln{\left[\sigma_r^2(1 - y^2)\right]} \right] &= 0 \\
+\ln{\left[\sigma_r^2(1 - y^2)\right]} \right] &= 0 \\\label{eqn:estimator}
 -r_A^\prime r_B^\prime \hat{y}^2 + ({r_A^\prime}^2 + {r_B^\prime}^2)\hat{y} - r_A^\prime
@@
 {\em user predictivity} to each user defined as
-\begin{align}
+\begin{align}\label{eqn:predictivity}
 p_u = e^{-\epsilon_u/\lambda},
@@
 where $\lambda$ is a parameter to be learned and
-\begin{align}
+\begin{align}\label{eqn:userError}
 \epsilon_u = \frac{1}{\left|I_u\right|}
@@
 Having chosen values for $\mu_s$, $\sigma_s$, $\mu_r$, and $\sigma_r$ we are now
-able to estimate the similarity of item pairs. We do this for each common user
-$u$ in the training set in order to determine their average absolute error
-$\epsilon_u$. Next, we apply supervised learning techniques to optimize the
-value of the parameter $\lambda$ that appears in the definition of the {\em user
-predictivity} $p_u$. We do this by minimizing the error in the similarity
-estimates made against the training set where each estimate is the weighted
-average of the common user's implied prediction. That is, our new estimate for
-the similarity of items $A$ and $B$ is
-\begin{align}
+able to estimate the similarity of item pairs. To do this, we must solve Eq.
+(\ref{eqn:estimator}) for $\hat{y}$ given a pair of ratings. This can be easily
+done in practice using a root finding algorithm\footnote{To solve Eq.
+(\ref{eqn:estimator}) for $\hat{y}$ given ratings $r_A$ and $r_B$, we use the
+{\tt brentq} root finding function of the widely available Python module {\tt
+scipy.optimize}.}. The implied predictions of individual users can then be
+aggregated to form a new estimate of the similarity of items $A$ and $B$ as
+\begin{align}\label{eqn:userPredSim}
 UserPredSim(A,B) = \frac{\sum_{u \in U_{AB}}{p_u\cdot\hat{y}(r_{u,A}, r_{u,B})}}
@@
 \end{align}
-where, as before, $U_{AB}$ is the set of common users for items $A$ and $B$.
+where, as before, $U_{AB}$ is the set of common users for items $A$ and $B$ and
+$p_u$ is the {\em user predictivity} defined in Eq. (\ref{eqn:predictivity}). We
+leverage the training set to find $p_u$ for each user $u$. This is done by
+recording the implied prediction $\hat{y}(r_{u,A},r_{u,B})$ for each common user
+$u$ for each item pair $(A, B)$ in the training set and then computing each
+user's average absolute error $\epsilon_u$ as per Eq. (\ref{eqn:userError}).
+Finally, we apply supervised learning techniques to optimize the value of the
+parameter $\lambda$ that appears in the definition of $p_u$. This is achieved by
+minimizing the error in Eq. {\ref{eqn:userPredSim} made against the training
+set.
 

### 2014-03-25 paper/paper.tex
Fix typo.
small changes

@@
 
-Fig. \ref{fig:RetingDist} shows density plots of rating pairs, for $N \geq 40$
+Figure \ref{fig:RatingDist} shows density plots of rating pairs, for $N \geq 40$
 and $N \geq 80$. These data, together with the similarity score histograms of Fig.
@@
 
-We observe that while the data of Fig. \ref{fig:RatingDist} clearly do not
+We observe that while the data of Fig. \ref{fig:RatingDist} do not
 conform precisely to a multivariate Gaussian distribution, this approximation is

### 2014-03-26 paper/paper.tex
Extend results of rating-based method.

--- 2014-03-25 paper/paper.tex
@@
 minimizing the error in Eq. {\ref{eqn:userPredSim} made against the training
-set.
+set. We find that $\lambda = 0.01$, which yields a strong preference for high
+{\em user predictivity}, works well for the data studied here.
+
+Figure \ref{fig:MSE_SimR} shows the results of our experiment using
+$UserPredSim$ to estimate $TrueSimRand$. To be clear, these plots are showing
+the mean squared error (MSE) given by
+\begin{align}
+\mathrm{MSE} = \frac{1}{\left|S\right|}\sum_{(A,B) \in S}
+{\left(Estimator(A,B) - GoldSim(A,B)\right)^2},
+\end{align}
+where $S$ is the test set of item pairs and $Estimator(A, B)$ represents an
+arbitrary estimator of the similarity between items $A$ and $B$. In the upper
+panel, we show the MSE for the $N \geq 40$ data. The black dashed line
+represents the estimates of $PearSim$. For example, with $n = 20$ common users,
+the MSE of $PearSim$ is about $0.07$, which corresponds to an average absolute
+error of $0.26$ in the similarity score---roughly the magnitude of $\sigma_s/2$
+(see Fig. \ref{fig:Histograms}). These data clearly show that the reliability of
+$PearSim$ in estimating $TrueSimRand$ deteriorates rapidly as $n$ decreases.
+Below about $n = 20$, the horizontal red line, which represents the MSE obtained
+from always guessing the average $GoldSimRand$ value from the training set,
+outperforms $PearSim$. These two estimators serve as simple benchmarks.
 

### 2014-03-26 paper/paper.tex
started future work section

--- 2014-03-25 paper/paper.tex
@@
 \section*{Future Work}
+The results achieved over this dataset were significant. In order to show that
+the techniques are general, though, we would like to run the same experiments on
+other datasets, with different types of items and user behavior. Most of the 
+assumptions made for this model were not specific to the data used, and so we 
+hope to achieve similar results. It would also be interesting to see how 
+different the learned parameters would be for other datasets and how well 
+the techniques could perform using the same parameters learned here.
+
+We would also like to investigate the use of latent variable
 

### 2014-03-26 paper/paper.tex
Add to results of rating-based method.

--- 2014-03-26 paper/paper.tex
@@
 
+The green curve in the upper panel of Fig. \ref{fig:MSE_SimR} is the MSE of
+$LinRegSim$, which was already shown in Fig. \ref{fig:Linear}, and is included
+here for comparison.  $LinRegSim$ outperforms always-guessing-the-average for
+all $n > 2$ and does better than $PearSim$ over the entire scope of the plot.
+$UserPredSim$ is shown in blue and for $3 < n < 33$ achieves a lower MSE than
+$LinRegSim$.
+
+The lower panel of Fig. \ref{fig:MSE_SimR} shows the MSE of estimators in the
+experiment carried out with item pairs with $N \geq 80$. The curves for
+$PearSim$, always guessing the average, and $LinRegSim$ exhibit behavior
+analogous to that observed in the $N \geq 40$ experiment, with addition of
+increased noise. The performance of $UserPredSim$, meanwhile, is significantly
+improved, achieving the MSE that $PearSim$ reaches at $n \approx 70$ by $n
+\approx 20$. This result demonstrates the efficacy, at least in some datasets,
+of measuring and exploiting {\em user predictivity} when estimating item-item
+similarity.
+
+Next, we consider the results of using $UserPredSim$ to estimate $TrueSimPref$.
+
 \begin{figure}[!htbp]

### 2014-03-26 paper/paper.tex
Finished first version of future work

--- 2014-03-26 paper/paper.tex
@@
 the techniques are general, though, we would like to run the same experiments on
-other datasets, with different types of items and user behavior. Most of the 
-assumptions made for this model were not specific to the data used, and so we 
+other datasets, with different types of items and user behavior. The 
+assumptions made for this model were not specific to the data used, so we 
 hope to achieve similar results. It would also be interesting to see how 
@@
 
-We would also like to investigate the use of latent variable
+We would also like to incorporate latent factors into our model.
+Latent factor techniques, such as the those presented in \cite{Koren2009}, have
+become very popular since the Netflix Challenge and have proved to be better for
+predicting user-item rating than classical CF. One experiment would be to
+construct a new $TrueSimLatent$ based on the similarity of two items in 
+latent factor space. Then we could apply both of the approaches in this
+paper analagously in latent factor space. Another idea is to use latent factors 
+as a new input to a model for predicting $TrueSimRand$ or $TrueSimPref$.
+
+One difficulty with each of these experiments is that it is time consuming to 
+compute the latent factors at different points in simulated time. Since we could
+not recompute the matrix factorization for every new user rating, we would 
+instead compute the latent factorization of the whole matrix at a few 
+``snap-shots'' in simulated time. We then calculate the latent factors of many
+different pairs at this time, each varying in their degrees of certainty. 
+Hopefully, this should give enough pairs with dfferent degrees of certainty 
+to train and test over.
+
+In a different direction, we believe there are a number of possible applications
+of the user predictivity score presented in this paper. Based on the results
+found here, it is reasonable to believe that adding weights to users by their
+predictivity would increase the performance of any user-item rating predictor.
+The predictivity could also have applications in the problem of Robust
+Collaborative Filtering \cite{Mehta2007}, where rating data is particularly noisy
+or riddled with spam.
 
@@
     \emph{Computer} 42.8, 2009. 30-37.
-
+\bibitem{Mehta2007}
+    Bhaskar Mehta, Thomas Hofmann, and Wolfgang Nejdl.
+    ``Robust collaborative filtering.'' 
+    \emph{Proceedings of the 2007 ACM conference on Recommender systems.}
+    ACM, 2007.
 \end{thebibliography}

### 2014-03-26 paper/paper.tex
Add to results.
Analyze results of TrueSimPref experiments.

@@
 
-Next, we consider the results of using $UserPredSim$ to estimate $TrueSimPref$.
-
 \begin{figure}[!htbp]
@@
 \end{figure}
+
+Next, we consider the results of using $UserPredSim$ to estimate $TrueSimPref$.
+Recall that $UserSimPref$ corresponds to $CosSim$ as time goes to infinity. As
+such, the relationship in our model between the covariance of the conditional
+probability distributions that generate ratings given a similarity does not hold
+for $TrueSimPref$. That is, in our model
+\begin{align}
+\mathbf{\Sigma_z} \neq \sigma_r^2 \cdot
+\left[ \begin{array}{cc}  1 & z \\ z & 1 \end{array} \right],
+\end{align}
+where $\mathbf{\Sigma_z}$ is the covariance matrix of $P(\vec{R} | Z=z)$ with
+$Z$ a random variable representing the $TrueSimPref$ of two items. Nonetheless,
+we can choose parameters and naively use $UserPredSim$ to estimate
+$TrueSimPref$. For this experiment, we choose $\mu_s = 0.054$, $\sigma_s =
+0.049$, $\mu_r = 0.25$, $\sigma_r = 0.15$, and $\lambda = 0.002$ using the same
+methods as before.
+
+In Fig. \ref{fig:MSE_SimP} we show the results of our experiments estimating
+$TrueSimPref$ using the $N \geq 80$ data set. The MSE of $CosSim$, like that of
+$PearSim$ in the case of $TrueSimRand$, rapidly diverges as the number of common
+users $n$ goes to zero. Interestingly, $CosSim$ outperforms
+always-guess-the-average for all $n > 25$, in contrast with $PearSim$, which
+does the same only after $n > 45$. The speaks to the greater stability of
+$CosSim$, which takes into consideration all ratings, not only those from common
+users. Meanwhile, the $LinRegSim$ performs surprisingly similar for estimations
+of $TrueSimPref$ and $TrueSimRand$. In both cases, the MSE begins at
+approximately that of always-guess-the-average for $n = 1$ and then decreases
+approximately linearly with increasing $n$. Indeed, this was also true for the
+$N \geq 40$ experiment. We attribute this robustness to the fact that the model
+from which $LinRegSim$ is derived abstracts away the underlying mechanism from
+which $TrueSim$ is computed. Finally, $UserPredSim$, perhaps surprisingly, has
+the lowest MSE for $n < 15$, although not by much. For all $n > 15$, $LinRegSim$
+outperforms $UserPredSim$ in estimating $TrueSimPref$. And for all $n > 40$,
+$CosSim$ achieves lower MSE than that of $UserPredSim$. This lackluster
+performance by $UserPredSim$ might be expected, however, given the model on
+which it is based is a dubious fit for $TrueSimPref$.
 

### 2014-03-26 paper/paper.tex
Small changes throughout. Full pass

--- 2014-03-26 paper/paper.tex
@@
 item vectors that contain provided ratings from all users. The true similarity
-in this model of {\em random} rating, denoted $TrueSimRand$, is then the vector
+in this model of {\em random rating}, denoted $TrueSimRand$, is then the vector
 product of two complete item vectors. That is,
@@
 in this report.}. At any point in time, we may estimate the true similarity in
-this model of {\em preferential} rating, by computing the cosine similarity of
+this model of {\em preferential rating}, by computing the cosine similarity of
 the two item vectors. Only in the limit of inifinite time does this estimate
@@
 Fortunately, we can achieve good approximations of either $TrueSim$ given
-sufficient data. We therefore define $GoldPearSim = PearSim_N$ and $GoldCosSim =
+sufficient data. We therefore define $GoldSimRand = PearSim_N$ and $GoldSimPref =
 CosSim_N$ where $N$ is a large enough number of common users that $GoldSim 
-\approx TrueSim$.\footnote{To calculate $GoldSim$ in practice, we use $Sim$ with
-all rating data, as long as the number of common users is greater than or equal
-to $N$} With a measurable $GoldSim$ as our goal, we can implement 
+\approx TrueSim$.  With a measurable $GoldSim$ as our goal, we can implement 
 a simple predictivity experiment.
@@
 of similarity scores of all pairs of items, respectively. Figure 1 shows the
-distributions of $GoldSim$, which motivates a Gaussina model for $TrueSim$.
+distributions of $GoldSim$, which motivates the Gaussian model for $TrueSim$.
 
@@
     \includegraphics[width=0.8\textwidth]{Histograms.png}
-	\caption{Histograms of {\em GoldPearSim} (top) and {\em GoldCosSim}
+	\caption{Histograms of {\em GoldSimRand} (top) and {\em GoldSimPref}
 (bottom) for training datasets containing at least 40 common reviewers (left)
@@
 of considering them independent of each other. Recall that $\sigma_{err,n}^2$ is
-the noisiness of the observed $Sim_n$ about $TrueSim$.  $Sim$ converges to
-$TrueSim$ as the number of users $n$ increases.  The Central Limit Theorem
-therefore motivates the intuition that the variance of $Sim$ would decrease as
-the square root of $n$. Thus we model $\sigma_{err,n}^2$ as
+the noisiness of the observed $Sim_n$ about $TrueSim$.  $Sim_n$ converges to
+$TrueSim$ as the number of users $n$ increases. Although $Sim$ is not simply a
+sample mean, The Central Limit Theorem motivates the intuition that the variance of 
+$Sim$ would decrease as the square root of $n$. Thus we model $\sigma_{err,n}^2$ as
 \begin{align}
@@
 \begin{align}
-\left(Y - \mu_s\right) = \frac{\sigma_{s}^2}{\sigma_{s}^2+\frac{\alpha}{\sqrt{n}}}
-\left(X_n-\mu_s\right).
+\hat{y}= \frac{\sigma_{s}^2}{\sigma_{s}^2+\frac{\alpha}{\sqrt{n}}}
+\left(x -\mu_s\right) + \mu_s.
 \end{align}
@@
 
-The graph of Fig. 3 compares the results of three linear fits, with varying
-degrees of fidelity to the probabilistic model. The most faithful fit uses
-$\mu_s$, $\sigma_{s}^2$ and $\alpha$ as calculated above to predict the $GoldSim$
-using Eqn (18). The next fit uses the calculated values of $\sigma_{err,n}^2$
-directly, following Eqn (16) for predictions. The third fit simply uses a series
+The graph of Fig. 3 compares the results of three linear predictors, with varying
+degrees of fidelity to the probabilistic model. The most faithful predictor uses
+$\mu_s$, $\sigma_{s}^2$ and $\alpha$ as calculated above to predict the $TrueSim$
+using Eqn (19). The next predictor uses the calculated values of $\sigma_{err,n}^2$
+directly, following Eqn (16) for predictions. The third predictor simply uses a series
 of linear regressions, one for each $n$. This model only assumes that for each
-$n$, $Sim_n$ and $GoldSim$ are linearly correlated. The linear regression
+$n$, $Sim_n$ and $TrueSim$ are linearly correlated. The linear regression
 technique trades the generality of the model for a better fit on the training
@@
     \includegraphics[width=0.7\textwidth]{LinearTechniques.png}
-	\caption{Comparison of different linear techniques. For each number of
+	\caption{Comparison of different linear predictors. For each number of
 reviewers in common, $n$, the graphs represent the average squared error of
@@
 $TrueSimRand$ and, to a lesser extent, $TrueSimPref$. In addition, this approach
-has the added benefit of allowing us to infer implied predictions of
-$TrueSimRand$ made by individual users each time they rate two items. This
-latter feature allows us to define a new metric for users that we call {\em
-user-predictivity}, which allows us to ascribe varying levels of confidence to
-their contributions to a similarity estimate.
+allows us to infer implied predictions of $TrueSimRand$ from individual users 
+each time they rate two items. This latter feature allows us to define a new 
+metric for users that we call {\em user-predictivity}. User-predictivity allows
+us to ascribe varying levels of confidence to a user's ratings and thus modify
+the contribution a user makes to a similarity estimate accordingly.
 

### 2014-03-26 paper/paper.tex
Fix typo.;

@@
 We would also like to incorporate latent factors into our model.
-Latent factor techniques, such as the those presented in \cite{Koren2009}, have
+Latent factor techniques, such as those presented in \cite{Koren2009}, have
 become very popular since the Netflix Challenge and have proved to be better for

### 2014-03-26 paper/paper.tex
Add Conclusion.

@@
 	\caption{Mean square error {\em vs} number of common reviewers, $n$, for
-estimates of {\em TrueSimR} for item pairs ultimately having at least 40 common
-reviewers (top) and 80 common reviewers (bottom). The dashed curve represents
-{\em SimR} computed using only the ratings from the first $n$ reviewers. The
-flat red line indicate the MSE that results from always guessing the average
-{\em TrueSimR} from the training set. The green curve is {\em LinRegSim} and the
-blue curve is {\em UserPredSim}. }
+estimates of {\em TrueSimRand} for item pairs ultimately having at least 40
+common reviewers (top) and 80 common reviewers (bottom). The dashed curve
+represents {\em SimR} computed using only the ratings from the first $n$
+reviewers. The flat red line indicate the MSE that results from always guessing
+the average {\em TrueSimR} from the training set. The green curve is {\em
+LinRegSim} and the blue curve is {\em UserPredSim}. }
     \label{fig:MSE_SimR}
@@
 	\caption{Mean square error {\em vs} number of common reviewers, $n$, for
-estimates of {\em TrueSimP} for item pairs ultimately having at least 80 common
-reviewers. The dashed curve represents {\em CosSim} computed using only the
-ratings from the first $n$ reviewers. The flat red line indicate the MSE that
-results from always guessing the average {\em TrueSimPref} from the training
-set. The green curve is {\em LinRegSim} and the blue curve is {\em UserPredSim}.
-}
+estimates of {\em TrueSimPref} for item pairs ultimately having at least 80
+common reviewers. The dashed curve represents {\em CosSim} computed using only
+the ratings from the first $n$ reviewers. The flat red line indicate the MSE
+that results from always guessing the average {\em TrueSimPref} from the
+training set. The green curve is {\em LinRegSim} and the blue curve is {\em
+UserPredSim}.  }
     \label{fig:MSE_SimP}
@@
 \section*{Conclusion}
+
+In this report, we explored the issue of item-item similarity in the context of
+collaborative filtering. Estimates of item-item similarity such as the Pearson
+correlation and cosine similarity can have very large variances as data is
+acquired, yielding large discrepancies between early results and those computed
+later after sufficient rating data is acquired. We described an off-line
+experiment that simulates the accumulation of rating data over time and the
+impact that this data has on the performance of various similarity estimators.
+Two new similarity estimators were described and analyzed.
+
+The first is based on a model of noisy score observations. Using a series of
+linear regressions, this estimator consistently performs as well or better than
+both the Pearson correlation or cosine similarity at estimating future values of
+these same functions. The second new estimator is based on a model of noisy
+rating pairs generated by an item pair having a true, but unknown, similarity.
+This method uses a parametrized surface to efficiently estimate item-item
+similarity given only a single pair of ratings. Combining these implied
+similarity ``predictions'' with the notion of {\em user predictivity}, we
+constructed a novel item-item similarity estimator that performs very well, but
+whose effectiveness appears limited to estimations of future Pearson
+correlation.
+
 \section*{Future Work}

### 2014-03-26 paper/paper.tex
Add abstract.

@@
 \section*{Abstract}
-\emph{Will write this when the paper is done}
+\emph{The problem of inaccurate item-item similarity estimates due to a lack of
+data is explored within the context of collaborative filtering. An off-line
+experiment that evaluates the performance of similarity estimators is described.
+Two novel item-item similarity estimators are described and analyzed. The first
+is based on a model of noisy similarity score generation and proves effective at
+estimating both future Pearson correlations and cosine similarities. The second
+models noisy ratings as generated by similarity score dependent distributions.
+Using the notion of user predictivity, accurate estimates of future Pearson
+correlation are made.}
 

### 2014-03-26 paper/paper.tex
Make minor tweaks.

@@
 models noisy ratings as generated by similarity score dependent distributions.
-Using the notion of user predictivity, accurate estimates of future Pearson
+Using the notion of user-predictivity, accurate estimates of future Pearson
 correlation are made.}
@@
 Determining the similarity of two users or items is valuable for a number of
-data-mining applications including recommender systems, product assortment and
-differentiation, as well as branding. Collaborative filtering is often used to
-exploit user-item rating data to make such estimates. This is typically done by
+data-mining applications including recommender systems as well as product
+assortment and differentiation. Collaborative filtering is often used to exploit
+user-item rating data to make such estimates. This is typically done by
 computing the Pearson correlation or cosine similarity of vector representations
@@
 
-\section*{Similarity}
-\subsection*{True Item-Item Similarity}
+\section*{True Item-Item Similarity}
 
@@
 this model of {\em preferential rating}, by computing the cosine similarity of
-the two item vectors. Only in the limit of inifinite time does this estimate
+the two item vectors. Only in the limit of infinite time does this estimate
 converge to the true similarity, which we denote $TrueSimPref$.  That is,
@@
 
-\subsection*{Estimating Item-Item Similarity}
+\section*{Estimating Item-Item Similarity}
 
@@
 subtracting an expectation value from each dimension. This is only approximately
-true in our case. We discuss the how we subtract rating biases later in this
+true in our case. We discuss how we subtract rating biases later in this
 report.} for this estimate, ignoring all ratings by users who have not rated
@@
 this unlikely to be an accurate assessment of the true similarity of $A$ and
-$B$, it also gives the most extreme results possible, without giving any
+$B$, it also gives the most extreme result possible, without giving any
 indication that this is a low-confidence calculation.
@@
 two subsets of this data. The first consists of the 13,752 item pairs having at
-least 40 common users (i.e. users that have provided a rating for the item). The
-second consists of the 1,312 item pairs having at least 80 common reviewers. As
-discussed below, we must confine ourselves to item pairs that have many common
-reviewers in order to determine a good approximation of the their true
+least 40 common users (i.e. users that have provided a rating for both items).
+The second consists of the 1,312 item pairs having at least 80 common reviewers.
+As discussed below, we must confine ourselves to item pairs that have many
+common reviewers in order to determine a good approximation of their true
 similarity.
@@
 To obtain good $GoldSim$ values, we work only with the item pairs that have at
-least either $N=40$ and $80$ common users. These choices are a compromise 
-between having enough data for statistically significant results and having 
-a sufficiently large $N$ that $GoldSim \approx TrueSim$.
+least either $N=40$ or $80$ common users. These choices are a compromise between
+having enough data for statistically significant results and having a
+sufficiently large $N$ that $GoldSim \approx TrueSim$.
 
-In order to test predicted $TrueSim$, we construct an off-line experiment to
-simulate observing sparse amounts of rating data. Our experiment reenacts  
-the passing of time. Starting with all ratings withheld, we apply ratings to 
-items in the order they actually occured. The similarity of item pairs is then
+In order to evaluate estimates of $TrueSim$, we construct an off-line experiment
+to simulate observing sparse amounts of rating data. Our experiment reenacts the
+passing of time. Starting with all ratings withheld, we apply ratings to items
+in the order they actually occured. The similarity of item pairs is then
 estimated each time the number of common users $n$ is incremented. In this way,
-we can quanitatively compare various estimators by calculating their mean
-squared error when approximating $GoldSim$ across pairs of items. Since 
-different estimators may perform better for different degrees of sparsity, 
-we compute the mean squared error for $n=1,...,N$ separately.
+we can quantitatively compare various estimators by calculating their mean
+squared error. Since different estimators may perform better for different
+degrees of sparsity, we compute the mean squared error for $n=1,...,N$
+separately.
 
@@
 Similarly, we consider $Sim_n$ to be a generic observed similarity score when
-the two items have $n$ common reviewers. Now, let $Y$ be a random variable 
+the two items have $n$ common reviewers. Let $Y$ be a random variable
 representing the $TrueSim$ of a randomly chosen pair of items. Let $X_{n}$ be a
@@
 $TrueSim$ as the number of users $n$ increases. Although $Sim$ is not simply a
-sample mean, The Central Limit Theorem motivates the intuition that the variance of 
-$Sim$ would decrease as the square root of $n$. Thus we model $\sigma_{err,n}^2$ as
+sample mean, The Central Limit Theorem motivates the intuition that the variance
+of $Sim$ would decrease as the square root of $n$. Thus we model
+$\sigma_{err,n}^2$ as
 \begin{align}
@@
 
-The graph of Fig. 3 compares the results of three linear predictors, with varying
-degrees of fidelity to the probabilistic model. The most faithful predictor uses
-$\mu_s$, $\sigma_{s}^2$ and $\alpha$ as calculated above to predict the $TrueSim$
-using Eqn (19). The next predictor uses the calculated values of $\sigma_{err,n}^2$
-directly, following Eqn (16) for predictions. The third predictor simply uses a series
-of linear regressions, one for each $n$. This model only assumes that for each
-$n$, $Sim_n$ and $TrueSim$ are linearly correlated. The linear regression
-technique trades the generality of the model for a better fit on the training
-data.  As the results show, the trade-off is worth it.
+The graph of Fig. 3 compares the results of three linear predictors, with
+varying degrees of fidelity to the probabilistic model. The most faithful
+predictor uses $\mu_s$, $\sigma_{s}^2$ and $\alpha$ as calculated above to
+predict the $TrueSim$ using Eqn (19). The next predictor uses the calculated
+values of $\sigma_{err,n}^2$ directly, following Eqn (16) for predictions. The
+third predictor simply uses a series of linear regressions, one for each $n$.
+This model only assumes that for each $n$, $Sim_n$ and $TrueSim$ are linearly
+correlated. The linear regression technique trades the generality of the model
+for a better fit on the training data.  As the results show, the trade-off is
+worth it.
 
@@
 technique of guessing the average $GoldSim$ is included. For these experiments,
-we use $Sim$ = $PearSim$ and $N=80$ for determining $GoldSim$.}
+we use $Sim$ = $PearSim$ and $N \geq 80$ for determining $GoldSim$.}
     \label{fig:Linear}
@@
 where $\mu_s$ and $\sigma_s^2$ are the average and variance of all true
-similarity scores, respectively. Now, let $\vec{R} = \left(R_A~~R_B\right)$ be a
+similarity scores, respectively. Let $\vec{R} = \left(R_A~~R_B\right)$ be a
 random two element vector representing the ratings given to items $A$ and $B$.
@@
 accuracy in a test set. We exploit this finding by ascribing a quantity we call
-{\em user predictivity} to each user defined as
+{\em user-predictivity} to each user defined as
 \begin{align}\label{eqn:predictivity}
@@
     \includegraphics[width=0.7\textwidth]{RatingDist_80.png}
-	\caption{Distribution of rating pairs by common reviewers for item pairs.
-Top figure shows distribution for item pairs having at least 40 common
-reviewers. Bottom figure shows distribution for item pairs having at least 80
-common reviewers. Rating1 and Rating2 are chosen in arbitrary order.}
+	\caption{Distribution of rating pairs by common users for item pairs.  Top
+figure shows distribution for item pairs having at least 40 common users. Bottom
+figure shows distribution for item pairs having at least 80 common users.
+Rating1 and Rating2 are chosen in arbitrary order.}
     \label{fig:RatingDist}
@@
 where, as before, $U_{AB}$ is the set of common users for items $A$ and $B$ and
-$p_u$ is the {\em user predictivity} defined in Eq. (\ref{eqn:predictivity}). We
+$p_u$ is the {\em user-predictivity} defined in Eq. (\ref{eqn:predictivity}). We
 leverage the training set to find $p_u$ for each user $u$. This is done by
@@
 set. We find that $\lambda = 0.01$, which yields a strong preference for high
-{\em user predictivity}, works well for the data studied here.
+{\em user-predictivity}, works well for the data studied here.
 
@@
 the MSE of $PearSim$ is about $0.07$, which corresponds to an average absolute
-error of $0.26$ in the similarity score---roughly the magnitude of $\sigma_s/2$
+error of $0.26$ in the similarity score---roughly the magnitude of $\sigma_s$
 (see Fig. \ref{fig:Histograms}). These data clearly show that the reliability of
@@
     \includegraphics[width=0.9\textwidth]{MSE_SimR_80.png}
-	\caption{Mean square error {\em vs} number of common reviewers, $n$, for
+	\caption{Mean square error {\em vs} number of common users, $n$, for
 estimates of {\em TrueSimRand} for item pairs ultimately having at least 40
-common reviewers (top) and 80 common reviewers (bottom). The dashed curve
-represents {\em SimR} computed using only the ratings from the first $n$
-reviewers. The flat red line indicate the MSE that results from always guessing
-the average {\em TrueSimR} from the training set. The green curve is {\em
+common users (top) and 80 common users (bottom). The dashed curve
+represents {\em PearSim} computed using only the ratings from the first $n$
+users. The flat red line indicate the MSE that results from always guessing
+the average {\em TrueSimRand} from the training set. The green curve is {\em
 LinRegSim} and the blue curve is {\em UserPredSim}. }
@@
 \approx 20$. This result demonstrates the efficacy, at least in some datasets,
-of measuring and exploiting {\em user predictivity} when estimating item-item
+of measuring and exploiting {\em user-predictivity} when estimating item-item
 similarity.
@@
 similarity given only a single pair of ratings. Combining these implied
-similarity ``predictions'' with the notion of {\em user predictivity}, we
+similarity ``predictions'' with the notion of {\em user-predictivity}, we
 constructed a novel item-item similarity estimator that performs very well, but
@@
 In a different direction, we believe there are a number of possible applications
-of the user predictivity score presented in this paper. Based on the results
+of the user-predictivity score presented in this paper. Based on the results
 found here, it is reasonable to believe that adding weights to users by their

### 2014-06-09 paper2/paper.tex
Add paper2 directory

\documentclass[11pt]{article}

% =========================================================================
% document style changes
% =========================================================================

\usepackage{amsmath}                    % AMS math packages
\usepackage{amssymb}                    %
\usepackage[]{graphpap}
\usepackage[T1]{fontenc}                % for \mathrm{}
\usepackage{courier}                    % for \texttt{}
\usepackage{bbm}                        % for \mathbbm{1} (indicator function)
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage[font={small}]{caption}
\usepackage{subcaption}

%\setlength{\parskip}{\baselineskip}     % skip line following paragraphs
%\pagestyle{empty}                       % No page numbers
\setlength{\topmargin}{-.5in}
\setlength{\textheight}{9in}
\setlength{\oddsidemargin}{.125in}
\setlength{\textwidth}{6.25in}

\newcommand{\spc}{\vspace{0.25in}}      % Shortcut commands
\newcommand{\ds}{\displaystyle}         %\newcommand{\ds}[1]{\displaystyle{#1}}
\newcommand{\ra}{\rightarrow}
\newcommand{\Cov}{\mathrm{Cov}}
\DeclareMathOperator*{\argmax}{arg\!\max}
\DeclareMathOperator*{\argmin}{arg\!\min}

\begin{document}                        % This is where the document begins

\title{A Technique for Inter-Catalogue Recommendations}
\author{Bill Chickering and Jamie Irvine\\
CS 399 with Anand Rajaraman\\
Stanford University}
\renewcommand{\today}{June 12, 2014}
\maketitle

\section*{Abstract}
\emph{Abstracts are abstract. They confront you. There was a reviewer a
while back who wrote that my papers didn't have any beginning or any end. He
didn't mean it as a compliment, but it was.} --Jackson Pollock

\section*{Introduction}
Recommender systems have become a critical component for online retailers in
nearly every space of retail. Whether the products are movies or clothes or research
articles, large retailer tend to have a way of recommending products based on
other previously viewed or purchased products.

These recomendations usually come from a combination of content-based
information and collaborative filtering techniques. The content-based approach
uses information about the products, such as category or descriptioin, to
connect related products. Alone, however, this approach is usually unable to
capture the subtle ways products may actually be related. Collaborative
filtering, on the other hand, learns relations from aggregated user behavior. 
Such recommendation often have explicit ttiles like ``People who liked 
this product also liked these products''. Collaborative filtering has proven to
be effective at recommending products, given that there is a large amount of user 
data.

Since each retailer only has user data within their site, these user-based
recommender systems are good at learning connections between products in
one catalogue, but they do little to learn how these products may relate to
other products outside of the retailer's inventory. This type of information is
valuable in a number of situations. Retailers such as department stores
want to know which new products make the most sense to add to their catalogue.
This is essentially a question of recommending the best outside products based
on current in-house ones. In a different scenario, two retailers may merge and 
want to construct a unified recommender system between the two catalgoues.

We model this as a problem of connecting recommender systems between disjoint
catalogues. The goal is to construct good recommendations from products in one
catalogue to products in the other. These catalogues may be from competing 
retailers who don't have access to each other's user data. Thus, to keep the 
problem general, we do not assume that we have any private internal data. 
Instead, we constrain ourselves to use only the product information and internal
recommendations that are posted publicly to all users.

This is a difficult problem. What makes a good recommendation is hard to define
objectively. One can say that if users co-purchase or co-view products together,
then these products should be recommended for each other. But without any user 
data, it is hard to simulate this kind of user behavior. The connections between 
co-viewed products are often less tangible than simple metrics of similarity.

In this paper, we outline a technique that leverages the intra-catalogue 
recommendations of each catalogue to construct inter-catalgoue recommendations
between them. We focus on the clothing retailers, since they tend to all have
public recommendations for each product. Specifically, the data used for this
project was all collected from the public product webpages for Macy's.

\section*{Experiment}
To measure the success of constructing recommendations, we design the following
experiment:

First, we scrape all of the clothing products from Macy's website at 
www.macys.com. For each product we get its image, what category of clothing it
is in, its description, and a list of other products recommended for it.

Next, we simulate the catalogues of two different retailers by partitioning all
of the products into two disjoint collections of products. To make this
partition realistic, we ensure that no brand has products in both catalogue. We
remove all recommendations between products from one partition to the other,
thus creating two entirely disjoint and unconnected catalogues.

Now, for each product in each catalogue, we choose which products to recommend
in the other catalogue, using the recommendation method to be tested. To make
this problem slightly simpler, we restrict recommendations to be within the same
category. In this way, connecting up entire catalogues becomes a series of
smaller tasks of connecting the products in each category.

We still have all of the true inter-catalogue recommendations from the
unpartitioned products. Although we don't use this information for constructing
recommendations between the partitioned graphs, we can use the original
recommendations as a gold-standard to test the construct ones. Using these
gold-standard recommendations, we can evaluate the precision and recall of the
predicted inter-catalogue recommendations.

For the Macy's catalogue, each product has four recommendations. This means that
on average, there are only two gold-standard inter-catalogue recommendations
from each product. However, since there potentially are other good
recommendations besides these two, we create another metric that expands our
notion of a correct recommendation. Here we look at recommendations of
recommendations. For example, if Product B is recommended for Product A and
Product C is recommended for Product B, we say that Product C is a good
recommendation for Product A. This leads us to the notion of 2-precision and
2-recall.

For a rigorous definition of 2-precision and 2-recall, we first propose the idea
of a recommender graph. A recommender graph has products as nodes and
recommendations as edges. If Product B is recommended for Product A, there is an
edge between node A and node B in the graph. For simplicity, we define our
recommender graphs to be undirected, but we note that they could be interpreted
as directed graphs as well. With this terminology, our problem becomes one of
connecting two disjoint recommender graphs by adding edges between them.

2-precision is defined as the percentage of predicted edges that are within two
steps in the original, unpartitioned recommender graph. In the example above,
Product A and Product C are within two steps in the original graph, thus an edge
between A and C is considered correct and increases precision. 2-recall is
defined analogously. Here, we consider the new recommender graph consisting of
the two partitioned recommender graphs connected by the predicted edges.
2-recall is then the percentage of gold-standard edges that are within two steps
in the predicted recommender graph.

\section*{Constructing Recommendations}
Recommending products between different catalogues is a daunting task.
Traditionally, constructing recommendations is done by leveraging user behavior
using a collaborative filtering technique. At the most basic level, If a large
number of users view or purchase both Product A and Product B, the two are
connected with a recommendation. Unfortunately, for this problem there are no
user data between the two unconnected catalogues.

\subsection*{Content Based Similarity}
A simple approach for this problem is to recommend products that have similar
content. The products have both a short and long description which we can use to
construct a content-based similarity function. To do this, we represent each
product as a vector of tf-idf values for each word in its descriptions. By using
tf-idf, we reduce the relevance of frequent words and boost that of rare ones.
The vectors are the length of the dictionary, with zeros for all absent words
from the descriptions. To determine the content-based similarity between two
products, we simply take the dot product of the two tf-idf vectors.

This similarity function yields the following recommendation algorithm: For each
product, scan through the products in the opposite catalogue and calculate the
content-based similarity score with each. Predict recommendations between this
product and each of the top k most similar products in the other catalogue.

While this is a quadric-time operation, it can be sped up using matrix
multiplication. Furthermore, the division into different categories reduces the
impact of the quadratic effect. For instance, Macy's entire catalogue consists
of over one million products, making a quadratic time algorithm infeasible.
However, the average category has 2000 products, which is significantly more
manageable.

\subsection*{Leveraging Recommendations}
The simple content-based recommender is limited for a few reasons. It attempts
to capture similarity between products by looking exclusively at the words used
to describe them. One limitation here is that different brands often use
different vocabularies to describe the same thing, such as ``off-white'' versus
``ivory.'' Another issue is that the best recommendation for a product is often
more subtle than simply the most similar other product. For instance, a striped
black-and-white button up shirt may be an excellent recommendation for a blue
checkered button-up, even though the two have ostensibly different patterns and
colors.

We can alleviate some of these issues by leveraging the recommendations we
already have between products in the same catalogue. One approach is to consider
the neighborhood of each product in their recommender graphs, that is a product
along with each of the products it is recommended with in its own catalogue.
When comparing two products in different catalogues, we use the same tf-idf
similarity function from before to compare each pair of products between the two
neighborhoods. Mathematically, we construct the similarity function
NeighborhoodSimilarity defined as:

where...

Assuming the recommendations are good within each of the catalogues, this
approach can help both of the issues brought up with the simple content-based
recommender. Since a neighborhood has many products, each describing themselves
with potentially different vocabularies, there is a higher likelihood that the
neighborhood contains multiple synonyms and thus decreases the potential of
missing due to different vocabularies. Furthermore, the recommender graphs
capture some of the user behavior not expressed in product descriptions. Looking
back on the dress-shirts example, the neighborhood of a striped shirt may
include checkered shirts and vise-versa. In this case, the two neighborhoods
would share the words ``striped'' and ``checkered'' and thus the two previously
dissimilar shirts would be considered similar.

\subsection*{Latent Concepts}
Although the NeighborhoodSimilarity method alleviates some of the issues of
content-similarity, it is still fundamentally based on the descriptions of
products. Ideally, we would like to recommend products that are similar in more
fundamental ways than their descriptions. If a dress is elegant and modern, we
want to recommend another dress that is elegant and modern, even if those words
are not explicitly in the descriptions of either. These types of latent concepts
can be intangible and difficult to define. However, if we assume that the given
recommendations are based on these concepts, we can attempt to learn the latent
concepts by reverse-engineering the graphs.

One approach to learning latent concepts is Latent Semantic Indexing or LSI. In
traditional LSI, a dataset consists of documents, each containing terms.
Singular Value Decomposition (SVD) is applied to the term-document matrix and
the resulting three matrices are truncated to the number of latent concepts
desired. The resulting left matrix is a term-concept matrix, where each row is a
term represented in concept space.

We apply the same idea to the product catalogues in order to learn the latent
concepts of each product. Analogous to documents containing terms, we look at
user sessions which contain the products viewed by that user. Although we do not
have actual user data, we can simulate user sessions by traversing the
recommender graphs.

To simulate user sessions, we use a random walk technique on the graph. We
assume that the edges of a recommender graph indeed connect products that a user
would co-view. Thus traversing the edges of a graph is a reasonable simulation
of a user session. For details on these random walks, see the section titled
Random Walks below.

With the simulated user sessions, we can implement LSI on each of the catalogues
to learn the latent concepts for each product. We want to find the products from
one catalogue that are most similar in latent concept space to products in the
other catalogue. However, since the user sessions are disjoint, we must learn
different latent spaces for each catalogue. The problem is therefore reduced to
finding a mapping from one concept space to the other. If this can be done, we
can map the concepts of each product into concepts of the other catalogue and
then find its nearest neighbors in that catalogue's concept space.

To map concepts from one catalogue to another, we must revisit
content-similarity. In this case, we represent each concept as a vector of
words. To calculate the value for a word w for concept c, we sum the tf-idf
value of w across all products, weighted by how much that product is represented
by concept c. Thus if a concept could be described as formal, the words
``sophisticated'' or ``official'' may have high value since products with those
descriptive words would also have a high amount of the formal concept in their
concept representations. Using these vector representation for concepts, we can
simply calculate the dot product between two concepts to find how similar they
are, in a relative sense. This can be used to create a mapping from one set of
concepts to another.

In all, the LSI approach consists of four phases. First we simulate user
sessions by randomly walking the recommender graphs. Second we learn concept
spaces for products in each catalogue via SVD. Third we find a mapping between
concept spaces. And lastly we find the nearest neighbors of each product in the
other catalogue's concept space. These nearest neighbors are the predicted
recommendations between products in different catalogues.

\section*{Implementation}
Here are some of the details for implementing the LSI method for constructing
inter-catalogue recommendations.

\subsection*{Random Walks}
Topic-specific random walk starting from each product
Parameters needed and the ones chosen

\subsection*{Singular Value Decomposition}
Subtract mean before hand (why this is good here but not usually)
Number of concepts to use

\subsection*{Concept Mapping}
Issue of normalization - normalize each column (L1 and L2), orthogonalize, map
to sphere
Bigrams

\subsection*{Popularity}
Important - definition, popularity breakdown of graphs
Model of popularity and similarity
Removing popularity from random walks
Adding popularity to nearest neighbors / modifying number of predictions
Baseline popularity methods
Number of products mapped to in different methods

\section*{Performance}
\subsection*{Baselines}
Random, tfidf, one model (explain), most popular

\subsection*{Results}
\begin{center}
\begin{tabular}{ | l | c | c | c |}
\hline
Normalization & 2-Precision & 2-Recall & 2-F1 Score \\ \hline\hline
L1 normalize each column of matrix &&&\\ \hline
L2 normalize each column of matrix &&&\\ \hline
Orthogonalize matrix &&&\\ \hline
Normalize products in concept space &&&\\ \hline
\end{tabular}
\end{center}

\begin{center}
\begin{tabular}{ | l | c | c | c |}
\hline
Method & 2-Precision & 2-Recall & 2-F1 Score \\ \hline\hline
Random &&&\\ \hline
Simple Content-based &&&\\ \hline
Neighborhood Content-based &&&\\ \hline
One Model &&&\\ \hline
Popularity Baselines &&&\\ \hline
\hspace*{0.5cm} Map to most popular &&&\\ \hline
\hspace*{0.5cm} Random concept mapping &&&\\ \hline
LSI &&&\\ \hline
\hspace*{0.5cm} + Normalized concept mapping &&&\\ \hline
\hspace*{0.5cm} + Zero-mean SVD &&&\\ \hline
\hspace*{0.5cm} + Bigrams from descriptions &&&\\ \hline
\hspace*{0.5cm} + Discounted popularity from random walks &&&\\ \hline
\hspace*{0.5cm} + Popularity for finding nearest neighbors &&&\\ \hline
\hspace*{0.5cm} + Number of recommendations based on popularity &&&\\ \hline

\end{tabular}
\end{center}

\begin{center}
\begin{tabular}{ | l | c | c | c | c | c | c |}
\hline
Category & 1-Precision & 1-Recall & 1-F1 Score & 2-Precision & 2-Recall & 2-F1 Score \\ \hline\hline
Men's Shirts &&&&&&\\ \hline
Women's Dresses &&&&&&\\ \hline
\end{tabular}
\end{center}

Analysis/descriptions for each of these tables

\section*{Conclusion}
We tried a few different methods for creating recommendations between
retailers. We find that this one is the best.
The last method uses an unconventional approach that adds some value in the
process. These results could be used in real life for these problems, since all
of the data exists.

\section*{Future Work}
Incorporate more data such as customer reviews. Use computer vision techniques.
Focus more on popular hubs model.

\begin{thebibliography}{9}

\end{thebibliography}

\end{document}

### 2014-06-10 common/batchExperiment.py
Add batchExperiment.py, CSV record to evalPredictedEdges.py, and minor related changes to runExperiment.sh.

@@ -1,0 +1,113 @@
+#!/usr/bin/env python
+
+"""
+Execute runExperiment.sh multiple times with various options and arguments.
+"""
+
+from optparse import OptionParser
+import os
+import sys
+import sqlite3
+from subprocess import call
+
+# local moduls
+from Util import getAndCheckFilename
+
+# params
+command = '../common/runExperiment.sh'
+
+# db params
+createProcessedCategoriesTableStmt =\
+    ('CREATE TABLE IF NOT EXISTS ProcessedCategories(ParentCategory TEXT, '
+     'Category TEXT, Processed INT, PRIMARY KEY (ParentCategory, Category))')
+createBadCategoriesTableStmt =\
+    ('CREATE TABLE IF NOT EXISTS BadCategories(ParentCategory TEXT, '
+     'Category TEXT, PRIMARY KEY (ParentCategory, Category))')
+selectProcessedCategoriesCountStmt =\
+    ('SELECT COUNT(*) '
+     'FROM ProcessedCategories')
+insertProcessedCategoriesStmt =\
+    ('INSERT INTO ProcessedCategories(ParentCategory, Category, Processed) '
+     'SELECT ParentCategory, Category, 0 '
+     'FROM Categories '
+     'GROUP BY ParentCategory, Category')
+selectProcessedCategoriesStmt =\
+    ('SELECT PC.ParentCategory, PC.Category '
+     'FROM ProcessedCategories AS PC '
+     'WHERE PC.Processed = 0 '
+     'AND NOT EXISTS '
+     '(SELECT * '
+     ' FROM BadCategories '
+     ' WHERE ParentCategory = PC.ParentCategory '
+     ' AND Category = PC.Category)')
+updateProcessedCategoryStmt =\
+    ('UPDATE ProcessedCategories '
+     'SET Processed = 1 '
+     'WHERE ParentCategory = :ParentCategory '
+     'AND Category = :Category')
+
+def getParser(usage=None):
+    parser = OptionParser(usage=usage)
+    parser.add_option('--options', dest='options', default=None,
+        help='Space-separated options for runExperiment.sh.', metavar='STR')
+    return parser
+
+def main():
+    # Parse options
+    usage = 'Usage: %prog [options] database sequence'
+    parser = getParser(usage=usage)
+    (options, args) = parser.parse_args()
+    if len(args) != 2:
+        parser.error('Wrong number of arguments')
+    dbname = getAndCheckFilename(args[0])
+    try:
[…]

### 2014-06-10 common/evalPredictedEdges.py

@@ -10,6 +10,7 @@ import os
 import sys
 import math
 import numpy as np
+import csv
 
 # local modules
 from Util import loadPickle, getAndCheckFilename
@@ -20,6 +21,10 @@ def getParser(usage=None):
         help='Distance away in original graph to consider correct')
     parser.add_option('--directed', action='store_true', dest='directed',
         default=False, help='Treat graphs as directed.')
+    parser.add_option('--data', dest='data', default=None,
+        help='Comma-separated data to include in CSV record.', metavar='CSV')
+    parser.add_option('--append-file', dest='appendFile', default=None,
+        help='Append CSV record to file.', metavar='FILE')
     return parser
 
 def getMatrix(filename):
@@ -82,24 +87,34 @@ def main():
                                sourceDict, options.k)
 
     # get results
-    numCorrectPredictions = len(predictedEdges)
+    totalPredictions = len(predictedEdges)
     numLostEdges = len(lostEdges)
     if not options.directed:
         recalledEdges /= 2
         numLostEdges /= 2
 
     # compute metrics
-    precision = float(correctPredictions) / numCorrectPredictions
+    precision = float(correctPredictions) / totalPredictions
     recall = float(recalledEdges) / numLostEdges
     if precision == 0.0 and recall == 0.0:
         f1score = 0.0
     else:
         f1score = 2*precision*recall/(precision + recall)
 
+    # append CSV record to file
+    if options.appendFile is not None:
+        if options.data is not None:
+            row = options.data.split(',')
+        row += [options.k, correctPredictions, totalPredictions, recalledEdges,
+                numLostEdges]
+        with open(options.appendFile, 'a') as csvfile:
+            writer = csv.writer(csvfile)
+            writer.writerow(row)
+
     # display results
     print '==============================================='
     print 'Correct predictions \t : %d' % correctPredictions
-    print 'Total predictions \t : %d' % len(predictedEdges)
+    print 'Total predictions \t : %d' % totalPredictions
     print '%d-Precision \t\t : %0.3f' % (options.k, precision)
     print 'Recalled Edges (k=%d)\t : %d' % (options.k, recalledEdges)
     print 'Withheld edges \t\t : %d' % numLostEdges

### 2014-06-10 common/runExperiment.sh

@@ -333,15 +333,22 @@ SEED_OPT="--seed=$SEED"
 ######
 # Setup environment
 ##############
-CMDTERM='-----'
 SRC=../common
-SEED_EXT=Seed${SEED}
-EXPMT=${MODEL_TYPE}_${NUM_TOPICS}_${CAT// /_}_$SEED_EXT
 DB=$DATA/macys.db
 STOPWORDS=$DATA/stopwords.txt
+SEED_EXT=Seed${SEED}
+CMDTERM='-----'
+
+# escape/replace special characters in category
+ESC_CAT=${PARENTCAT}_${CAT}
+ESC_CAT=${ESC_CAT// /_}
+ESC_CAT=${ESC_CAT//,/}
+ESC_CAT=${ESC_CAT//&/AND}
+
+EXPMT=${MODEL_TYPE}_${NUM_TOPICS}_${ESC_CAT}_$SEED_EXT
 
 # graphs
-GRAPH_BASE=$DATA/graph${CAT// /_}
+GRAPH_BASE=$DATA/graph${ESC_CAT}
 GRAPH=${GRAPH_BASE}.pickle
 GRAPH1=${GRAPH_BASE}${SEED_EXT}_1.pickle
 GRAPH2=${GRAPH_BASE}${SEED_EXT}_2.pickle
@@ -354,7 +361,7 @@ ONE_GRAPH=${GRAPH_BASE}${SEED_EXT}_one.pickle
 SOURCE_GRAPH=${GRAPH_BASE}${SEED_EXT}_${NUM_TOPICS}_src.pickle
 
 # random walks
-RWALK_BASE=$DATA/randomWalk${CAT// /_}
+RWALK_BASE=$DATA/randomWalk${ESC_CAT}
 RWALK=${RWALK_BASE}.npz
 RWALK1=${RWALK_BASE}${SEED_EXT}_1.npz
 RWALK2=${RWALK_BASE}${SEED_EXT}_2.npz
@@ -390,7 +397,7 @@ else
 fi
 
 # TF-IDF and maps
-TFIDF=$DATA/tfidf${CAT// /_}.pickle
+TFIDF=$DATA/tfidf${ESC_CAT}.pickle
 MAP=$DATA/topicMap_${EXPMT}.pickle
 IDENT_MAP=$DATA/identMap_${NUM_TOPICS}.pickle
 RAND_MAP=$DATA/randMap_${NUM_TOPICS}.pickle
[…]

### 2014-06-10 paper2/paper.tex

@@
 First, we scrape all of the clothing products from Macy's website at 
-www.macys.com. For each product we get its image, what category of clothing it
+www.macys.com. For each product we save its image, what category of clothing it
 is in, its description, and a list of other products recommended for it.
 
-Next, we simulate the catalogues of two different retailers by partitioning all
-of the products into two disjoint collections of products. To make this
-partition realistic, we ensure that no brand has products in both catalogue. We
-remove all recommendations between products from one partition to the other,
-thus creating two entirely disjoint and unconnected catalogues.
+Next, we partition all of the Macy's products into two disjoint sets to simulate 
+the catalogues of two different retailers. To make this partition realistic, 
+we ensure that no brand has products in both catalogue. We remove all 
+recommendations between products from one partition to the other, thus creating 
+two entirely disjoint and unconnected catalogues.
 
-Now, for each product in each catalogue, we choose which products to recommend
-in the other catalogue, using the recommendation method to be tested. To make
-this problem slightly simpler, we restrict recommendations to be within the same
-category. In this way, connecting up entire catalogues becomes a series of
-smaller tasks of connecting the products in each category.
+Now, we can use this partition to test a method for constructing recommendation.
+For each product in each of the two catalogues, we use our recommender to choose 
+which products to recommend in the other catalogue. To make this problem slightly
+simpler, we restrict recommendations to be within the same category, such as
+Women's Dresses or Men's Shoes. In this way, connecting up entire catalogues 
+becomes a series of smaller tasks of connecting the products in each category.
 
-We still have all of the true inter-catalogue recommendations from the
-unpartitioned products. Although we don't use this information for constructing
-recommendations between the partitioned graphs, we can use the original
-recommendations as a gold-standard to test the construct ones. Using these
-gold-standard recommendations, we can evaluate the precision and recall of the
-predicted inter-catalogue recommendations.
+For this experiment, we have access to all of the true inter-catalogue 
+recommendations from the unpartitioned Macy's catalogue. Although we don't
+use this information for constructing recommendations between the partitioned 
+graphs, we can use the original recommendations as a gold-standard to test the 
+construct ones. Using these gold-standard recommendations, we can evaluate 
+the precision and recall of the predicted inter-catalogue recommendations.
 
@@
 on average, there are only two gold-standard inter-catalogue recommendations
-from each product. However, since there potentially are other good
+from each product. Since there potentially are other good
 recommendations besides these two, we create another metric that expands our
-notion of a correct recommendation. Here we look at recommendations of
-recommendations. For example, if Product B is recommended for Product A and
-Product C is recommended for Product B, we say that Product C is a good
-recommendation for Product A. This leads us to the notion of 2-precision and
+notion of a correct recommendation. Here we extend the definition of a good
+connection to include recommendations of recommendations in the original
+catalgoue. For example, if Product B is recommended for Product A and 
+Product C is recommended for Product B, we say that Product C is still a good 
+recommendation for Product A. This leads us to the notions of 2-precision and 
 2-recall.
@@
 between A and C is considered correct and increases precision. 2-recall is
-defined analogously. Here, we consider the new recommender graph consisting of
-the two partitioned recommender graphs connected by the predicted edges.
-2-recall is then the percentage of gold-standard edges that are within two steps
-in the predicted recommender graph.
+defined analogously. Here, we consider the new recommender graph, consisting of
+the two partitioned recommender graphs connected by the predicted
+recommendations. 2-recall is then the percentage of gold-standard edges that 
+are within two steps in the predicted recommender graph.
 
@@
 content. The products have both a short and long description which we can use to
-construct a content-based similarity function. To do this, we represent each
+construct a textual content-based similarity function. To do this, we represent each
 product as a vector of tf-idf values for each word in its descriptions. By using
-tf-idf, we reduce the relevance of frequent words and boost that of rare ones.
-The vectors are the length of the dictionary, with zeros for all absent words
-from the descriptions. To determine the content-based similarity between two
-products, we simply take the dot product of the two tf-idf vectors.
+tf-idf, we reduce the relevance of frequent words and boost that of rare ones,
+thus emphasizing the more words that are more descriptive of a product versus
+all the other ones in the catalogue. The vector representations are the length 
+of the dictionary, with zeros for all words not in the descriptions. To determine
+the content-based similarity between two products, we simply take the dot 
+product of the two tf-idf vectors.
+
+Mathematical definition of content-based similairty...
 
@@
 impact of the quadratic effect. For instance, Macy's entire catalogue consists
-of over one million products, making a quadratic time algorithm infeasible.
+of over ONE MILLION products, making a quadratic time algorithm infeasible.
 However, the average category has 2000 products, which is significantly more
@@
 more subtle than simply the most similar other product. For instance, a striped
-black-and-white button up shirt may be an excellent recommendation for a blue
+black-and-white button-up shirt may be an excellent recommendation for a blue
 checkered button-up, even though the two have ostensibly different patterns and
@@
 
-where...
+NeighborhoodSimilarity definition...
 
@@
 missing due to different vocabularies. Furthermore, the recommender graphs
-capture some of the user behavior not expressed in product descriptions. Looking
-back on the dress-shirts example, the neighborhood of a striped shirt may
+capture some of the user behavior not expressed in product descriptions.
+Re-examining the button-up shirts example, the neighborhood of a striped shirt may
 include checkered shirts and vise-versa. In this case, the two neighborhoods
 would share the words ``striped'' and ``checkered'' and thus the two previously
-dissimilar shirts would be considered similar.
+dissimilar shirts would now be considered similar.
 
@@
 can be intangible and difficult to define. However, if we assume that the given
-recommendations are based on these concepts, we can attempt to learn the latent
-concepts by reverse-engineering the graphs.
+recommendations are truly based on these underlying concepts, we can attempt to 
+learn the latent concepts by analyzing the graphs.
 
 One approach to learning latent concepts is Latent Semantic Indexing or LSI. In
-traditional LSI, a dataset consists of documents, each containing terms.
-Singular Value Decomposition (SVD) is applied to the term-document matrix and
-the resulting three matrices are truncated to the number of latent concepts
-desired. The resulting left matrix is a term-concept matrix, where each row is a
-term represented in concept space.
+traditional LSI, used for Natural Language Processing, concepts are learned over 
+a dataset consists of documents, each containing terms. Singular Value 
+Decomposition (SVD) is applied to the term-document matrix and the resulting 
+three matrices are truncated to the number of latent concepts desired. The 
+resulting left matrix is a term-concept matrix, where each row is a term 
+represented in concept space.
 
@@
 of a user session. For details on these random walks, see the section titled
-Random Walks below.
+Random Walks in the Implementation section below.
 
 With the simulated user sessions, we can implement LSI on each of the catalogues
-to learn the latent concepts for each product. We want to find the products from
+and learn the latent concepts for each product. We want to find the products from
 one catalogue that are most similar in latent concept space to products in the
-other catalogue. However, since the user sessions are disjoint, we must learn
-different latent spaces for each catalogue. The problem is therefore reduced to
-finding a mapping from one concept space to the other. If this can be done, we
-can map the concepts of each product into concepts of the other catalogue and
-then find its nearest neighbors in that catalogue's concept space.
+other catalogue. However, since the user sessions (random walks) include
+products either from one catalogue or the other, we learn different latent 
+spaces for each catalogue. The problem is therefore reduced to finding a mapping
+from one concept space to the other. If this can be accomplished, we
+can map the concepts of each product into concepts of the other catalogue. This
+creates a concept representation in the other catalouge's concept space, meaning
+that we can then find nearest neighbors in that catalogue's concept space.
 

### 2014-06-10 common/runExperiment.sh
Add TFIDF NOPOP benchmark, ability to run individual benchmarks, and macys/BadCategories.dat file.

@@ -50,6 +50,13 @@ while test $# -gt 0; do
             echo "--no-weight-in            Exclude popularity from KNN search"
             echo "--no-weight-out           Predict outgoing edges uniformaly"
             echo "--no-benchmarks           Don't run benchmark predictors"
+            echo "--benchmark-random        Only do random benchmark"
+            echo "--benchmark-popular       Only do popular benchmark"
+            echo "--benchmark-tfidf-nopop   Only do TF-IDF NoPop benchmark"
+            echo "--benchmark-tfidf         Only do TF-IDF benchmark"
+            echo "--benchmark-randommap     Only do random-map benchmark"
+            echo "--benchmark-onemodel      Only do OneModel benchmark"
+            echo "--no-main                 Don't do main experiment"
             echo "--no-popularity-added     Don't add popularity back into graph"
             echo "--no-popularity-removed   Don't remove popularity in graph traversal"
             echo "--lda (BROKEN!)           Use lda instead of lsi for latent space"
@@ -167,7 +174,70 @@ while test $# -gt 0; do
             shift
             ;;
         --no-benchmarks)
-            export BENCHMARKS=0
+            export BENCHMARK_RANDOM=0
+            export BENCHMARK_POPULAR=0
+            export BENCHMARK_TFIDF_NOPOP=0
+            export BENCHMARK_TFIDF=0
+            export BENCHMARK_RANDMAP=0
+            export BENCHMARK_ONE=0
+            shift
+            ;;
+        --benchmark-random)
+            export BENCHMARK_RANDOM=1
+            export BENCHMARK_POPULAR=0
+            export BENCHMARK_TFIDF_NOPOP=0
+            export BENCHMARK_TFIDF=0
[…]

### 2014-06-10 paper2/paper.tex
Fill out Implementation section

@@
 \section*{Implementation}
-Here are some of the details for implementing the LSI method for constructing
-inter-catalogue recommendations.
+Let us dig deeper into the details of the LSI approach described above. There 
+are a number of implementation decisions that, in all, dramatically affect the
+performance of this method.
 
 \subsection*{Random Walks}
-Topic-specific random walk starting from each product
-Parameters needed and the ones chosen
+The first step of the LSI approach is to simulate user sessions by randomly
+walking the recommender graphs. There are a few ways this could be implemented.
+One method is to randomly assign a random walker a starting node and then allow
+it to randomly walk the edges, with a probability of termination at each step.
+Running this many times is a straight-forward model of user sessions.
+
+To speed up the simulation, we calculate a series of product-specific stationary
+distributions. As with any random walk technique, we first interpret the 
+recommender graph as a transition matrix. In this walk, for each product we create a
+product-specific transition matrix which has added teleportation edges back the
+that product. We calculate the stationary distributions of each of these
+matrices, one per product. We use these product-specific stationary
+distributions as our user sessions. Since there is only one user session per 
+product, we have many fewer sessions than we would with a real dataset. However,
+each of these stationary distributions can be interpreted as an aggregation of
+an infinite number of random walkers starting at each node and travelling until
+a random termination time, as in the original model.
+
+This process requires a couple of parameters. The first is the number of steps used to
+approximate each product-specific statinary distribution. The second parameter
+is the likelihood of teleporting to the special product. For our purposes we use
+30 steps and a teleportation probability of 5\%.
 
 \subsection*{Singular Value Decomposition}
-Subtract mean before hand (why this is good here but not usually)
-Number of concepts to use
+Unlike Principal Compenent Analysis, standard LSI does not include any
+preprocessing of the data, such as subtracting the mean from each column of the
+product-session matrix. A major reason for this is that LSI is typically applied
+to natural language processing problems, where the rows are terms, and there is
+one term per word in the vocabulary. This otherwise sparse matrix becomes dense
+if the mean is subtracted, making many NLP datasets untractable to work with. 
+Our datast, on the other hand, has significantly fewer rows and columns and is
+already dense. Therefore, there is little added cost to subtracting the
+means. We include this preprocessing step and find that we get better results
+because of it.
+
+Another design decision is in the size of the truncation of the matrices learned
+from SVD. This number dictates the number of latent concepts learned for LSI.
+Too few concepts fail to capture enough detail to make accurate representations
+of each product. Too many make mapping between concepts challenging. We found a
+good balance at 16 concepts.
+
+[Graphs of concepts]
 
 \subsection*{Concept Mapping}
-Issue of normalization - normalize each column (L1 and L2), orthogonalize, map
-to sphere
-Bigrams
+A tricky operation of this procedure is the concept mapping. We attempt to find
+the best linear transformation from one concept space to the other. We do this
+by representing each concept as a vector of words in the vocabulary, each
+weighted with how well it represents the concept. We construct this by weighing
+the tfidf of each word in a product's description by how well that word
+represents the concept. Note that products can have a positive or negative value
+for each concept, thus the word representation of concpets are also positive and
+negative.
+
+With these concept representations, we construct an innner-product matrix. The 
+value of a cell (i, j) in the matrix is the dot product of concept i in the first 
+space with concept j in the second. The columns and rows of the resulting matrix
+represent the relative similarity of the descriptive words between two concepts.
+But the scale of this matrix can be very large. If the inner-product matrix is
+used without any normalization as the linear transformation between the concept
+spaces, the scale of the transformed products will be inconsistent with those
+originally in that space.
+
+One way to normalize is to normalize each column of the inner-product matrix by
+the L1 norm. This preserves the L1 norm of products after the transformation,
+but does not preserve the L2 norm. Similarly, we can normalize the columns by
+the L2 norm. Neither of these methods preserve the L2 norm, which would be ideal
+as we find the nearest neighbors in the other concept space by their L2
+distance.
+
+Another approach is to orthogonalize the inner-product matrix. That is, find the
+orthogonal matrix nearest to the inner-product matrix. Nearness is defined by
+the NORM FUNCTION on the difference of the two matrices. This orthogonal matrix
+does perserve the L2 norm of the vectors it transforms, however the
+transformation is an approximation of the ideal inner-product matrix and can
+distort the relative mapping from one concept to the others.
+
+A third approach to normalizing aims to preserve the angle of the inner-product
+transformation. For this approach, we remove the issue of scale by normalizing
+all products in the mapped-to space to be unit lenght. Then, after transforming
+the products from the other catalogue, we normmalize them as well. Thus the
+scales for both sets of products are the same and the angles of the
+inner-product transformation are preserved, at the expense of discarding the
+norms of all products in their product spaces. This is equivalent to calculating
+the cosine similarity between products without any normalization, as cosine
+similarity is indeifferent to scale. We find that this method of normalization
+performs the best.
+
+Further gains can be seen if we consider bigrams in the description as well as
+the unigrams. Here, we treat bigrams as a separate set dimensions to represent
+each concept. These are included in the inner product between two concepts,
+weighted compared to the unigrams by a parameter $\beta$. $\beta = 0.3$ improves the
+performance of the concept mapping.
 
 \subsection*{Popularity}
-Important - definition, popularity breakdown of graphs
-Model of popularity and similarity
-Removing popularity from random walks
-Adding popularity to nearest neighbors / modifying number of predictions
-Baseline popularity methods
-Number of products mapped to in different methods
+The most effective addition to the LSI procedure is incorporating the popularity
+of products. We do not know a priori how popular an item is as we do not have
+explicit user data. However, we can use the recommendation graphs to model the
+popularity of each item. The assumption is that a retailer will recommend
+popular products more frequently than unpopular ones.
+
+We imagine that recommender graphs are constructed as follows: A retailer first
+builds a similarity matrix, using a collaborative filtering method. In this
+matrix, all similarity relationships are symmetric. Next, the retailer modifies
+the values of the matrix to include popularity. As a simple model, we say that
+the popularity-modified similarity score from Product i to Proudct j is the
+simple similarity score at (i, j) multiplied by the popularity of Product j. In
+this way, more popular items are more frequently recommended.
+
+We see signs of this in the undirected recommender graphs. Although each product
+has at most four outgoing recommendations on macys.com, we see that some
+products have over one hundred incoming recommendations. With this in mind, we
+approximate the popularity of each product as the logarithm of the degree of
+that product in its undirected recommender graph. We choose to model it as the
+logarithm because a few of the products tend to have degrees in the hundreds.
+Without some way of scaling back these popularities, these few very popular
+products tend to take over all recommendations.
+
+[Graph of popularity]
+
+Assuming this model of recommender graph construction, which separates similarity
+from popularity, we seek to isolate similarity from the recommender graph. Since
+popular products were promoted for popularity, we demote them when walking the
+graphs to create user sessions. In practice, we divide the likelihood of walking
+an edge by the popularity of the product on the other side of that edge. Thus
+the concepts learned on that matrix better represent products independent of
+popularity.
+
+Proceeding as before, we find the nearest products in concept space. However,
+before deeming the nearest neighbors to be the best recommendations, we divide
+the distance to each product by its popularity. Since we model distance in 
+concept space to be inversely proportional to similarity, this reincorporation
+of popularity is analagous to multiplying similarity by popularity, as done in
+the original construction of the recommender graph.
+
+A final way in which we incorporate popularity is in the number of
+recommendations alotted per product. Since we model the recommender graph as an
+undirected graph, different nodes have different degrees. A product that is
+popular within its catalogue is likely to be more popular across catalogues as
+well. Therefore, we allow popular products to have more recommendations across
+the two catalogues, proportional to their popularity.
+
+In all, these popularity additions significantly boost performance in the
+experiment. The results of this, along with the other modifications outlined in
+this section, can be seen in the following section on performance.
 
@@
 \subsection*{Baselines}
+As in product-product content-similarity, it is difficult to capture similarity using
+descriptions alone. It is not even clear that a mapping necessarily exists
+between two seperately learned concept spaces. The concepts may represent
+fundamentally different things, which cannot be well mapped to each other. We
+can measure the effectiveness of the mapping by comparing it to a couple of
+baselines. The upper limit = One Model, lower limit = random mapping.
+Indeed, we find that this is a bottle neck of the operation.
+
+Baseline popularity methods
+[Number of products mapped to in different methods]
+
 Random, tfidf, one model (explain), most popular
@@
 
+\section*{Appendix}
+
 \begin{thebibliography}{9}

### 2014-06-10 common/predictEdges.py
Incorporate popularity into the --tau option of predictEdges.py.

@@ -146,8 +146,6 @@ def main():
             [item for idx, item in sorted(dictionary2.items(),
                                           key=lambda tup: tup[0])]
         simCache2.preComputeSims(transformedItems1, items2)
-        # scale similarities by options.tau ceofficient
-        simCache2.sims *= options.tau
         if options.symmetric:
             simCache1 = TFIDF_SimilarityCache(
                 tfidfs,
@@ -163,8 +161,6 @@ def main():
                 [item for idx, item in sorted(dictionary1.items(),
                                               key=lambda tup: tup[0])]
             simCache1.preComputeSims(transformedItems2, items1)
-            # scale similarities by options.tau ceofficient
-            simCache1.sims *= options.tau
 
     # make predictions
     if popDictionary is not None and options.weightOut:
@@ -190,7 +186,8 @@ def main():
                 # create addTerms dictionary
                 if tfidfs is not None and options.tau is not None:
                     addTerms = dict(
-                        [(dictionary2[idx], sim)\
+                        [(dictionary2[idx],
+                          options.tau*popDictionary[dictionary2[idx]]*sim)\
                          for (idx, sim) in enumerate(simCache2.sims[i,:])]
                     )
                 else:
@@ -217,7 +214,8 @@ def main():
                     # create addTerms dictionary
                     if tfidfs is not None and options.tau is not None:
                         addTerms = dict(
-                            [(dictionary1[idx], sim)\
+                            [(dictionary1[idx],
+                              options.tau*popDictionary[dictionary1[idx]]*sim)\
                              for (idx, sim) in enumerate(simCache1.sims[i,:])]
                         )
                     else:

### 2014-06-10 paper2/paper.tex
Add Baseline section to paper

@@
 \subsection*{Baselines}
-As in product-product content-similarity, it is difficult to capture similarity using
-descriptions alone. It is not even clear that a mapping necessarily exists
-between two seperately learned concept spaces. The concepts may represent
-fundamentally different things, which cannot be well mapped to each other. We
-can measure the effectiveness of the mapping by comparing it to a couple of
-baselines. The upper limit = One Model, lower limit = random mapping.
-Indeed, we find that this is a bottle neck of the operation.
+To understand how well our LSI recommender performs, we need good baselines
+for comparison.  The simplest and easiest to beat baseline is the Random 
+Recommender. This recommender randomly picks pairs of products, one from 
+each catalogue, and recommends them together. This should obviously be the 
+worst performing recommender but it gives us some sense of whether we are 
+doing anything intelligent in our choices of recommendation.
 
-Baseline popularity methods
-[Number of products mapped to in different methods]
+A more impressive baseline is that of the Simple Content-Based approach, 
+outlined earlier in the paper. This recommender uses information exclusively
+from the descriptions of the products and ignores the recommender graphs 
+entirely. A third, somewhat similar baseline comes from the Neighborhood 
+Content-based variant, which looks at the content similarity between 
+neighborhoods of products.
 
-Random, tfidf, one model (explain), most popular
+We construct a couple of other baselines to show that our performance is not
+exclusively due to any single feature of the recommender. One such baseline
+tests the effects of popularity alone. Incorporating popularity has a large 
+impact on the performance of any recommender and part of this is due simply 
+to the evaluation metric. Any time we connect to a popular product, we are 
+now two away from many more products, increasing the likelihood of getting a
+correct recommendation by the 2-precision standard. To ensure that these kinds 
+of effects are not the sole reason that our recommender performs well, 
+we construct the Popularity Recommender. This recommender simply recommends 
+that every product in one catalogue be connected to the two most popular 
+products from the other catalogue. This baseline shows how much of the LSI 
+recommender's performance is due simply to leveraging popularity and how 
+much is due to other factors, such as intelligently learning and mapping 
+latent concepts.
+
+It is also informative to test the efficacy of the concept mapping, since this
+such a hard task. As in product-product content-similarity, it is difficult to
+capture similarity using descriptions alone. It is not even clear that a mapping 
+necessarily exists between two seperately learned concept spaces. The concepts 
+may represent fundamentally different things, which cannot be well mapped to 
+each other. We can measure the efficacy of the mapping by comparing it to a 
+couple of baselines.
+
+The first baseline is in fact an upper bound of sorts on the mapping. This
+baseline recommender is called the One Model Recommender. Unlike the other
+baselines, this recommender is not a valid one, as it cheats by using
+information from the unpartitioned graph. The One Model Recommender learns a
+latent concept space over the entire recommender graph by allowing the random
+walkers to traverse it before it is partitioned. In every other respect, it is
+the same as the LSI Recommender. However, since products from both catalogues
+are represented in the same concept space, there is no need for a mapping
+between concept spaces.
+
+The second mapping-related baseline is the Random Mapper Recommender. As the
+name suggests, this recommender is identical to the LSI Recommender, except that
+it uses a randomly generated linear transformation to map between concept
+spaces. These two baselines represent (not strict) upper and lower bounds on how
+well the concept mapping performs. The results shows how much of a bottleneck
+this step is.
 
 \subsection*{Results}
+[Number of products mapped to in different methods]
 \begin{center}

### 2014-06-10 paper2/paper.tex
Added Data section and re-worked part of Experiment section.

--- 2014-06-10 paper2/paper.tex
@@
 Recommender systems have become a critical component for online retailers in
-nearly every space of retail. Whether the products are movies or clothes or research
-articles, large retailer tend to have a way of recommending products based on
-other previously viewed or purchased products.
+nearly every space of retail. Whether the products are movies or clothes or
+research articles, large retailers tend to have a way of recommending products
+based on others previously viewed or purchased products.
 
@@
 information and collaborative filtering techniques. The content-based approach
-uses information about the products, such as category or descriptioin, to
-connect related products. Alone, however, this approach is usually unable to
-capture the subtle ways products may actually be related. Collaborative
-filtering, on the other hand, learns relations from aggregated user behavior. 
-Such recommendation often have explicit ttiles like ``People who liked 
-this product also liked these products''. Collaborative filtering has proven to
-be effective at recommending products, given that there is a large amount of user 
-data.
+uses information about the products, such as category or description, to connect
+related products. Alone, however, this approach is usually unable to capture the
+subtle ways products may actually be related. Collaborative filtering, on the
+other hand, learns relations from aggregated user behavior.  Such
+recommendations often have explicit titles like ``People who liked this product
+also liked these products''. Collaborative filtering has proven to be effective
+at recommending products, given that there is a large amount of user data.
 
 Since each retailer only has user data within their site, these user-based
-recommender systems are good at learning connections between products in
-one catalogue, but they do little to learn how these products may relate to
-other products outside of the retailer's inventory. This type of information is
-valuable in a number of situations. Retailers such as department stores
-want to know which new products make the most sense to add to their catalogue.
-This is essentially a question of recommending the best outside products based
-on current in-house ones. In a different scenario, two retailers may merge and 
-want to construct a unified recommender system between the two catalgoues.
+recommender systems are good at learning connections between products in one
+catalogue, but they do little to learn how these products may relate to other
+products outside of the retailer's inventory. This type of information is
+valuable in a number of situations. Retailers such as department stores want to
+know which new products make the most sense to add to their catalogue.  This is
+essentially a question of recommending the best outside products based on
+current in-house ones. In a different scenario, two retailers may merge and want
+to construct a unified recommender system between the two catalgoues.
 
@@
 
-\section*{Experiment}
-To measure the success of constructing recommendations, we design the following
-experiment:
+\section*{Data}
+For this work, we utilize most of the product catalog of Macy's as presented at
+www.macys.com. This catalog consists of numerous categories of mostly apparel,
+each of which includes hundreds to thousands of items. We confine this study to
+the 49 categories within the two main parent categories ``Men'' and ``Women''
+that contain at least 200 items that are associated to multiple brands. In
+total, this dataset consists of 66,071 items in 49 categories with some items
+listed in multiple categories. For each item we record its description, all
+categories within which it is listed, and its associated recommendations. By
+{\em recommendations}, we are referring to the items---there are typically four
+in the case of Macy's---that are displayed on a product's details page under the
+heading ``Customers Also Shopped''. The presence of these recommendations
+implicitly forms a directed graph over the product catalogue, in which the nodes
+are the {\em products} or {\em items} and the edges are the {\em
+recommendations}. These graphs play a central role in this study.
 
-First, we scrape all of the clothing products from Macy's website at 
-www.macys.com. For each product we save its image, what category of clothing it
-is in, its description, and a list of other products recommended for it.
+\section*{Experiments}
+Our goal is to develop methods for connecting initially disconnected
+recommendation graphs. That is, given two online retailers, each with a website
+that displays items in their respective catalogue, where each item is
+accompanied by several {\em recommended} items, we are presented with two
+disconnected recommendation graphs. We would like to associate ``good''
+recommendations for items in one catalogue with items in the other catalogue. In
+this way, we are effectively introducing edges that connect the two
+recommendation graphs. Formulating this inter-retailer recommendations problem
+in terms of graphs offers insight in how to evaluate our recommendation choices
+as well as how to choose ``good'' recommendations. Leveraging the graph
+structure for evaluation is discussed in this section while exploiting the graph
+for the purpose of making better recommendations is addressed in a subsequent
+section.
 
-Next, we partition all of the Macy's products into two disjoint sets to simulate 
-the catalogues of two different retailers. To make this partition realistic, 
-we ensure that no brand has products in both catalogue. We remove all 
+To evaluate our inter-retailer recommendation methods, we simulate the problem
+by randomly partitioning Macy's products into two disjoint sets. Given the
+original recommendation graph, each partition therefore corresponds to a graph
+cut. Our premise is that an ideal recommendation method could guess the edges in
+this cut with high precision and recall.
+
+Precision, recall, and $F_1$ score are well-known metrics for evaluating
+prediction algorithms.  Together, these metrics capture how effective one is at
+guessing the items within a target set. Consider a typical Macy's category, for
+example, Dresses, which contains about $3,500$ items. Randomly partitioning
+results in two sets, each with about $1,750$ items. For each interset
+recommendation, we must therefore choose two nodes, one from each of the two
+sets. That is, we have $1,750 \times 1,750 > 3 \times 10^6$ choices. Choosing
+the correct edges is a formidable challenge to say the least. It is therefore
+worth asking: Are some prediction errors better or worse than others? The graph
+nature of our problem reveals that the answer is yes. It is better to guess an
+edge that connects two items that are separated by two edges in the original
+graph than to connect two items that are separated by five edges. We therefore
+introduce the notion of {\em 2-precision}, {\em 2-recall}, and {\em 2-}$F_1$ are
+defined as
+
+\begin{align}
+\text{\em 2-precision} &= \left|\left\{(u,v) \in P | d_L(u,v) \leq 2
+\right\}\right|
+\\\text{\em 2-recall} &= \left|\left\{(u,v) \in L | d_P(u,v) \leq 2
+\right\}\right|
+\\\text{\em 2-}F_1 &= 2\cdot \frac{\text{\em 2-precision} \cdot \text{\em
+2-recall}}{\text{\em 2-precision} + \text{\em 2-recall}}
+\end{align}
+where $P$ is the set of predicted edges, $L$ is the set of lost edges (i.e.
+those in the cut resulting from the partition), and $d_L(u,v)$ is the distance
+between nodes $u$ and $v$ in the original unpartitioned graph, and $d_P(u,v)$ is
+the distance between nodes $u$ and $v$ in the new graph formed using the
+predicted edges in $P$.
+
+In these experiments, we perform the partitioning on a per-category basis.
+
+Next, for each category (e.g. Women Activewear, Men Dress Shirts, etc.), we
+partition the products into two disjoint sets to simulate the catalogues of two
+distinct retailers. Importantly, we ensure that the items of each brand are
+entirely within a single partition. In this way, we preclude the easiest method
+of associating two items: recongnizing a common brand. This is done to increase
+the difficulty and realism of our experiment.
+
+To make this partition realistic, we
+ensure that no brand has products in both catalogue. We remove all
 recommendations between products from one partition to the other, thus creating
@@
 For each product in each of the two catalogues, we use our recommender to choose
-which products to recommend in the other catalogue. To make this problem slightly
-simpler, we restrict recommendations to be within the same category, such as
-Women's Dresses or Men's Shoes. In this way, connecting up entire catalogues 
-becomes a series of smaller tasks of connecting the products in each category.
+which products to recommend in the other catalogue. To make this problem
+slightly simpler, we restrict recommendations to be within the same category,
+such as Women's Dresses or Men's Shoes. In this way, connecting up entire
+catalogues becomes a series of smaller tasks of connecting the products in each
+category.
 
 For this experiment, we have access to all of the true inter-catalogue
-recommendations from the unpartitioned Macy's catalogue. Although we don't
-use this information for constructing recommendations between the partitioned 
+recommendations from the unpartitioned Macy's catalogue. Although we don't use
+this information for constructing recommendations between the partitioned
 graphs, we can use the original recommendations as a gold-standard to test the
-construct ones. Using these gold-standard recommendations, we can evaluate 
-the precision and recall of the predicted inter-catalogue recommendations.
+construct ones. Using these gold-standard recommendations, we can evaluate the
+precision and recall of the predicted inter-catalogue recommendations.
 

### 2014-06-10 common/joinEveryNLines.sh
Add joinEveryNLines.sh.

@@ -1,0 +1,18 @@
+#!/bin/bash
+
+# Joins every N lines of a file.
+
+verify_number () {
+    re='^[0-9]+$'
+    if ! [[ $1 =~ $re ]] ; then
+        die "error: Not a number: $1"
+    fi
+}
+
+# get args
+N=$1
+FILE=$2
+
+verify_number $N
+
+perl -pi -e "s/\r\n/ / if $.%$N" $FILE 2> /dev/null

### 2014-06-10 paper2/paper.tex

@@
 Stanford University}
-\renewcommand{\today}{June 12, 2014}
+\renewcommand{\today}{June 11, 2014}
 \maketitle
@@
 retailers who don't have access to each other's user data. Thus, to keep the
-problem general, we do not assume that we have any private internal data.
+problem general, we do not assume that we have any private data.
 Instead, we constrain ourselves to use only the product information and internal
@@
 recommendations of each catalogue to construct inter-catalgoue recommendations
-between them. We focus on the clothing retailers, since they tend to all have
+between them. We focus on clothing retailers, since they tend to all have
 public recommendations for each product. Specifically, the data used for this
@@
 are the {\em products} or {\em items} and the edges are the {\em
-recommendations}. These graphs play a central role in this study.
+recommendations}. We call these types of graphs {\em recommendation graphs} and they play a central role in this study.
 
@@
 Our goal is to develop methods for connecting initially disconnected
-recommendation graphs. That is, given two online retailers, each with a website
-that displays items in their respective catalogue, where each item is
-accompanied by several {\em recommended} items, we are presented with two
-disconnected recommendation graphs. We would like to associate ``good''
-recommendations for items in one catalogue with items in the other catalogue. In
-this way, we are effectively introducing edges that connect the two
-recommendation graphs. Formulating this inter-retailer recommendations problem
-in terms of graphs offers insight in how to evaluate our recommendation choices
-as well as how to choose ``good'' recommendations. Leveraging the graph
-structure for evaluation is discussed in this section while exploiting the graph
-for the purpose of making better recommendations is addressed in a subsequent
-section.
+recommendation graphs. Given two online retailers, each with a website
+displaying their items along with several other {\em recommended} items from
+their catalogue, we are presented with two disconnected recommendation graphs. 
+We would like to associate ``good'' recommendations for items in one catalogue 
+with items in the other catalogue. In this way, we are effectively introducing 
+edges that connect the two recommendation graphs. Formulating this inter-retailer
+recommendations problem in terms of graphs offers insight in how to evaluate our 
+recommendation choices as well as how to choose ``good'' recommendations. 
+Leveraging the graph structure for evaluation is discussed in this section while 
+exploiting the graph for the purpose of making better recommendations is addressed 
+in a subsequent section.
 
@@
 prediction algorithms.  Together, these metrics capture how effective one is at
-guessing the items within a target set. Consider a typical Macy's category, for
-example, Dresses, which contains about $3,500$ items. Randomly partitioning
-results in two sets, each with about $1,750$ items. For each interset
-recommendation, we must therefore choose two nodes, one from each of the two
+guessing items within a target set. Consider a typical Macy's category, for
+example, Dresses, which contains approximately $3,500$ items. Randomly 
+partitioning results in two sets, each with about $1,750$ items. For each 
+interset recommendation, we must therefore choose two nodes, one from each of the two
 sets. That is, we have $1,750 \times 1,750 > 3 \times 10^6$ choices. Choosing
@@
 where $P$ is the set of predicted edges, $L$ is the set of lost edges (i.e.
-those in the cut resulting from the partition), and $d_L(u,v)$ is the distance
-between nodes $u$ and $v$ in the original unpartitioned graph, and $d_P(u,v)$ is
-the distance between nodes $u$ and $v$ in the new graph formed using the
-predicted edges in $P$.
+those in the cut resulting from the partition), $d_L(u,v)$ is the shortest
+distance between nodes $u$ and $v$ in the original unpartitioned graph, and 
+$d_P(u,v)$ is the shortest distance between nodes $u$ and $v$ in the new graph 
+formed using the predicted edges in $P$.
 
@@
 spaces. These two baselines represent (not strict) upper and lower bounds on how
-well the concept mapping performs. The results shows how much of a bottleneck
+well the concept mapping performs. The results show how much of a bottleneck
 this step is.
@@
 One Model &&&\\ \hline
-Popularity Baselines &&&\\ \hline
-\hspace*{0.5cm} Map to most popular &&&\\ \hline
-\hspace*{0.5cm} Random concept mapping &&&\\ \hline
+Random concept mapping &&&\\ \hline
+Popularity &&&\\ \hline
 LSI &&&\\ \hline

### 2014-06-10 common/joinEveryNLines.sh
Add --delim option and help to joinEveryNLines.sh.

@@ -9,10 +9,38 @@ verify_number () {
     fi
 }
 
+COMMAND=${0##*/}
+
+while test $# -gt 0; do
+    case "$1" in
+        -h|--help)
+            echo "$COMMAND - join every N lines"
+            echo
+            echo "Usage: $COMMAND [options] N file"
+            echo
+            echo "options:"
+            echo "-h, --help                Show brief help"
+            echo "--delim=STRING            Delimiter for joined lines"
+            exit 0
+            ;;
+        --delim*)
+            DELIM=`echo $1 | sed -e 's/^[^=]*=//g'`
+            shift
+            ;;
+        *)
+            break
+            ;;
+    esac
+done
+
+if [ $# -lt 2 ]; then
+    die "Usage: $COMMAND [options] N file"
+fi
+
 # get args
 N=$1
 FILE=$2
 
 verify_number $N
 
-perl -pi -e "s/\r\n/ / if $.%$N" $FILE 2> /dev/null
+perl -pi -e "s/\r\n/$DELIM/ if $.%$N" $FILE 2> /dev/null

### 2014-06-10 paper2/paper.tex
Finished re-working Experiment section and added Evaluation section.

@@
 are the {\em products} or {\em items} and the edges are the {\em
-recommendations}. We call these types of graphs {\em recommendation graphs} and they play a central role in this study.
+recommendations}. We call these types of graphs {\em recommendation graphs} and
+they play a central role in this study.
+
+To simplify the present study, we transform these directed recommendation graphs
+into undirected graphs. This is consistently done using the following policy:
+For each pair of items connected by a single directed edge, replace this edge
+with a single undirected edge. For each pair of items connected by two
+oppositely directed edges, replace both edges with a single undirected edge.
+Working with undirected graphs simplifies both the prediction algorithms as well
+as the evaluation of their performance. At the same time, by ignoring the
+directionality of the original edges, we are discarding important information.
+It would therefore be worthwhile to consider the more difficult problem of
+connecting directed recommendation graphs in a future study.
 
@@
 with items in the other catalogue. In this way, we are effectively introducing
-edges that connect the two recommendation graphs. Formulating this inter-retailer
-recommendations problem in terms of graphs offers insight in how to evaluate our 
-recommendation choices as well as how to choose ``good'' recommendations. 
-Leveraging the graph structure for evaluation is discussed in this section while 
-exploiting the graph for the purpose of making better recommendations is addressed 
-in a subsequent section.
+edges that connect the two recommendation graphs. Formulating this
+inter-retailer recommendations problem in terms of graphs offers insight in how
+to evaluate our recommendation choices as well as how to choose ``good''
+recommendations.  Leveraging the graph structure for evaluation is discussed in
+this section while exploiting the graph for the purpose of making better
+recommendations is addressed in a subsequent section.
 
@@
 original recommendation graph, each partition therefore corresponds to a graph
-cut. Our premise is that an ideal recommendation method could guess the edges in
-this cut with high precision and recall.
+cut. The edges in this graph cut form a {\em gold-standard} since it is assumed
+that these are indeed ``good'' recommendations. Our premise is that an ideal
+recommendation method could guess these gold-standard edges with high precision
+and recall.
 
+For each experiment, we choose a particular category from the Macy's catalogue
+(e.g. Women Activewear, Men Dress Shirts, etc.). Since most recommendations are
+between items in a common category, very few, though some, edges are lost by
+confining ourselves to an individual category. Edges that are lost in this way
+do not participate in the experiment and are not considered during evaluation.
+Also, we randomly partition the items in the category such that all items
+associated with a particular brand are entirely within a single partition. In
+this way, we preclude the easiest method of associating two items: recognizing
+a common brand. This is done to increase the difficulty and realism of the
+experiments.
+
+Having partitioned the graph of items, we now predict recommendation relations
+between items across the partitions. Each prediction algorithm has knowledge of
+all items in each partition, including their descriptions and all
+recommendations (i.e.  edges) between items in a common partition. In addition,
+the algorithms may exploit the fact that most items listed at www.macys.com are
+accompanied by four recommended items. The algorithms do not have knowledge of
+the recommendations (i.e. edges) between items across partitions. Each
+prediction algorithm is then free to choose an arbitrary number of edges as long
+as 1) the items connected by the predicted edge are not in the same partition
+and 2) at most one edge is predicted for a particular pair of items.
+
+\section*{Evaluation}
 Precision, recall, and $F_1$ score are well-known metrics for evaluating
-prediction algorithms.  Together, these metrics capture how effective one is at
-guessing items within a target set. Consider a typical Macy's category, for
-example, Dresses, which contains approximately $3,500$ items. Randomly 
-partitioning results in two sets, each with about $1,750$ items. For each 
-interset recommendation, we must therefore choose two nodes, one from each of the two
-sets. That is, we have $1,750 \times 1,750 > 3 \times 10^6$ choices. Choosing
-the correct edges is a formidable challenge to say the least. It is therefore
-worth asking: Are some prediction errors better or worse than others? The graph
-nature of our problem reveals that the answer is yes. It is better to guess an
-edge that connects two items that are separated by two edges in the original
-graph than to connect two items that are separated by five edges. We therefore
-introduce the notion of {\em 2-precision}, {\em 2-recall}, and {\em 2-}$F_1$ are
-defined as
+prediction algorithms \cite{Powers2011}. Together, these metrics capture how
+effective one is at guessing items within a target set. Consider a typical
+Macy's category, for example, Dresses, which contains approximately $3,500$
+items. Suppose that randomly partitioning results in two sets, each with about
+$1,750$ items. For each interset recommendation, we must therefore choose two
+nodes, one from each of the two sets. That is, we have $1,750 \times 1,750 > 3
+\times 10^6$ choices. Choosing the correct edges is a formidable challenge to
+say the least. It is therefore worth asking: Are some prediction errors better
+or worse than others?
+
+The graph nature of our problem reveals that the answer to this question is
+yes. For instance, it is better to guess an edge that connects two items that
+are separated by two edges in the original graph than to connect two items that
+are separated by five edges. We therefore introduce the notion of {\em
+2-precision}, {\em 2-recall}, and {\em 2-}$F_1$ are defined as
 
@@
 
-In these experiments, we perform the partitioning on a per-category basis.
+The following definitions will prove helpful. Let $G$ be the original
+unpartitioned graph formed from the partitions together with the lost edges $L$
+and let $G^{\prime}$ be the new graph formed from the partitions together with
+the predicted edges $P$.  In the context of {\em 2-precision}, an edge in $P$
+that connects items $A$ and $B$ is considered correct if 1) $A$ and $B$ share an
+edge within $G$ or 2) $A$ shares an edge with a direct neighbor of $B$ within
+$G$ or 3) a direct neighbor of $A$ shares an edge with $B$ within $G$.
+Meanwhile, in the context of {\em 2-recall}, an edge in $L$ that connects items
+$C$ and $D$ is considered recalled if 1) $C$ and $D$ share an edge within
+$G^{\prime}$ or 2) $C$ shares an edge with a direct neighbor of $D$ within
+$G^{\prime}$ or 3) a direct neighbor of $C$ shares an edge with $D$ within
+$G^{\prime}$.
 
-Next, for each category (e.g. Women Activewear, Men Dress Shirts, etc.), we
-partition the products into two disjoint sets to simulate the catalogues of two
-distinct retailers. Importantly, we ensure that the items of each brand are
-entirely within a single partition. In this way, we preclude the easiest method
-of associating two items: recongnizing a common brand. This is done to increase
-the difficulty and realism of our experiment.
-
-To make this partition realistic, we
-ensure that no brand has products in both catalogue. We remove all
-recommendations between products from one partition to the other, thus creating
-two entirely disjoint and unconnected catalogues.
-
-Now, we can use this partition to test a method for constructing recommendation.
-For each product in each of the two catalogues, we use our recommender to choose
-which products to recommend in the other catalogue. To make this problem
-slightly simpler, we restrict recommendations to be within the same category,
-such as Women's Dresses or Men's Shoes. In this way, connecting up entire
-catalogues becomes a series of smaller tasks of connecting the products in each
-category.
-
-For this experiment, we have access to all of the true inter-catalogue
-recommendations from the unpartitioned Macy's catalogue. Although we don't use
-this information for constructing recommendations between the partitioned
-graphs, we can use the original recommendations as a gold-standard to test the
-construct ones. Using these gold-standard recommendations, we can evaluate the
-precision and recall of the predicted inter-catalogue recommendations.
-
-For the Macy's catalogue, each product has four recommendations. This means that
-on average, there are only two gold-standard inter-catalogue recommendations
-from each product. Since there potentially are other good
-recommendations besides these two, we create another metric that expands our
-notion of a correct recommendation. Here we extend the definition of a good
-connection to include recommendations of recommendations in the original
-catalgoue. For example, if Product B is recommended for Product A and 
-Product C is recommended for Product B, we say that Product C is still a good 
-recommendation for Product A. This leads us to the notions of 2-precision and 
-2-recall.
-
-For a rigorous definition of 2-precision and 2-recall, we first propose the idea
-of a recommender graph. A recommender graph has products as nodes and
-recommendations as edges. If Product B is recommended for Product A, there is an
-edge between node A and node B in the graph. For simplicity, we define our
-recommender graphs to be undirected, but we note that they could be interpreted
-as directed graphs as well. With this terminology, our problem becomes one of
-connecting two disjoint recommender graphs by adding edges between them.
-
-2-precision is defined as the percentage of predicted edges that are within two
-steps in the original, unpartitioned recommender graph. In the example above,
-Product A and Product C are within two steps in the original graph, thus an edge
-between A and C is considered correct and increases precision. 2-recall is
-defined analogously. Here, we consider the new recommender graph, consisting of
-the two partitioned recommender graphs connected by the predicted
-recommendations. 2-recall is then the percentage of gold-standard edges that 
-are within two steps in the predicted recommender graph.
+We use both traditional precision, recall, and $F_1$ score as well as {\em
+2-precision}, {\em 2-recall}, and {\em 2-}$F_1$ score to evaluate our prediction
+algorithms. Including the latter set of metrics provides a more complete picture
+of algorithmic performance. These additional metrics also accommodate the fact
+that the edges of $G$ do not necessarily correspond to the only or even the best
+item-item recommendations. We therefore suggest that while traditional precision
+and recall are relevant metrics, relative performance as measured by {\em
+2-precision} and {\em 2-recall} better captures the effectiveness of these
+prediction algorithms.
 
@@
 
+\bibitem{Powers2011}
+    Powers, David M W.
+    ``Evaluation: From Precision, Recall and F-Measure to ROC, Informedness,
+Markedness \& Correlation.''
+    \emph{Journal of Machine Learning Technologies} \textbf{2}, 37-63, 2011.
+
 \end{thebibliography}

### 2014-06-10 paper2/paper.tex
Rework the introduction.

@@
 \section*{Introduction}
-Recommender systems have become a critical component for online retailers in
-nearly every space of retail. Whether the products are movies or clothes or
-research articles, large retailers tend to have a way of recommending products
-based on others previously viewed or purchased products.
+A key challenge in retail is choosing which items to incorporate into one's
+collection of offered goods and services. These decisions can significantly
+impact a retailer's revenue in the short run as well as affect customer
+experience, and hence, their relationship with the retailer. Several factors
+must be considered including the number of product lines, the variety of
+products in each line, as well as the consistency and relationships between
+products. In this study, we focus on the latter and explore techniques that
+leverage publicly available product recommendation information for the purpose
+of improving product assortment decisions.
 
-These recomendations usually come from a combination of content-based
-information and collaborative filtering techniques. The content-based approach
-uses information about the products, such as category or description, to connect
-related products. Alone, however, this approach is usually unable to capture the
-subtle ways products may actually be related. Collaborative filtering, on the
-other hand, learns relations from aggregated user behavior.  Such
-recommendations often have explicit titles like ``People who liked this product
-also liked these products''. Collaborative filtering has proven to be effective
-at recommending products, given that there is a large amount of user data.
+It has become standard practice for online retailers to recommend one or more
+products to prospective customers who have viewed or purchased an item. These
+recommendations are generally derived from either content-based approaches or
+via collaborative filtering, utilizing data-mining techniques \cite{Ricci2011}.
+Importantly, these recommendations provide a source of publicly available
+business intelligence by relating the items in their respective catalogue.
+Specifically, these online recommendations logically form a directed graph in
+which the nodes are products and an edge pointing from item $A$ to item $B$
+indicates that customers who view or purchase item $A$ are recommended item $B$.
+Given the recommendation graphs of two or more distinct retailers, our goal is
+to determine new, meaningful edges that connect these graphs in a way that would
+allow one retailer to relate their products to those of another.
 
-Since each retailer only has user data within their site, these user-based
-recommender systems are good at learning connections between products in one
-catalogue, but they do little to learn how these products may relate to other
-products outside of the retailer's inventory. This type of information is
-valuable in a number of situations. Retailers such as department stores want to
-know which new products make the most sense to add to their catalogue.  This is
-essentially a question of recommending the best outside products based on
-current in-house ones. In a different scenario, two retailers may merge and want
-to construct a unified recommender system between the two catalgoues.
+Content-based approaches and collaborative filtering have distinct challenges,
+and therefore, offer unique advantages as solutions to the general recommender
+problem.  An effective content-based solution requires a detailed analysis of
+item and/or user profile information. A well-known, successful example is the
+Music Genome Project, which powers the music service Pandora \cite{mgp}. Such an
+approach can be difficult, however, since necessary user and/or item information
+might be unavailable or the required domain expertise might be too costly. The
+alternative method is that of collaborative filtering (CF), which leverages the
+correlations within user purchase or rating data. The idea is that two items
+that tend to be purchased or highly rated by the same users can be considered
+related such that if a new user likes one they will probably like the other. A
+key advantage of this approach is that it effectively outsources the
+recommendation problem to the retailer's customers. Online retailers such as
+Amazon.com and Netflix are known to employ recommender systems based on CF
+\cite{Koren2009}. At the same time, CF notoriously suffers from what is known as
+the {\em cold start}---without user-item data of sufficient quantity and
+diversity these systems fail to yield accurate results.
 
-We model this as a problem of connecting recommender systems between disjoint
-catalogues. The goal is to construct good recommendations from products in one
-catalogue to products in the other. These catalogues may be from competing
-retailers who don't have access to each other's user data. Thus, to keep the
-problem general, we do not assume that we have any private data.
-Instead, we constrain ourselves to use only the product information and internal
-recommendations that are posted publicly to all users.
+The problem we examine here is, in many ways, more difficult than that of the
+standard recommendation problem. Essentially, we would like to recommend an item
+from retailer $X$ based solely on a known preference for an item from retailer
+$Y$. For the product assortment problem, user purchase and rating information is
+presumably known for one of the two retailers. In this study, however, we
+simplify the scenario by making it symmetric. No user purchase or rating
+information will be leveraged for either retailer. Instead, the algorithms
+investigated here are limited to publicly available information, which includes
+product descriptions and online product recommendations. Given the lack of
+inter-retailer user data, a CF technique is ostensibly not an option. The
+obvious approach is therefore content-based.
 
-This is a difficult problem. What makes a good recommendation is hard to define
-objectively. One can say that if users co-purchase or co-view products together,
-then these products should be recommended for each other. But without any user
-data, it is hard to simulate this kind of user behavior. The connections between
-co-viewed products are often less tangible than simple metrics of similarity.
-
-In this paper, we outline a technique that leverages the intra-catalogue
-recommendations of each catalogue to construct inter-catalgoue recommendations
-between them. We focus on clothing retailers, since they tend to all have
-public recommendations for each product. Specifically, the data used for this
-project was all collected from the public product webpages for Macy's.
+In this report. . . . (BREIFLY OUTLINE THE PAPER HERE.)
 
 \section*{Data}
-For this work, we utilize most of the product catalog of Macy's as presented at
-www.macys.com. This catalog consists of numerous categories of mostly apparel,
-each of which includes hundreds to thousands of items. We confine this study to
-the 49 categories within the two main parent categories ``Men'' and ``Women''
-that contain at least 200 items that are associated to multiple brands. In
-total, this dataset consists of 66,071 items in 49 categories with some items
-listed in multiple categories. For each item we record its description, all
-categories within which it is listed, and its associated recommendations. By
-{\em recommendations}, we are referring to the items---there are typically four
-in the case of Macy's---that are displayed on a product's details page under the
-heading ``Customers Also Shopped''. The presence of these recommendations
-implicitly forms a directed graph over the product catalogue, in which the nodes
-are the {\em products} or {\em items} and the edges are the {\em
+For this work, we utilize most of the product catalogue of Macy's as presented
+at www.macys.com. This catalogue consists of numerous categories of mostly
+apparel, each of which includes hundreds to thousands of items. We confine this
+study to the 49 categories within the two main parent categories ``Men'' and
+``Women'' that contain at least 200 items that are associated to multiple
+brands. In total, this dataset consists of 66,071 items in 49 categories with
+some items listed in multiple categories. For each item we record its
+description, all categories within which it is listed, and its associated
+recommendations. By {\em recommendations}, we are referring to the items---there
+are typically four in the case of Macy's---that are displayed on a product's
+details page under the heading ``Customers Also Shopped''. The presence of these
+recommendations implicitly forms a directed graph over the product catalogue, in
+which the nodes are the {\em products} or {\em items} and the edges are the {\em
 recommendations}. We call these types of graphs {\em recommendation graphs} and
@@
 
+\bibitem{Ricci2011}
+    Francesco Ricci, Lior Rokach, and Bracha Shapira.
+    \emph{Introduction to Recommender Systems Handbook}, Springer, 1-35, 2011.
+
+\bibitem{mgp}
+    For information on the Music Genome Project, visit
+    http://www.pandora.com/about/mgp
+
+\bibitem{Koren2009}
+    Yehuda Koren, Robert Bell, and Chris Volinsky.
+    ``Matrix factorization techniques for recommender systems.''
+    \emph{Computer} 42.8, 2009. 30-37.
+
 \bibitem{Powers2011}
-    Powers, David M W.
+    David M W Powers.
     ``Evaluation: From Precision, Recall and F-Measure to ROC, Informedness,

### 2014-06-10 paper2/paper.tex
Small edits

--- 2014-06-10 paper2/paper.tex
@@
 \section*{Constructing Recommendations}
-Recommending products between different catalogues is a daunting task.
-Traditionally, constructing recommendations is done by leveraging user behavior
-using a collaborative filtering technique. At the most basic level, If a large
-number of users view or purchase both Product A and Product B, the two are
-connected with a recommendation. Unfortunately, for this problem there are no
+Recommending products between different catalogues is a difficult task.
+Typically, recommendations are constructed by leveraging user behavior
+with a collaborative filtering technique. At the most basic level, If a large
+number of users view or purchase both Product A and Product B, the two products
+are recommended for each other. Unfortunately, for this problem there are no
 user data between the two unconnected catalogues.
@@
 \subsection*{Content Based Similarity}
-A simple approach for this problem is to recommend products that have similar
-content. The products have both a short and long description which we can use to
-construct a textual content-based similarity function. To do this, we represent each
+A simple approach is to recommend products that have similar content. Each product
+has includes a short description which we can use to construct a textual 
+content-based similarity function. To do this, we represent each
 product as a vector of tf-idf values for each word in its descriptions. By using
-tf-idf, we reduce the relevance of frequent words and boost that of rare ones,
-thus emphasizing the more words that are more descriptive of a product versus
-all the other ones in the catalogue. The vector representations are the length 
-of the dictionary, with zeros for all words not in the descriptions. To determine
-the content-based similarity between two products, we simply take the dot 
-product of the two tf-idf vectors.
+tf-idf, we reduce the relevance of frequent words and boost that of rare ones.
+This emphasizes the words that better explain this product versus all the other 
+ones in the catalogue. The vector representations are the length of the
+vocabulary. They are very sparse as most words do not appear in any given
+product description. With this tf-idf represenation, we can define the 
+$ContentSimiarity$ functoin as follows:
 
-Mathematical definition of content-based similairty...
+\begin{align}
+ContentSimilarity(P_1, P_2) &= v_1 \cdot v_2
+\end{align}
+where $P_1$ and $P_2$ are the products being compared and $v_1$ and $v_2$ are
+their representations as tf-idf vectors.
 
-This similarity function yields the following recommendation algorithm: For each
+This similarity function yields the following recommendation algorithm which we
+call the $ContentBased Recommender$: For each
 product, scan through the products in the opposite catalogue and calculate the
@@
 multiplication. Furthermore, the division into different categories reduces the
-impact of the quadratic effect. For instance, Macy's entire catalogue consists
-of over ONE MILLION products, making a quadratic time algorithm infeasible.
-However, the average category has 2000 products, which is significantly more
-manageable.
+impact of the quadratic effect. In our case, Macy's entire catalogue consists
+of over $60,000$ products, making a quadratic time algorithm very slow.
+However, the average category has only one or two thousand products, which is 
+significantly more manageable size for quadratic operations.
 
 \subsection*{Leveraging Recommendations}
-The simple content-based recommender is limited for a few reasons. It attempts
+This $ContentBased Recommender$ is limited in a few ways. For one, it attempts
 to capture similarity between products by looking exclusively at the words used

### 2014-06-11 paper2/paper.tex
Rewrite content based similarity sections

@@
 \subsection*{Content Based Similarity}
-A simple approach is to recommend products that have similar content. Each product
-has includes a short description which we can use to construct a textual 
-content-based similarity function. To do this, we represent each
-product as a vector of tf-idf values for each word in its descriptions. By using
-tf-idf, we reduce the relevance of frequent words and boost that of rare ones.
-This emphasizes the words that better explain this product versus all the other 
-ones in the catalogue. The vector representations are the length of the
-vocabulary. They are very sparse as most words do not appear in any given
-product description. With this tf-idf represenation, we can define the 
-$ContentSimiarity$ functoin as follows:
+A simple approach that does not rely on user data is to recommend products that
+have similar content. Each product page includes a short description which we 
+can use to construct a textual content-based similarity function. To do this, 
+we represent each product as a vector of tf-idf values for each word in its 
+description. By using tf-idf, we reduce the relevance of frequent words and 
+boost that of rare ones. This emphasizes the words that better represent this 
+product versus all the other ones in the catalogue. The vector representations 
+are the length of the vocabulary. These are sparse vectors as most words do 
+not appear in any given product description. With this tf-idf represenation,
+we can define the $ContentSimiarity$ functoin as follows:
 
@@
 This similarity function yields the following recommendation algorithm which we
-call the $ContentBased Recommender$: For each
-product, scan through the products in the opposite catalogue and calculate the
-content-based similarity score with each. Predict recommendations between this
-product and each of the top k most similar products in the other catalogue.
+call the $ContentBased Recommender$: For each product, calculate its
+$ContentSimilarity$ score with each of the products in the other catalogue.
+Predict recommendations between this product and the top 2 most similar products 
+in the other catalogue.
 
-While this is a quadric-time operation, it can be sped up using matrix
-multiplication. Furthermore, the division into different categories reduces the
-impact of the quadratic effect. In our case, Macy's entire catalogue consists
-of over $60,000$ products, making a quadratic time algorithm very slow.
-However, the average category has only one or two thousand products, which is 
-significantly more manageable size for quadratic operations.
+While this is a quadric-time operation, it can be executed efficiently using 
+matrix multiplication. Furthermore, the division into different categories 
+reduces the impact of the quadratic effect. In our case, the entire Macy's 
+catalogue consists of over $60,000$ products, making a quadratic time algorithm 
+slow. However, the average category has only one to two thousand products, which
+is a significantly more manageable size for quadratic operations.
 
@@
 This $ContentBased Recommender$ is limited in a few ways. For one, it attempts
-to capture similarity between products by looking exclusively at the words used
-to describe them. One limitation here is that different brands often use
-different vocabularies to describe the same thing, such as ``off-white'' versus
-``ivory.'' Another issue is that the best recommendation for a product is often
-more subtle than simply the most similar other product. For instance, a striped
-black-and-white button-up shirt may be an excellent recommendation for a blue
-checkered button-up, even though the two have ostensibly different patterns and
-colors.
+to capture similarity between products by looking exclusively at the words in
+their descriptions. One issue here is that brands tend to use different 
+vocabularies to describe the their products. One may refer to a color as 
+``off-white'' while another calls the same color ``ivory.'' To the $ContentBased
+Recommender$, these words are as different as ``black'' and ``white.'' Another 
+issue is that the best recommendation for a product may not be the most similar
+other product. For instance, a striped black-and-white button-up shirt could be 
+an excellent recommendation for a blue checkered button-up, even though the two 
+have ostensibly different descriptions.
 
 We can alleviate some of these issues by leveraging the recommendations we
-already have between products in the same catalogue. One approach is to consider
-the neighborhood of each product in their recommender graphs, that is a product
-along with each of the products it is recommended with in its own catalogue.
-When comparing two products in different catalogues, we use the same tf-idf
-similarity function from before to compare each pair of products between the two
-neighborhoods. Mathematically, we construct the similarity function
-NeighborhoodSimilarity defined as:
+already have between products within the same catalogue. One approach is to 
+consider the {\emph neighborhood} of each product, that is that product along 
+with each of its neighbors in the recommendation graph. With this we create a
+new similarity function, $NeighborhoodSimilarity$. This function uses the 
+$ContentSimilarity$ function from before to compare every pair of products 
+between the two neighborhoods:
 
-NeighborhoodSimilarity definition...
+\begin{align}
+NeighborhoodSimilarity(P_1, P_2) &= \sum\limits_{S\in
+Neigh(P_1)}\sum\limits_{T\in Neigh(P_2)}
+ContentSimilarity(S, T) 
+\end{align}
+where $Neigh(P)$ is the neighborhood of Product P. We use this similarity
+function to construct the $Neighborhood Recommender$, which is analgous to the
+$ContentBased Recommender$.
 
-Assuming the recommendations are good within each of the catalogues, this
-approach can help both of the issues brought up with the simple content-based
-recommender. Since a neighborhood has many products, each describing themselves
-with potentially different vocabularies, there is a higher likelihood that the
-neighborhood contains multiple synonyms and thus decreases the potential of
-missing due to different vocabularies. Furthermore, the recommender graphs
-capture some of the user behavior not expressed in product descriptions.
-Re-examining the button-up shirts example, the neighborhood of a striped shirt may
-include checkered shirts and vise-versa. In this case, the two neighborhoods
-would share the words ``striped'' and ``checkered'' and thus the two previously
-dissimilar shirts would now be considered similar.
+Assuming that the edges of the recommendation graphs are in fact good
+recommendations, this approach can help both of the issues brought up with the
+$ContentBased Recommender$. Since a neighborhood has many products, each 
+describing themselves using potentially different vocabularies, there is a 
+higher likelihood that the neighborhood contains multiple synonyms for the
+underlying concept that describes their similarity. This decreases the
+likelihood of false negatives due to different vocabularies. Furthermore,
+the recommendation graphs capture some of the user behavior that cannot be
+expressed in product descriptions. Re-examining the button-up shirts example, 
+the neighborhood of a striped shirt may include checkered shirts and vise-versa.
+In this case, the two neighborhoods would share the words ``striped'' and 
+``checkered'' and thus the two previously dissimilar shirts would now correctly
+be deemed similar.
 
 \subsection*{Latent Concepts}
-Although the NeighborhoodSimilarity method alleviates some of the issues of
-content-similarity, it is still fundamentally based on the descriptions of
+Although the $Neighborhood Recommender$ alleviates some of the issues of 
+content-based similarity, it is still fundamentally based on the descriptions of
 products. Ideally, we would like to recommend products that are similar in more
-fundamental ways than their descriptions. If a dress is elegant and modern, we
-want to recommend another dress that is elegant and modern, even if those words
-are not explicitly in the descriptions of either. These types of latent concepts
-can be intangible and difficult to define. However, if we assume that the given
-recommendations are truly based on these underlying concepts, we can attempt to 
-learn the latent concepts by analyzing the graphs.
+fundamental ways. If a dress is elegant and modern, we would like to recommend 
+another dress that is elegant and modern, even if those words are not explicitly
+in the descriptions of either. These types of underlying concepts can be 
+intangible and difficult to define. However, if we assume that the given
+recommendations are truly based on such latent concepts, we can attempt to learn 
+them by analyzing the graphs.
 
-One approach to learning latent concepts is Latent Semantic Indexing or LSI. In
+One common approach to learning latent semantic factors is to apply Singular Value
+Decomposition to . In
 traditional LSI, used for Natural Language Processing, concepts are learned over 
@@
 have actual user data, we can simulate user sessions by traversing the
-recommender graphs.
+recommendation graphs.
 
 To simulate user sessions, we use a random walk technique on the graph. We
-assume that the edges of a recommender graph indeed connect products that a user
+assume that the edges of a recommendation graph indeed connect products that a user
 would co-view. Thus traversing the edges of a graph is a reasonable simulation
@@
 In all, the LSI approach consists of four phases. First we simulate user
-sessions by randomly walking the recommender graphs. Second we learn concept
+sessions by randomly walking the recommendation graphs. Second we learn concept
 spaces for products in each catalogue via SVD. Third we find a mapping between
@@
 The first step of the LSI approach is to simulate user sessions by randomly
-walking the recommender graphs. There are a few ways this could be implemented.
+walking the recommendation graphs. There are a few ways this could be implemented.
 One method is to randomly assign a random walker a starting node and then allow
@@
 distributions. As with any random walk technique, we first interpret the 
-recommender graph as a transition matrix. In this walk, for each product we create a
+recommendation graph as a transition matrix. In this walk, for each product we create a
 product-specific transition matrix which has added teleportation edges back the
@@
 
-We imagine that recommender graphs are constructed as follows: A retailer first
+We imagine that recommendation graphs are constructed as follows: A retailer first
 builds a similarity matrix, using a collaborative filtering method. In this
@@
 
-We see signs of this in the undirected recommender graphs. Although each product
+We see signs of this in the undirected recommendation graphs. Although each product
 has at most four outgoing recommendations on macys.com, we see that some
@@
 approximate the popularity of each product as the logarithm of the degree of
-that product in its undirected recommender graph. We choose to model it as the
+that product in its undirected recommendation graph. We choose to model it as the
 logarithm because a few of the products tend to have degrees in the hundreds.
@@
 
-Assuming this model of recommender graph construction, which separates similarity
-from popularity, we seek to isolate similarity from the recommender graph. Since
+Assuming this model of recommendation graph construction, which separates similarity
+from popularity, we seek to isolate similarity from the recommendation graph. Since
 popular products were promoted for popularity, we demote them when walking the
@@
 of popularity is analagous to multiplying similarity by popularity, as done in
-the original construction of the recommender graph.
+the original construction of the recommendation graph.
 
 A final way in which we incorporate popularity is in the number of
-recommendations alotted per product. Since we model the recommender graph as an
+recommendations alotted per product. Since we model the recommendation graph as an
 undirected graph, different nodes have different degrees. A product that is
@@
 outlined earlier in the paper. This recommender uses information exclusively
-from the descriptions of the products and ignores the recommender graphs 
+from the descriptions of the products and ignores the recommendation graphs 
 entirely. A third, somewhat similar baseline comes from the Neighborhood 
@@
 information from the unpartitioned graph. The One Model Recommender learns a
-latent concept space over the entire recommender graph by allowing the random
+latent concept space over the entire recommendation graph by allowing the random
 walkers to traverse it before it is partitioned. In every other respect, it is

### 2014-06-11 paper2/paper.tex
Rework ContentBasedRecommender subsection.

@@
 
-\section*{Experiments}
+\section*{Experiment}
 Our goal is to develop methods for connecting initially disconnected
@@
 
-\section*{Constructing Recommendations}
-Recommending products between different catalogues is a difficult task.
-Typically, recommendations are constructed by leveraging user behavior
-with a collaborative filtering technique. At the most basic level, If a large
-number of users view or purchase both Product A and Product B, the two products
-are recommended for each other. Unfortunately, for this problem there are no
-user data between the two unconnected catalogues.
+\section*{Algorithms}
 
-\subsection*{Content Based Similarity}
-A simple approach that does not rely on user data is to recommend products that
-have similar content. Each product page includes a short description which we 
-can use to construct a textual content-based similarity function. To do this, 
-we represent each product as a vector of tf-idf values for each word in its 
-description. By using tf-idf, we reduce the relevance of frequent words and 
-boost that of rare ones. This emphasizes the words that better represent this 
-product versus all the other ones in the catalogue. The vector representations 
-are the length of the vocabulary. These are sparse vectors as most words do 
-not appear in any given product description. With this tf-idf represenation,
-we can define the $ContentSimiarity$ functoin as follows:
+\subsection*{\em ContentBasedRecommender}
+The most straightforward approach to inter-retailer recommendations is a
+content-based approach. Retail products typically have a description that can be
+used to construct a textual content-based similarity function. One can then
+consider all pairwise product combinations across retailers, but within a common
+category, and choose recommendations between items with sufficiently high
+similarity.
 
+Toward this end, we borrow the concept of {\em Term Frequency--Inverse Document
+Frequency} (tfidf) from the information retrieval community. In using tf-idf, we
+are considering each item description to be a ``document''.  Thus we define the
+tfidf of a term $t$ within an item description $d$ that's part of a combined
+category catalogue $C$ (i.e. the catalogue consisting of all items in both
+retailers' catalogues for a particular common category) as
 \begin{align}
-ContentSimilarity(P_1, P_2) &= v_1 \cdot v_2
+\mathrm{tfidf}(t,d,C) = f(t,d) \times
+\mathrm{log}\frac{\left|C\right|}{\left|\left\{d \in C : t \in
+d\right\}\right|},
 \end{align}
-where $P_1$ and $P_2$ are the products being compared and $v_1$ and $v_2$ are
-their representations as tf-idf vectors.
+where $f(t,d)$ is the number of occurances of term $t$ in item description $d$,
+and $\left|\left\{d \in C : t \in d\right\}\right|$ is the number of items in
+$C$ with descriptions containing at least one instance of term $t$.  We can now
+represent each item as a sparse vector of tfidf values, which has a nonzero
+element for each term in its description.
 
-This similarity function yields the following recommendation algorithm which we
-call the $ContentBased Recommender$: For each product, calculate its
-$ContentSimilarity$ score with each of the products in the other catalogue.
-Predict recommendations between this product and the top 2 most similar products 
-in the other catalogue.
+A key feature of tfidf that is useful here is how terms that appear in many item
+description are discounted via the idf factor. Conversely, rare terms will have
+larger tfidf values. We now define the $ContentSimilarity$ between items $A$ and
+$B$ as
+\begin{align}
+ContentSimilarity(A,B) = v_A \cdot v_B
+\end{align}
+where $v_A$ and $v_B$ are the tfidf vector representations of items $A$ and $B$,
+respectively.
 
-While this is a quadric-time operation, it can be executed efficiently using 
-matrix multiplication. Furthermore, the division into different categories 
-reduces the impact of the quadratic effect. In our case, the entire Macy's 
+Using this content-based similarity function, we construct the following
+inter-catalogue recommendation algorithm, which we call the $ContentBased
+Recommender$. For each product in each catalogue, calculate its
+$ContentSimilarity$ with each and every product in the other catalogue. Predict
+recommendations between this product and the top two most similar products in
+the other catalogue. The reason for choosing the top two most similar products
+originates from our knowledge that most items listed on www.macy.com are
+associated with four recommended items. Thus, by choosing two edges per item per
+partition, we will predict approximately the same number of edges that were lost
+during the partition. This relatively straightforward algorithm serves as a
+baseline against which to compare the performance of other more sophisticated
+algorithms.
+
+The $ContentBasedRecommender$ deserves a few more comments. For starters, the
+use of an inner product, which is bounded only by the length of the tfidf
+vectors, is not fundamental to this algorithm. A bounded similarity function
+such as cosine similarity could have been used instead without qualitatively
+changing the results discussed later in this report. This is because the lengths
+of the product descriptions have little variation, and therefore, the vector
+norms of the tfidf vectors have little variation. Finally, we mention that while
+this algorithm is quadratic in the number items, it can be executed relatively
+efficiently using matrix multiplication, which trades time for space (i.e.
+memory). Furthermore, confining our experiments to individual categories reduces
+the impact of the quadratic time complexity. In our case, Macy's entire
 catalogue consists of over $60,000$ products, making a quadratic time algorithm
-slow. However, the average category has only one to two thousand products, which
-is a significantly more manageable size for quadratic operations.
+challenging. However, the average category has about two thousand products,
+which is a significantly more manageable.
 

### 2014-06-11 paper2/paper.tex
Modify Latent Concept section

--- 2014-06-11 paper2/paper.tex
@@
 issue is that the best recommendation for a product may not be the most similar
-other product. For instance, a striped black-and-white button-up shirt could be 
+product. For instance, a striped black-and-white button-up shirt could be 
 an excellent recommendation for a blue checkered button-up, even though the two 
@@
 already have between products within the same catalogue. One approach is to 
-consider the {\emph neighborhood} of each product, that is that product along 
-with each of its neighbors in the recommendation graph. With this we create a
-new similarity function, $NeighborhoodSimilarity$. This function uses the 
+consider the {\em neighborhood} of each product, that is a product along 
+with each of its neighbors in the recommendation graph. With this, we create a
+new similarity function called $NeighborhoodSimilarity$. This function uses the 
 $ContentSimilarity$ function from before to compare every pair of products 
@@
 Although the $Neighborhood Recommender$ alleviates some of the issues of 
-content-based similarity, it is still fundamentally based on the descriptions of
+content-based similarity, it is still primarily based on the descriptions of
 products. Ideally, we would like to recommend products that are similar in more
@@
 
-One common approach to learning latent semantic factors is to apply Singular Value
-Decomposition to . In
-traditional LSI, used for Natural Language Processing, concepts are learned over 
-a dataset consists of documents, each containing terms. Singular Value 
-Decomposition (SVD) is applied to the term-document matrix and the resulting 
-three matrices are truncated to the number of latent concepts desired. The 
-resulting left matrix is a term-concept matrix, where each row is a term 
-represented in concept space.
+One common approach to learning latent semantic factors is to apply {\em
+Singular Value Decomposition} (SVD) to a matrix of user data, as in many 
+collaborative filtering methods [CITE]. This method requires a product-user
+matrix, where each column represents the products viewed in a user session.
+After SVD is applied to the matrix, the resulting three matrices are truncated
+to the number of latent concepts desired. The resulting left matrix becomes a
+product-concept matrix, where each row is a product represented in 
+{\em concept-space}
 
-We apply the same idea to the product catalogues in order to learn the latent
-concepts of each product. Analogous to documents containing terms, we look at
-user sessions which contain the products viewed by that user. Although we do not
-have actual user data, we can simulate user sessions by traversing the
-recommendation graphs.
+In this experiment, there is no such product-user matrix to apply SVD.
+However, since the recommendation graphs are constructed from user behavior,
+we can use them to simulate user sessions. To do this, we use a random walk 
+technique on the graph. We assume that the edges of a recommendation graph 
+indeed connect products that a user would co-view. Thus traversing the edges 
+of a graph is a reasonable simulation of a user session. For details on these
+random walks, see the section titled Random Walks in the Implementation section
+below.
 
-To simulate user sessions, we use a random walk technique on the graph. We
-assume that the edges of a recommendation graph indeed connect products that a user
-would co-view. Thus traversing the edges of a graph is a reasonable simulation
-of a user session. For details on these random walks, see the section titled
-Random Walks in the Implementation section below.
-
-With the simulated user sessions, we can implement LSI on each of the catalogues
-and learn the latent concepts for each product. We want to find the products from
-one catalogue that are most similar in latent concept space to products in the
-other catalogue. However, since the user sessions (random walks) include
-products either from one catalogue or the other, we learn different latent 
-spaces for each catalogue. The problem is therefore reduced to finding a mapping
-from one concept space to the other. If this can be accomplished, we
-can map the concepts of each product into concepts of the other catalogue. This
-creates a concept representation in the other catalouge's concept space, meaning
-that we can then find nearest neighbors in that catalogue's concept space.
+With simulated user data for each recommendation graph, we can apply SVD on the
+matrices from the two catalogues to learn the latent concepts for each product.
+We want to find the products from one catalogue that are most similar in latent
+concept-space to products in the other catalogue. However, since each user
+session/random walk includes products exclusively from one of the catalogues, we
+learn different latent spaces for each catalogue. Note that this would be an
+issue even if we had the actual user data used to construct the two
+recommendation graphs. The problem therefore becomes finding a mapping from one
+concept-space to the other. If this can be accomplished, we can represent each
+product of one catalogue in terms of concepts of the other catalogue and then
+find the nearest neighbors of that catalogue by distance in concept-space.
 
 To map concepts from one catalogue to another, we must revisit
-content-similarity. In this case, we represent each concept as a vector of
-words. To calculate the value for a word w for concept c, we sum the tf-idf
-value of w across all products, weighted by how much that product is represented
-by concept c. Thus if a concept could be described as formal, the words
-``sophisticated'' or ``official'' may have high value since products with those
-descriptive words would also have a high amount of the formal concept in their
-concept representations. Using these vector representation for concepts, we can
-simply calculate the dot product between two concepts to find how similar they
-are, in a relative sense. This can be used to create a mapping from one set of
-concepts to another.
+content-similarity. In this case, we represent each concept (rather than each
+product) as a vector of words. To represent a concept in terms of the
+vocabulary, we calculate the value of each word by summing the
+tf-idf of that word for each product $P$, weighted by the projection of $P$ on
+our concept in concept-space. That is:
 
-In all, the LSI approach consists of four phases. First we simulate user
-sessions by randomly walking the recommendation graphs. Second we learn concept
-spaces for products in each catalogue via SVD. Third we find a mapping between
-concept spaces. And lastly we find the nearest neighbors of each product in the
-other catalogue's concept space. These nearest neighbors are the predicted
-recommendations between products in different catalogues.
+\begin{align}
+C_w = \sum\limits_{P} \text{tfidf}(w)\times P_C
+\end{align}
+where $C_w$ is the value of concept $C$'s representation in dimension $w$, the sum 
+is over all products $P$ in the catalgoue, tfidf($w$) is the tf-idf score for
+word $w$, and $P_C$ is the projection of $P$ on concpet $C$'s axis in
+concept-space.
+
+Now for a concept such as {\em formal}, the words ``sophisticated'' or
+``official'' may have high value since products with those descriptive words
+would also have a large amount of the {\em formal} concept in their
+concept-representations. Using these vector representation, we calculate the
+inner product between two concepts to find how similar they are to eachother.
+This can be used to create a mapping from one set of concepts to another as it
+captures the relative breakdown of a concept from one space to the concepts in the
+other.
+
+In all, the latent-concept approach consists of four phases. First, we simulate user
+sessions by randomly walking the recommendation graphs. Second, we learn concept
+spaces for products in each catalogue via SVD. Third, we find a mapping between
+concept-spaces using the aggregated descriptions of all products. And lastly, we
+find the nearest neighbors of each product in the concept-space of the other 
+catalogue concept-space. These nearest neighbors become the predicted
+recommendations between products across the two catalogues. We call this method
+the $LatentConcept Recommender$
 
@@
 A tricky operation of this procedure is the concept mapping. We attempt to find
-the best linear transformation from one concept space to the other. We do this
+the best linear transformation from one concept-space to the other. We do this
 by representing each concept as a vector of words in the vocabulary, each
@@
 the L2 norm. Neither of these methods preserve the L2 norm, which would be ideal
-as we find the nearest neighbors in the other concept space by their L2
+as we find the nearest neighbors in the other concept-space by their L2
 distance.
@@
 
-Proceeding as before, we find the nearest products in concept space. However,
+Proceeding as before, we find the nearest products in concept-space. However,
 before deeming the nearest neighbors to be the best recommendations, we divide
 the distance to each product by its popularity. Since we model distance in 
-concept space to be inversely proportional to similarity, this reincorporation
+concept-space to be inversely proportional to similarity, this reincorporation
 of popularity is analagous to multiplying similarity by popularity, as done in
@@
 capture similarity using descriptions alone. It is not even clear that a mapping 
-necessarily exists between two seperately learned concept spaces. The concepts 
+necessarily exists between two seperately learned concept-spaces. The concepts 
 may represent fundamentally different things, which cannot be well mapped to 
@@
 information from the unpartitioned graph. The One Model Recommender learns a
-latent concept space over the entire recommendation graph by allowing the random
+latent concept-space over the entire recommendation graph by allowing the random
 walkers to traverse it before it is partitioned. In every other respect, it is
 the same as the LSI Recommender. However, since products from both catalogues
-are represented in the same concept space, there is no need for a mapping
-between concept spaces.
+are represented in the same concept-space, there is no need for a mapping
+between concept-spaces.
 
@@
 Orthogonalize matrix &&&\\ \hline
-Normalize products in concept space &&&\\ \hline
+Normalize products in concept-space &&&\\ \hline
 \end{tabular}

### 2014-06-11 paper2/paper.tex
Edits to intro

@@
 A key challenge in retail is choosing which items to incorporate into one's
-collection of offered goods and services. These decisions can significantly
-impact a retailer's revenue in the short run as well as affect customer
-experience, and hence, their relationship with the retailer. Several factors
-must be considered including the number of product lines, the variety of
-products in each line, as well as the consistency and relationships between
-products. In this study, we focus on the latter and explore techniques that
-leverage publicly available product recommendation information for the purpose
-of improving product assortment decisions.
+collection of offered goods and services.  Several factors must be considered
+including the number of product lines, the variety of products in each line, as
+well as the consistency and relationships between products. In this study, we
+focus on the latter and explore techniques that leverage publicly available
+product recommendation information for the purpose of improving product
+assortment decisions.
 
@@
 product descriptions and online product recommendations. Given the lack of
-inter-retailer user data, a CF technique is ostensibly not an option. The
-obvious approach is therefore content-based.
+inter-retailer user data, a traditional CF technique is ostensibly not
+applicable. Instead, we propose a new technique that leverages the recomender
+graphs as well as the content from each product to create inter-retail
+recommendations.
 

### 2014-06-11 paper2/paper.tex
Minor changes to NeighborhoodRecommender subsection.

--- 2014-06-11 paper2/paper.tex
@@
 
-\section*{Algorithms}
+\section*{Recommender Algorithms}
 
@@
 A key feature of tfidf that is useful here is how terms that appear in many item
-description are discounted via the idf factor. Conversely, rare terms will have
+descriptions are discounted via the idf factor. Conversely, rare terms will have
 larger tfidf values. We now define the $ContentSimilarity$ between items $A$ and
@@
 challenging. However, the average category has about two thousand products,
-which is a significantly more manageable.
+which is significantly more manageable.
 
-\subsection*{Leveraging Recommendations}
-This $ContentBased Recommender$ is limited in a few ways. For one, it attempts
-to capture similarity between products by looking exclusively at the words in
-their descriptions. One issue here is that brands tend to use different 
-vocabularies to describe the their products. One may refer to a color as 
-``off-white'' while another calls the same color ``ivory.'' To the $ContentBased
-Recommender$, these words are as different as ``black'' and ``white.'' Another 
-issue is that the best recommendation for a product may not be the most similar
-product. For instance, a striped black-and-white button-up shirt could be 
-an excellent recommendation for a blue checkered button-up, even though the two 
-have ostensibly different descriptions.
+\subsection*{\em NeighborhoodRecommender}
+The $ContentBasedRecommender$ is limited in that it attempts to capture
+similarity between products by looking exclusively at the terms in their
+descriptions. One issue here is that brands tend to use different vocabularies
+to describe their products. One may refer to a color as ``off-white'' while
+another calls the same color ``ivory.'' To the $ContentBasedRecommender$ these
+terms are as different as ``black'' and ``white.'' Another issue is that the
+best recommendation for a product may not even be the most similar product. For
+instance, a striped black-and-white button-up shirt could be an excellent
+recommendation associated with a blue checkered button-up, even though the two
+products are clearly different and would presumably use signficantly different
+terms their descriptions.
 
-We can alleviate some of these issues by leveraging the recommendations we
-already have between products within the same catalogue. One approach is to 
-consider the {\em neighborhood} of each product, that is a product along 
-with each of its neighbors in the recommendation graph. With this, we create a
-new similarity function called $NeighborhoodSimilarity$. This function uses the 
-$ContentSimilarity$ function from before to compare every pair of products 
-between the two neighborhoods:
-
+We can begin to address these issues by leveraging the recommendations we
+already have between products within a single partition. One approach is to
+consider the {\em neighborhood} of each product, that is a product along with
+each of its immediate neighbors in the recommendation graph. With this, we
+create a new similarity function called $NeighborhoodSimilarity$. This function
+uses the $ContentSimilarity$ function from before to compare every pair of
+products between the two neighborhoods of a pair of products $A$ and $B$:
 \begin{align}
-NeighborhoodSimilarity(P_1, P_2) &= \sum\limits_{S\in
-Neigh(P_1)}\sum\limits_{T\in Neigh(P_2)}
+NeighborhoodSimilarity(A, B) = \sum\limits_{S\in
+Neigh(A)}\sum\limits_{T\in Neigh(B)}
 ContentSimilarity(S, T) 
 \end{align}
-where $Neigh(P)$ is the neighborhood of Product P. We use this similarity
-function to construct the $Neighborhood Recommender$, which is analgous to the
-$ContentBased Recommender$.
+where $Neigh(A)$ is the set of items that share edges with item $A$. We use this
+similarity function to construct the $NeighborhoodRecommender$, which is
+an extension of the $ContentBasedRecommender$. The algorithm is the same as
+before except now we choose a recommendation associated with items $A$ and $B$
+by maximnizing
+\begin{align}
+ContentSimilarity(A,B) + \eta \cdot NeighborhoodSimilarity(A, B),
+\end{align}
+where $\eta$ is a parameter. We find that $\eta = 0.1$ yields good results for
+this algorithm.
 
 Assuming that the edges of the recommendation graphs are in fact good
-recommendations, this approach can help both of the issues brought up with the
-$ContentBased Recommender$. Since a neighborhood has many products, each 
-describing themselves using potentially different vocabularies, there is a 
-higher likelihood that the neighborhood contains multiple synonyms for the
-underlying concept that describes their similarity. This decreases the
-likelihood of false negatives due to different vocabularies. Furthermore,
-the recommendation graphs capture some of the user behavior that cannot be
-expressed in product descriptions. Re-examining the button-up shirts example, 
-the neighborhood of a striped shirt may include checkered shirts and vise-versa.
-In this case, the two neighborhoods would share the words ``striped'' and 
-``checkered'' and thus the two previously dissimilar shirts would now correctly
-be deemed similar.
+recommendations, this approach can help both of the shortcomings of
+$ContentBasedRecommender$ previously mentioned. Since a neighborhood includes
+several items, each with descriptions that potentially use different
+vocabularies, there is a higher likelihood that the neighborhood contains
+multiple synonyms for the underlying concept that captures their similarity. At
+the same time, this decreases the likelihood of false negatives due to different
+vocabularies. Presumably, a retailer's recommendation graph include information
+from user behavior that is not expressed in product descriptions. Re-examining
+the button-up shirts example, the neighborhood of a striped shirt may include
+checkered shirts and vice versa.  In this case, the two neighborhoods might
+share the words ``striped'' and ``checkered'' and thus the two previously
+dissimilar shirts would now be considered more similar.
 
-\subsection*{Latent Concepts}
-Although the $Neighborhood Recommender$ alleviates some of the issues of 
+\subsection*{\em LatentConceptRecommender}
+Although the $NeighborhoodRecommender$ alleviates some of the issues of 
 content-based similarity, it is still primarily based on the descriptions of
@@
 \end{align}
-where $C_w$ is the value of concept $C$'s representation in dimension $w$, the sum 
-is over all products $P$ in the catalgoue, tfidf($w$) is the tf-idf score for
-word $w$, and $P_C$ is the projection of $P$ on concpet $C$'s axis in
+where $C_w$ is the value of concept $C$'s representation in dimension $w$, the
+sum is over all products $P$ in the catalgoue, tfidf($w$) is the tf-idf score
+for word $w$, and $P_C$ is the projection of $P$ on concpet $C$'s axis in
 concept-space.
@@
 This can be used to create a mapping from one set of concepts to another as it
-captures the relative breakdown of a concept from one space to the concepts in the
-other.
+captures the relative breakdown of a concept from one space to the concepts in
+the other.
 
-In all, the latent-concept approach consists of four phases. First, we simulate user
-sessions by randomly walking the recommendation graphs. Second, we learn concept
-spaces for products in each catalogue via SVD. Third, we find a mapping between
-concept-spaces using the aggregated descriptions of all products. And lastly, we
-find the nearest neighbors of each product in the concept-space of the other 
-catalogue concept-space. These nearest neighbors become the predicted
+In all, the latent-concept approach consists of four phases. First, we simulate
+user sessions by randomly walking the recommendation graphs. Second, we learn
+concept spaces for products in each catalogue via SVD. Third, we find a mapping
+between concept-spaces using the aggregated descriptions of all products. And
+lastly, we find the nearest neighbors of each product in the concept-space of
+the other catalogue concept-space. These nearest neighbors become the predicted
 recommendations between products across the two catalogues. We call this method
-the $LatentConcept Recommender$
+the $LatentConceptRecommender$
 

### 2014-06-11 paper2/paper.tex
Edits to Data section

--- 2014-06-11 paper2/paper.tex
@@
 For this work, we utilize most of the product catalogue of Macy's as presented
-at www.macys.com. This catalogue consists of numerous categories of mostly
-apparel, each of which includes hundreds to thousands of items. We confine this
-study to the 49 categories within the two main parent categories ``Men'' and
-``Women'' that contain at least 200 items that are associated to multiple
-brands. In total, this dataset consists of 66,071 items in 49 categories with
-some items listed in multiple categories. For each item we record its
-description, all categories within which it is listed, and its associated
-recommendations. By {\em recommendations}, we are referring to the items---there
-are typically four in the case of Macy's---that are displayed on a product's
-details page under the heading ``Customers Also Shopped''. The presence of these
-recommendations implicitly forms a directed graph over the product catalogue, in
-which the nodes are the {\em products} or {\em items} and the edges are the {\em
+at www.macys.com. This catalogue consists of numerous categories, which are
+mostly apparel. Each category includes hundreds to thousands of items. We
+confine this study to the 49 categories under the two main parent categories
+``Men'' and ``Women'' that contain at least 200 items from multiple brands.  In
+total, this dataset consists of 66,071 items in 49 categories with some items
+listed in multiple categories. For each item we record its description, all
+categories within which it is listed, and its associated recommendations. By
+{\em recommendations}, we are referring to the items---there are typically four
+in the case of Macy's---that are displayed on a product's details page under the
+heading ``Customers Also Shopped''. The presence of these recommendations
+implicitly forms a directed graph over the product catalogue, in which the nodes
+are the {\em products} or {\em items} and the edges are the {\em
 recommendations}. We call these types of graphs {\em recommendation graphs} and

### 2014-06-11 paper2/paper.tex
Edits to Experiment section

@@
 with items in the other catalogue. In this way, we are effectively introducing
-edges that connect the two recommendation graphs. Formulating this
-inter-retailer recommendations problem in terms of graphs offers insight in how
-to evaluate our recommendation choices as well as how to choose ``good''
-recommendations.  Leveraging the graph structure for evaluation is discussed in
-this section while exploiting the graph for the purpose of making better
-recommendations is addressed in a subsequent section.
+edges that connect the two recommendation graphs. Formulating the inter-retailer
+recommendation problem in terms of graphs offers insight on how to evaluate our
+recommendation choices as well as how to choose ``good'' recommendations.
+Leveraging the graph structure for evaluation is discussed in this section while
+exploiting the graph for the purpose of making better recommendations is
+addressed in a subsequent section, Recommender Algorithms.
 
-To evaluate our inter-retailer recommendation methods, we simulate the problem
-by randomly partitioning Macy's products into two disjoint sets. Given the
-original recommendation graph, each partition therefore corresponds to a graph
-cut. The edges in this graph cut form a {\em gold-standard} since it is assumed
-that these are indeed ``good'' recommendations. Our premise is that an ideal
-recommendation method could guess these gold-standard edges with high precision
-and recall.
+To evaluate our inter-retailer recommendation algorithms, we simulate the
+problem by randomly partitioning Macy's products into two disjoint sets. Given
+the original recommendation graph, each partition therefore corresponds to a
+graph cut. The edges in this graph cut form a {\em gold-standard} since it is
+assumed that these are indeed ``good'' recommendations. Our premise is that an
+ideal recommendation method could guess these gold-standard edges with high
+precision and recall.
 
@@
 (e.g. Women Activewear, Men Dress Shirts, etc.). Since most recommendations are
-between items in a common category, very few, though some, edges are lost by
-confining ourselves to an individual category. Edges that are lost in this way
-do not participate in the experiment and are not considered during evaluation.
-Also, we randomly partition the items in the category such that all items
-associated with a particular brand are entirely within a single partition. In
-this way, we preclude the easiest method of associating two items: recognizing
-a common brand. This is done to increase the difficulty and realism of the
-experiments.
+between items in a common category, very few edges are lost by confining
+ourselves to an individual category. Edges that are lost in this way do not
+participate in the experiment and are not considered during evaluation.  Also,
+we randomly partition the items in the category such that all items associated
+with a particular brand are entirely within a single partition. In this way, we
+preclude the easiest method of associating two items: recognizing a common
+brand. This is done to increase the difficulty and realism of the experiment.
 
@@
 all items in each partition, including their descriptions and all
-recommendations (i.e.  edges) between items in a common partition. In addition,
+recommendations (i.e.  edges) within a common partition. In addition,
 the algorithms may exploit the fact that most items listed at www.macys.com are

### 2014-06-11 paper2/paper.tex
Edits to Evaluation section

@@
 Macy's category, for example, Dresses, which contains approximately $3,500$
-items. Suppose that randomly partitioning results in two sets, each with about
-$1,750$ items. For each interset recommendation, we must therefore choose two
-nodes, one from each of the two sets. That is, we have $1,750 \times 1,750 > 3
-\times 10^6$ choices. Choosing the correct edges is a formidable challenge to
-say the least. It is therefore worth asking: Are some prediction errors better
-or worse than others?
+items. Suppose that a random partition results in two sets with approximately
+$1,750$ items in each. Such a partition would cut approximately half of all the
+edges in the unpartioned graph. Since each node has on average eight edges
+before partitioning (four in and four out in the original directed graph) and
+each edge is shared by two nodes, there are originally a total of approximately
+$\left(8 \times 3,500\right) / 2 = 14,000$ edges. This implies that there are
+$7,000$ withheld edges in the cut. For each predicted recommendation, we must
+choose two nodes, one from each of the two sets. Therefore we have $\left(1,750 \times
+1,750\right)/2 > 1 \times 10^6$ choices and only $7,000$ of them are correct. Choosing the
+correct edges is a formidable challenge to say the least. It is therefore worth
+asking: Are some prediction errors better or worse than others?
 
@@
 The following definitions will prove helpful. Let $G$ be the original
-unpartitioned graph formed from the partitions together with the lost edges $L$
-and let $G^{\prime}$ be the new graph formed from the partitions together with
-the predicted edges $P$.  In the context of {\em 2-precision}, an edge in $P$
-that connects items $A$ and $B$ is considered correct if 1) $A$ and $B$ share an
-edge within $G$ or 2) $A$ shares an edge with a direct neighbor of $B$ within
+unpartitioned graph, formed from the partitions together with the lost edges
+$L$, and let $G^{\prime}$ be the new graph, formed from the partitions together
+with the predicted edges $P$.  In the context of {\em 2-precision}, an edge in
+$P$ that connects items $A$ and $B$ is considered correct if 1) $A$ and $B$
+share an edge in $G$ or 2) $A$ shares an edge with a direct neighbor of $B$ in
 $G$ or 3) a direct neighbor of $A$ shares an edge with $B$ within $G$.
 Meanwhile, in the context of {\em 2-recall}, an edge in $L$ that connects items
-$C$ and $D$ is considered recalled if 1) $C$ and $D$ share an edge within
-$G^{\prime}$ or 2) $C$ shares an edge with a direct neighbor of $D$ within
-$G^{\prime}$ or 3) a direct neighbor of $C$ shares an edge with $D$ within
+$C$ and $D$ is considered recalled if 1) $C$ and $D$ share an edge in
+$G^{\prime}$ or 2) $C$ shares an edge with a direct neighbor of $D$ in
+$G^{\prime}$ or 3) a direct neighbor of $C$ shares an edge with $D$ in
 $G^{\prime}$.
 
-We use both traditional precision, recall, and $F_1$ score as well as {\em
+We use traditional precision, recall, and $F_1$ score as well as {\em
 2-precision}, {\em 2-recall}, and {\em 2-}$F_1$ score to evaluate our prediction

### 2014-06-11 paper2/paper.tex
Edits to the two Content Recommender sections

@@
 used to construct a textual content-based similarity function. One can then
-consider all pairwise product combinations across retailers, but within a common
-category, and choose recommendations between items with sufficiently high
-similarity.
+consider all pairwise product combinations across the two retailers, but within
+a common category, and choose recommendations between items with sufficiently
+high similarity.
 
 Toward this end, we borrow the concept of {\em Term Frequency--Inverse Document
-Frequency} (tfidf) from the information retrieval community. In using tf-idf, we
+Frequency} (tfidf) from the information retrieval community. In using tfidf, we
 are considering each item description to be a ``document''.  Thus we define the
@@
 
-A key feature of tfidf that is useful here is how terms that appear in many item
-descriptions are discounted via the idf factor. Conversely, rare terms will have
-larger tfidf values. We now define the $ContentSimilarity$ between items $A$ and
-$B$ as
+A key feature of tfidf is that terms appearing in many item descriptions are
+discounted via the idf factor. Conversely, rare terms will have larger tfidf
+values. These features help capture the most relevant words in each description.
+We now define the $ContentSimilarity$ between items $A$ and $B$ as
 \begin{align}
@@
 Recommender$. For each product in each catalogue, calculate its
-$ContentSimilarity$ with each and every product in the other catalogue. Predict
-recommendations between this product and the top two most similar products in
+$ContentSimilarity$ with every product in the other catalogue. Predict
+recommendations between this product and the two most similar products in
 the other catalogue. The reason for choosing the top two most similar products
@@
 vectors, is not fundamental to this algorithm. A bounded similarity function
-such as cosine similarity could have been used instead without qualitatively
+such as cosine similarity could have been used instead, without qualitatively
 changing the results discussed later in this report. This is because the lengths
@@
 norms of the tfidf vectors have little variation. Finally, we mention that while
-this algorithm is quadratic in the number items, it can be executed relatively
+this algorithm is quadratic in the number of items, it can be executed relatively
 efficiently using matrix multiplication, which trades time for space (i.e.
@@
 recommendation associated with a blue checkered button-up, even though the two
-products are clearly different and would presumably use signficantly different
-terms their descriptions.
+products are different and would presumably use signficantly different terms in
+their descriptions.
 
@@
 vocabularies, there is a higher likelihood that the neighborhood contains
-multiple synonyms for the underlying concept that captures their similarity. At
-the same time, this decreases the likelihood of false negatives due to different
-vocabularies. Presumably, a retailer's recommendation graph include information
-from user behavior that is not expressed in product descriptions. Re-examining
-the button-up shirts example, the neighborhood of a striped shirt may include
-checkered shirts and vice versa.  In this case, the two neighborhoods might
-share the words ``striped'' and ``checkered'' and thus the two previously
-dissimilar shirts would now be considered more similar.
+multiple synonyms for the underlying concept that captures their similarity.
+This therefore decreases the likelihood of false negatives due to different
+vocabularies between a pair of items. Presumably, a retailer's recommendation
+graph includes information from user behavior that is not expressed in product
+descriptions. Re-examining the button-up shirts example, the neighborhood of a
+striped shirt may include checkered shirts and vice versa.  In this case, the
+two neighborhoods might share the words ``striped'' and ``checkered'' and thus
+the two previously dissimilar shirts would now be considered more similar.
 
@@
 vocabulary, we calculate the value of each word by summing the
-tf-idf of that word for each product $P$, weighted by the projection of $P$ on
+tfidf of that word for each product $P$, weighted by the projection of $P$ on
 our concept in concept-space. That is:
@@
 where $C_w$ is the value of concept $C$'s representation in dimension $w$, the
-sum is over all products $P$ in the catalgoue, tfidf($w$) is the tf-idf score
+sum is over all products $P$ in the catalgoue, tfidf($w$) is the tfidf score
 for word $w$, and $P_C$ is the projection of $P$ on concpet $C$'s axis in

### 2014-06-11 paper2/paper.tex
Rework LatentConceptRecommender subsection, but not finished.

--- 2014-06-11 paper2/paper.tex
@@
 $7,000$ withheld edges in the cut. For each predicted recommendation, we must
-choose two nodes, one from each of the two sets. Therefore we have $\left(1,750 \times
-1,750\right)/2 > 1 \times 10^6$ choices and only $7,000$ of them are correct. Choosing the
-correct edges is a formidable challenge to say the least. It is therefore worth
-asking: Are some prediction errors better or worse than others?
+choose two nodes, one from each of the two sets. Therefore we have $\left(1,750
+\times 1,750\right)/2 > 1 \times 10^6$ choices and only $7,000$ of them are
+correct. Choosing the correct edges is a formidable challenge to say the least.
+It is therefore worth asking: Are some prediction errors better or worse than
+others?
 
-The graph nature of our problem reveals that the answer to this question is
-yes. For instance, it is better to guess an edge that connects two items that
-are separated by two edges in the original graph than to connect two items that
-are separated by five edges. We therefore introduce the notion of {\em
-2-precision}, {\em 2-recall}, and {\em 2-}$F_1$ are defined as
+The graph nature of our problem reveals that the answer to this question is yes.
+For instance, it is better to guess an edge that connects two items that are
+separated by two edges in the original graph than to connect two items that are
+separated by five edges. We therefore introduce the notion of {\em 2-precision},
+{\em 2-recall}, and {\em 2-}$F_1$ are defined as
 
@@
 \subsection*{\em LatentConceptRecommender}
-Although the $NeighborhoodRecommender$ alleviates some of the issues of 
-content-based similarity, it is still primarily based on the descriptions of
-products. Ideally, we would like to recommend products that are similar in more
-fundamental ways. If a dress is elegant and modern, we would like to recommend 
-another dress that is elegant and modern, even if those words are not explicitly
-in the descriptions of either. These types of underlying concepts can be 
-intangible and difficult to define. However, if we assume that the given
-recommendations are truly based on such latent concepts, we can attempt to learn 
-them by analyzing the graphs.
+Although the $NeighborhoodRecommender$ leverages the information contained in
+the recommendation graphs to some extent, it is still primarily based on product
+descriptions. Presumably, the graphs are derivative of a collaborative filtering
+scheme, and hence, reveal associations between products that are not evident
+from the language of their descriptions. For example, a dress's description
+might not capture if it is more conservative or more risque. Yet, this could be
+the salient feature it has in common with an associated recommendation.  These
+sort of abstract, perhaps ineffable, features can be essential to item-item
+relations. We must therefore ask whether we can more fully utilize the given
+graphical information to discover these latent features in order to choose
+better inter-catalogue associations.
 
-One common approach to learning latent semantic factors is to apply {\em
-Singular Value Decomposition} (SVD) to a matrix of user data, as in many 
-collaborative filtering methods [CITE]. This method requires a product-user
-matrix, where each column represents the products viewed in a user session.
-After SVD is applied to the matrix, the resulting three matrices are truncated
-to the number of latent concepts desired. The resulting left matrix becomes a
-product-concept matrix, where each row is a product represented in 
-{\em concept-space}
+A common approach to learning latent features is to apply matrix factorization
+techniques to user-item data, as in many CF methods \cite{Koren2009}. But,
+user-item data, a prerequisite for such techniques, in not available to us here.
+A key insight of the present work is that publicly available recommendation
+graphs, such as those found on Macy's website, can be used to simulate user-item
+data through the use of special random walks known as {\em Topic-Sensitive
+PageRank}, a technique typically associated with web search
+\cite{Haveliwala2002}.
 
-In this experiment, there is no such product-user matrix to apply SVD.
-However, since the recommendation graphs are constructed from user behavior,
-we can use them to simulate user sessions. To do this, we use a random walk 
-technique on the graph. We assume that the edges of a recommendation graph 
-indeed connect products that a user would co-view. Thus traversing the edges 
-of a graph is a reasonable simulation of a user session. For details on these
-random walks, see the section titled Random Walks in the Implementation section
-below.
+The idea is that a user with a preference for a particular product is more
+likely to also have a preference for one of its associated recommendations than
+another arbitrary product. We can simulate this with a random walker that, by
+construction has a preference for an item $A$, and therefore, begins her walk at
+that item on the recommendation graph. At each step in the walk, she is equally
+likely to visit any of the items that share an edge with the presently occupied
+item or stay at the present item. There is also a small chance at each step that
+the walker will teleport back to item $A$. In the language of Topic-Sensitive
+PageRank, item $A$ is the sole nonzero entry in the walker's personalization
+vector. We create our synthetic user-item data by performing such a random walk
+for each node in the graph. That is, each node has a random walk in which it is
+the sole nonzero entry in the walker's personalization vector. In this way, we
+simulate a user for each item in each graph partition.
 
-With simulated user data for each recommendation graph, we can apply SVD on the
-matrices from the two catalogues to learn the latent concepts for each product.
-We want to find the products from one catalogue that are most similar in latent
-concept-space to products in the other catalogue. However, since each user
-session/random walk includes products exclusively from one of the catalogues, we
-learn different latent spaces for each catalogue. Note that this would be an
-issue even if we had the actual user data used to construct the two
-recommendation graphs. The problem therefore becomes finding a mapping from one
-concept-space to the other. If this can be accomplished, we can represent each
-product of one catalogue in terms of concepts of the other catalogue and then
-find the nearest neighbors of that catalogue by distance in concept-space.
+Constructing our random walk in this way provides an important benefit. In the
+context of the random walk, the recommendation graph may be considered a {\em
+discrete-time Markov chain}. By providing the walker a finite probability of
+staying at each item, our chain becomes {\em aperiodic}. And since our walker
+begins at $A$ and can only teleport to $A$, she is effectively confined to a
+strongly connected subgraph of the recommendation graph. This subgraph is
+therefore, by construction, {\em irreducible}.  Importantly, according to the
+{\em Fundamental Theorem of Markov Chains}, any finite, aperiodic, irreducible
+Markov chain has a unique stationary distribution \cite{Norris1998}.
 
-To map concepts from one catalogue to another, we must revisit
-content-similarity. In this case, we represent each concept (rather than each
+We may then construct a synthetic $N \times N$ user-item matrix $M$ with rows
+representing items and columns representing users, which are in fact the
+stationary distributions discovered from our random walks. This matrix has at
+least one notable advantage over most real user-item matrices---it is not
+sparse.  Rather, each synthetic user has a finite value associated with every
+item that is reachable from its origin item in the graph. Unlike the typical CF
+scenario, we may therefore employ SVD as our matrix factorization technique
+since our user-item matrix is virtually complete (i.e. nearly void of empty
+entries).  Using SVD, we factor our user-item matrix as
+\begin{align}
+M = U \Sigma V^T,
+\end{align}
+where the rows of the $N \times N$ matrix $U$ and the $N \times N$ matrix $V$
+are the items and users represented in an $N$-dimensional concept-space,
+respectively, and the $N$ diagonal entries of $\Sigma$ are the singular values
+associated with each concept.
+
+Having transformed our items into concept-space, we now discard all but the $k$
+leftmost columns of the matrix $U$. The idea is that these leftmost columns are
+the most significant eigenvectors of concept-space and capture most, if not all,
+of the relevant latent features. Meanwhile, the discarded columns are presumed
+to contain mostly noise. This rank-reduction process yields matrices that are
+typically written as
+\begin{align}
+M \approx U_k \Sigma_k V^T_k,
+\end{align}
+where the rows of $U_k$ and $V_k$ are the items and users represented in a
+$k$-dimensional concept-space, respectively.  Rank-reducing following the
+application of SVD in this way is a common technique used in information
+retrieval and is often referred to as {\em Latent Semantec Indexing} (LSI)
+\cite{Deerwester1990}.
+
+Using random walks to simulate user data, followed by the application of LSI, is
+indeed an effective way to encode information from a recommendation graph into a
+latent feature space. But what is the relationship between latent feature
+vectors derived from disconnected graph partitions? In principle, there is no
+guarantee of any relationship between such latent spaces in the same way that
+there is no guarantee of any relationship between two arbitrary, disconnected
+recommendation graphs. Nonetheless, our hypothesis is that a relationship
+between such latent spaces often will exist due to the effectiveness of the
+underlying CF techniques that generate the graphs to capture meaningful
+item-item associations. Put another way, we believe that the most fundamental
+latent dimensions along which products in a category such as Women's swimsuits
+vary will be common to many swimsuit catalogues. If this is indeed true, then it
+might be possible to learn a mapping between these latent spaces.
+
+If a mapping between the latent spaces associated with two catalogues were
+known, we could translate all items into a common space and then choose
+inter-catalogue edges between item pairs that are nearest to one another in
+concept-space. To learn these mapping, we must revisit the notion of content
+similarity. Once again, we utilize tf-idf vectors. This time, however, we weight
+an item's tf-idf vector by the elements in the item's latent feature vector. In
+this way, we construct a dense tf-idf vector for each concept for each graph
+partition. Formally, we say that each concept $j$ is represented by 
+\begin{align}
+EXPRESSION FOR A CONCEPT'S TF-IDF VECTOR
+\end{align}
+
+In this case, we represent each concept (rather than each
 product) as a vector of words. To represent a concept in terms of the
-vocabulary, we calculate the value of each word by summing the
-tf-idf of that word for each product $P$, weighted by the projection of $P$ on
-our concept in concept-space. That is:
+vocabulary, we calculate the value of each word by summing the tf-idf of that
+word for each product $P$, weighted by the projection of $P$ on our concept in
+concept-space. That is:
 
@@
 The first step of the LSI approach is to simulate user sessions by randomly
-walking the recommendation graphs. There are a few ways this could be implemented.
-One method is to randomly assign a random walker a starting node and then allow
-it to randomly walk the edges, with a probability of termination at each step.
-Running this many times is a straight-forward model of user sessions.
+walking the recommendation graphs. There are a few ways this could be
+implemented.  One method is to randomly assign a random walker a starting node
+and then allow it to randomly walk the edges, with a probability of termination
+at each step.  Running this many times is a straight-forward model of user
+sessions.
 
@@
 distributions. As with any random walk technique, we first interpret the
-recommendation graph as a transition matrix. In this walk, for each product we create a
-product-specific transition matrix which has added teleportation edges back the
-that product. We calculate the stationary distributions of each of these
-matrices, one per product. We use these product-specific stationary
+recommendation graph as a transition matrix. In this walk, for each product we
+create a product-specific transition matrix which has added teleportation edges
+back the that product. We calculate the stationary distributions of each of
+these matrices, one per product. We use these product-specific stationary
 distributions as our user sessions. Since there is only one user session per
@@
 
-This process requires a couple of parameters. The first is the number of steps used to
-approximate each product-specific statinary distribution. The second parameter
-is the likelihood of teleporting to the special product. For our purposes we use
-30 steps and a teleportation probability of 5\%.
+This process requires a couple of parameters. The first is the number of steps
+used to approximate each product-specific statinary distribution. The second
+parameter is the likelihood of teleporting to the special product. For our
+purposes we use 30 steps and a teleportation probability of 5\%.
 
@@
 
+\bibitem{Haveliwala2002}
+    Taher Haveliwala.
+    ``Topic-Sensitive PageRank.''
+    \emph{Proceedings of the Eleventh International World Wide Web Conference}
+(Honolulu, Hawaii) 2002.
+
+\bibitem{Norris1998}
+    James R. Norris.
+    \emph{Markov Chains}. Cambridge University Press, 1998.
+
+\bibitem{Deerwester1990}
+    Scott Deerwester, Susan T. Dumais, George W. Furnas, Thomas K. Landauer,
+Richard Harshman.
+    ``Indexing by Latent Semantex Analysis.''
+    \emph{American Society for Information Science} \textbf{41}, 391-407, 1990.
+
 \end{thebibliography}

### 2014-06-11 paper2/paper.tex

@@
 
-In this case, we represent each concept (rather than each
-product) as a vector of words. To represent a concept in terms of the
-vocabulary, we calculate the value of each word by summing the tf-idf of that
-word for each product $P$, weighted by the projection of $P$ on our concept in
-concept-space. That is:
-
+In this case, we represent each concept (rather than each product) as a vector
+of words. To represent a concept in terms of the vocabulary, we calculate the
+value of each word by summing the tf-idf of that word for each product $P$,
+weighted by the projection of $P$ on our concept in concept-space. That is:
 \begin{align}

### 2014-06-11 paper2/paper.tex
Finish reworking the LatentFeatureRecommender subsection.

@@
 Toward this end, we borrow the concept of {\em Term Frequency--Inverse Document
-Frequency} (tfidf) from the information retrieval community. In using tfidf, we
+Frequency} (tf-idf) from the information retrieval community. In using tf-idf, we
 are considering each item description to be a ``document''.  Thus we define the
-tfidf of a term $t$ within an item description $d$ that's part of a combined
+tf-idf of a term $t$ within an item description $d$ that's part of a combined
 category catalogue $C$ (i.e. the catalogue consisting of all items in both
@@
 \begin{align}
-\mathrm{tfidf}(t,d,C) = f(t,d) \times
+\textrm{tf-idf}(t,d,C) = f(t,d) \times
 \mathrm{log}\frac{\left|C\right|}{\left|\left\{d \in C : t \in
@@
 $C$ with descriptions containing at least one instance of term $t$.  We can now
-represent each item as a sparse vector of tfidf values, which has a nonzero
+represent each item as a sparse vector of tf-idf values, which has a nonzero
 element for each term in its description.
 
-A key feature of tfidf is that terms appearing in many item descriptions are
-discounted via the idf factor. Conversely, rare terms will have larger tfidf
+A key feature of tf-idf is that terms appearing in many item descriptions are
+discounted via the idf factor. Conversely, rare terms will have larger tf-idf
 values. These features help capture the most relevant words in each description.
@@
 \begin{align}
-ContentSimilarity(A,B) = v_A \cdot v_B
+ContentSimilarity(A,B) = \vec{\tau}(A) \cdot \vec{\tau}(B)
 \end{align}
-where $v_A$ and $v_B$ are the tfidf vector representations of items $A$ and $B$,
-respectively.
+where $\vec{\tau}(A)$ is the tf-idf vector representation of item $A$.
 
@@
 The $ContentBasedRecommender$ deserves a few more comments. For starters, the
-use of an inner product, which is bounded only by the length of the tfidf
+use of an inner product, which is bounded only by the length of the tf-idf
 vectors, is not fundamental to this algorithm. A bounded similarity function
@@
 of the product descriptions have little variation, and therefore, the vector
-norms of the tfidf vectors have little variation. Finally, we mention that while
-this algorithm is quadratic in the number of items, it can be executed relatively
-efficiently using matrix multiplication, which trades time for space (i.e.
-memory). Furthermore, confining our experiments to individual categories reduces
-the impact of the quadratic time complexity. In our case, Macy's entire
+norms of the tf-idf vectors have little variation. Finally, we mention that
+while this algorithm is quadratic in the number of items, it can be executed
+relatively efficiently using matrix multiplication, which trades time for space
+(i.e.  memory). Furthermore, confining our experiments to individual categories
+reduces the impact of the quadratic time complexity. In our case, Macy's entire
 catalogue consists of over $60,000$ products, making a quadratic time algorithm
@@
 \begin{align}
-NeighborhoodSimilarity(A, B) = \sum\limits_{S\in
-Neigh(A)}\sum\limits_{T\in Neigh(B)}
-ContentSimilarity(S, T) 
+NeighborhoodSimilarity(A, B) = \sum\limits_{i\in
+Neigh(A)}\sum\limits_{j\in Neigh(B)}
+ContentSimilarity(i, j) 
 \end{align}
@@
 
-\subsection*{\em LatentConceptRecommender}
+\subsection*{\em LatentFeatureRecommender}
 Although the $NeighborhoodRecommender$ leverages the information contained in
@@
 where the rows of the $N \times N$ matrix $U$ and the $N \times N$ matrix $V$
-are the items and users represented in an $N$-dimensional concept-space,
+are the items and users represented in an $N$-dimensional concept space,
 respectively, and the $N$ diagonal entries of $\Sigma$ are the singular values
@@
 
-Having transformed our items into concept-space, we now discard all but the $k$
+Having transformed our items into concept space, we now discard all but the $k$
 leftmost columns of the matrix $U$. The idea is that these leftmost columns are
-the most significant eigenvectors of concept-space and capture most, if not all,
+the most significant eigenvectors of concept space and capture most, if not all,
 of the relevant latent features. Meanwhile, the discarded columns are presumed
@@
 where the rows of $U_k$ and $V_k$ are the items and users represented in a
-$k$-dimensional concept-space, respectively.  Rank-reducing following the
+$k$-dimensional concept space, respectively.  Rank-reducing following the
 application of SVD in this way is a common technique used in information
@@
 inter-catalogue edges between item pairs that are nearest to one another in
-concept-space. To learn these mapping, we must revisit the notion of content
+concept space. To learn these mapping, we must revisit the notion of content
 similarity. Once again, we utilize tf-idf vectors. This time, however, we weight
-an item's tf-idf vector by the elements in the item's latent feature vector. In
-this way, we construct a dense tf-idf vector for each concept for each graph
-partition. Formally, we say that each concept $j$ is represented by 
+an item $i$'s tf-idf vector $\vec{\tau}(i)$ by the elements in its latent
+feature vector $\vec{v}(i)$. In this way, we construct a dense {\em
+super-tf-idf} vector for each concept for each catalogue. Formally, we say that
+a concept $j$ of a catalogue $C$ is represented by the super-tf-idf vector
 \begin{align}
-EXPRESSION FOR A CONCEPT'S TF-IDF VECTOR
+\vec{s}(j,C) = \sum_{i \in C}{v(i)_j \cdot \vec{\tau}(i)},
 \end{align}
+where $v(i)_j$ is the $j$\textsuperscript{th} component of item $i$'s latent
+feature vector. These super-tf-idf vectors act as a proxy for the concept. We
+approximate the mapping between two concept spaces by taking inner products of
+super-tf-idf vectors. The details of this mapping process are provided in a
+subsequent section.
 
-In this case, we represent each concept (rather than each product) as a vector
-of words. To represent a concept in terms of the vocabulary, we calculate the
-value of each word by summing the tf-idf of that word for each product $P$,
-weighted by the projection of $P$ on our concept in concept-space. That is:
-\begin{align}
-C_w = \sum\limits_{P} \text{tfidf}(w)\times P_C
-\end{align}
-where $C_w$ is the value of concept $C$'s representation in dimension $w$, the
-sum is over all products $P$ in the catalgoue, tfidf($w$) is the tfidf score
-for word $w$, and $P_C$ is the projection of $P$ on concpet $C$'s axis in
-concept-space.
+Together, these steps comprise the $LatentFeatureRecommender$. First, we
+simulate user sessions by randomly walking the recommendation graphs. Second, we
+learn concept spaces for products in each catalogue via LSI. Third, we find a
+mapping between concept spaces using weighted tf-idf vector representations of
+the products. And lastly, we find the nearest neighbors of each product from one
+catalogue within the concept space of another catalogue. These nearest neighbors
+are then the predicted recommendations between products across the two
+catalogues.
 
-Now for a concept such as {\em formal}, the words ``sophisticated'' or
-``official'' may have high value since products with those descriptive words
-would also have a large amount of the {\em formal} concept in their
-concept-representations. Using these vector representation, we calculate the
-inner product between two concepts to find how similar they are to eachother.
-This can be used to create a mapping from one set of concepts to another as it
-captures the relative breakdown of a concept from one space to the concepts in
-the other.
-
-In all, the latent-concept approach consists of four phases. First, we simulate
-user sessions by randomly walking the recommendation graphs. Second, we learn
-concept spaces for products in each catalogue via SVD. Third, we find a mapping
-between concept-spaces using the aggregated descriptions of all products. And
-lastly, we find the nearest neighbors of each product in the concept-space of
-the other catalogue concept-space. These nearest neighbors become the predicted
-recommendations between products across the two catalogues. We call this method
-the $LatentConceptRecommender$
+%Now for a concept such as {\em formalness}, the words ``sophisticated'' or
+%``classy'' may have high value since products with those descriptive words would
+%also have a large amount of the {\em formalness} concept in their latent feature
+%vectors.
 
@@
 A tricky operation of this procedure is the concept mapping. We attempt to find
-the best linear transformation from one concept-space to the other. We do this
+the best linear transformation from one concept space to the other. We do this
 by representing each concept as a vector of words in the vocabulary, each
 weighted with how well it represents the concept. We construct this by weighing
-the tfidf of each word in a product's description by how well that word
+the tf-idf of each word in a product's description by how well that word
 represents the concept. Note that products can have a positive or negative value
@@
 With these concept representations, we construct an innner-product matrix. The
-value of a cell (i, j) in the matrix is the dot product of concept i in the first 
-space with concept j in the second. The columns and rows of the resulting matrix
-represent the relative similarity of the descriptive words between two concepts.
-But the scale of this matrix can be very large. If the inner-product matrix is
-used without any normalization as the linear transformation between the concept
-spaces, the scale of the transformed products will be inconsistent with those
-originally in that space.
+value of a cell (i, j) in the matrix is the dot product of concept i in the
+first space with concept j in the second. The columns and rows of the resulting
+matrix represent the relative similarity of the descriptive words between two
+concepts.  But the scale of this matrix can be very large. If the inner-product
+matrix is used without any normalization as the linear transformation between
+the concept spaces, the scale of the transformed products will be inconsistent
+with those originally in that space.
 
@@
 the L2 norm. Neither of these methods preserve the L2 norm, which would be ideal
-as we find the nearest neighbors in the other concept-space by their L2
+as we find the nearest neighbors in the other concept space by their L2
 distance.
@@
 
-Proceeding as before, we find the nearest products in concept-space. However,
+Proceeding as before, we find the nearest products in concept space. However,
 before deeming the nearest neighbors to be the best recommendations, we divide
 the distance to each product by its popularity. Since we model distance in 
-concept-space to be inversely proportional to similarity, this reincorporation
+concept space to be inversely proportional to similarity, this reincorporation
 of popularity is analagous to multiplying similarity by popularity, as done in
@@
 capture similarity using descriptions alone. It is not even clear that a mapping 
-necessarily exists between two seperately learned concept-spaces. The concepts 
+necessarily exists between two seperately learned concept spaces. The concepts 
 may represent fundamentally different things, which cannot be well mapped to 
@@
 information from the unpartitioned graph. The One Model Recommender learns a
-latent concept-space over the entire recommendation graph by allowing the random
+latent concept space over the entire recommendation graph by allowing the random
 walkers to traverse it before it is partitioned. In every other respect, it is
 the same as the LSI Recommender. However, since products from both catalogues
-are represented in the same concept-space, there is no need for a mapping
-between concept-spaces.
+are represented in the same concept space, there is no need for a mapping
+between concept spaces.
 
@@
 Orthogonalize matrix &&&\\ \hline
-Normalize products in concept-space &&&\\ \hline
+Normalize products in concept space &&&\\ \hline
 \end{tabular}

### 2014-06-11 paper2/paper.tex
Change references to "latent space" to "concept space" to be more consistent.

@@
 vectors derived from disconnected graph partitions? In principle, there is no
-guarantee of any relationship between such latent spaces in the same way that
+guarantee of any relationship between such concept spaces in the same way that
 there is no guarantee of any relationship between two arbitrary, disconnected
 recommendation graphs. Nonetheless, our hypothesis is that a relationship
-between such latent spaces often will exist due to the effectiveness of the
+between these concept spaces often will exist due to the effectiveness of the
 underlying CF techniques that generate the graphs to capture meaningful
@@
 vary will be common to many swimsuit catalogues. If this is indeed true, then it
-might be possible to learn a mapping between these latent spaces.
+might be possible to learn a mapping between these concept spaces.
 
-If a mapping between the latent spaces associated with two catalogues were
+If a mapping between the concept spaces associated with two catalogues were
 known, we could translate all items into a common space and then choose

### 2014-06-11 paper2/paper.tex
Absorb Implementation section into Algorithms section.

@@
 values. These features help capture the most relevant words in each description.
-We now define the $ContentSimilarity$ between items $A$ and $B$ as
+
+We can further improve our content similarity function by considering the
+bigrams that appear in item descriptions. To exploit this fact, we construct
+another tf-idf vector for each item, this one corresponding to the observed
+bigrams. Combining these tf-idf vectors, we define the $ContentSimilarity$
+between items $A$ and $B$ as
 \begin{align}
-ContentSimilarity(A,B) = \vec{\tau}(A) \cdot \vec{\tau}(B)
+ContentSimilarity(A,B) = \vec{\tau}_1(A) \cdot \vec{\tau}_1(B) + \beta \cdot
+\vec{\tau}_2(A) \cdot \vec{\tau}_2(B),
 \end{align}
-where $\vec{\tau}(A)$ is the tf-idf vector representation of item $A$.
+where $\vec{\tau}_1(A)$ is the unigram tf-idf vector and $\vec{\tau}_2(A)$ is
+the bigram tf-idf vector representations of item $A$, and $\beta$ is a
+parameter. We find that that $\beta=0.3$ works well for our data.
 
@@
 
+\subsubsection*{Simulating User Data}
 A common approach to learning latent features is to apply matrix factorization
@@
 likely to visit any of the items that share an edge with the presently occupied
-item or stay at the present item. There is also a small chance at each step that
-the walker will teleport back to item $A$. In the language of Topic-Sensitive
-PageRank, item $A$ is the sole nonzero entry in the walker's personalization
-vector. We create our synthetic user-item data by performing such a random walk
-for each node in the graph. That is, each node has a random walk in which it is
-the sole nonzero entry in the walker's personalization vector. In this way, we
-simulate a user for each item in each graph partition.
+item or stay at the present item.  There is also a small chance at each step
+that the walker will teleport back to item $A$. In the language of
+Topic-Sensitive PageRank, item $A$ is the sole nonzero entry in the walker's
+personalization vector. We create our synthetic user-item data by performing
+such a random walk for each node in the graph. That is, each node has a random
+walk in which it is the sole nonzero entry in the walker's personalization
+vector. In this way, we simulate a user for each item in each graph partition.
 
@@
 {\em Fundamental Theorem of Markov Chains}, any finite, aperiodic, irreducible
-Markov chain has a unique stationary distribution \cite{Norris1998}.
+Markov chain has a unique stationary distribution \cite{Norris1998}. For our
+data, we find that 30 steps with a 5\% probability of teleporting back to the
+origin item yields a good approximation of the true stationary distribution. We
+then interpret the distribution associated with each item as the preference data
+assoicated with a particular user.
 
-We may then construct a synthetic $N \times N$ user-item matrix $M$ with rows
+\subsubsection*{Extracting Latent Features}
+We may now construct a synthetic $N \times N$ user-item matrix $M$ with rows
 representing items and columns representing users, which are in fact the
@@
 where the rows of $U_k$ and $V_k$ are the items and users represented in a
-$k$-dimensional concept space, respectively.  Rank-reducing following the
-application of SVD in this way is a common technique used in information
-retrieval and is often referred to as {\em Latent Semantec Indexing} (LSI)
-\cite{Deerwester1990}.
+$k$-dimensional concept space, respectively. For our data, we find that $k=16$
+yields good results across all categories, striking a balance between
+information loss and the tractability of mapping concept spaces (see below).
+Rank-reducing following the application of SVD in this way is a common technique
+used in information retrieval and is often referred to as {\em Latent Semantec
+Indexing} (LSI) \cite{Deerwester1990}.
 
+\subsubsection*{Concept Space Mapping}
 Using random walks to simulate user data, followed by the application of LSI, is
@@
 where $v(i)_j$ is the $j$\textsuperscript{th} component of item $i$'s latent
-feature vector. These super-tf-idf vectors act as a proxy for the concept. We
-approximate the mapping between two concept spaces by taking inner products of
-super-tf-idf vectors. The details of this mapping process are provided in a
-subsequent section.
+feature vector. Note that since the components $v(i)_j$ may be positive or
+negative so too can the elements of a concept's super-tf-idf vector, in
+constrast to the elements of an individual item's tf-idf vector which are
+necessarily greater than or equal to zero. Also note that we compute both
+$\vec{s}_1$ and $\vec{s}_2$, which correpond to unigram and bigram super-tf-idf
+vectors, respectively, for each concept. These super-tf-idf vectors then act as
+proxies for the concept.
+
+With these proxy representations of the concepts for the concept spaces of two
+catalogues $C_1$ and $C_2$, we construct a matrix $A$ of inner-products of
+super-tf-idf vectors.  Specifically, the matrix elements of $A$ are given by
+\begin{align}
+A_{ij}(C_1,C_2) = \vec{s}_1(i,C_1) \cdot \vec{s}_1(j,C_2) + \beta \cdot
+\vec{s}_2(i,C_1) \cdot \vec{s}_2(j,C_2),
+\end{align}
+where $i$ is a concept from $C_1$ and $j$ is a concept from $C_2$ and $\beta$ is
+the same parameters as before (again, we find $\beta=0.3$ works well).  We may
+now use this matrix to transform the vector representation of an item in one
+concept space into a vector representation in another concept space.
+
+It is important to note that the transformation matrix $A$ is not orthogonal, and
+therefore, does not preserve vector norms. We address this issue by normalizing
+all feature vectors, both foreign and domestic, in the concept space following
+the transformation. This radially projects all items onto the surface of sphere
+within the 16-dimensional concept space. Now when ranking an item's neighbors by
+distance, we are actually ranking neighbors by angular separation. That is, we
+are ranking neighbors by their cosine similarity in concept space. After
+experimenting with various normalization schemes we find that this approach
+works best.
 
@@
 
-%Now for a concept such as {\em formalness}, the words ``sophisticated'' or
-%``classy'' may have high value since products with those descriptive words would
-%also have a large amount of the {\em formalness} concept in their latent feature
-%vectors.
+\subsubsection*{Popularity}
+The $LatentFeatureRecommender$ can be significantly improved by incorporating
+the notion of item {\em popularity}. The idea is that a more popular item is
+more likely to appear as a recommendation independent of how similar it is.  We
+see signs of this in the undirected recommendation graphs from www.macys.com.
+Although each product has at most four outgoing recommendations on its details
+page, we see that some products have over one hundred incoming recommendations.
+But how might we define {\em popularity}? A reasonable definition would
+introduce a quantity that is correlated with the number of users that indicate a
+preference for an item. Like all user data, this is not publicly available
+information.  Once again, however, we can leverage the recommendation graph to
+infer an approximate popularity metric for each item.
 
-\section*{Implementation}
-Let us dig deeper into the details of the LSI approach described above. There 
-are a number of implementation decisions that, in all, dramatically affect the
-performance of this method.
+We define item popularity as
+\begin{align}
+\rho(A) = \mathrm{log}(1 + \mathrm{deg}(A)),
+\end{align}
+where $\mathrm{deg}(A)$ is the degree of the item $A$ within its recommendation
+graph.  We find this quanity to be useful in improving the
+$LatentFeatureRecommender$ in several ways.
 
-\subsection*{Random Walks}
-The first step of the LSI approach is to simulate user sessions by randomly
-walking the recommendation graphs. There are a few ways this could be
-implemented.  One method is to randomly assign a random walker a starting node
-and then allow it to randomly walk the edges, with a probability of termination
-at each step.  Running this many times is a straight-forward model of user
-sessions.
+We begin with an attempt to seprate the notion of item similarity from that of
+item popularity. The idea that more popular items are more likely to be
+recommended than less popular items motivates an alternative mode for conducting
+our random walks over a recommendation graph in order to better isolate item
+similarity. We do this by adjusting the probability of transitioning to an item
+on the graph by making it inversely proportional to that item's popularity
+$\rho$. Stationary distributions learned from these alternative random walks
+better reflect item-item similarity. We find that removing item popularity from
+our simulated user data results in better success when mapping between the
+derived concept spaces (as evidenced by the improved precision and recall
+discussed in the next section).
 
-To speed up the simulation, we calculate a series of product-specific stationary
-distributions. As with any random walk technique, we first interpret the
-recommendation graph as a transition matrix. In this walk, for each product we
-create a product-specific transition matrix which has added teleportation edges
-back the that product. We calculate the stationary distributions of each of
-these matrices, one per product. We use these product-specific stationary
-distributions as our user sessions. Since there is only one user session per
-product, we have many fewer sessions than we would with a real dataset. However,
-each of these stationary distributions can be interpreted as an aggregation of
-an infinite number of random walkers starting at each node and travelling until
-a random termination time, as in the original model.
+Next, we reincorporate popularity by modifying how we search for nearest
+neighbors in concept space. Instead of simply ranking neighbors by distance, we
+now rank by distance divided by $\rho$.  In this context, having defined
+popularity as the logorithm of node degree reduces the liklihood that one or two
+very popular items will always appear as the best recommendations when using
+this weighted ranking policy.
 
-This process requires a couple of parameters. The first is the number of steps
-used to approximate each product-specific statinary distribution. The second
-parameter is the likelihood of teleporting to the special product. For our
-purposes we use 30 steps and a teleportation probability of 5\%.
+The third way in which popularity is used is determining the number
+recommendations alotted per product. We assume that a popular product in one
+catalogue is likely to be a popular product in another catalogue. Specifically,
+the number of predicted inter-catalogue recommendations made per item is
+made directly proportional to the item's value of $\rho$.
 
-\subsection*{Singular Value Decomposition}
-Unlike Principal Compenent Analysis, standard LSI does not include any
-preprocessing of the data, such as subtracting the mean from each column of the
-product-session matrix. A major reason for this is that LSI is typically applied
-to natural language processing problems, where the rows are terms, and there is
-one term per word in the vocabulary. This otherwise sparse matrix becomes dense
-if the mean is subtracted, making many NLP datasets untractable to work with. 
-Our datast, on the other hand, has significantly fewer rows and columns and is
-already dense. Therefore, there is little added cost to subtracting the
-means. We include this preprocessing step and find that we get better results
-because of it.
-
-Another design decision is in the size of the truncation of the matrices learned
-from SVD. This number dictates the number of latent concepts learned for LSI.
-Too few concepts fail to capture enough detail to make accurate representations
-of each product. Too many make mapping between concepts challenging. We found a
-good balance at 16 concepts.
-
-[Graphs of concepts]
-
-\subsection*{Concept Mapping}
-A tricky operation of this procedure is the concept mapping. We attempt to find
-the best linear transformation from one concept space to the other. We do this
-by representing each concept as a vector of words in the vocabulary, each
-weighted with how well it represents the concept. We construct this by weighing
-the tf-idf of each word in a product's description by how well that word
-represents the concept. Note that products can have a positive or negative value
-for each concept, thus the word representation of concpets are also positive and
-negative.
-
-With these concept representations, we construct an innner-product matrix. The
-value of a cell (i, j) in the matrix is the dot product of concept i in the
-first space with concept j in the second. The columns and rows of the resulting
-matrix represent the relative similarity of the descriptive words between two
-concepts.  But the scale of this matrix can be very large. If the inner-product
-matrix is used without any normalization as the linear transformation between
-the concept spaces, the scale of the transformed products will be inconsistent
-with those originally in that space.
-
-One way to normalize is to normalize each column of the inner-product matrix by
-the L1 norm. This preserves the L1 norm of products after the transformation,
-but does not preserve the L2 norm. Similarly, we can normalize the columns by
-the L2 norm. Neither of these methods preserve the L2 norm, which would be ideal
-as we find the nearest neighbors in the other concept space by their L2
-distance.
-
-Another approach is to orthogonalize the inner-product matrix. That is, find the
-orthogonal matrix nearest to the inner-product matrix. Nearness is defined by
-the NORM FUNCTION on the difference of the two matrices. This orthogonal matrix
-does perserve the L2 norm of the vectors it transforms, however the
-transformation is an approximation of the ideal inner-product matrix and can
-distort the relative mapping from one concept to the others.
-
-A third approach to normalizing aims to preserve the angle of the inner-product
-transformation. For this approach, we remove the issue of scale by normalizing
-all products in the mapped-to space to be unit lenght. Then, after transforming
-the products from the other catalogue, we normmalize them as well. Thus the
-scales for both sets of products are the same and the angles of the
-inner-product transformation are preserved, at the expense of discarding the
-norms of all products in their product spaces. This is equivalent to calculating
-the cosine similarity between products without any normalization, as cosine
-similarity is indeifferent to scale. We find that this method of normalization
-performs the best.
-
-Further gains can be seen if we consider bigrams in the description as well as
-the unigrams. Here, we treat bigrams as a separate set dimensions to represent
-each concept. These are included in the inner product between two concepts,
-weighted compared to the unigrams by a parameter $\beta$. $\beta = 0.3$ improves the
-performance of the concept mapping.
-
-\subsection*{Popularity}
-The most effective addition to the LSI procedure is incorporating the popularity
-of products. We do not know a priori how popular an item is as we do not have
-explicit user data. However, we can use the recommendation graphs to model the
-popularity of each item. The assumption is that a retailer will recommend
-popular products more frequently than unpopular ones.
-
-We imagine that recommendation graphs are constructed as follows: A retailer first
-builds a similarity matrix, using a collaborative filtering method. In this
-matrix, all similarity relationships are symmetric. Next, the retailer modifies
-the values of the matrix to include popularity. As a simple model, we say that
-the popularity-modified similarity score from Product i to Proudct j is the
-simple similarity score at (i, j) multiplied by the popularity of Product j. In
-this way, more popular items are more frequently recommended.
-
-We see signs of this in the undirected recommendation graphs. Although each product
-has at most four outgoing recommendations on macys.com, we see that some
-products have over one hundred incoming recommendations. With this in mind, we
-approximate the popularity of each product as the logarithm of the degree of
-that product in its undirected recommendation graph. We choose to model it as the
-logarithm because a few of the products tend to have degrees in the hundreds.
-Without some way of scaling back these popularities, these few very popular
-products tend to take over all recommendations.
-
-[Graph of popularity]
-
-Assuming this model of recommendation graph construction, which separates similarity
-from popularity, we seek to isolate similarity from the recommendation graph. Since
-popular products were promoted for popularity, we demote them when walking the
-graphs to create user sessions. In practice, we divide the likelihood of walking
-an edge by the popularity of the product on the other side of that edge. Thus
-the concepts learned on that matrix better represent products independent of
-popularity.
-
-Proceeding as before, we find the nearest products in concept space. However,
-before deeming the nearest neighbors to be the best recommendations, we divide
-the distance to each product by its popularity. Since we model distance in 
-concept space to be inversely proportional to similarity, this reincorporation
-of popularity is analagous to multiplying similarity by popularity, as done in
-the original construction of the recommendation graph.
-
-A final way in which we incorporate popularity is in the number of
-recommendations alotted per product. Since we model the recommendation graph as an
-undirected graph, different nodes have different degrees. A product that is
-popular within its catalogue is likely to be more popular across catalogues as
-well. Therefore, we allow popular products to have more recommendations across
-the two catalogues, proportional to their popularity.
-
-In all, these popularity additions significantly boost performance in the
-experiment. The results of this, along with the other modifications outlined in
-this section, can be seen in the following section on performance.
+In all, these popularity additions significantly boost the performance of the
+\linebreak $LatentFeatureRecommender$. The results of this, along with
+comparisons to the other inter-catalogue  recommeder algorithms described in
+this section, are provided in the following section.
 

### 2014-06-11 paper2/paper.tex
Edits...
Restructure Baselines subsection.

@@
 
-\title{A Technique for Inter-Catalogue Recommendations}
+\title{A Technique for Inter-Retailer Recommendations}
 \author{Bill Chickering and Jamie Irvine\\
@@
 graphical information to discover these latent features in order to choose
-better inter-catalogue associations.
+better inter-catalogue recommendations.
 
@@
 \subsection*{Baselines}
-To understand how well our LSI recommender performs, we need good baselines
-for comparison.  The simplest and easiest to beat baseline is the Random 
-Recommender. This recommender randomly picks pairs of products, one from 
-each catalogue, and recommends them together. This should obviously be the 
-worst performing recommender but it gives us some sense of whether we are 
-doing anything intelligent in our choices of recommendation.
+In addition to the $ContentBasedRecommender$ and the $NeighborhoodRecommender$,
+we constructed and tested several other baseline algorithms to compare the
+$LatentFeatureRecommender$ against. These baselines show that the performance of
+$LatentFeatureRecommender$ is not due to a single mechanism but rather is the
+result of the mult-stage algorithm. In this section, we briefly outline these
+additional baselines.
 
-A more impressive baseline is that of the Simple Content-Based approach, 
-outlined earlier in the paper. This recommender uses information exclusively
-from the descriptions of the products and ignores the recommendation graphs 
-entirely. A third, somewhat similar baseline comes from the Neighborhood 
-Content-based variant, which looks at the content similarity between 
-neighborhoods of products.
+\subsubsection*{\em RandomRecommender}
+The simplest and least challenging baseline. This recommender randomly chooses a
+fixed number of recommendations for each item across catalogues.
 
-We construct a couple of other baselines to show that our performance is not
-exclusively due to any single feature of the recommender. One such baseline
-tests the effects of popularity alone. Incorporating popularity has a large 
-impact on the performance of any recommender and part of this is due simply 
-to the evaluation metric. Any time we connect to a popular product, we are 
-now two away from many more products, increasing the likelihood of getting a
-correct recommendation by the 2-precision standard. To ensure that these kinds 
-of effects are not the sole reason that our recommender performs well, 
-we construct the Popularity Recommender. This recommender simply recommends 
-that every product in one catalogue be connected to the two most popular 
-products from the other catalogue. This baseline shows how much of the LSI 
-recommender's performance is due simply to leveraging popularity and how 
-much is due to other factors, such as intelligently learning and mapping 
-latent concepts.
+\subsubsection*{\em PopularityRecommender}
+A baseline that exclusively exploits popularity $\rho$, without any other
+knowledge of the recommendation graphs or of item content. This recommender
+simply chooses, at random, one of the top three most popular items from the
+other catalogue as recommendations for each item in a catalogue. This baseline
+aims to demonstrate the efficacy of popularity alone.
 
-It is also informative to test the efficacy of the concept mapping, since this
-such a hard task. As in product-product content-similarity, it is difficult to
-capture similarity using descriptions alone. It is not even clear that a mapping 
-necessarily exists between two seperately learned concept spaces. The concepts 
-may represent fundamentally different things, which cannot be well mapped to 
-each other. We can measure the efficacy of the mapping by comparing it to a 
-couple of baselines.
+\subsubsection*{\em RandomMapRecommender}
+This is the $LatentFeatureRecommender$ without the mapping stage. Instead it
+generates a random value between -1.0 and 1.0 for each element of the mapping
+matrix $A$. As such, it is still leveraging popularity $\rho$ in the same way as
+the $LatentFeatureRecommender$. This baseline aims to reveal the importance of
+the mapping stage.
 
-The first baseline is in fact an upper bound of sorts on the mapping. This
-baseline recommender is called the One Model Recommender. Unlike the other
-baselines, this recommender is not a valid one, as it cheats by using
-information from the unpartitioned graph. The One Model Recommender learns a
-latent concept space over the entire recommendation graph by allowing the random
-walkers to traverse it before it is partitioned. In every other respect, it is
-the same as the LSI Recommender. However, since products from both catalogues
-are represented in the same concept space, there is no need for a mapping
-between concept spaces.
+\subsubsection*{\em OneModelRecommender}
+This is not baseline but rather a cheat. This is the $LatentFeatureRecommender$
+except, for this experiment, we defer the graph partitioning until after we have
+performed our random walks and constructed a single concept space. Only then do
+we partition the items such that their latent feature vectors represent all
+items in a single concept space. The goal of this experiment is to determine how
+well our algorithm might do if a perfect concept mapping existed and we could
+somehow compute it.
 
-The second mapping-related baseline is the Random Mapper Recommender. As the
-name suggests, this recommender is identical to the LSI Recommender, except that
-it uses a randomly generated linear transformation to map between concept
-spaces. These two baselines represent (not strict) upper and lower bounds on how
-well the concept mapping performs. The results show how much of a bottleneck
-this step is.
+%Incorporating popularity has a large impact on the performance of any
+%recommender and part of this is due simply to the evaluation metric. Any time we
+%connect to a popular product, we are now two away from many more products,
+%increasing the likelihood of getting a correct recommendation by the 2-precision
+%standard. To ensure that these kinds of effects are not the sole reason that our
+%recommender performs well, we construct the Popularity Recommender. This
+%recommender simply recommends that every product in one catalogue be connected
+%to the two most popular products from the other catalogue. This baseline shows
+%how much of the LSI recommender's performance is due simply to leveraging
+%popularity and how much is due to other factors, such as intelligently learning
+%and mapping latent concepts.
 

### 2014-06-11 paper2/paper.tex
Add first draft abstract.

@@
 \section*{Abstract}
-\emph{Abstracts are abstract. They confront you. There was a reviewer a
-while back who wrote that my papers didn't have any beginning or any end. He
-didn't mean it as a compliment, but it was.} --Jackson Pollock
+\emph{The problem of making recommendation across retailers using only publicly
+available information is explored. Experiments are conducted using real data in
+which several recommender algorithms are compared. A novel latent feature
+inter-retailer recommender algorithm that leverages public intra-retailer
+recommendation information is described and shown to outperform a nontrivial
+content-based approach.}
 
@@
 from the language of their descriptions. For example, a dress's description
-might not capture if it is more conservative or more risque. Yet, this could be
-the salient feature it has in common with an associated recommendation.  These
-sort of abstract, perhaps ineffable, features can be essential to item-item
-relations. We must therefore ask whether we can more fully utilize the given
-graphical information to discover these latent features in order to choose
+might not capture whether it is more conservative or more risque. Yet, this
+could be the salient feature it has in common with an associated recommendation.
+These sort of abstract, perhaps ineffable, features can be essential to
+item-item relations. We must therefore ask whether we can more fully utilize the
+given graphical information to discover these latent features in order to choose
 better inter-catalogue recommendations.

### 2014-06-11 paper2/paper.tex
Add popularity histogram and explanation

--- 2014-06-11 paper2/paper.tex
@@
 from the language of their descriptions. For example, a dress's description
-might not capture if it is more conservative or more risque. Yet, this could be
-the salient feature it has in common with an associated recommendation.  These
-sort of abstract, perhaps ineffable, features can be essential to item-item
-relations. We must therefore ask whether we can more fully utilize the given
-graphical information to discover these latent features in order to choose
+might not capture whether it is more conservative or more risque. Yet, this
+could be the salient feature it has in common with an associated recommendation.
+These sort of abstract, perhaps ineffable, features can be essential to
+item-item relations. We must therefore ask whether we can more fully utilize the
+given graphical information to discover these latent features in order to choose
 better inter-catalogue associations.
@@
 the notion of item {\em popularity}. The idea is that a more popular item is
-more likely to appear as a recommendation independent of how similar it is.  We
-see signs of this in the undirected recommendation graphs from www.macys.com.
-Although each product has at most four outgoing recommendations on its details
-page, we see that some products have over one hundred incoming recommendations.
-But how might we define {\em popularity}? A reasonable definition would
-introduce a quantity that is correlated with the number of users that indicate a
-preference for an item. Like all user data, this is not publicly available
-information.  Once again, however, we can leverage the recommendation graph to
-infer an approximate popularity metric for each item.
+more likely to be recommended, independent of how similar it is. We see signs of
+this in the undirected recommendation graphs from www.macys.com. Although each
+product has at most four outgoing recommendations on its details page, we see
+that some products have over one hundred incoming recommendations. But how
+might we define {\em popularity}? A reasonable definition would introduce a
+quantity that is correlated with the number of users that indicate a preference
+for an item. Like all user data, this is not publicly available information.
+Once again, however, we can leverage the recommendation graph to infer an
+approximate popularity metric for each item.
 
@@
 similarity. We do this by adjusting the probability of transitioning to an item
-on the graph by making it inversely proportional to that item's popularity
+on the graph by making it inversely proportional to that item's popularity,
 $\rho$. Stationary distributions learned from these alternative random walks
@@
 made directly proportional to the item's value of $\rho$.
+
+This assumption is corroborated in the data. The following histogram shows the
+number of gold-standard recommendations connected to the top 25 most popular
+items, ranked by their value of $\rho$. The histogram also compares the
+distribution of predicted recommendations to these items by the
+$LatentFeatureRecomender$ with and without popularity taken into account. With
+popularity, the predictions follow a similar distribution to the gold-standard
+recommendations over these popular items.
+
+\begin{figure}[!htbp]
+    \centering
+    \includegraphics[width=1.0\textwidth]{PopularityHistogram.png}
+	\caption{The top 25 most popular items and how many times they appear in
+    recommendations. Here we compare the distribution of gold-standard
+    recommendations (red) to the recommendations predicted by the
+    $LatentFeatureRecommender$ with (blue) and without (yellow) popularity. The
+    yellow bars are small and barely visible.}
+    \label{fig:PopularityHistogram}
+\end{figure}
+
+Note that without popularity, the recommendations are scattered more uniformly
+over the thousands of items in the two catalogues, and so the number of times
+these 25 items are recommended is significantly smaller (hardly visible in the
+plot). This is because the recommendations are soley based on the similarity in
+latent concept space and popular items are not inherently more similar to other
+items. This recommender in fact still performs better than the 
+$ContentBasedRecommender$, however it fails to exploit this trend in the data.
 

### 2014-06-11 paper2/paper.tex
Add paper outline at end of intro.

--- 2014-06-11 paper2/paper.tex
@@
 
-In this report. . . . (BREIFLY OUTLINE THE PAPER HERE.)
+This report is organized as follows. In the next next section we describe our
+dataset and some simple preprocessing that was performed. Next, we outline an
+experiment used to simulate the inter-retailer recommendation problem. This is
+followed by a section that explains and justifies our evaluation metrics.
+Several pages are then dedicated to the details of a novel latent feature
+algorithm along with two nontrivial primarily content-based, baseline
+recommenders. We then summarize the results of our experiments with these
+recommenders along with several additional baseline algorithms.
 

### 2014-06-11 paper2/paper.tex
Tweak discussion of popularity plot.

--- 2014-06-11 paper2/paper.tex
@@
 catalogue is likely to be a popular product in another catalogue. Specifically,
-the number of predicted inter-catalogue recommendations made per item is
+the number of predicted inter-catalogue recommendations predicted per item is
 made directly proportional to the item's value of $\rho$.
+
+This assumption is corroborated by the data. The following histogram shows the
+number of gold-standard recommendations associated with the top 25 most popular
+items, ranked by their value of $\rho$. The histogram also includes the
+distribution of predicted recommendations for these items by the
+$LatentFeatureRecommender$ both with and without popularity mechanisms. With
+popularity, the predictions follow a similar distribution to that of the
+gold-standard recommendations. Without popularity, the recommendations are
+scattered more uniformly over the thousands of items of each catalogue such that
+the number of times these 25 items are recommended is significantly smaller
+(barely visible in the plot). This is because the recommendations are based
+solely on similarity in concept space (after mapping) and popular items are not
+inherently more similar to other items. Importantly, the
+$LatentFeatureRecommender$ {\em without popularity} still outperforms the
+$ContentBasedRecommender$ demonstrating that while popularity is important, it
+is only one of several features that useful in choosing good inter-catalogue
+recommendations.
+
+\begin{figure}[!htbp]
+    \centering
+    \includegraphics[width=1.0\textwidth]{PopularityHistogram.png}
+	\caption{The top 25 most popular items and how many times they appear in
+    recommendations. Here we compare the distribution of gold-standard
+    recommendations (red) to the recommendations predicted by the
+    $LatentFeatureRecommender$ with (blue) and without (yellow) popularity. The
+    yellow bars are small and barely visible.}
+    \label{fig:PopularityHistogram}
+\end{figure}
 

### 2014-06-12 paper2/paper.tex
Add images of representative items for concepts

--- 2014-06-11 paper2/paper.tex
@@
 
-This report is organized as follows. In the next next section we describe our
-dataset and some simple preprocessing that was performed. Next, we outline an
-experiment used to simulate the inter-retailer recommendation problem. This is
-followed by a section that explains and justifies our evaluation metrics.
-Several pages are then dedicated to the details of a novel latent feature
-algorithm along with two nontrivial primarily content-based, baseline
-recommenders. We then summarize the results of our experiments with these
-recommenders along with several additional baseline algorithms.
+In this report. . . . (BREIFLY OUTLINE THE PAPER HERE.)
 
@@
 
+Figures 1 and 2 show the top and bottom most relevant items for two of the sixteen
+concepts learned for the Women's Dresses category. Note that concepts learned by
+SVD range from positive to negative, representing the two extremes of the
+concept.
+
+\begin{figure}
+\centering
+\begin{subfigure}{.18\textwidth}
+\centering
+\includegraphics[width=\linewidth]{concepts/concept1_pos1.jpg}
+\label{fig:sub1}
+\end{subfigure}%
+\begin{subfigure}{.18\textwidth}
+\centering
+\includegraphics[width=\linewidth]{concepts/concept1_pos2.jpg}
+\label{fig:sub2}
+\end{subfigure}
+\begin{subfigure}{.18\textwidth}
+\centering
+\includegraphics[width=\linewidth]{concepts/concept1_pos3.jpg}
+\label{fig:sub2}
+\end{subfigure}
+\begin{subfigure}{.18\textwidth}
+\centering
+\includegraphics[width=\linewidth]{concepts/concept1_pos4.jpg}
+\label{fig:sub2}
+\end{subfigure}
+\begin{subfigure}{.18\textwidth}
+\centering
+\includegraphics[width=\linewidth]{concepts/concept1_pos5.jpg}
+\label{fig:sub2}
+\end{subfigure}
+\subcaption{The top 5 items, representing the positive direction}
+
+\centering
+\begin{subfigure}{.18\textwidth}
+\centering
+\includegraphics[width=\linewidth]{concepts/concept1_neg1.jpg}
+\label{fig:sub1}
+\end{subfigure}%
+\begin{subfigure}{.18\textwidth}
+\centering
+\includegraphics[width=\linewidth]{concepts/concept1_neg2.jpg}
+\label{fig:sub2}
+\end{subfigure}
+\begin{subfigure}{.18\textwidth}
+\centering
+\includegraphics[width=\linewidth]{concepts/concept1_neg3.jpg}
+\label{fig:sub2}
+\end{subfigure}
+\begin{subfigure}{.18\textwidth}
+\centering
+\includegraphics[width=\linewidth]{concepts/concept1_neg4.jpg}
+\label{fig:sub2}
+\end{subfigure}
+\begin{subfigure}{.18\textwidth}
+\centering
+\includegraphics[width=\linewidth]{concepts/concept1_neg5.jpg}
+\label{fig:sub2}
+\end{subfigure}
+\subcaption{The bottom 5 items, representing the negative direction}
+\label{fig:conceptA}
+\title{Figure 1: The top and bottom items representing a concept in the Dresses
+category.  This concept could be characterized as modern and tiered in the
+positive direction, and gown-like and elegant in the negative direction.}
+\end{figure}
+
+\begin{figure}
+\centering
+\begin{subfigure}{.18\textwidth}
+\centering
+\includegraphics[width=\linewidth]{concepts/concept2_pos1.jpg}
+\label{fig:sub1}
+\end{subfigure}%
+\begin{subfigure}{.18\textwidth}
+\centering
+\includegraphics[width=\linewidth]{concepts/concept2_pos2.jpg}
+\label{fig:sub2}
+\end{subfigure}
+\begin{subfigure}{.18\textwidth}
+\centering
+\includegraphics[width=\linewidth]{concepts/concept2_pos3.jpg}
+\label{fig:sub2}
+\end{subfigure}
+\begin{subfigure}{.18\textwidth}
+\centering
+\includegraphics[width=\linewidth]{concepts/concept2_pos4.jpg}
+\label{fig:sub2}
+\end{subfigure}
+\begin{subfigure}{.18\textwidth}
+\centering
+\includegraphics[width=\linewidth]{concepts/concept2_pos5.jpg}
+\end{subfigure}
+\subcaption{The top 5 items, representing the positive direction}
+
+\centering
+\begin{subfigure}{.18\textwidth}
+\centering
+\includegraphics[width=\linewidth]{concepts/concept2_neg1.jpg}
+\label{fig:sub1}
+\end{subfigure}%
+\begin{subfigure}{.18\textwidth}
+\centering
+\includegraphics[width=\linewidth]{concepts/concept2_neg2.jpg}
+\label{fig:sub2}
+\end{subfigure}
+\begin{subfigure}{.18\textwidth}
+\centering
+\includegraphics[width=\linewidth]{concepts/concept2_neg3.jpg}
+\label{fig:sub2}
+\end{subfigure}
+\begin{subfigure}{.18\textwidth}
+\centering
+\includegraphics[width=\linewidth]{concepts/concept2_neg4.jpg}
+\label{fig:sub2}
+\end{subfigure}
+\begin{subfigure}{.18\textwidth}
+\centering
+\includegraphics[width=\linewidth]{concepts/concept2_neg5.jpg}
+\label{fig:sub2}
+\end{subfigure}
+\subcaption{The bottom 5 items, representing the negative direction}
+\title{Figure 2: The top and bottom items representing another different Dresses
+concept. This concept could be characterized as two-toned and playful in the
+positive direction, and belted and light in the negative direction.}
+\label{fig:conceptB} \end{figure}
+
 \subsubsection*{Concept Space Mapping}
@@
     $LatentFeatureRecommender$ with (blue) and without (yellow) popularity. The
-    yellow bars are small and barely visible.}
+    yellow bars are small and barely visible. This plot was generated on the
+    Women's Dresses category.}
     \label{fig:PopularityHistogram}

### 2014-06-12 paper2/paper.tex
Add results table and graph

@@
 \subsection*{Results}
-[Number of products mapped to in different methods]
-\begin{center}
-\begin{tabular}{ | l | c | c | c |}
-\hline
-Normalization & 2-Precision & 2-Recall & 2-F1 Score \\ \hline\hline
-L1 normalize each column of matrix &&&\\ \hline
-L2 normalize each column of matrix &&&\\ \hline
-Orthogonalize matrix &&&\\ \hline
-Normalize products in concept space &&&\\ \hline
-\end{tabular}
-\end{center}
 
 \begin{center}
-\begin{tabular}{ | l | c | c | c |}
+\begin{tabular}{ | l || c | c | c || c | c | c |}
 \hline
-Method & 2-Precision & 2-Recall & 2-F1 Score \\ \hline\hline
-Random &&&\\ \hline
-Simple Content-based &&&\\ \hline
-Neighborhood Content-based &&&\\ \hline
-One Model &&&\\ \hline
-Random concept mapping &&&\\ \hline
-Popularity &&&\\ \hline
-LSI &&&\\ \hline
-\hspace*{0.5cm} + Normalized concept mapping &&&\\ \hline
-\hspace*{0.5cm} + Zero-mean SVD &&&\\ \hline
-\hspace*{0.5cm} + Bigrams from descriptions &&&\\ \hline
-\hspace*{0.5cm} + Discounted popularity from random walks &&&\\ \hline
-\hspace*{0.5cm} + Popularity for finding nearest neighbors &&&\\ \hline
-\hspace*{0.5cm} + Number of recommendations based on popularity &&&\\ \hline
+Recommender & Precision & Recall & F1 Score & 2-Precision & 2-Recall & 2-F1 Score \\ \hline\hline
+$RandomRecommender$ & .002 & .001 & .001 & .021 & .008 & .012 \\ \hline
+$PopularityRecommender$ & .016 & .019 & .017 & .080 & .096 & .087 \\ \hline
+$RandomMapRecommender$ & .014 & .019 & .016 & .070 & .095 & .081 \\ \hline
+$OneModelRecommender$ & .103 & .132 & .116 & .404 & .410 & .407 \\ \hline
+$ContentBasedRecommender$ & .014 & .020 & .016 & .115 & .186 & .142 \\ \hline
+$NeighborhoodRecommneder$ & .040 & .060 & .048 & .137 & .240 & .175 \\ \hline
+$LatentFeatureRecommender$ & .036 & .060 & .045 & .174 & .228 & .197 \\ \hline
 
@@
 
-\begin{center}
-\begin{tabular}{ | l | c | c | c | c | c | c |}
-\hline
-Category & 1-Precision & 1-Recall & 1-F1 Score & 2-Precision & 2-Recall & 2-F1 Score \\ \hline\hline
-Men's Shirts &&&&&&\\ \hline
-Women's Dresses &&&&&&\\ \hline
-\end{tabular}
-\end{center}
-
-Analysis/descriptions for each of these tables
+\begin{figure}[!htbp]
+    \centering
+    \includegraphics[width=0.65\textwidth]{RecommenderPerformance.png}
+	\caption{The performance of the three main recommenders in terms of
+    2-metrics across all categories. We emphasize the 2-F1 Score as our most
+    important single metric of comparison. We see that the $LatentFeatureRecommender$
+    outperforms the other recommenders in this score.} 
+    \label{fig:RecommenderPerformance} 
+\end{figure}
 

### 2014-06-12 paper2/paper.tex
Add first draft conclusion.

--- 2014-06-12 paper2/paper.tex
@@
 \\\text{\em 2-}F_1 &= 2\cdot \frac{\text{\em 2-precision} \cdot \text{\em
-2-recall}}{\text{\em 2-precision} + \text{\em 2-recall}}
+2-recall}}{\text{\em 2-precision} + \text{\em 2-recall}},
 \end{align}
@@
 
-Figures 1 and 2 show the top and bottom most relevant items for two of the sixteen
-concepts learned for the Women's Dresses category. Note that concepts learned by
-SVD range from positive to negative, representing the two extremes of the
-concept.
-
-\begin{figure}
-\centering
-\begin{subfigure}{.18\textwidth}
-\centering
-\includegraphics[width=\linewidth]{concepts/concept1_pos1.jpg}
-\label{fig:sub1}
-\end{subfigure}%
-\begin{subfigure}{.18\textwidth}
-\centering
-\includegraphics[width=\linewidth]{concepts/concept1_pos2.jpg}
-\label{fig:sub2}
-\end{subfigure}
-\begin{subfigure}{.18\textwidth}
-\centering
-\includegraphics[width=\linewidth]{concepts/concept1_pos3.jpg}
-\label{fig:sub2}
-\end{subfigure}
-\begin{subfigure}{.18\textwidth}
-\centering
-\includegraphics[width=\linewidth]{concepts/concept1_pos4.jpg}
-\label{fig:sub2}
-\end{subfigure}
-\begin{subfigure}{.18\textwidth}
-\centering
-\includegraphics[width=\linewidth]{concepts/concept1_pos5.jpg}
-\label{fig:sub2}
-\end{subfigure}
-\subcaption{The top 5 items, representing the positive direction}
-
-\centering
-\begin{subfigure}{.18\textwidth}
-\centering
-\includegraphics[width=\linewidth]{concepts/concept1_neg1.jpg}
-\label{fig:sub1}
-\end{subfigure}%
-\begin{subfigure}{.18\textwidth}
-\centering
-\includegraphics[width=\linewidth]{concepts/concept1_neg2.jpg}
-\label{fig:sub2}
-\end{subfigure}
-\begin{subfigure}{.18\textwidth}
-\centering
-\includegraphics[width=\linewidth]{concepts/concept1_neg3.jpg}
-\label{fig:sub2}
-\end{subfigure}
-\begin{subfigure}{.18\textwidth}
-\centering
-\includegraphics[width=\linewidth]{concepts/concept1_neg4.jpg}
-\label{fig:sub2}
-\end{subfigure}
-\begin{subfigure}{.18\textwidth}
-\centering
-\includegraphics[width=\linewidth]{concepts/concept1_neg5.jpg}
-\label{fig:sub2}
-\end{subfigure}
-\subcaption{The bottom 5 items, representing the negative direction}
-\label{fig:conceptA}
-\title{Figure 1: The top and bottom items representing a concept in the Dresses
-category.  This concept could be characterized as modern and tiered in the
-positive direction, and gown-like and elegant in the negative direction.}
-\end{figure}
-
-\begin{figure}
-\centering
-\begin{subfigure}{.18\textwidth}
-\centering
-\includegraphics[width=\linewidth]{concepts/concept2_pos1.jpg}
-\label{fig:sub1}
-\end{subfigure}%
-\begin{subfigure}{.18\textwidth}
-\centering
-\includegraphics[width=\linewidth]{concepts/concept2_pos2.jpg}
-\label{fig:sub2}
-\end{subfigure}
-\begin{subfigure}{.18\textwidth}
-\centering
-\includegraphics[width=\linewidth]{concepts/concept2_pos3.jpg}
-\label{fig:sub2}
-\end{subfigure}
-\begin{subfigure}{.18\textwidth}
-\centering
-\includegraphics[width=\linewidth]{concepts/concept2_pos4.jpg}
-\label{fig:sub2}
-\end{subfigure}
-\begin{subfigure}{.18\textwidth}
-\centering
-\includegraphics[width=\linewidth]{concepts/concept2_pos5.jpg}
-\end{subfigure}
-\subcaption{The top 5 items, representing the positive direction}
-
-\centering
-\begin{subfigure}{.18\textwidth}
-\centering
-\includegraphics[width=\linewidth]{concepts/concept2_neg1.jpg}
-\label{fig:sub1}
-\end{subfigure}%
-\begin{subfigure}{.18\textwidth}
-\centering
-\includegraphics[width=\linewidth]{concepts/concept2_neg2.jpg}
-\label{fig:sub2}
-\end{subfigure}
-\begin{subfigure}{.18\textwidth}
-\centering
-\includegraphics[width=\linewidth]{concepts/concept2_neg3.jpg}
-\label{fig:sub2}
-\end{subfigure}
-\begin{subfigure}{.18\textwidth}
-\centering
-\includegraphics[width=\linewidth]{concepts/concept2_neg4.jpg}
-\label{fig:sub2}
-\end{subfigure}
-\begin{subfigure}{.18\textwidth}
-\centering
-\includegraphics[width=\linewidth]{concepts/concept2_neg5.jpg}
-\label{fig:sub2}
-\end{subfigure}
-\subcaption{The bottom 5 items, representing the negative direction}
-\title{Figure 2: The top and bottom items representing another different Dresses
-concept. This concept could be characterized as two-toned and playful in the
-positive direction, and belted and light in the negative direction.}
-\label{fig:conceptB} \end{figure}
-
 \subsubsection*{Concept Space Mapping}
@@
     $LatentFeatureRecommender$ with (blue) and without (yellow) popularity. The
-    yellow bars are small and barely visible. This plot was generated on the
-    Women's Dresses category.}
+    yellow bars are small and barely visible.}
     \label{fig:PopularityHistogram}
@@
 \section*{Conclusion}
-We tried a few different methods for creating recommendations between
-retailers. We find that this one is the best.
-The last method uses an unconventional approach that adds some value in the
-process. These results could be used in real life for these problems, since all
-of the data exists.
+In this study, we address the problem of determining inter-retailer, or
+inter-catalogue, item-item associations. We frame this problem as a recommender
+problem, which motivates a relatively straightforward experiment in which a
+candidate algorithm must connect two recommendation graphs by guessing edges
+between them.
 
-\section*{Future Work}
-Incorporate more data such as customer reviews. Use computer vision techniques.
-Focus more on popular hubs model.
+In this work, we introduced a novel technique for predicting item-item
+associations across retailers. This technique employs {\em Topic-Sensitive
+PageRank} on a publicly available recommendation graph to generate simulated
+user data. With this data, we perform {\em Latent Semantec Indexing} to learn a
+latent feature space for items from in a particular online catalogue. By
+generalizing the notion of {\em Term Frequency---Inverse Document Frequency}
+score, we construct proxy tf-idf vectors for the latent dimensions, or {\em
+concepts}, associated with a particular catalogue. This, in turn, allows us to
+construct a transformation matrix such that vectors in one catalogue's concept
+space can be transformed into another catalogue's concept space. Inter-catalogue
+associations can then be made by finding pairs of items that are a minimal
+distance apart in a common concept space.
 
-\section*{Appendix}
+This technique is improved upon by incorporating the notion of item {\em
+popularity} $\rho$, which we define in this context to be the logarithm of the
+item's degree within its associated recommendation graph. We show that $\rho$
+can be used in multiple ways to improve the performance of an inter-retailer
+recommender system.
+
+Finally, we compared our latent inter-catalogue recommender system to several
+nontrival baseline algorithms and find that it outperforms all of them in both
+standard precision and recall of gold-standard inter-catalogue edges as well as
+another combination of informative metrics we call {\em 2-precision} and {\em
+2-recall}.
+
+Making accurate item-item associations across retailers using only publicly
+available information is a challenging problem. The ability to accurately
+identity these relationships, however, could potentially provide a retailer or
+other service provider with valuable business intelligence.
 
@@
 
+\section*{Appendix}
+
 \end{document}

### 2014-06-12 paper2/paper.tex
Add pagebreak before conclusion.
Clean table

--- 2014-06-12 paper2/paper.tex
@@
 Recommender & Precision & Recall & F1 Score & 2-Precision & 2-Recall & 2-F1 Score \\ \hline\hline
-$RandomRecommender$ & .002 & .001 & .001 & .021 & .008 & .012 \\ \hline
-$PopularityRecommender$ & .016 & .019 & .017 & .080 & .096 & .087 \\ \hline
-$RandomMapRecommender$ & .014 & .019 & .016 & .070 & .095 & .081 \\ \hline
-$OneModelRecommender$ & .103 & .132 & .116 & .404 & .410 & .407 \\ \hline
-$ContentBasedRecommender$ & .014 & .020 & .016 & .115 & .186 & .142 \\ \hline
-$NeighborhoodRecommneder$ & .040 & .060 & .048 & .137 & .240 & .175 \\ \hline
-$LatentFeatureRecommender$ & .036 & .060 & .045 & .174 & .228 & .197 \\ \hline
+$Random$ & .002 & .001 & .001 & .021 & .008 & .012 \\ \hline
+$Popularity$ & .016 & .019 & .017 & .080 & .096 & .087 \\ \hline
+$RandomMap$ & .014 & .019 & .016 & .070 & .095 & .081 \\ \hline
+$OneModel$ & .103 & .132 & .116 & .404 & .410 & .407 \\ \hline
+$ContentBased$ & .014 & .020 & .016 &
+\textbf{.115} & \textbf{.186} & \textbf{.142} \\ \hline 
+$Neighborhood$ & .040 & .060 & .048 &
+\textbf{.137} & \textbf{.240} & \textbf{.175} \\ \hline 
+$LatentFeature$ & .036 & .060 & .045 &
+\textbf{.174} & \textbf{.228} & \textbf{.197} \\ \hline
 

### 2014-06-12 paper2/paper.tex

--- 2014-06-12 paper2/paper.tex
@@
 
+\pagebreak
 \section*{Conclusion}

### 2014-06-12 paper2/paper.tex
Made plot labels bigger

@@
     \centering
-    \includegraphics[width=1.0\textwidth]{PopularityHistogram.png}
+    \includegraphics[width=1.0\textwidth]{PopularityHistogram2.png}
 	\caption{The top 25 most popular items and how many times they appear in
@@
 
+\begin{table}
 \begin{center}
@@
 \textbf{.174} & \textbf{.228} & \textbf{.197} \\ \hline
-
 \end{tabular}
+\caption{The performance of each of the recommenders across all categories. The
+first five are the baseline recommenders outlined in the previous section. The
+final three are the recommenders of interest, described in the Recommender
+Algorithms section. The scores in bold are the most important results and are 
+further illustrated in the following figure.}
 \end{center}
-
+\end{table}
 \begin{figure}[!htbp]
     \centering
-    \includegraphics[width=0.65\textwidth]{RecommenderPerformance.png}
+    \includegraphics[width=0.65\textwidth]{RecommenderPerformance2.png}
 	\caption{The performance of the three main recommenders in terms of
@@
 \end{figure}
-
 \pagebreak
@@
 
-\section*{Appendix}
-
 \end{document}

### 2014-06-12 paper2/paper.tex
Make minor tweaks/corrections.

--- 2014-06-12 paper2/paper.tex
@@
 \section*{Abstract}
-\emph{The problem of making recommendation across retailers using only publicly
+\emph{The problem of making recommendations across retailers using only publicly
 available information is explored. Experiments are conducted using real data in
 which several recommender algorithms are compared. A novel latent feature
-inter-retailer recommender algorithm that leverages public intra-retailer
+inter-retailer recommender system that leverages public intra-retailer
 recommendation information is described and shown to outperform a nontrivial
@@
 inter-retailer user data, a traditional CF technique is ostensibly not
-applicable. Instead, we propose a new technique that leverages the recomender
-graphs as well as the content from each product to create inter-retail
+applicable. Instead, we propose a new technique that leverages the recommender
+graphs as well as the content from each product to create inter-retailer
 recommendations.

### 2014-06-12 paper2/paper.tex
BIGGER
Add paragraph to Results section.

@@
 
-%Incorporating popularity has a large impact on the performance of any
-%recommender and part of this is due simply to the evaluation metric. Any time we
-%connect to a popular product, we are now two away from many more products,
-%increasing the likelihood of getting a correct recommendation by the 2-precision
-%standard. To ensure that these kinds of effects are not the sole reason that our
-%recommender performs well, we construct the Popularity Recommender. This
-%recommender simply recommends that every product in one catalogue be connected
-%to the two most popular products from the other catalogue. This baseline shows
-%how much of the LSI recommender's performance is due simply to leveraging
-%popularity and how much is due to other factors, such as intelligently learning
-%and mapping latent concepts.
-
 \subsection*{Results}
@@
 \end{figure}
+
+Table (1) shows the results of all inter-catalogue recommenders across all
+categories. For example, the {\em precision} column lists the sum, across all
+categories, of correctly predicted edges divided by the sum, across all
+categories, of predicted edges. We first point out that both $Neighborhood$ and
+$LatentFeature$ outperform $ContentBased$ by all metrics. This makes clear the
+utility of leveraging information from the recommendation when making
+inter-catalogue recommendations. Next, we notice that $Neighborhood$ actually
+outperforms $LatentFeature$ in {\em precision} by 10\% and in {\em 2-recall} by
+about 5\%. Importantly, however, $LatentFeature$ bests $Neighborhood$ in the key
+metric {\em 2-}$F_1$ by about 12\%.
+
 \pagebreak

### 2014-06-12 paper2/paper.tex
Tweak abstract

--- 2014-06-12 paper2/paper.tex
@@
 \section*{Abstract}
-\emph{The problem of making recommendations across retailers using only publicly
-available information is explored. Experiments are conducted using real data in
-which several recommender algorithms are compared. A novel latent feature
-inter-retailer recommender system that leverages public intra-retailer
+\emph{The problem of making recommendation across retailers using only publicly
+available information is explored. Several inter-retailer recommender algortihms
+are presented. Each is compared in an experiment conducted with real data. A 
+novel latent feature recommender that leverages public intra-retailer
 recommendation information is described and shown to outperform a nontrivial
@@
     \centering
-    \includegraphics[width=0.65\textwidth]{RecommenderPerformance2.png}
+    \includegraphics[width=0.75\textwidth]{RecommenderPerformance2.png}
 	\caption{The performance of the three main recommenders in terms of

### 2014-06-12 paper2/paper.tex
Edits up to Concept Mapping

@@
 \subsubsection*{Extracting Latent Features}
-We may now construct a synthetic $N \times N$ user-item matrix $M$ with rows
+We may now construct a synthetic $N \times N$ item-user matrix $M$ with rows
 representing items and columns representing users, which are in fact the
 stationary distributions discovered from our random walks. This matrix has at
-least one notable advantage over most real user-item matrices---it is not
+least one notable advantage over most real item-user matrices---it is not
 sparse.  Rather, each synthetic user has a finite value associated with every
@@
 scenario, we may therefore employ SVD as our matrix factorization technique
-since our user-item matrix is virtually complete (i.e. nearly void of empty
-entries).  Using SVD, we factor our user-item matrix as
+since our item-user matrix is virtually complete (i.e. nearly void of empty
+entries).  Using SVD, we factor our item-user matrix as
 \begin{align}
@@
 Figures 1 and 2 show the top and bottom most relevant items for two of the sixteen
-concepts learned for the Women's Dresses category. Note that concepts learned by
-SVD range from positive to negative, representing the two extremes of the
-concept.
+concepts learned in this way on the Women's Dresses category. Note that concepts
+learned by SVD range from positive to negative, representing the two extremes of
+the concept, therefore it is relevant to look at both the top and bottom items.
 

### 2014-06-12 paper2/paper.tex
Add paragraphs to Results section.

--- 2014-06-12 paper2/paper.tex
@@
 
+\pagebreak
 \section*{Performance}
@@
 
+\pagebreak
 \subsection*{Results}
+Table (1) shows the results of all inter-catalogue recommenders across all
+categories. For example, the {\em precision} column lists the sum, across all
+categories, of correctly predicted edges divided by the sum, across all
+categories, of predicted edges. We first point out that both $Neighborhood$ and
+$LatentFeature$ outperform $ContentBased$ by all metrics. This makes clear the
+utility of leveraging information from the recommendation when making
+inter-catalogue recommendations. Next, we notice that $Neighborhood$ actually
+outperforms $LatentFeature$ in {\em precision} by 10\% and in {\em 2-recall} by
+about 5\%. Importantly, however, $LatentFeature$ bests $Neighborhood$ in the key
+metric {\em 2-}$F_1$ by about 12\%.
+
+Of course, $Random$ does quite poorly. Interestingly, $Popularity$ and
+$RandomMap$ have comparable performance and score almost half as well as
+$LatentFeature$. The fact that their results are similar is not suprising,
+however, since they are both relying entirely on item popularity $\rho$.
+$Popularity$ utilizes $\rho$ explicity by connecting each item to one of the
+most popular items in the other catalogue. $RandomMap$, on the otherhand,
+employs $\rho$ in an identical manner to that by $LatentFeature$. This reveals
+that popularity alone accounts for nearly half of $LatentFeature$'s success.
 
@@
 
-Table (1) shows the results of all inter-catalogue recommenders across all
-categories. For example, the {\em precision} column lists the sum, across all
-categories, of correctly predicted edges divided by the sum, across all
-categories, of predicted edges. We first point out that both $Neighborhood$ and
-$LatentFeature$ outperform $ContentBased$ by all metrics. This makes clear the
-utility of leveraging information from the recommendation when making
-inter-catalogue recommendations. Next, we notice that $Neighborhood$ actually
-outperforms $LatentFeature$ in {\em precision} by 10\% and in {\em 2-recall} by
-about 5\%. Importantly, however, $LatentFeature$ bests $Neighborhood$ in the key
-metric {\em 2-}$F_1$ by about 12\%.
+As expected, we find that $OneModel$ outperforms $LatentFeature$. Indeed, one
+might ask why $OneModel$ does not score even higher. The performance of
+$OneModel$ is limited by the rank reduction stage of the algorithm, which
+discards a large amount of information. $LatentFeature$ benefits from this stage
+since by discarding dimensions, the subsequent mapping problem becomes more
+tractable. Nonetheless, it is interesting that $OneModel$ outperforms
+$LatentFeature$ by a about a factor of 2. Given that $OneModel$ is a cheat that
+involves perfectly mapped concept spaces, we can consider it a loose upperbound
+for $LatentFeature$ (at least when working with 16 dimensions). It might,
+therefore, be possible to signficantly improve the performance of
+$LatentFeature$ by improving its concept space mapping stage.  At the same time,
+it is not at all clear that such a mapping in general even exists. Rather, the
+common existence of an approximate mapping is a hypothesis of the authors that
+is corroborated by our experimental results for the present dataset.
 
-\pagebreak
 \section*{Conclusion}

### 2014-06-12 paper2/paper.tex
Small edits

--- 2014-06-12 paper2/paper.tex
@@
 recommendation graphs. Nonetheless, our hypothesis is that a relationship
-between these concept spaces often will exist due to the effectiveness of the
+between these concept spaces will often exist due to the effectiveness of the
 underlying CF techniques that generate the graphs to capture meaningful
@@
 inter-catalogue edges between item pairs that are nearest to one another in
-concept space. To learn these mapping, we must revisit the notion of content
+concept space. To learn these mappings, we must revisit the notion of content
 similarity. Once again, we utilize tf-idf vectors. This time, however, we weight
@@
 feature vector $\vec{v}(i)$. In this way, we construct a dense {\em
-super-tf-idf} vector for each concept for each catalogue. Formally, we say that
+super-tf-idf} vector for each concept of each catalogue. Formally, we say that
 a concept $j$ of a catalogue $C$ is represented by the super-tf-idf vector
@@
 therefore, does not preserve vector norms. We address this issue by normalizing
-all feature vectors, both foreign and domestic, in the concept space following
-the transformation. This radially projects all items onto the surface of sphere
-within the 16-dimensional concept space. Now when ranking an item's neighbors by
-distance, we are actually ranking neighbors by angular separation. That is, we
-are ranking neighbors by their cosine similarity in concept space. After
-experimenting with various normalization schemes we find that this approach
-works best.
+all feature vectors, both foreign and domestic, in the concept space followingj
+the transformation. This radially projects all items onto the surface of a
+hypersphere in the 16-dimensional concept space. Now, when ranking an item's
+neighbors by distance, we are actually ranking neighbors by angular separation.
+That is, we are ranking neighbors by their cosine similarity in concept space.
+After experimenting with various normalization schemes we find that this
+approach works best.
 

### 2014-06-12 paper2/paper.tex
Edits - bigger ticks

@@
 neighbors in concept space. Instead of simply ranking neighbors by distance, we
-now rank by distance divided by $\rho$.  In this context, having defined
-popularity as the logorithm of node degree reduces the liklihood that one or two
-very popular items will always appear as the best recommendations when using
-this weighted ranking policy.
+now rank by distance divided by $\rho$, thus promoting popular items.  In this
+context, having defined popularity as the logorithm of node degree reduces the
+liklihood that one or two very popular items will always appear as the best
+recommendations when using this weighted ranking policy.
 
-The third way in which popularity is used is determining the number
+The third way in which popularity is used is determining the number of
 recommendations alotted per product. We assume that a popular product in one
@@
 number of gold-standard recommendations associated with the top 25 most popular
-items, ranked by their value of $\rho$. The histogram also includes the
+items, as determinded by $\rho$. The histogram also includes the
 distribution of predicted recommendations for these items by the
@@
 $LatentFeatureRecommender$ {\em without popularity} still outperforms the
-$ContentBasedRecommender$ demonstrating that while popularity is important, it
-is only one of several features that useful in choosing good inter-catalogue
+$ContentBasedRecommender$, demonstrating that while popularity is important, it
+is only one of several features that are useful in choosing good inter-catalogue
 recommendations.

### 2014-06-12 paper2/paper.tex
Minor corrections.

--- 2014-06-12 paper2/paper.tex
@@
 \emph{The problem of making recommendation across retailers using only publicly
-available information is explored. Several inter-retailer recommender algortihms
+available information is explored. Several inter-retailer recommender algorithms
 are presented. Each is compared in an experiment conducted with real data. A 
@@
 \subsubsection*{Extracting Latent Features}
-We may now construct a synthetic $N \times N$ item-user matrix $M$ with rows
+We may now construct a synthetic $N \times N$ user-item matrix $M$ with rows
 representing items and columns representing users, which are in fact the
 stationary distributions discovered from our random walks. This matrix has at
-least one notable advantage over most real item-user matrices---it is not
+least one notable advantage over most real user-item matrices---it is not
 sparse.  Rather, each synthetic user has a finite value associated with every
@@
 scenario, we may therefore employ SVD as our matrix factorization technique
-since our item-user matrix is virtually complete (i.e. nearly void of empty
-entries).  Using SVD, we factor our item-user matrix as
+since our user-item matrix is virtually complete (i.e. nearly void of empty
+entries).  Using SVD, we factor our user-item matrix as
 \begin{align}
@@
 Figures 1 and 2 show the top and bottom most relevant items for two of the sixteen
-concepts learned in this way on the Women's Dresses category. Note that concepts
-learned by SVD range from positive to negative, representing the two extremes of
-the concept, therefore it is relevant to look at both the top and bottom items.
+concepts learned for the Women's Dresses category. Note that concepts learned by
+SVD range from positive to negative, representing the two extremes of the
+concept.
 
@@
 recommendation graphs. Nonetheless, our hypothesis is that a relationship
-between these concept spaces will often exist due to the effectiveness of the
+between these concept spaces often will exist due to the effectiveness of the
 underlying CF techniques that generate the graphs to capture meaningful
@@
 inter-catalogue edges between item pairs that are nearest to one another in
-concept space. To learn these mappings, we must revisit the notion of content
+concept space. To learn these mapping, we must revisit the notion of content
 similarity. Once again, we utilize tf-idf vectors. This time, however, we weight
@@
 feature vector $\vec{v}(i)$. In this way, we construct a dense {\em
-super-tf-idf} vector for each concept of each catalogue. Formally, we say that
+super-tf-idf} vector for each concept for each catalogue. Formally, we say that
 a concept $j$ of a catalogue $C$ is represented by the super-tf-idf vector
@@
 where $i$ is a concept from $C_1$ and $j$ is a concept from $C_2$ and $\beta$ is
-the same parameters as before (again, we find $\beta=0.3$ works well).  We may
+the same parameter as before (again, we find $\beta=0.3$ works well).  We may
 now use this matrix to transform the vector representation of an item in one
@@
 therefore, does not preserve vector norms. We address this issue by normalizing
-all feature vectors, both foreign and domestic, in the concept space followingj
-the transformation. This radially projects all items onto the surface of a
-hypersphere in the 16-dimensional concept space. Now, when ranking an item's
-neighbors by distance, we are actually ranking neighbors by angular separation.
-That is, we are ranking neighbors by their cosine similarity in concept space.
-After experimenting with various normalization schemes we find that this
-approach works best.
+all feature vectors, both foreign and domestic, in the concept space following
+the transformation. This radially projects all items onto the surface of sphere
+within the 16-dimensional concept space. Now when ranking an item's neighbors by
+distance, we are actually ranking neighbors by angular separation. That is, we
+are ranking neighbors by their cosine similarity in concept space. After
+experimenting with various normalization schemes we find that this approach
+works best.
 
@@
 now rank by distance divided by $\rho$.  In this context, having defined
-popularity as the logorithm of node degree reduces the liklihood that one or two
-very popular items will always appear as the best recommendations when using
-this weighted ranking policy.
+popularity as the logorithm of node degree, instead of simply node degree,
+reduces the liklihood that one or two very popular items will always appear as
+the best recommendations when using this weighted ranking policy.
 
-The third way in which popularity is used is determining the number
+The third way in which popularity is used is determining the number of
 recommendations alotted per product. We assume that a popular product in one

### 2014-06-12 paper2/paper.tex

--- 2014-06-12 paper2/paper.tex
@@
 now rank by distance divided by $\rho$, thus promoting popular items.  In this
-context, having defined popularity as the logorithm of node degree reduces the
-liklihood that one or two very popular items will always appear as the best
-recommendations when using this weighted ranking policy.
+context, having defined popularity as the logorithm of node degree, instead of
+simply node degree, reduces the liklihood that one or two very popular items
+will always appear as the best recommendations when using this weighted ranking
+policy.
 

### 2014-06-12 paper2/paper.tex
Final edits

--- 2014-06-12 paper2/paper.tex
@@
 
-Of course, $Random$ does quite poorly. Interestingly, $Popularity$ and
-$RandomMap$ have comparable performance and score almost half as well as
-$LatentFeature$. The fact that their results are similar is not suprising,
-however, since they are both relying entirely on item popularity $\rho$.
-$Popularity$ utilizes $\rho$ explicity by connecting each item to one of the
-most popular items in the other catalogue. $RandomMap$, on the otherhand,
-employs $\rho$ in an identical manner to that by $LatentFeature$. This reveals
-that popularity alone accounts for nearly half of $LatentFeature$'s success.
+Of course, $Random$ does quite poorly, emphasizing the difficulty of the
+problem. Interestingly, $Popularity$ and $RandomMap$ have comparable performance
+and score almost half as well as $LatentFeature$. The fact that their results
+are similar is not suprising, however, since they are both relying entirely on
+item popularity $\rho$.  $Popularity$ utilizes $\rho$ explicity by connecting
+each item to one of the most popular items in the other catalogue. $RandomMap$,
+on the otherhand, employs $\rho$ in an identical manner to that by
+$LatentFeature$. This reveals that popularity alone accounts for nearly half of
+$LatentFeature$'s success.
 
@@
 tractable. Nonetheless, it is interesting that $OneModel$ outperforms
-$LatentFeature$ by a about a factor of 2. Given that $OneModel$ is a cheat that
+$LatentFeature$ by about a factor of 2. Given that $OneModel$ is a cheat that
 involves perfectly mapped concept spaces, we can consider it a loose upperbound
@@
 user data. With this data, we perform {\em Latent Semantec Indexing} to learn a
-latent feature space for items from in a particular online catalogue. By
-generalizing the notion of {\em Term Frequency---Inverse Document Frequency}
-score, we construct proxy tf-idf vectors for the latent dimensions, or {\em
-concepts}, associated with a particular catalogue. This, in turn, allows us to
-construct a transformation matrix such that vectors in one catalogue's concept
-space can be transformed into another catalogue's concept space. Inter-catalogue
-associations can then be made by finding pairs of items that are a minimal
-distance apart in a common concept space.
-
-This technique is improved upon by incorporating the notion of item {\em
-popularity} $\rho$, which we define in this context to be the logarithm of the
-item's degree within its associated recommendation graph. We show that $\rho$
-can be used in multiple ways to improve the performance of an inter-retailer
-recommender system.
+latent feature space for items in a particular online catalogue. By generalizing
+the notion of {\em Term Frequency---Inverse Document Frequency}, we construct
+proxy tf-idf vectors for the latent dimensions, or {\em concepts}, associated
+with a particular catalogue. This, in turn, allows us to construct a
+transformation matrix such that vectors in one catalogue's concept space can be
+transformed into another catalogue's concept space. Inter-catalogue associations
+can then be made by finding pairs of items that are a minimal distance apart in
+a common concept space. This technique is improved upon by incorporating the
+notion of item {\em popularity} $\rho$, which we define in this context to be
+the logarithm of the item's degree within its associated recommendation graph.
+We show that $\rho$ can be used in multiple ways to improve the performance of
+an inter-retailer recommender system.
 
 Finally, we compared our latent inter-catalogue recommender system to several
-nontrival baseline algorithms and find that it outperforms all of them in both
+nontrival baseline algorithms. We showed that it outperforms all of them in
 standard precision and recall of gold-standard inter-catalogue edges as well as
-another combination of informative metrics we call {\em 2-precision} and {\em
-2-recall}.
+in two more informative metrics we call {\em 2-precision} and {\em 2-recall}.
 
@@
 available information is a challenging problem. The ability to accurately
-identity these relationships, however, could potentially provide a retailer or
+identify these relationships, however, could potentially provide a retailer or
 other service provider with valuable business intelligence.

### 2014-06-12 paper2/paper.tex
Minor corrections.

--- 2014-06-12 paper2/paper.tex
@@
 
-We begin with an attempt to seprate the notion of item similarity from that of
+We begin with an attempt to separate the notion of item similarity from that of
 item popularity. The idea that more popular items are more likely to be
@@
 now rank by distance divided by $\rho$, thus promoting popular items.  In this
-context, having defined popularity as the logorithm of node degree, instead of
+context, having defined popularity as the logarithm of node degree, instead of
 simply node degree, reduces the liklihood that one or two very popular items
@@
 
-In this work, we introduced a novel technique for predicting item-item
-associations across retailers. This technique employs {\em Topic-Sensitive
-PageRank} on a publicly available recommendation graph to generate simulated
-user data. With this data, we perform {\em Latent Semantec Indexing} to learn a
-latent feature space for items from in a particular online catalogue. By
-generalizing the notion of {\em Term Frequency---Inverse Document Frequency}
-score, we construct proxy tf-idf vectors for the latent dimensions, or {\em
-concepts}, associated with a particular catalogue. This, in turn, allows us to
-construct a transformation matrix such that vectors in one catalogue's concept
-space can be transformed into another catalogue's concept space. Inter-catalogue
-associations can then be made by finding pairs of items that are a minimal
-distance apart in a common concept space.
+We introduced a novel technique for predicting item-item associations across
+retailers. This technique employs {\em Topic-Sensitive PageRank} on a publicly
+available recommendation graph to generate simulated user data. With this data,
+we perform {\em Latent Semantec Indexing} to learn a latent feature space for
+items in a particular online catalogue. By generalizing the notion of {\em Term
+Frequency---Inverse Document Frequency} score, we construct proxy tf-idf vectors
+for the latent dimensions, or {\em concepts}, associated with a particular
+catalogue. This, in turn, allows us to construct a transformation matrix such
+that vectors in one catalogue's concept space can be transformed into another
+catalogue's concept space. Inter-catalogue associations can then be made by
+finding pairs of items that are a minimal distance apart in a common concept
+space.
 

### 2014-06-12 paper2/paper.tex

--- 2014-06-12 paper2/paper.tex
@@
 
-In this work, we introduced a novel technique for predicting item-item
-associations across retailers. This technique employs {\em Topic-Sensitive
-PageRank} on a publicly available recommendation graph to generate simulated
-user data. With this data, we perform {\em Latent Semantec Indexing} to learn a
-latent feature space for items in a particular online catalogue. By generalizing
-the notion of {\em Term Frequency---Inverse Document Frequency}, we construct
-proxy tf-idf vectors for the latent dimensions, or {\em concepts}, associated
-with a particular catalogue. This, in turn, allows us to construct a
-transformation matrix such that vectors in one catalogue's concept space can be
-transformed into another catalogue's concept space. Inter-catalogue associations
-can then be made by finding pairs of items that are a minimal distance apart in
-a common concept space. This technique is improved upon by incorporating the
-notion of item {\em popularity} $\rho$, which we define in this context to be
-the logarithm of the item's degree within its associated recommendation graph.
-We show that $\rho$ can be used in multiple ways to improve the performance of
-an inter-retailer recommender system.
+We introduced a novel technique for predicting item-item associations across
+retailers. This technique employs {\em Topic-Sensitive PageRank} on a publicly
+available recommendation graph to generate simulated user data. With this data,
+we perform {\em Latent Semantec Indexing} to learn a latent feature space for
+items in a particular online catalogue. By generalizing the notion of {\em Term
+Frequency---Inverse Document Frequency}, we construct proxy tf-idf vectors for
+the latent dimensions, or {\em concepts}, associated with a particular
+catalogue. This, in turn, allows us to construct a transformation matrix such
+that vectors in one catalogue's concept space can be transformed into another
+catalogue's concept space. Inter-catalogue associations can then be made by
+finding pairs of items that are a minimal distance apart in a common concept
+space. This technique is improved upon by incorporating the notion of item {\em
+popularity} $\rho$, which we define in this context to be the logarithm of the
+item's degree within its associated recommendation graph.  We show that $\rho$
+can be used in multiple ways to improve the performance of an inter-retailer
+recommender system.
 

### 2014-06-12 paper2/paper.tex
Moded last sentence

--- 2014-06-12 paper2/paper.tex
@@
 Making accurate item-item associations across retailers using only publicly
-available information is a challenging problem. The ability to accurately
-identify these relationships, however, could potentially provide a retailer or
-other service provider with valuable business intelligence.
+available information is a challenging problem. The problem is relevent,
+however, and the ability to accurately identify these relationships could 
+provide a retailer or other service provider with valuable business
+intelligence.
 
